### 2018-12-17 paper2018/paper.tex
paper stub

\documentclass{article}
\usepackage{todonotes}

\title{FraCaS: 80 percent complete}

\begin{document}

% - Two-phase system (Dyn semantics from theorem proving)
% - Anaphora (as in Jolli paper) --- but this is the first time that this is done for FraCas.
% - Multiple readings
% - Threshold-based interpretation of (some) adjectives + Linear arithmetic proofs
% - New handling of prepositions and adverbs
% - Definites handled by adding an assumption at the top-level. (instead as inline existential) -- Enabled by dyn. semantics
% - Plurals/Quantifiers improved
% - Genders are properties of nouns
% - Handling of comparatives (with numbers)

\begin{figure}
  \centering
{
  \tiny
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3.7cm,
                    semithick]
  \tikzstyle{every state}=[draw=black,text=black,shape=rectangle,text width=2cm]

  \node[state]         (txt)                {text};
  \node[state]         (ast) [right of=txt] {AST};
  \node[state]         (pp)  [right of=ast] {Probabilistic program};
  \node[state]         (res) [right of=pp] {result in $(0,1)$};

  \path (txt) edge              node (GF)  {GF}                  (ast)
        (ast) edge              node (Hask) [text width=1.5cm]{Haskell interpreter} (pp)
        (pp) edge               node (Coq) {Coq} (res);

  \node (Syntax) [above of=GF,node distance=1cm] {Syntax};
  \node (Sem) [above of=Hask,node distance=1cm] {Dynamic Semantics};
  \node (Hyper) [above of=Coq,node distance=1cm] {Additional Semantics$^\ast$};

  \path
  (Syntax) edge (GF)
  (Sem) edge (Hask)
  (Hyper) edge (Coq);
\end{tikzpicture}
}  
\caption{Phases in our system. ($\ast$) At the level of Coq, we handle
  the details of the adverbial (veridicality properties) and
  adjectival semantics (division into subsective, extentional,
  non-committal, etc. categories.) }
  \label{fig:overview}
\end{figure}

\end{document}

### 2019-01-15 MS.hs
comment code

@@ -1436,22 +1436,30 @@ data Quant = PossQuant Pron | UniversalQuant Pol | IndefQuant | ExistQuant | ETy
 possPron :: Pron -> Quant
 possPron = PossQuant
 
+-- | given a quantity θ of cn in the env, return, exactly θ cn
+-- (In the example, "exactly θ orders")
 elliptic_NP_Pl :: NP
 elliptic_NP_Pl = do
   (amount,cn') <- getQuantity
   return $ MkNP [ExactlyPredet] (Cardinal amount) IndefQuant cn'
 
+-- Consider the example "ITEL won more orders than APCOM did"
+-- f(x) = AtLeast (x+1)
+-- cn = orders
+-- s = APCOM <elliptic_VP>
 relativeAmountQuant :: Logic.Pol -> (Nat -> Amount) -> CN -> S -> NP
 relativeAmountQuant pol f cn s = do
   (cn',gender) <- cn
-  threshold <- getFresh
-  modify (pushQuantity threshold (cn',gender)) -- priming "ellptic_NP_Pl"
-  modify (pushVP (complSlash elliptic_VPSlash elliptic_NP_Pl)) -- in 's', there may be an elliptic vp, referring to this
-  s' <- s
+  threshold <- getFresh -- invent an abstract quantity.
+  modify (pushQuantity threshold (cn',gender)) -- priming "ellptic_NP_Pl", so that we find the above quantity
+  modify (pushVP (complSlash elliptic_VPSlash elliptic_NP_Pl)) -- in 's', there is an elliptic vp, referring to this
+  s' <- s -- in our example: s evaluates to "APCOM <elliptic_VP>", then "APCOM (<elliptic_VPSlash> <elliptic_NP_Pl>)", then "APCOM won exactly θ orders"
   let q :: Quant'
       q _num _cn _role = do
         x <- getFresh
-        return (\vp' extraObjs -> noExtraObjs s' ∧ Quant (f (Nat (Var threshold))) pol x (noExtraObjs (cn' (Var x))) (vp' (Var x) extraObjs))
+        return (\vp' extraObjs -> noExtraObjs s' ∧ -- "APCOM won exactly θ orders"
+                                  Quant (f (Nat (Var threshold))) pol x (noExtraObjs (cn' (Var x))) (vp' (Var x) extraObjs)) -- Itel won at least θ+1 orders
+      -- quantifier that implements "there exists (f threshold)"
[…]

### 2019-01-18 paper2018/paper.tex
packages
paper structure
sec. 2

@@
 \usepackage{todonotes}
+\usepackage{tikz}
+\usetikzlibrary{arrows,automata}
+
+\usepackage[backend=bibtex,citestyle=authoryear,style=alphabetic,natbib]{biblatex}
+\addbibresource{PaperTools/bibtex/jp.bib}
+\addbibresource{phorae.bib}
 
 \title{FraCaS: 80 percent complete}
-
-\begin{document}
 
@@
 % - Genders are properties of nouns
-% - Handling of comparatives (with numbers)
+% - Handling of comparatives (with threshold BUT not those of adjectives), relying largely on the dyn. semantic system
+
+\begin{document}
+
+\section{Introduction}
+
+\paragraph{Contributions}
+(see list above)
+\section{Background}
+
+\section{Overview of the system}
+
+Our system consists of two main parts.
+\begin{enumerate}
+\item A converter from syntax trees to types. The syntax trees follow
+  the GF formalism, and the types follow the Coq formalism. The
+  converter itself is a Haskell program. It implements a dynamic
+  semantics. It comprises the bigger part of our system.
+\item A number of type-theoretical combinators. These take care of the
+  natural semantics which have no influence on the dynamical
+  aspects. This includes for example the treatment of adjectives
+  (intersective, subsective, etc.) and adverbs (veridical or not).
+\end{enumerate}
+This architecture is represented schematically in Figure \ref{fig:overview}.
 
@@
 
+All underlying systems (GF, Haskell, Coq) are based on lambda calculi
+with types. We ensure that each translation preserve typing, locally:
+\begin{enumerate}
+\item Every GF syntactic category $C$ is mapped to a type noted $⟦C⟧$.
+\item GF Functional types are mapped compositionally : $⟦A → B⟧ = ⟦A⟧ → ⟦B⟧$
+\item Every GF syntactic construction function $f : X$ is mapped to a function $⟦f⟧$ such that $⟦f⟧ : ⟦X⟧$.
+\item GF function applications are mapped compositionally: $⟦t(u)⟧ = ⟦t⟧ (⟦u⟧)$.
+\end{enumerate}
+Because all system embed the simply-typed lambda calculus, ensuring
+type-preservation locally means that types are preserved globally.
+Therefor, we are certain that every GF syntax tree can be mapped to
+Haskell, and eventuall Coq, without error.
+
+The dynamic semantics follows a monadic structure, as pioneered by
+\citet{Shan:2002}. There are two kinds of effects carried by the
+monad.  The first one comprises a series of state updates and queries,
+corresponding to all elements that can be referred to by anaphoric
+expressions. These can be usual ones (like NPs), but also less usual
+ones (like 2-place verbs, or a quantity --- which we illustrate
+below). The other kind of effects is \emph{non-determinism}. This
+models the property that linguistic expressions can have several
+interpretations. The monadic structure allows to locally express that
+a given expression has several meaning; the monadic bind ensures that
+all combinations of meanings are considered at the top-level,
+combinatorially. This dynamic semantics allows us to model many
+phenomena in a precise way:
+
+\paragraph{anaphora} We can handle anaphora % wait so see what JoLLi says here
+
+  % - Dynamic Semantics/Anaphora (as in Jolli paper) --- but this is the first time that this is done for FraCas.
+%   -> Gender is a property of nouns
+\paragraph{ellipsis} Ellipsis is handled in essentially the same way
+as anaphora. This method is made especially straightforward thanks to
+using GF syntax tree, which requires an explicit argument for each
+predicate. Thus, ellpsis are made explicit by the parsing phase.
+
+\paragraph{definites} A naive way to handle definites is as sigma (or
+exsistential types). \todo{add example} However, if the semantics
+does not feature a dynamic element, then the existential
+quantification is introduced locally. This means that the quantifier
+can be introduced in the wrong context. Consider the phrase
+``everyone pets the dog''. The structure of the interpretation would
+be be $∀x. person(x) → ∃y. dog(y) ∧ pet(x,y)$.
+
+Instead, our take is that definites should be treated as anaphoric
+expressions. If the referent is not found in the discourse, then it
+will be introduced using a sigma type, \emph{at the top-level} of
+the expression. Therefore we record all definites without referent,
+using a monadic effect. When the complete sentence has been
+interpreted, we re-introduced referent using existentials. For the
+above example, we obtain the desired interpretation:
+be be $∃y. dog(y) ∧ ∀x. person(x) → pet(x,y)$.
+
+\paragraph{Phrasal comparatives}
+Previous attempts to tackle section of the FraCaS test suite showed
+that handling comparatives is not easy. Our strategy here is to
+leverage our dynamic semantics, revealing an anaphoric element of
+comparatives. Consider the hypothesis of FraCaS 239: ``ITEL won more
+orders than APCOM lost.''  We postulate that this sentence is
+equivalent to the following two separate parts: ``APCOM lost zero or
+more orders. ITEL won more orders [than some elliptic quantity].''
+The elliptic quantity is introduced every time we talk about some
+quantity (indexed by a CN, in this case ``orders''), and it can be
+referred to by a comparative, later in the discourse. Using this idea,
+we can go one level deeper in the interpretation of our example:
+``APCOM lost θ orders. θ >= 0.  ITEL won at least θ+1 orders.''. We
+see here how the quantities are introduced. They are added to the
+environment so that, they can be referred to as elliptic quantity
+expressions. Finally, ``more'' is intepreted as ``at least <elliptic
+quantity>+1''.
+
+\paragraph{Adjectives}
+In our system, we interpret gradable adjectives using a pair of a
+measure $m : objects → Z$ and and a threshold $τ : Z$, where $Z$ is
+treated as an abstract ordered ring by Coq. (This structure has no
+dynamic aspect in our model, and thus is entirely handled within Coq.)
+For subsective adjectives, $τ$ will additionally depend on the class
+of the object in question. This structure has the benefit that
+opposite adjectives can be easily represented (measures are opposites
+$∀x. m_1 x = -m_2 x$ and thresholds do not overlap $τ_1 + τ_2 > 0$).
+\todo{reference to this view in the litterature}.  Additionally,
+adjectival assumptions as given in FraCaS are interpreted as linear
+equations in $Z$. Because solving such equations is decidable, solving
+such problems can be made efficient. Indeed, the tactic that Coq
+offers for this purpose can solve all such problems in FraCaS,
+automatically.
+
+\paragraph{Adverbs}
+Another point, of minor theoretical importance but major practical
+one, is our handling of adverbial phrases. We interpret adverbs (and
+in general all adverbial and prepositional phrases) are VP-modifiers:
+$Adv = VP -> VP$, where $VP = object -> Prop$. However, applying
+adverbs to verb-phrases heavily complicates the Coq proofs, because
+such phrases can contain quantifiers. Therefore, we instead move the
+adverbs, so that they apply to (atomic) verbs only, and can simplify
+proofs accordingly. \todo{theoretical justification in the literature}
+
+\section{Results and evaluation}
+% - show overall numbers
+% - compare with others
+% - error analysis (classify, show examples)
+
+\section{Future work}
+\section{Conclusion}
+
+\printbibliography
+
 \end{document}

### 2019-01-18 paper2018/paper.tex
typography
sec. 3

@@
 \addbibresource{phorae.bib}
+\addbibresource{IWCSbib.bib}
+
+\usepackage{stmaryrd}
+\usepackage{newunicodechar}
+\input{PaperTools/latex/newunicodedefs}
 
@@
 in general all adverbial and prepositional phrases) are VP-modifiers:
-$Adv = VP -> VP$, where $VP = object -> Prop$. However, applying
+$Adv = VP → VP$, where $VP = object → Prop$. However, applying
 adverbs to verb-phrases heavily complicates the Coq proofs, because
@@
 \section{Results and evaluation}
-% - show overall numbers
-% - compare with others
-% - error analysis (classify, show examples)
+
+We evaluated FraCoq against 8 sections of the FraCaS test suite, for a
+total of 259 examples. We excluded only one section, 7. The reason for
+doing so is that, in our view, it contains too many examples which
+require \textit{ad-hoc} treatment, and thus makes little sense to
+include without complementing it with a more thorough suite which
+captures a more complete landscape of the phenomena that section 7
+touches.
+
+We classify as YES if a proof can be constructed from the premises to
+the hypothesis, NO if a proof of the negated hypothesis can be
+constructed and UNK otherwise.\footnote{Because we work with a non-decidable logic, that poses a theoretical question that we address below.}
+
+%  ------ Section 1
+% complete: 74/74   []
+% correct: 71/74   [13,14,46]
+% score: 0.9594594594594594
+%  ------ Section 2
+% complete: 33/33   []
+% correct: 27/33   [84,91,94,95,109,111]
+% score: 0.8181818181818182
+%  ------ Section 3
+% complete: 27/28   [137]
+% correct: 24/27   [124,126,127]
+% score: 0.8571428571428571
+%  ------ Section 4
+% complete: 47/52   [171,172,191,193,195]
+% correct: 45/47   [170,188]
+% score: 0.8653846153846154
+%  ------ Section 5
+% complete: 20/20   []
+% correct: 19/20   [216]
+% score: 0.95
+%  ------ Section 6
+% complete: 29/31   [244,245]
+% correct: 27/29   [242,250]
+% score: 0.8709677419354839
+%  ------ Section 7
+% complete: 0/67   [251,252,253,254,255,258,259,260,261,262,263,264,265,266,267,268,269,270,271,272,273,274,275,277,278,279,280,281,282,283,284,285,286,287,288,289,290,292,293,294,295,296,297,298,299,300,301,302,303,304,306,307,311,312,313,314,315,316,317,318,319,320,321,322,323,324,325]
+% correct: 0/0   []
+% score: 0.0
+%  ------ Section 8
+% complete: 8/8   []
+% correct: 6/8   [326,333]
+% score: 0.75
+%  ------ Section 9
+% complete: 13/13   []
+% correct: 12/13   [346]
+% score: 0.9230769230769231
+%  ------ Section 0
+% complete: 251/259   [137,171,172,191,193,195,244,245]
+% correct: 231/251   [13,14,46,84,91,94,95,109,111,124,126,127,170,188,216,242,250,326,333,346]
+% score: 0.8918918918918919
+
+\todo{rejected and changed cases in FraCaS}
+\providecommand\forcecenter{\multicolumn{1}{c}}
+\begin{table}
+  \centering
+\begin{tabular}{rlrrrrrr}
+  & Section      & \# examples & Ours      & FraCoq & MINE & Nut  & Langpro  \\ \hline
+1 & Quantifiers  & 75          & .96 (74)  & .96    & .77  & .53  & .93 (44) \\
+2 & Plurals      & 33          & .82       & .76    & .67  & .52  & .73 (24) \\
+3 & Anaphora     & 28          & .86       &   -    & -    & -    &  -       \\
+4 & Ellipsis     & 52          & .87       &   -    & -    & -    &  -       \\
+5 & Adjectives   & 22          & .95 (20)  & .95    & .68  & .32  & .73 (12) \\
+6 & Comparatives & 31          & .87       & .56    & .48  & .45  &  -       \\
+8 & Verbs        & 8           & .75       &   -    & -    & -    &  -       \\
+9 & Attitudes    & 13          & .92       & .85    & .77  & .46  & .92 (9)  \\ \hline
+  & Total        & 262         & .89 (259) & .83    & .69  & .50  & .85 (89)
+  \end{tabular}
+  \caption{Overview of the accuracy of our system compared to others.
+    ``Ours" refers to the approach presented in this paper. When a
+    system does not handle the nominal number of examples shown in the
+    third column, the actual number of examples attempted is shown in
+    parentheses.  ``FraCoq'' refers to the work of
+    \citet{bernardy_type_2017}. ``MINE" refers to the approach of
+    \citet{Mineshima:2015}, ``NUT" to the CCG system that utilizes the
+    first-order automated theorem prover \textit{nutcracker} described
+    by \citet{bos:2008}, and ``Langpro" to the system presented by
+    \citet{Abzianidze:2015}.  }
+\end{table}
+
+The above result reveals a considerable improvement over earlier
+approaches in terms of coverage, with three more sections covered over
+previous state of the art. Additionally, our system is generally best
+in terms of accuracy overall. In particular, section 6 sees a large
+improvement in accuracy, which we attribute to our new analysis of
+comparatives, enabled by our dynamic semantics.
+
+\todo{error analysis}
 

### 2019-01-18 paper2018/paper.tex
typography

@@
 9 & Attitudes    & 13          & .92      & .85    & .77  & .46  & .92 (9)  \\ \hline
-  & Total        & 262         & .89 (259) & .83    & .69  & .50  & .85 (89)
+  & Total        & 262         & .89      & .83    & .69  & .50  & .85  \\
+  &              &             & (259)    & (174)  & (174)& (174)& (89)
   \end{tabular}

### 2019-01-18 paper2018/paper.tex
clarification on decidability

@@
 the hypothesis, NO if a proof of the negated hypothesis can be
-constructed and UNK otherwise.\footnote{Because we work with a non-decidable logic, that poses a theoretical question that we address below.}
+constructed and UNK otherwise.\footnote{Because we work with a
+  non-decidable logic, one can not \emph{in general} conclude
+  decisively that no proof exists. Thus, we consider here that no
+  proof exists if it cannot be constructed with reasonable effort. In
+  particular, we test at the minimum that the automatic proof search
+  built in Coq does not succeed before classifying a problem as UNK.}
 

### 2019-01-18 paper2018/paper.tex
overruled cases

--- 2019-01-18 paper2018/paper.tex
@@
 
-We classify as YES if a proof can be constructed from the premises to
-the hypothesis, NO if a proof of the negated hypothesis can be
-constructed and UNK otherwise.\footnote{Because we work with a non-decidable logic, that poses a theoretical question that we address below.}
+FraCaS intends to classify each problem as either entailment (YES),
+entailment of the opposite (NO) or no entailment (UNK).  For this
+work, we have amended FraCaS to address a few problems. First, certain
+test case are not formed correctly. Those were identified by Bill
+MacCartney as ``undef'', and we removed those. Second, a few test
+cases occurs twice in the suite, but with two different answers (one
+YES and one UNK), with the annotation that those correspond to
+different readings. However, elsewhere in the suite, if a problem has
+several readings but only one has entailment, it occurs only once and
+is marked as YES. To make the test suite consistent, if one reading
+yields entailment we have always considered it as YES. Finally, we
+have removed case 199 (which appears to be vacuous), and changed the
+following cases, which appeared to have been misclassified. We note
+that the majority of the mistaken classifications occur in sections 3
+and 4, which have not been previously attempted and thus, we propose,
+have not been properly scrutinized.
+
+\begin{center}
+\begin{tabular}{ccc}
+ test case & new class & comment \\ \hline
+    005 & UNK & missing hypothesis: there are italian tenors \\
+   056 &  Yes & Already identified as such by MacCartney \\
+  069 & Unk & Mary could have used someone else's workstation \\
+  119 & Unk & \textit{ibid.} \\
+  181 & Yes & for the same reason as 180 \\
+  226 & Yes &
+\end{tabular}
+\end{center}
+
+Our system classifies a case as YES if a proof can be constructed from
+the premises to the hypothesis, NO if a proof of the negated
+hypothesis can be constructed and UNK otherwise. Because we work with
+a non-decidable logic, one can not \emph{in general} conclude
+decisively that no proof exists. Thus, we consider here that no proof
+exists if it cannot be constructed with reasonable effort. In
+particular, we test at the minimum that the automatic proof search
+built in Coq does not succeed before classifying a problem as UNK.
 
@@
 
-\todo{rejected and changed cases in FraCaS}
 \providecommand\forcecenter{\multicolumn{1}{c}}

### 2019-01-18 paper2018/paper.tex
error analysis

@@
 
-\todo{error analysis}
+\paragraph{error analysis}
+Our system fails to classify correctly 28 problems of of 259 attempts.
+Unfortunately, it isn't easy to concisely characterize the areas where
+it fails.  The biggest source of error can be attributed to its
+incorrect handling of groups
+(013,014,046,084,111,124,126,127,137,171,172,191,193,195,243,250,333,346). These
+are cases where a conjunction of individuals is treated as group, or
+where counting members of a group is necessary.  Other notable
+problematic cases are that definite plural does not have universal
+reading are incorrectly classified (091,094,095). Neither measure
+phrases (242) nor attributive comparatives (244,245) are handled.
 

### 2019-01-21 paper2018/paper.tex
revision

--- 2019-01-18 paper2018/paper.tex
@@
 \usepackage{stmaryrd}
+\usepackage{enumitem}
 \usepackage{newunicodechar}
@@
 % - Handling of comparatives (with threshold BUT not those of adjectives), relying largely on the dyn. semantic system
+
+\newlist{lingex}{enumerate}{3}
+\setlist[lingex,1]{parsep=0pt,itemsep=1pt,label=(\arabic*),resume=lingexcount}
+\newcommand\onelingex[1]{\begin{lingex}\item #1 \end{lingex}}
 
@@
 \citet{Shan:2002}. There are two kinds of effects carried by the
-monad.  The first one comprises a series of state updates and queries,
-corresponding to all elements that can be referred to by anaphoric
-expressions. These can be usual ones (like NPs), but also less usual
-ones (like 2-place verbs, or a quantity --- which we illustrate
-below). The other kind of effects is \emph{non-determinism}. This
-models the property that linguistic expressions can have several
+monad.  The first one comprises a series of updates and queries of
+stateful elements.  There is one piece of updateable state for every
+element which can be referred to by anaphoric expressions. These can
+be the usual ones (like NPs), but also less usual ones (like 2-place
+verbs, or a quantity --- which we illustrate below). The other kind of
+effects is \emph{non-determinism}. We use non-determinism to model the
+property that linguistic expressions can have several
 interpretations. The monadic structure allows to locally express that
-a given expression has several meaning; the monadic bind ensures that
+a given expression has several meanings; the monadic bind ensures that
 all combinations of meanings are considered at the top-level,
 combinatorially. This dynamic semantics allows us to model many
-phenomena in a precise way:
+phenomena in a precise way.
 
-\paragraph{anaphora} We can handle anaphora % wait so see what JoLLi says here
+\paragraph{anaphora} Thanks to the above system, we can handle many
+anaphoric cases, including E-Type and Donkey pronouns. Indeed, even
+objects which have no syntactic representation can be added to the
+environment. We follow here the general pattern outlined by
+\citet{unger:2011}. However, we use a more general scope-extension
+mechanism, which allows us to support examples such as the following:
+\item[FraCaS 122]
+P1	Every committee has a chairman.\\
+P2	He is appointed its members.\\
+H 	Every committee has a chairman appointed by members of the committee. 
+\end{lingex}
+In the above example, the pronoun ``he'' is allowed to refer to the
+object quantified over by ``every'', whose scope is extended
+accordingly.
 
-  % - Dynamic Semantics/Anaphora (as in Jolli paper) --- but this is the first time that this is done for FraCas.
-%   -> Gender is a property of nouns
 \paragraph{ellipsis} Ellipsis is handled in essentially the same way
@@
 using GF syntax tree, which requires an explicit argument for each
-predicate. Thus, ellpsis are made explicit by the parsing phase.
+predicate. Thus, ellpsis are made explicit by the parsing phase. Such
+ellptic expressions are handled in the same way as anaphora.
+\begin{lingex}
+\item[FraCas 173]
+  P1	Bill spoke to everyone that John did [elliptic V2].\\
+  P2	John spoke to Mary.\\
+  Q 	Did Bill speak to Mary?\\
+  H 	Bill spoke to Mary.
+\end{lingex}
+For example, in (FraCaS 173, P1), the argument of ``did'' is
+explicitly marked as an elliptic V2, which we resolve to ``speak'' in
+that context.
+\paragraph{definites} A naive way to handle definites is an
+existential (or sigma). However, if the semantics does not feature a
+dynamic element, then the existential quantification is introduced
+locally. This means that the quantifier can be introduced in the wrong
+context. Consider the phrase ``everyone pets the dog''. The structure
+of the interpretation would be
+$∀x. person(x) → ∃y. dog(y) ∧ pet(x,y)$.
 
-\paragraph{definites} A naive way to handle definites is as sigma (or
-exsistential types). \todo{add example} However, if the semantics
-does not feature a dynamic element, then the existential
-quantification is introduced locally. This means that the quantifier
-can be introduced in the wrong context. Consider the phrase
-``everyone pets the dog''. The structure of the interpretation would
-be be $∀x. person(x) → ∃y. dog(y) ∧ pet(x,y)$.
-
-Instead, our take is that definites should be treated as anaphoric
-expressions. If the referent is not found in the discourse, then it
-will be introduced using a sigma type, \emph{at the top-level} of
-the expression. Therefore we record all definites without referent,
-using a monadic effect. When the complete sentence has been
-interpreted, we re-introduced referent using existentials. For the
+Instead, our take is that definites should be treated as an anaphoric
+expression with an implicit referrent. That is, if the referent is not
+found in the discourse, then it will be forcibly introduced, using an
+existential type, \emph{at the top-level} of the expression. To be
+able to do this, we record all definites without referent, using
+another portion of the environment (using a monadic effect). For the
 above example, we obtain the desired interpretation:
-be be $∃y. dog(y) ∧ ∀x. person(x) → pet(x,y)$.
+$∃y. dog(y) ∧ (∀x. person(x) → pet(x,y))$.
 
 \paragraph{Phrasal comparatives}
-Previous attempts to tackle section of the FraCaS test suite showed
-that handling comparatives is not easy. Our strategy here is to
-leverage our dynamic semantics, revealing an anaphoric element of
-comparatives. Consider the hypothesis of FraCaS 239: ``ITEL won more
-orders than APCOM lost.''  We postulate that this sentence is
-equivalent to the following two separate parts: ``APCOM lost zero or
-more orders. ITEL won more orders [than some elliptic quantity].''
-The elliptic quantity is introduced every time we talk about some
-quantity (indexed by a CN, in this case ``orders''), and it can be
-referred to by a comparative, later in the discourse. Using this idea,
-we can go one level deeper in the interpretation of our example:
-``APCOM lost θ orders. θ >= 0.  ITEL won at least θ+1 orders.''. We
-see here how the quantities are introduced. They are added to the
-environment so that, they can be referred to as elliptic quantity
-expressions. Finally, ``more'' is intepreted as ``at least <elliptic
-quantity>+1''.
+Previous attempts to tackle the section of the FraCaS test suite
+devoted to comparatives showed that handling them is not easy. Our
+strategy here is to leverage our dynamic semantics, revealing an
+anaphoric element of comparatives. Indeed, Consider the hypothesis of
+(FraCaS 239): ``ITEL won more orders than APCOM lost.''  We postulate
+that this sentence is equivalent to the following two separate parts:
+``APCOM lost zero or more orders. ITEL won more orders [than some
+elliptic quantity].''  A quantity is introduced every time we talk
+about some quantity (indexed by a CN, in this case ``orders''), and it
+can be referred to by a comparative, later in the discourse. Using
+this idea, we can go one level deeper in the interpretation of our
+example: ``APCOM lost θ orders. θ >= 0.  ITEL won at least θ+1
+orders.''. We see here how the quantities are introduced. They are
+added to the environment so that, they can be referred to as elliptic
+quantity expressions. Finally, ``more'' is intepreted as ``at least
+<elliptic quantity>+1''. This treatment, which we illustrated here on
+an example, is systematic in our implementation.
 
@@
 \todo{reference to this view in the litterature}.  Additionally,
-adjectival assumptions as given in FraCaS are interpreted as linear
-equations in $Z$. Because solving such equations is decidable, solving
-such problems can be made efficient. Indeed, the tactic that Coq
-offers for this purpose can solve all such problems in FraCaS,
-automatically.
+adjectival phrases, as present in the FraCaS suite, are interpreted as
+linear inequations in $Z$. Solving systems of such inequations is
+decidable. Indeed, the tactic that Coq offers for this purpose can
+solve all such problems in the FraCaS suite, automatically.
 
@@
 one, is our handling of adverbial phrases. We interpret adverbs (and
-in general all adverbial and prepositional phrases) are VP-modifiers:
-$Adv = VP → VP$, where $VP = object → Prop$. However, applying
-adverbs to verb-phrases heavily complicates the Coq proofs, because
-such phrases can contain quantifiers. Therefore, we instead move the
-adverbs, so that they apply to (atomic) verbs only, and can simplify
-proofs accordingly. \todo{theoretical justification in the literature}
+in general all adverbial and prepositional phrases) as VP-modifiers:
+$Adv = VP → VP$, where $VP = object → Prop$. However, applying adverbs
+to verb-phrases heavily complicates the Coq proofs, because such
+phrases can contain quantifiers. Therefore, we instead move the
+adverbs, so that they apply to (atomic) verbs only. Proofs can then be
+simplify accordingly. \todo{theoretical justification in the
+  literature?}
 
@@
 We evaluated FraCoq against 8 sections of the FraCaS test suite, for a
-total of 259 examples. We excluded only one section, 7. The reason for
+total of 259 examples. We excluded only section 7, ``temporal reference''. The reason for
 doing so is that, in our view, it contains too many examples which
@@
 FraCaS intends to classify each problem as either entailment (YES),
-entailment of the opposite (NO) or no entailment (UNK).  For this
-work, we have amended FraCaS to address a few problems. First, certain
-test case are not formed correctly. Those were identified by Bill
-MacCartney as ``undef'', and we removed those. Second, a few test
-cases occurs twice in the suite, but with two different answers (one
-YES and one UNK), with the annotation that those correspond to
-different readings. However, elsewhere in the suite, if a problem has
-several readings but only one has entailment, it occurs only once and
-is marked as YES. To make the test suite consistent, if one reading
-yields entailment we have always considered it as YES. Finally, we
-have removed case 199 (which appears to be vacuous), and changed the
-following cases, which appeared to have been misclassified. We note
-that the majority of the mistaken classifications occur in sections 3
-and 4, which have not been previously attempted and thus, we propose,
-have not been properly scrutinized.
+entailment of the opposite (NO) or no entailment (UNK).  In this work,
+we have amended the FraCaS suite to correct a few problems. First,
+certain test case are not formed correctly. Those were already
+identified by Bill MacCartney\todo{cite} as such (using an ``undef''
+labelling), and we removed those. Second, a few test cases occur twice
+in the suite, but with two different labelling (one YES and one UNK),
+with an annotation that those labellings correspond to different
+readings. However, elsewhere in the suite, if a problem has several
+readings but only one has entailment, it occurs only once and is
+marked as YES. To make the test suite consistent, if one reading
+yields entailment we have always considered it as YES. We have also
+removed case 199 (which appears to be vacuous). Finally we changed the
+labelling of the following cases, which appeared to have been
+misclassified. We note that the majority of the mistaken
+classifications occur in sections 3 and 4, which have not been
+previously attempted and thus, we propose, have not been properly
+scrutinized.
 
@@
 hypothesis can be constructed and UNK otherwise. Because we work with
-a non-decidable logic, one can not \emph{in general} conclude
+a non-decidable logic, one cannot \emph{in general} conclude
 decisively that no proof exists. Thus, we consider here that no proof
@@
 6 & Comparatives & 31          & .87      & .56    & .48  & .45  &  -       \\
+7 & Temporal     & 75          &  -       &   -    &   -  &  -   &  -       \\
 8 & Verbs        & 8           & .75      &   -    & -    & -    &  -       \\
 9 & Attitudes    & 13          & .92      & .85    & .77  & .46  & .92 (9)  \\ \hline
-  & Total        & 262         & .89      & .83    & .69  & .50  & .85  \\
+  & Total        & 337         & .89      & .83    & .69  & .50  & .85  \\
   &              &             & (259)    & (174)  & (174)& (174)& (89)
@@
     by \citet{bos:2008}, and ``Langpro" to the system presented by
-    \citet{Abzianidze:2015}.  }
+    \citet{Abzianidze:2015}. A dash indicates that no attempt was made
+    for a given section. }
+  \label{tbl:results}
 \end{table}
 
-The above result reveals a considerable improvement over earlier
+Table \ref{tbl:results} shows a considerable improvement over earlier
 approaches in terms of coverage, with three more sections covered over
-previous state of the art. Additionally, our system is generally best
-in terms of accuracy overall. In particular, section 6 sees a large
-improvement in accuracy, which we attribute to our new analysis of
-comparatives, enabled by our dynamic semantics.
+previous state of the art. We thus cover 259 out of 337 cases (77\%),
+compared to at most 174 cases (52\%) in previous work. Additionally,
+our system performs generally the best in terms of accuracy. In
+particular, section 6 sees a large improvement in accuracy, which we
+attribute to our new analysis of comparatives, enabled by our dynamic
+semantics.
 
-\todo{error analysis}
+\paragraph{error analysis}
+Our system fails to classify correctly 28 cases out of 259.  We give
+here a summary of the missing features which are responsible for the
+biggest number of failures. The one biggest source of error is an
+incomplete handling of group readings. (FraCaS
+013,014,046,084,111,124,126,127,137,171,172,191,193,195,243,250,333,346). These
+are cases where a syntactic conjunction of individuals is treated as a
+semantic group, or where precise counting of the members of a group is
+necessary.  Other problematic cases are where a definite plural does
+not have a universal reading (091,094,095) --- these require an
+analysis at the pragmatic level, which our system lacks. Additionally,
+neither measure phrases (242) nor attributive comparatives (244,245)
+are handled.
 

### 2019-01-21 paper2018/paper.tex
oops, compilation
typography
more typography

@@
 mechanism, which allows us to support examples such as the following:
+\begin{lingex}
 \item[FraCaS 122]
@@
 We evaluated FraCoq against 8 sections of the FraCaS test suite, for a
-total of 259 examples. We excluded only section 7, ``temporal reference''. The reason for
+total of 259 cases. We excluded only section 7, ``temporal reference''. The reason for
 doing so is that, in our view, it contains too many examples which
@@
 \begin{tabular}{rlrrrrrr}
-  & Section      & \# examples & Ours     & FraCoq & MINE & Nut  & Langpro  \\ \hline
+  & Section      & {\kern -1em} \# cases & Ours     & FraCoq & MINE & Nut  & Langpro  \\ \hline
 1 & Quantifiers  & 75          & .96 (74) & .96    & .77  & .53  & .93 (44) \\
@@
     ``Ours" refers to the approach presented in this paper. When a
-    system does not handle the nominal number of examples shown in the
-    third column, the actual number of examples attempted is shown in
+    system does not handle the nominal number of test cases shown in the
+    third column, the actual number of test cases attempted is shown in
     parentheses.  ``FraCoq'' refers to the work of
@@
 incomplete handling of group readings. (FraCaS
-013,014,046,084,111,124,126,127,137,171,172,191,193,195,243,250,333,346). These
+013, 014, 046, 084, 111, 124, 126, 127, 137, 171, 172, 191, 193, 195, 243, 250, 333, 346). These
 are cases where a syntactic conjunction of individuals is treated as a

### 2019-02-11 paper2018/paper.tex
intro additions

@@
 \section{Introduction}
+Natural Language Inference (NLI) is task of determining of whether an NL hypothesis H follows from an NL premise(s) P.
+NLI has received a lot of attention in the Computational Semantics literature and has been approached using a variety of techniques, ranging from logical approaches  \citet{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy_type_2017} all the way to the recent Deep Learning (DL) models for NLP.
 

### 2019-02-11 paper2018/paper.tex
intro additions

@@
 NLI has received a lot of attention in the Computational Semantics literature and has been approached using a variety of techniques, ranging from logical approaches  \citet{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy_type_2017} all the way to the recent Deep Learning (DL) models for NLP.
+The latter approaches, following the general trend in NLP, have been dominating NLI and indeed a number of impressive results have been produced.
+State-of-the-art DL systems achieve an accuracy of around 0.9 when tested on suitable datasets.
+However, the datasets that are used are assuming a definition of inference that can be thought to be looser compared to the definition assumed in platforms that used to test logical approaches.
+Consider the following two examples, one from the SNLI dataset, predominatly used to test DL approaches, and the FraCaS test suite, the dataset \textit{par excellence} to test logical approaches.
+
+\begin{itemize}
+	\item[] \textbf{Premise:} A man selling donuts to a customer during a world
+exhibition event held in the city of Angeles.
+\item[] \textbf{Hypothesis:} A woman drinks her coffee in a small cafe.
+\item[] \textbf{Label:} Contradiction
+
+\end{itemize}
+
+\begin{itemize}
+
+\item[\textbf{P1}] A Scandinavian won the Nobel Prize.
+
+	\item[\textbf{P2}]	Every Swede is  Scandinavian.
+
+	\item[\textbf{H.}] Did a Swede win the Nobel prize?
+
+	\item[\textbf{H.}] A Swede won the Nobel prize.
+
+\item[\textbf{Label}] 	YES [FraCaS 173] 	\end{itemize}
+
+In the SNLI case, one has to make a number of non-trivial assumptions in order to arrive at a contradiction:  a)  the two situations described have to be taken to refer to  the same situation and, thus, decide  whether one description contradicts the other, b) the indefinite article in the premise has to be identified with the indefinite article in the hypothesis, i.e.
+\textit{man} contradicting \textit{woman} (a person cannot be a man and a woman at the same time).
+This is part of the reasoning humans perform, but it is not the only one.
+More precise, logical reasoning can be also performed.
+Such an example is the FraCaS example, as well as very fine-grained cases like the Antecedent Contained Deletion  ellipsis example below:
+
+\begin{itemize}
+
+\item[\textbf{P1}]	Bill spoke to everyone that John did.
+
+	\item[\textbf{P2}]		John spoke to Mary.
+
+	\item[\textbf{H.}] 	Did Bill speak to Mary?
+
+	\item[\textbf{H.}] 	Bill spoke to Mary.
+
+\item[\textbf{Label}] 	UNK [FraCaS 065] 	\end{itemize}
+
+In this paper, we concentrate on this sort of fine-grained reasoning.
+In particular, we present the first system that deals with the sections on ellipsis and anaphora in the FraCaS and has both the most coverage, as well as accuracy compared to other logical approaches.
+More specifically, the contributions of the paper are:
 
 \paragraph{Contributions}
-(see list above)
+
+\begin{itemize}
+ \item  A two-phase system (rather than one-phase as in  \citet{bernardy_type_2017}): a) using Haskell as a converter between Grammatical Framework trees and Coq types, b) using various type combinators to reach the desired NL semantics.
+  \item  The first run of the Anaphora and Ellipsis sections of the FraCaS (based on work by \citet{bernardy_jolli})
+ \item  More coverage than any previous logical approach
+  \item  State-of-the-art results in terms of accuracy compared to all previous approaches, for every section of the FraCaS
+  \item  Handing of multiple readings
+  \item  Threshold-based interpretation of (some) adjectives plus Linear arithmetic proofs
+  \item  New handling of prepositions and adverbs compared to the earlier system in  \citet{bernardy_type_2017}, as well as new handling of definitions, enabled by the dynamic semantics system. 
+  \item  Improvement over the previous tre
+atment of Plurals/Quantifiers
+ \item  Big improvement on the comparatives section  enabled by handling  comparatives using a threshold (BUT one different to the one used for adjectives).
+Again, the dynamic semantics system is crucial here.
+
+\end{itemize}
 \section{Background}
@@
   the GF formalism, and the types follow the Coq formalism. The
-  converter itself is a Haskell program. It implements a dynamic
+  converter itself is a Haskell Program. It implements a dynamic
   semantics. It comprises the bigger part of our system.

### 2019-02-15 paper2018/paper.tex
intro additions

@@
 
-\begin{itemize}
-	\item[] \textbf{Premise:} A man selling donuts to a customer during a world
-exhibition event held in the city of Angeles.
-\item[] \textbf{Hypothesis:} A woman drinks her coffee in a small cafe.
-\item[] \textbf{Label:} Contradiction
+\begin{lingex}
+\item An SNLI example \\
+\textbf{P}  A  man selling donuts to a customer during a world
+exhibition event held in the city of Angeles. \\
+ \textbf{H}  A woman drinks her coffee in a small cafe.\\
+\textbf{Label: Contradiction}
+\end{lingex}
 
-\end{itemize}
+\begin{lingex}
+\item An example from FraCaS\\
 
-\begin{itemize}
-
-\item[\textbf{P1}] A Scandinavian won the Nobel Prize.
-
-	\item[\textbf{P2}]	Every Swede is  Scandinavian.
-
-	\item[\textbf{H.}] Did a Swede win the Nobel prize?
-
-	\item[\textbf{H.}] A Swede won the Nobel prize.
-
-\item[\textbf{Label}] 	YES [FraCaS 173] 	\end{itemize}
+\textbf{P1}] A Scandinavian won the Nobel Prize.\\
+\textbf{P2}	Every Swede is  Scandinavian.\\
+\textbf{Q}  Did a Swede win the Nobel prize? \\
+\textbf{H} A Swede won the Nobel prize.\\
+\textbf{Label: Yes} [FraCaS 173]
+\end{lingex}
 
@@
 
-\begin{itemize}
+\begin{lingex}
 
-\item[\textbf{P1}]	Bill spoke to everyone that John did.
+\item An example from FraCaS involving fine-grained reasoning\\
+\textbf{P1}	Bill spoke to everyone that John did.	\\	
+\textbf{P2}	John spoke to Mary.	\\
+\textbf{Q} 	Did Bill speak to Mary?\\
+\textbf{H} 	Bill spoke to Mary.\\
+\textbf{Label:	UNK} [FraCaS 065] 	\end{lingex}
 
-	\item[\textbf{P2}]		John spoke to Mary.
+For this reason, and despite the dominance of DL approaches in pretty much all NLP tasks, logical approaches have been continuing to be developed and evaluated on datasets like the FraCaS test suite and the SICK dataset \cite{marelli:2014}.
+\citet{bernardy:2017}  define  a correspondence between abstract syntax parse trees of the FraCas examples, parsed using the Grammatical Framework (GF, \citet{Ranta:GF}), and modern type-theoretic semantics that are output in the Coq proof assistant (the FraCoq system).
+The accuracy is 0.85 for 5 ections of the FraCaS test suite.
+The \textsc{Langpro} system presented by \citet{abzianidze:2017} is based on a Natural Logic tableau theorem prover.
+It achieves an accuracy of .82 on the SICK dataset.
 
-	\item[\textbf{H.}] 	Did Bill speak to Mary?
-
-	\item[\textbf{H.}] 	Bill spoke to Mary.
-
-\item[\textbf{Label}] 	UNK [FraCaS 065] 	\end{itemize}
-
-In this paper, we concentrate on this sort of fine-grained reasoning.
-In particular, we present the first system that deals with the sections on ellipsis and anaphora in the FraCaS and has both the most coverage, as well as accuracy compared to other logical approaches.
+In this paper, we concentrate on this sort of fine-grained, logical reasoning.
+In particular, we present the first system that deals with the sections on ellipsis and anaphora in the FraCaS test suite and has both the best coverage, as well as accuracy compared to other logical approaches.
 More specifically, the contributions of the paper are:
-
-\paragraph{Contributions}
 
@@
 opposite adjectives can be easily represented (measures are opposites
-$∀x. m_1 x = -m_2 x$ and thresholds do not overlap $τ_1 + τ_2 > 0$).
-\todo{reference to this view in the litterature}.  Additionally,
+$∀x.
+m_1 x = -m_2 x$ and thresholds do not overlap $τ_1 + τ_2 > 0$).
+Formalization aside, this idea is reminiscent of degree-based approaches to gradable adjectives like \citet{Cresswell:1976,Heim:1985,Kennedy:2007}.
+Additionally,
 adjectival phrases, as present in the FraCaS suite, are interpreted as

### 2019-02-20 paper2018/paper.tex
background part added

@@
 \section{Introduction}
-Natural Language Inference (NLI) is task of determining of whether an NL hypothesis H follows from an NL premise(s) P.
-NLI has received a lot of attention in the Computational Semantics literature and has been approached using a variety of techniques, ranging from logical approaches  \citet{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy_type_2017} all the way to the recent Deep Learning (DL) models for NLP.
-The latter approaches, following the general trend in NLP, have been dominating NLI and indeed a number of impressive results have been produced.
-State-of-the-art DL systems achieve an accuracy of around 0.9 when tested on suitable datasets.
-However, the datasets that are used are assuming a definition of inference that can be thought to be looser compared to the definition assumed in platforms that used to test logical approaches.
-Consider the following two examples, one from the SNLI dataset, predominatly used to test DL approaches, and the FraCaS test suite, the dataset \textit{par excellence} to test logical approaches.
+Natural Language Inference (NLI) is the task of determining of whether an NL hypothesis H follows from an NL premise(s) P.
+NLI has received a lot of attention in the Computational Semantics literature and has been approached using a variety of techniques, ranging from logical approaches  \citet{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy_type_2017} all the way to the recent Deep Learning (DL) models for NLI.
+The latter approaches, following the general trend in NLP, have been dominating NLI and indeed a number of impressive results have been produced \citet{kim:2018,radford:2018,liu:2019}.\footnote{These are the three systems with the best results on SNLI in increasing order.} State-of-the-art DL systems achieve an accuracy of around 0.9 when tested on suitable datasets.
+However, the datasets that are used are assuming a definition of inference that can be thought to be `looser' or less precise compared to the definition assumed in platforms that used to test logical approaches.
+Consider the following two examples, one from the SNLI dataset, predominatly used to test DL approaches and which has a state-of-the-art performace of 0.911 \citet{lui:2019'}, and the FraCaS test suite, the dataset \textit{par excellence} to test logical approaches.
 
@@
 
-In the SNLI case, one has to make a number of non-trivial assumptions in order to arrive at a contradiction:  a)  the two situations described have to be taken to refer to  the same situation and, thus, decide  whether one description contradicts the other, b) the indefinite article in the premise has to be identified with the indefinite article in the hypothesis, i.e.
+In the SNLI example, one has to make a number of non-trivial assumptions in order to arrive at a contradiction:  a)  the two situations described have to be taken to refer to  the same situation and, thus, decide  whether one description contradicts the other, b) the indefinite article in the premise has to be identified with the indefinite article in the hypothesis, i.e.
 \textit{man} contradicting \textit{woman} (a person cannot be a man and a woman at the same time).
@@
 \citet{bernardy:2017}  define  a correspondence between abstract syntax parse trees of the FraCas examples, parsed using the Grammatical Framework (GF, \citet{Ranta:GF}), and modern type-theoretic semantics that are output in the Coq proof assistant (the FraCoq system).
-The accuracy is 0.85 for 5 ections of the FraCaS test suite.
+The accuracy is 0.85 for 5 sections of the FraCaS test suite.
 The \textsc{Langpro} system presented by \citet{abzianidze:2017} is based on a Natural Logic tableau theorem prover.
@@
 \end{itemize}
-\section{Background}
+\section{Background: GF, Monadic Dynamic Semantics, Coq}
+
+\section{GF}
+
+\subsection{GF}
+In GF, abstract syntax is comprised of: a) a number of syntactic
+categories, and b) a number of syntactic construction functions. The latter
+provide the means to compose basic syntactic categories into more complex
+ones.  For example, consider the constructor: $AdjCN : AP → CN → CN$. This  expresses
+that one can append an adjectival phrase to a common noun and obtain
+a new common noun. Furthermore,  GF is equipped  with a library of mappings from abstract
+syntax to the concrete s usyntax of varionatural languages. These mappings can be inverted by GF, thus offering
+parsing from natural text into abstract syntax. However, in this project we skip the
+parsing phase and use the parse trees constructed by \citet{Ljunglof:2012},
+thereby avoiding any syntactic ambiguity. \textbf{Do you want to further comment here?}
+
+\subsection{Coq}
+Coq is an  interactive theorem prover (proof assistant)  based on the calculus of inductive
+constructions (CiC), i.e.  a lambda calculus with dependent
+types. Coq is arguably one of the leading proof assistants. It has
+famously been instrumental in formalising a proof of the four-color
+theorem \citep{Gonthier:2008}, a proof of the odd order theorem \citep{Gonthier:2013}, as well as developing CompCert, a formally verified compiler for C \citep{Leroy}.
+Coq is a very powerful reasoning engine that makes it fit for the task of NLI, when the latter is formalized as theorem proving task.
+Its underlying higher order logic is furthermore very expressive to express fine-grained logical semantics.
+Furthermore, Coq supports a number of constructions that we will be using notably dependent typing and subtyping.
+These features are used in order to take care of specific NL phenomena (e.g.
+subtyping is used in defining different adjectival inference classes).
+
+\subsection{Dynamic Monadic Semantics}
 

### 2019-02-22 paper2018/paper.tex
background addition

@@
 \subsection{Dynamic Monadic Semantics}
+Dynamic Monadic Semantics have been proven to be an effective way of dealing with anaphora and ellipsis.
+There are a number of  approaches using monads or other similar ideas (e.g.
+continuations as in \cite{de2006}) for anaphora and ellipsis  \citet{Shan:2002,unger:2011,Barker:04,de2016,charlow:2017}.
+In this paper, we follow the approach  as described in \citet{bernardy_jolli}.
+More details on this will be given in the next section, where the actual system is presented.
 

### 2019-02-25 paper2018/paper.tex
acl formatting

@@
-\documentclass{article}
+\documentclass[11pt]{article}
+\usepackage{acl2019}
 \usepackage{todonotes}
@@
 \usetikzlibrary{arrows,automata}
-
-\usepackage[backend=bibtex,citestyle=authoryear,style=alphabetic,natbib]{biblatex}
-\addbibresource{PaperTools/bibtex/jp.bib}
-\addbibresource{phorae.bib}
-\addbibresource{IWCSbib.bib}
 
@@
 
+\maketitle{}
+
+\begin{abstract}
+  TODO
+\end{abstract}
 \section{Introduction}
@@
 
-\begin{figure}
+\begin{figure*}
   \centering
@@
   \label{fig:overview}
-\end{figure}
+\end{figure*}
 
@@
 removed case 199 (which appears to be vacuous). Finally we changed the
-labelling of the following cases, which appeared to have been
+labelling of the cases listed in Table \ref{tab:overrulings}, which appeared to have been
 misclassified. We note that the majority of the mistaken
@@
 scrutinized.
-
-\begin{center}
+\begin{table*}
+  \centering
 \begin{tabular}{ccc}
@@
 \end{tabular}
-\end{center}
+  \caption{Overruled FraCaS cases}
+  \label{tab:overrulings}
+\end{table*}
 
@@
 \providecommand\forcecenter{\multicolumn{1}{c}}
-\begin{table}
+\begin{table*}
   \centering
@@
     for a given section. }
-  \label{tbl:results}
-\end{table}
+  \label{tab:results}
+\end{table*}
 
-Table \ref{tbl:results} shows a considerable improvement over earlier
+Table \ref{tab:results} shows a considerable improvement over earlier
 approaches in terms of coverage, with three more sections covered over
@@
 
-\printbibliography
+\bibliography{PaperTools/bibtex/jp.bib,phorae.bib,IWCSbib.bib}
+
+\bibliographystyle{acl_natbib}
 

### 2019-02-26 paper2018/paper.tex
various minor revisions

@@
 \section{Introduction}
-Natural Language Inference (NLI) is the task of determining of whether an NL hypothesis H follows from an NL premise(s) P.
-NLI has received a lot of attention in the Computational Semantics literature and has been approached using a variety of techniques, ranging from logical approaches  \citet{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy_type_2017} all the way to the recent Deep Learning (DL) models for NLI.
-The latter approaches, following the general trend in NLP, have been dominating NLI and indeed a number of impressive results have been produced \citet{kim:2018,radford:2018,liu:2019}.\footnote{These are the three systems with the best results on SNLI in increasing order.} State-of-the-art DL systems achieve an accuracy of around 0.9 when tested on suitable datasets.
-However, the datasets that are used are assuming a definition of inference that can be thought to be `looser' or less precise compared to the definition assumed in platforms that used to test logical approaches.
-Consider the following two examples, one from the SNLI dataset, predominatly used to test DL approaches and which has a state-of-the-art performace of 0.911 \citet{lui:2019'}, and the FraCaS test suite, the dataset \textit{par excellence} to test logical approaches.
-
+Natural Language Inference (NLI) is the task of determining of whether
+an NL hypothesis H follows from an NL premise(s) P. NLI has received a
+lot of attention in the Computational Semantics literature and has
+been approached using a variety of techniques, ranging from logical
+approaches
+\citep{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy_type_2017},
+all the way to the recent Deep Learning (DL) models for NLI. The
+latter approaches, following a general trend in NLP, have been
+dominating NLI and a number of impressive results have been
+produced \citep{kim:2018,radford:2018,liu:2019}.\footnote{These are
+  the three systems with the best results on SNLI in increasing
+  order at the time of writing.}  State-of-the-art DL systems achieve an accuracy of around
+0.9 when tested on suitable datasets. However, the datasets that are
+used are assuming a definition of inference that can be thought to be
+`looser' or less precise compared to the definition assumed in
+platforms based in logical approaches. Consider the following
+two examples, one from the SNLI dataset, predominatly used to test DL
+approaches and which has a state-of-the-art performace of 0.911
+\citet{lui:2019'}, and the FraCaS test suite, the dataset \textit{par
+  excellence} to test logical approaches.
+%
+The following two examples illustrate the issue.
 \begin{lingex}
-\item An SNLI example \\
+\item % An SNLI example \\
 \textbf{P}  A  man selling donuts to a customer during a world
@@
  \textbf{H}  A woman drinks her coffee in a small cafe.\\
-\textbf{Label: Contradiction}
-\end{lingex}
-
-\begin{lingex}
-\item An example from FraCaS\\
-
+\textbf{Label: Contradiction} [SNLI]
+\item % An example from FraCaS\\
 \textbf{P1}] A Scandinavian won the Nobel Prize.\\
@@
 \end{lingex}
-
-In the SNLI example, one has to make a number of non-trivial assumptions in order to arrive at a contradiction:  a)  the two situations described have to be taken to refer to  the same situation and, thus, decide  whether one description contradicts the other, b) the indefinite article in the premise has to be identified with the indefinite article in the hypothesis, i.e.
-\textit{man} contradicting \textit{woman} (a person cannot be a man and a woman at the same time).
-This is part of the reasoning humans perform, but it is not the only one.
-More precise, logical reasoning can be also performed.
-Such an example is the FraCaS example, as well as very fine-grained cases like the Antecedent Contained Deletion  ellipsis example below:
+%
+In the SNLI example, one has to make a number of non-trivial
+assumptions in order to arrive at a contradiction: a) the two
+situations described have to be taken to refer to the same situation
+and, thus, decide whether one description contradicts the other, b)
+the indefinite article in the premise has to be identified with the
+indefinite article in the hypothesis. (Additionally considering that a
+person cannot be a man selling donuts and a woman drinking coffee at
+the same time.) While this can be part of the reasoning humans
+perform, it is not the only possibility. More precise, logical
+reasoning can be also performed. Such an example is the FraCaS
+example, as well as very fine-grained cases like the Antecedent
+Contained Deletion ellipsis example below:
 
@@
 
-For this reason, and despite the dominance of DL approaches in pretty much all NLP tasks, logical approaches have been continuing to be developed and evaluated on datasets like the FraCaS test suite and the SICK dataset \cite{marelli:2014}.
-\citet{bernardy:2017}  define  a correspondence between abstract syntax parse trees of the FraCas examples, parsed using the Grammatical Framework (GF, \citet{Ranta:GF}), and modern type-theoretic semantics that are output in the Coq proof assistant (the FraCoq system).
-The accuracy is 0.85 for 5 sections of the FraCaS test suite.
-The \textsc{Langpro} system presented by \citet{abzianidze:2017} is based on a Natural Logic tableau theorem prover.
-It achieves an accuracy of .82 on the SICK dataset.
+For this reason, and despite the dominance of DL approaches in pretty
+much all NLP tasks, logical approaches continue to be
+developed and evaluated on datasets like the FraCaS test suite and the
+SICK dataset \cite{marelli:2014}.  \citet{bernardy:2017} define a
+correspondence between abstract syntax parse trees of the FraCas
+examples, parsed using the Grammatical Framework (GF,
+\citet{Ranta:GF}), and modern type-theoretic semantics that are output
+in the Coq proof assistant (the FraCoq system).  The accuracy is 0.85
+for 5 sections of the FraCaS test suite. The \textsc{Langpro} system
+presented by \citet{abzianidze:2017} is based on a Natural Logic
+tableau theorem prover. It achieves an accuracy of .82 on the SICK
+dataset.
 
-In this paper, we concentrate on this sort of fine-grained, logical reasoning.
-In particular, we present the first system that deals with the sections on ellipsis and anaphora in the FraCaS test suite and has both the best coverage, as well as accuracy compared to other logical approaches.
-More specifically, the contributions of the paper are:
+In this paper, we concentrate on this sort of fine-grained, logical
+reasoning. In particular, we present the first system that deals with
+the sections on ellipsis and anaphora in the FraCaS test suite and has
+both the best coverage, as well as accuracy compared to other logical
+approaches. More specifically, the contributions of the paper are:
 
 \begin{itemize}
- \item  A two-phase system (rather than one-phase as in  \citet{bernardy_type_2017}): a) using Haskell as a converter between Grammatical Framework trees and Coq types, b) using various type combinators to reach the desired NL semantics.
-  \item  The first run of the Anaphora and Ellipsis sections of the FraCaS (based on work by \citet{bernardy_jolli})
+\item A two-phase system (rather than a one-phase as described by
+  \citet{bernardy_type_2017}): a) using Haskell as a converter between
+  Grammatical Framework trees and Coq types, b) using various type
+  combinators to reach the desired NL semantics.
+\item The first run of the Anaphora and Ellipsis sections of the
+  FraCaS (based on work by \citet{bernardy_jolli})
  \item  More coverage than any previous logical approach
-  \item  State-of-the-art results in terms of accuracy compared to all previous approaches, for every section of the FraCaS
+ \item State-of-the-art results in terms of accuracy compared to all
+   previous approaches, for every section of the FraCaS
   \item Handing of multiple readings
   \item  Threshold-based interpretation of (some) adjectives plus Linear arithmetic proofs
-  \item  New handling of prepositions and adverbs compared to the earlier system in  \citet{bernardy_type_2017}, as well as new handling of definitions, enabled by the dynamic semantics system. 
-  \item  Improvement over the previous tre
-atment of Plurals/Quantifiers
- \item  Big improvement on the comparatives section  enabled by handling  comparatives using a threshold (BUT one different to the one used for adjectives).
-Again, the dynamic semantics system is crucial here.
+  \item New handling of prepositions and adverbs compared to the
+    earlier system in \citet{bernardy_type_2017}, as well as new
+    handling of definitions, enabled by the dynamic semantics system.
+  \item Improvement over the previous treatment of Plurals/Quantifiers
+  \item Big improvement on the comparatives section enabled by
+    handling comparatives using a threshold (BUT one different to the
+    one used for adjectives). Again, the dynamic semantics system is
+    crucial here.
 
@@
 parsing phase and use the parse trees constructed by \citet{Ljunglof:2012},
-thereby avoiding any syntactic ambiguity. \textbf{Do you want to further comment here?}
+thereby avoiding any syntactic ambiguity. \todo{Do you want to further comment here?}
 
 \subsection{Coq}
-Coq is an  interactive theorem prover (proof assistant)  based on the calculus of inductive
-constructions (CiC), i.e.  a lambda calculus with dependent
-types. Coq is arguably one of the leading proof assistants. It has
-famously been instrumental in formalising a proof of the four-color
-theorem \citep{Gonthier:2008}, a proof of the odd order theorem \citep{Gonthier:2013}, as well as developing CompCert, a formally verified compiler for C \citep{Leroy}.
-Coq is a very powerful reasoning engine that makes it fit for the task of NLI, when the latter is formalized as theorem proving task.
-Its underlying higher order logic is furthermore very expressive to express fine-grained logical semantics.
-Furthermore, Coq supports a number of constructions that we will be using notably dependent typing and subtyping.
-These features are used in order to take care of specific NL phenomena (e.g.
-subtyping is used in defining different adjectival inference classes).
+Coq is an interactive theorem prover (proof assistant) based on the
+calculus of inductive constructions (CiC), i.e.  a lambda calculus
+with dependent types. Coq is arguably one of the leading proof
+assistants. It has famously been instrumental in formalising a proof
+of the four-color theorem \citep{Gonthier:2008}, a proof of the odd
+order theorem \citep{Gonthier:2013}, as well as developing CompCert, a
+formally verified compiler for C \citep{Leroy}. Coq is a very powerful
+reasoning engine that makes it fit for the task of NLI, when the
+latter is formalized as theorem proving task. Its underlying higher
+order logic is furthermore powerful enough to express fine-grained
+logical semantics. Furthermore, Coq supports a number of constructions
+that we will be using notably dependent typing and subtyping. These
+features allow us to take care of specific NL phenomena
+(e.g. subtyping is used in defining different adjectival inference
+classes).
 
 \subsection{Dynamic Monadic Semantics}
-Dynamic Monadic Semantics have been proven to be an effective way of dealing with anaphora and ellipsis.
-There are a number of  approaches using monads or other similar ideas (e.g.
-continuations as in \cite{de2006}) for anaphora and ellipsis  \citet{Shan:2002,unger:2011,Barker:04,de2016,charlow:2017}.
-In this paper, we follow the approach  as described in \citet{bernardy_jolli}.
-More details on this will be given in the next section, where the actual system is presented.
+Dynamic Monadic Semantics have been proven to be an effective way of
+dealing with anaphora and ellipsis. There are a number of approaches
+using monads or other equivalent constructions (e.g. continuations as in the work of
+\citet{de2006}) for anaphora and ellipsis
+\citet{Shan:2002,unger:2011,Barker:04,de2016,charlow:2017}. In this
+paper, we follow the approach described by
+\citet{bernardy_jolli}. More details on this will be given in the next
+section, where the actual system is presented.
 
@@
   the GF formalism, and the types follow the Coq formalism. The
-  converter itself is a Haskell Program. It implements a dynamic
-  semantics. It comprises the bigger part of our system.
+  converter itself is a Haskell Program, which implements a dynamic
+  semantics and comprises the bulk of our system.
 \item A number of type-theoretical combinators. These take care of the
   natural semantics which have no influence on the dynamical
-  aspects. This includes for example the treatment of adjectives
+  aspects. Such aspects include for example the treatment of adjectives
   (intersective, subsective, etc.) and adverbs (veridical or not).
 \end{enumerate}
-This architecture is represented schematically in Figure \ref{fig:overview}.
+The architecture is represented schematically in Figure \ref{fig:overview}.
 
@@
 
-All underlying systems (GF, Haskell, Coq) are based on lambda calculi
-with types. We ensure that each translation preserve typing, locally:
+All the underlying systems (GF, Haskell, Coq) are based on lambda
+calculi with types. We take advantage of typing, ensuring that each
+translation preserve typing, locally:
 \begin{enumerate}
@@
 \item GF Functional types are mapped compositionally : $⟦A → B⟧ = ⟦A⟧ → ⟦B⟧$
-\item Every GF syntactic construction function $f : X$ is mapped to a function $⟦f⟧$ such that $⟦f⟧ : ⟦X⟧$.
+\item Every GF syntactic construction function ($f : X$) is mapped to a function $⟦f⟧$ such that $⟦f⟧ : ⟦X⟧$.
 \item GF function applications are mapped compositionally: $⟦t(u)⟧ = ⟦t⟧ (⟦u⟧)$.
 \end{enumerate}
-Because all system embed the simply-typed lambda calculus, ensuring
+Because all systems embed the simply-typed lambda calculus, ensuring
 type-preservation locally means that types are preserved globally.
-Therefor, we are certain that every GF syntax tree can be mapped to
+Therefore, we are certain that every GF syntax tree can be mapped to
 Haskell, and eventuall Coq, without error.
@@
 opposite adjectives can be easily represented (measures are opposites
-$∀x.
-m_1 x = -m_2 x$ and thresholds do not overlap $τ_1 + τ_2 > 0$).
-Formalization aside, this idea is reminiscent of degree-based approaches to gradable adjectives like \citet{Cresswell:1976,Heim:1985,Kennedy:2007}.
-Additionally,
+$∀x. m_1 x = -m_2 x$ and thresholds do not overlap $τ_1 + τ_2 >
+0$). Formalization aside, this idea is reminiscent of degree-based
+approaches to gradable adjectives of
+\citet{Cresswell:1976,Heim:1985,Kennedy:2007}.  Additionally,
 adjectival phrases, as present in the FraCaS suite, are interpreted as
@@
 necessary.  Other problematic cases are where a definite plural does
-not have a universal reading (091,094,095) --- these require an
+not have a universal reading (091, 094, 095) --- these require an
 analysis at the pragmatic level, which our system lacks. Additionally,
-neither measure phrases (242) nor attributive comparatives (244,245)
+neither measure phrases (242) nor attributive comparatives (244, 245)
 are handled.
@@
 
-\bibliography{PaperTools/bibtex/jp.bib,phorae.bib,IWCSbib.bib}
+\bibliography{PaperTools/bibtex/jp.bib,phorae.bib,IWCSbib.bib,NLI.bib}
 

### 2019-02-26 paper2018/paper.tex
paper
fix cite

@@
 approaches and which has a state-of-the-art performace of 0.911
-\citet{lui:2019'}, and the FraCaS test suite, the dataset \textit{par
+\citet{liu:2019}, and the FraCaS test suite, the dataset \textit{par
   excellence} to test logical approaches.
@@
 for 5 sections of the FraCaS test suite. The \textsc{Langpro} system
-presented by \citet{abzianidze:2017} is based on a Natural Logic
+presented by \citet{Abzianidze:2015} is based on a Natural Logic
 tableau theorem prover. It achieves an accuracy of .82 on the SICK
@@
 approaches to gradable adjectives of
-\citet{Cresswell:1976,Heim:1985,Kennedy:2007}.  Additionally,
+\citet{Cresswell:1976,Kennedy:2007}.  Additionally,
 adjectival phrases, as present in the FraCaS suite, are interpreted as

### 2019-02-26 paper2018/paper.tex
paper
compress result table

@@
 \usepackage{todonotes}
+\usepackage{tabularx}
 \usepackage{tikz}
@@
 for 5 sections of the FraCaS test suite. The \textsc{Langpro} system
-presented by \citet{Abzianidze:2015} is based on a Natural Logic
+presented by \citet{abzianidze:2015} is based on a Natural Logic
 tableau theorem prover. It achieves an accuracy of .82 on the SICK
@@
 \providecommand\forcecenter{\multicolumn{1}{c}}
-\begin{table*}
+\providecommand\ncases[1]{{\ensuremath{^{#1}}}}
+\begin{table}
   \centering
-\begin{tabular}{rlrrrrrr}
-  & Section      & {\kern -1em} \# cases & Ours     & FraCoq & MINE & Nut  & Langpro  \\ \hline
-1 & Quantifiers  & 75          & .96 (74) & .96    & .77  & .53  & .93 (44) \\
-2 & Plurals      & 33          & .82      & .76    & .67  & .52  & .73 (24) \\
-3 & Anaphora     & 28          & .86      &   -    & -    & -    &  -       \\
-4 & Ellipsis     & 52          & .87      &   -    & -    & -    &  -       \\
-5 & Adjectives   & 22          & .95 (20) & .95    & .68  & .32  & .73 (12) \\
-6 & Comparatives & 31          & .87      & .56    & .48  & .45  &  -       \\
-7 & Temporal     & 75          &  -       &   -    &   -  &  -   &  -       \\
-8 & Verbs        & 8           & .75      &   -    & -    & -    &  -       \\
-9 & Attitudes    & 13          & .92      & .85    & .77  & .46  & .92 (9)  \\ \hline
-  & Total        & 337         & .89      & .83    & .69  & .50  & .85  \\
-  &              &             & (259)    & (174)  & (174)& (174)& (89)
-  \end{tabular}
+  \small
+\begin{tabularx}{\columnwidth}{Xr@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r}
+Section      & {\kern -2em} \#cases & Ours     & FC & MINE & Nut  & Langpro  \\ \hline
+Quantifiers  & 75          & .96  & .96    & .77  & .53  & .93  \\
+             &             & \ncases{74}     &        &      &      &     \ncases{44} \\
+Plurals      & 33          & .82      & .76    & .67  & .52  & .73 \\
+      &          &       &     &   &   & \ncases{24} \\
+Anaphora     & 28          & .86      &   -    & -    & -    &  -       \\
+Ellipsis     & 52          & .87      &   -    & -    & -    &  -       \\
+Adjectives   & 22          & .95 & .95    & .68  & .32  & .73 \\
+             &          &  \ncases{20} &     &   &   &  \ncases{12} \\
+Comparatives {\kern -1em}& 31          & .87      & .56    & .48  & .45  &  -       \\
+Temporal     & 75          &  -       &   -    &   -  &  -   &  -       \\
+Verbs        & 8           & .75      &   -    & -    & -    &  -       \\
+Attitudes    & 13          & .92      & .85    & .77  & .46  & .92  \\ 
+    &          &       &     &   &   & \ncases {9}  \\ \hline
+Total        & 337         & .89      & .83    & .69  & .50  & .85  \\
+             &             & \ncases{259}    & \ncases{174}  & \ncases{174}& \ncases{174}& \ncases{89}
+  \end{tabularx}
   \caption{Overview of the accuracy of our system compared to others.
     ``Ours" refers to the approach presented in this paper. When a
-    system does not handle the nominal number of test cases shown in the
-    third column, the actual number of test cases attempted is shown in
-    parentheses.  ``FraCoq'' refers to the work of
-    \citet{bernardy_type_2017}. ``MINE" refers to the approach of
-    \citet{Mineshima:2015}, ``NUT" to the CCG system that utilizes the
-    first-order automated theorem prover \textit{nutcracker} described
-    by \citet{bos:2008}, and ``Langpro" to the system presented by
-    \citet{Abzianidze:2015}. A dash indicates that no attempt was made
-    for a given section. }
+    system does not handle the nominal number of test cases (shown in
+    the third column), the actual number of test cases attempted is
+    shown below the accuracy figure, in smaller font.  ``FraCoq''
+    refers to the work of \citet{bernardy_type_2017}. ``MINE" refers
+    to the approach of \citet{Mineshima:2015}, ``NUT" to the CCG
+    system that utilizes the first-order automated theorem prover
+    \textit{nutcracker} described by \citet{bos:2008}, and ``Langpro"
+    to the system presented by \citet{abzianidze:2015}. A dash
+    indicates that no attempt was made for a given section. }
   \label{tab:results}
-\end{table*}
+\end{table}
 

### 2019-02-26 paper2018/paper.tex
fix cite

@@
 for 5 sections of the FraCaS test suite. The \textsc{Langpro} system
-presented by \citet{abzianidze:2015} is based on a Natural Logic
+presented by \citet{Abzianidze:2015} is based on a Natural Logic
 tableau theorem prover. It achieves an accuracy of .82 on the SICK
@@
     \textit{nutcracker} described by \citet{bos:2008}, and ``Langpro"
-    to the system presented by \citet{abzianidze:2015}. A dash
+    to the system presented by \citet{Abzianidze:2015}. A dash
     indicates that no attempt was made for a given section. }

### 2019-02-26 paper2018/paper.tex
compress obviously compressible things

@@
 %
-The following two examples illustrate the issue.
 \begin{lingex}
 \item % An SNLI example \\
+  \label{ex:snli}
 \textbf{P}  A  man selling donuts to a customer during a world
@@
 \textbf{Label: Contradiction} [SNLI]
-\item % An example from FraCaS\\
-\textbf{P1}] A Scandinavian won the Nobel Prize.\\
-\textbf{P2}	Every Swede is  Scandinavian.\\
-\textbf{Q}  Did a Swede win the Nobel prize? \\
-\textbf{H} A Swede won the Nobel prize.\\
-\textbf{Label: Yes} [FraCaS 173]
+% \item % An example from FraCaS\\
+% \textbf{P1} A Scandinavian won the Nobel Prize.\\
+% \textbf{P2}	Every Swede is  Scandinavian.\\
+% \textbf{Q}  Did a Swede win the Nobel prize? \\
+% \textbf{H} A Swede won the Nobel prize.\\
+% \textbf{Label: Yes} [FraCaS 173]
 \end{lingex}
 %
-In the SNLI example, one has to make a number of non-trivial
+In \ref{ex:snli}, one has to make a number of non-trivial
 assumptions in order to arrive at a contradiction: a) the two
@@
 example, as well as very fine-grained cases like the Antecedent
-Contained Deletion ellipsis example below:
+Contained Deletion ellipsis example \ref{ex:acd} below.\todo{It'd be nice to have a more clear parallel with indefinite articles as well; but I could not find this in fracas.}
 
-\begin{lingex}
-
-\item An example from FraCaS involving fine-grained reasoning\\
-\textbf{P1}	Bill spoke to everyone that John did.	\\	
-\textbf{P2}	John spoke to Mary.	\\
-\textbf{Q} 	Did Bill speak to Mary?\\
-\textbf{H} 	Bill spoke to Mary.\\
-\textbf{Label:	UNK} [FraCaS 065] 	\end{lingex}
+% \begin{lingex}
+% \item % An example from FraCaS involving fine-grained reasoning\\
+% \textbf{P1}	Bill spoke to everyone that John did.	\\	
+% \textbf{P2}	John spoke to Mary.	\\
+% % \textbf{Q} 	Did Bill speak to Mary?\\
+% \textbf{H} 	Bill spoke to Mary.\\
+% \textbf{Label:	UNK} [FraCaS 065] 	\end{lingex}
 
@@
 In this paper, we concentrate on this sort of fine-grained, logical
-reasoning. In particular, we present the first system that deals with
-the sections on ellipsis and anaphora in the FraCaS test suite and has
-both the best coverage, as well as accuracy compared to other logical
-approaches. More specifically, the contributions of the paper are:
+reasoning. In particular, we present a logic-based system that deals
+with many linguistic phenomena \emph{at the same time}. It is the
+first systems covering the sections on ellipsis and anaphora in the
+FraCaS test suite and has the best coverage and accuracy on the
+overall test suite.
+% This is listed below; we can let the reader figure it out.
+% \begin{itemize}
+% \item A two-phase system (rather than a one-phase as described by
+%   \citet{bernardy_type_2017}): a) using Haskell as a converter between
+%   Grammatical Framework trees and Coq types, b) using various type
+%   combinators to reach the desired NL semantics.
+% \item The first run of the Anaphora and Ellipsis sections of the
+%   FraCaS (based on work by \citet{bernardy_jolli})
+%  \item  More coverage than any previous logical approach
+%  \item State-of-the-art results in terms of accuracy compared to all
+%    previous approaches, for every section of the FraCaS
+%   \item Handing of multiple readings
+%   \item  Threshold-based interpretation of (some) adjectives plus Linear arithmetic proofs
+%   \item New handling of prepositions and adverbs compared to the
+%     earlier system in \citet{bernardy_type_2017}, as well as new
+%     handling of definitions, enabled by the dynamic semantics system.
+%   \item Improvement over the previous treatment of Plurals/Quantifiers
+%   \item Big improvement on the comparatives section enabled by
+%     handling comparatives using a threshold (BUT one different to the
+%     one used for adjectives). Again, the dynamic semantics system is
+%     crucial here.
+% \end{itemize}
 
-\begin{itemize}
-\item A two-phase system (rather than a one-phase as described by
-  \citet{bernardy_type_2017}): a) using Haskell as a converter between
-  Grammatical Framework trees and Coq types, b) using various type
-  combinators to reach the desired NL semantics.
-\item The first run of the Anaphora and Ellipsis sections of the
-  FraCaS (based on work by \citet{bernardy_jolli})
- \item  More coverage than any previous logical approach
- \item State-of-the-art results in terms of accuracy compared to all
-   previous approaches, for every section of the FraCaS
-  \item Handing of multiple readings
-  \item  Threshold-based interpretation of (some) adjectives plus Linear arithmetic proofs
-  \item New handling of prepositions and adverbs compared to the
-    earlier system in \citet{bernardy_type_2017}, as well as new
-    handling of definitions, enabled by the dynamic semantics system.
-  \item Improvement over the previous treatment of Plurals/Quantifiers
-  \item Big improvement on the comparatives section enabled by
-    handling comparatives using a threshold (BUT one different to the
-    one used for adjectives). Again, the dynamic semantics system is
-    crucial here.
-
-\end{itemize}
 \section{Background: GF, Monadic Dynamic Semantics, Coq}
-
-\section{GF}
 
@@
 \end{enumerate}
-The architecture is represented schematically in Figure \ref{fig:overview}.
 
-\begin{figure*}
-  \centering
-{
-  \tiny
-\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3.7cm,
-                    semithick]
-  \tikzstyle{every state}=[draw=black,text=black,shape=rectangle,text width=2cm]
+% The architecture is represented schematically in Figure \ref{fig:overview}.
+% \begin{figure*}
+%   \centering
+% {
+%   \tiny
+% \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3.7cm,
+%                     semithick]
+%   \tikzstyle{every state}=[draw=black,text=black,shape=rectangle,text width=2cm]
 
-  \node[state]         (txt)                {text};
-  \node[state]         (ast) [right of=txt] {AST};
-  \node[state]         (pp)  [right of=ast] {Probabilistic program};
-  \node[state]         (res) [right of=pp] {result in $(0,1)$};
+%   \node[state]         (txt)                {text};
+%   \node[state]         (ast) [right of=txt] {AST};
+%   \node[state]         (pp)  [right of=ast] {Probabilistic program};
+%   \node[state]         (res) [right of=pp] {result in $(0,1)$};
 
-  \path (txt) edge              node (GF)  {GF}                  (ast)
-        (ast) edge              node (Hask) [text width=1.5cm]{Haskell interpreter} (pp)
-        (pp) edge               node (Coq) {Coq} (res);
+%   \path (txt) edge              node (GF)  {GF}                  (ast)
+%         (ast) edge              node (Hask) [text width=1.5cm]{Haskell interpreter} (pp)
+%         (pp) edge               node (Coq) {Coq} (res);
 
-  \node (Syntax) [above of=GF,node distance=1cm] {Syntax};
-  \node (Sem) [above of=Hask,node distance=1cm] {Dynamic Semantics};
-  \node (Hyper) [above of=Coq,node distance=1cm] {Additional Semantics$^\ast$};
+%   \node (Syntax) [above of=GF,node distance=1cm] {Syntax};
+%   \node (Sem) [above of=Hask,node distance=1cm] {Dynamic Semantics};
+%   \node (Hyper) [above of=Coq,node distance=1cm] {Additional Semantics$^\ast$};
 
-  \path
-  (Syntax) edge (GF)
-  (Sem) edge (Hask)
-  (Hyper) edge (Coq);
-\end{tikzpicture}
-}
-\caption{Phases in our system. ($\ast$) At the level of Coq, we handle
-  the details of the adverbial (veridicality properties) and
-  adjectival semantics (division into subsective, extentional,
-  non-committal, etc. categories.) }
-  \label{fig:overview}
-\end{figure*}
+%   \path
+%   (Syntax) edge (GF)
+%   (Sem) edge (Hask)
+%   (Hyper) edge (Coq);
+% \end{tikzpicture}
+% }
+% \caption{Phases in our system. ($\ast$) At the level of Coq, we handle
+%   the details of the adverbial (veridicality properties) and
+%   adjectival semantics (division into subsective, extentional,
+%   non-committal, etc. categories.) }
+%   \label{fig:overview}
+% \end{figure*}
 
@@
 \begin{lingex}
-\item[FraCaS 122]
-P1	Every committee has a chairman.\\
-P2	He is appointed its members.\\
-H 	Every committee has a chairman appointed by members of the committee. 
+\item
+    \label{ex:anaphora-scope}
+\textbf{P1}	Every committee has a chairman.\\
+\textbf{P2}	He is appointed its members.\\
+\textbf{H} 	Every committee has a chairman appointed by members of the committee. \\ \textbf{Label:	YES} [FraCaS 122]
 \end{lingex}
@@
 \begin{lingex}
-\item[FraCas 173]
-  P1	Bill spoke to everyone that John did [elliptic V2].\\
-  P2	John spoke to Mary.\\
-  Q 	Did Bill speak to Mary?\\
-  H 	Bill spoke to Mary.
+\item
+    \label{ex:acd}
+\textbf{P1}	Bill spoke to everyone that John did [elliptic V2].\\
+\textbf{P2}	John spoke to Mary.\\
+\textbf{Q} 	Did Bill speak to Mary?\\
+\textbf{H} 	Bill spoke to Mary. \\ \textbf{Label:	Yes} [FraCas 173]
 \end{lingex}
@@
 this idea, we can go one level deeper in the interpretation of our
-example: ``APCOM lost θ orders. θ >= 0.  ITEL won at least θ+1
+example: ``APCOM lost θ orders. θ $\ge$ 0.  ITEL won at least θ+1
 orders.''. We see here how the quantities are introduced. They are
@@
 scrutinized.
-\begin{table*}
+\begin{table}
   \centering
-\begin{tabular}{ccc}
+  \small
+\begin{tabularx}{\columnwidth}{ccp{3.8cm}}
  test case & new class & comment \\ \hline
@@
   226 & Yes &
-\end{tabular}
+\end{tabularx}
   \caption{Overruled FraCaS cases}
   \label{tab:overrulings}
-\end{table*}
+\end{table}
 

### 2019-02-26 paper2018/paper.tex
more compression

@@
 
-\paragraph{Adverbs}
-Another point, of minor theoretical importance but major practical
-one, is our handling of adverbial phrases. We interpret adverbs (and
-in general all adverbial and prepositional phrases) as VP-modifiers:
-$Adv = VP → VP$, where $VP = object → Prop$. However, applying adverbs
-to verb-phrases heavily complicates the Coq proofs, because such
-phrases can contain quantifiers. Therefore, we instead move the
-adverbs, so that they apply to (atomic) verbs only. Proofs can then be
-simplify accordingly. \todo{theoretical justification in the
-  literature?}
+% \paragraph{Adverbs}
+% Another point, of minor theoretical importance but major practical
+% one, is our handling of adverbial phrases. We interpret adverbs (and
+% in general all adverbial and prepositional phrases) as VP-modifiers:
+% $Adv = VP → VP$, where $VP = object → Prop$. However, applying adverbs
+% to verb-phrases heavily complicates the Coq proofs, because such
+% phrases can contain quantifiers. Therefore, we instead move the
+% adverbs, so that they apply to (atomic) verbs only. Proofs can then be
+% simplify accordingly. \todo{theoretical justification in the
+%   literature?}
 
@@
 removed case 199 (which appears to be vacuous). Finally we changed the
-labelling of the cases listed in Table \ref{tab:overrulings}, which appeared to have been
-misclassified. We note that the majority of the mistaken
+labelling of 6 cases 
+% the cases listed in Table \ref{tab:overrulings},
+which appeared to have been
+misclassified (as justified in supplementary material). We note that the majority of the mistaken
 classifications occur in sections 3 and 4, which have not been
@@
 scrutinized.
-\begin{table}
-  \centering
-  \small
-\begin{tabularx}{\columnwidth}{ccp{3.8cm}}
- test case & new class & comment \\ \hline
-    005 & UNK & missing hypothesis: there are italian tenors \\
-   056 &  Yes & Already identified as such by MacCartney \\
-  069 & Unk & Mary could have used someone else's workstation \\
-  119 & Unk & \textit{ibid.} \\
-  181 & Yes & for the same reason as 180 \\
-  226 & Yes &
-\end{tabularx}
-  \caption{Overruled FraCaS cases}
-  \label{tab:overrulings}
-\end{table}
+% \begin{table}
+%   \centering
+%   \small
+% \begin{tabularx}{\columnwidth}{ccp{3.8cm}}
+%  test case & new class & comment \\ \hline
+%     005 & UNK & missing hypothesis: there are italian tenors \\
+%    056 &  Yes & Already identified as such by MacCartney \\
+%   069 & Unk & Mary could have used someone else's workstation \\
+%   119 & Unk & \textit{ibid.} \\
+%   181 & Yes & for the same reason as 180 \\
+%   226 & Yes &
+% \end{tabularx}
+%   \caption{Overruled FraCaS cases}
+%   \label{tab:overrulings}
+% \end{table}
 
@@
 
-\paragraph{error analysis}
-Our system fails to classify correctly 28 cases out of 259.  We give
-here a summary of the missing features which are responsible for the
-biggest number of failures. The one biggest source of error is an
-incomplete handling of group readings. (FraCaS
-013, 014, 046, 084, 111, 124, 126, 127, 137, 171, 172, 191, 193, 195, 243, 250, 333, 346). These
-are cases where a syntactic conjunction of individuals is treated as a
-semantic group, or where precise counting of the members of a group is
-necessary.  Other problematic cases are where a definite plural does
-not have a universal reading (091, 094, 095) --- these require an
-analysis at the pragmatic level, which our system lacks. Additionally,
-neither measure phrases (242) nor attributive comparatives (244, 245)
-are handled.
+% \paragraph{error analysis}
+% Our system fails to classify correctly 28 cases out of 259.  We give
+% here a summary of the missing features which are responsible for the
+% biggest number of failures. The one biggest source of error is an
+% incomplete handling of group readings. (FraCaS
+% 013, 014, 046, 084, 111, 124, 126, 127, 137, 171, 172, 191, 193, 195, 243, 250, 333, 346). These
+% are cases where a syntactic conjunction of individuals is treated as a
+% semantic group, or where precise counting of the members of a group is
+% necessary.  Other problematic cases are where a definite plural does
+% not have a universal reading (091, 094, 095) --- these require an
+% analysis at the pragmatic level, which our system lacks. Additionally,
+% neither measure phrases (242) nor attributive comparatives (244, 245)
+% are handled.
 

### 2019-02-26 paper2018/paper.tex
tiny compression attemps

@@
 
-\section{Background: GF, Monadic Dynamic Semantics, Coq}
+\section{Background: GF, Coq, Monadic Dynamic Semantics}
 
-\subsection{GF}
+\paragraph{GF}
 In GF, abstract syntax is comprised of: a) a number of syntactic
@@
 
-\subsection{Coq}
+\paragraph{Coq}
 Coq is an interactive theorem prover (proof assistant) based on the
@@
 of the four-color theorem \citep{Gonthier:2008}, a proof of the odd
-order theorem \citep{Gonthier:2013}, as well as developing CompCert, a
-formally verified compiler for C \citep{Leroy}. Coq is a very powerful
+order theorem \citep{Gonthier:2013}. Coq is a very powerful
 reasoning engine that makes it fit for the task of NLI, when the
@@
 
-\subsection{Dynamic Monadic Semantics}
+\paragraph{Dynamic Monadic Semantics}
 Dynamic Monadic Semantics have been proven to be an effective way of

### 2019-02-26 paper2018/paper.tex
oops

@@
     system does not handle the nominal number of test cases (shown in
-    the third column), the actual number of test cases attempted is
+    the second column), the actual number of test cases attempted is
     shown below the accuracy figure, in smaller font.  ``FraCoq''

### 2019-02-27 paper2018/paper.tex
abstract added and shortened - see SC comments

@@
 \begin{abstract}
-  TODO
+ In this paper, we present a system for Natural Language Inference which uses a Haskell dynamic semantics  converter from abstract syntax trees to Coq types.
+The system makes use of the fine-grainedness of a dynamic semantics system with the powerfulness of a state-of-the-art proof assistant, like Coq.
+We evaluate the system on all sections of the FraCaS test suite, excluding section 6.
+This is the first system that does a complete run on the anaphora and ellipsis sections of the FraCaS and has the best accuracy overall when compared with previous systems.
 \end{abstract}
@@
 `looser' or less precise compared to the definition assumed in
-platforms based in logical approaches. Consider the following
-two examples, one from the SNLI dataset, predominatly used to test DL
-approaches and which has a state-of-the-art performace of 0.911
-\citet{liu:2019}, and the FraCaS test suite, the dataset \textit{par
-  excellence} to test logical approaches.
+platforms based in logical approaches. For example, consider the following
+ example  from the SNLI dataset, predominatly used to test DL
+approaches:
 %
@@
 %
-In \ref{ex:snli}, one has to make a number of non-trivial
-assumptions in order to arrive at a contradiction: a) the two
+In \ref{ex:snli}, a number of non-trivial
+assumptions have to be made in order to arrive at a contradiction: a) the two
 situations described have to be taken to refer to the same situation
-and, thus, decide whether one description contradicts the other, b)
+in order to judge that the latter  contradicts the former, b)
 the indefinite article in the premise has to be identified with the
@@
 perform, it is not the only possibility. More precise, logical
-reasoning can be also performed. Such an example is the FraCaS
-example, as well as very fine-grained cases like the Antecedent
-Contained Deletion ellipsis example \ref{ex:acd} below.\todo{It'd be nice to have a more clear parallel with indefinite articles as well; but I could not find this in fracas.}
+reasoning is also a possibility, and will render the above label as unknown. Furthermore, reasoning can get very fine-grained 
+as the Contained Deletion ellipsis example \ref{ex:acd} below shows:
+
+\begin{lingex}
+\item
+    \label{ex:acd}
+\textbf{P1}	Bill spoke to everyone that John did [elliptic V2].\\
+\textbf{P2}	John spoke to Mary.\\
+\textbf{Q} 	Did Bill speak to Mary?\\
+\textbf{H} 	Bill spoke to Mary. \\ \textbf{Label:	Yes} [FraCas 173]
+\end{lingex}
 
@@
 a new common noun. Furthermore,  GF is equipped  with a library of mappings from abstract
-syntax to the concrete s usyntax of varionatural languages. These mappings can be inverted by GF, thus offering
-parsing from natural text into abstract syntax. However, in this project we skip the
+syntax to the concrete syntax of various natural languages.
+% These mappings can be inverted by GF, thus offering parsing from natural text into abstract syntax.
+However, in this project we skip the
 parsing phase and use the parse trees constructed by \citet{Ljunglof:2012},
-thereby avoiding any syntactic ambiguity. \todo{Do you want to further comment here?}
+thereby avoiding any syntactic ambiguity.
 
@@
 calculus of inductive constructions (CiC), i.e.  a lambda calculus
-with dependent types. Coq is arguably one of the leading proof
+with dependent types. Coq is  one of the leading proof
 assistants. It has famously been instrumental in formalising a proof
@@
 reasoning engine that makes it fit for the task of NLI, when the
-latter is formalized as theorem proving task. Its underlying higher
-order logic is furthermore powerful enough to express fine-grained
-logical semantics. Furthermore, Coq supports a number of constructions
-that we will be using notably dependent typing and subtyping. These
-features allow us to take care of specific NL phenomena
-(e.g. subtyping is used in defining different adjectival inference
-classes).
+latter is formalized as theorem proving task.
+It involves a very higher order logic  and supports a number of constructions, notably dependent typing and subtyping, that are instrumental in expressing NL semantics.
 
@@
 \citet{Shan:2002,unger:2011,Barker:04,de2016,charlow:2017}. In this
-paper, we follow the approach described by
+paper, we follow the approach described in
 \citet{bernardy_jolli}. More details on this will be given in the next
-section, where the actual system is presented.
+section.
 
@@
   semantics and comprises the bulk of our system.
-\item A number of type-theoretical combinators. These take care of the
-  natural semantics which have no influence on the dynamical
-  aspects. Such aspects include for example the treatment of adjectives
+\item A number of type-theoretical combinators, that encode aspects of NL semantics which have no influence on the dynamic
+  part. Such aspects include, e.g.,  the treatment of adjectives
   (intersective, subsective, etc.) and adverbs (veridical or not).
@@
 \citet{unger:2011}. However, we use a more general scope-extension
-mechanism, which allows us to support examples such as the following:
+mechanism, which allows us to support examples like  the following:
 \begin{lingex}
@@
 as anaphora. This method is made especially straightforward thanks to
-using GF syntax tree, which requires an explicit argument for each
-predicate. Thus, ellpsis are made explicit by the parsing phase. Such
+using GF syntax trees, which require an explicit argument for each
+predicate. Thus, ellipsis are made explicit by the parsing phase. Such
 ellptic expressions are handled in the same way as anaphora.
-\begin{lingex}
-\item
-    \label{ex:acd}
-\textbf{P1}	Bill spoke to everyone that John did [elliptic V2].\\
-\textbf{P2}	John spoke to Mary.\\
-\textbf{Q} 	Did Bill speak to Mary?\\
-\textbf{H} 	Bill spoke to Mary. \\ \textbf{Label:	Yes} [FraCas 173]
-\end{lingex}
-For example, in (FraCaS 173, P1), the argument of ``did'' is
+
+For example, in \ref{ex:acd}, the argument of ``did'' is
 explicitly marked as an elliptic V2, which we resolve to ``speak'' in
@@
 able to do this, we record all definites without referent, using
-another portion of the environment (using a monadic effect). For the
-above example, we obtain the desired interpretation:
-$∃y. dog(y) ∧ (∀x. person(x) → pet(x,y))$.
+another portion of the environment (using a monadic effect).
+%SC
+%For the
+%above example, we obtain the desired interpretation:
+%$∃y. dog(y) ∧ (∀x. person(x) → pet(x,y))$.
 
@@
 added to the environment so that, they can be referred to as elliptic
-quantity expressions. Finally, ``more'' is intepreted as ``at least
-<elliptic quantity>+1''. This treatment, which we illustrated here on
-an example, is systematic in our implementation.
+quantity expressions.
+%SC
+%Finally, ``more'' is intepreted as ``at least
+%<elliptic quantity>+1''. This treatment, which we illustrated here on
+%an example, is systematic in our implementation.
 
@@
 linear inequations in $Z$. Solving systems of such inequations is
-decidable. Indeed, the tactic that Coq offers for this purpose can
-solve all such problems in the FraCaS suite, automatically.
+decidable.
+% SC
+%Indeed, the tactic that Coq offers for this purpose can
+%solve all such problems in the FraCaS suite, automatically.
 
@@
 certain test case are not formed correctly. Those were already
-identified by Bill MacCartney\todo{cite} as such (using an ``undef''
+identified by Bill MacCartney \citet{maccartney:2007} as such (using an ``undef''
 labelling), and we removed those. Second, a few test cases occur twice
@@
 labelling of 6 cases
+%(SC: Shall we remove this to the supplementary material and just say to have a look at how we amended the suite?)
 % the cases listed in Table \ref{tab:overrulings},
@@
 
-\section{Future work}
-\section{Conclusion}
+\section{Conclusions and Future Work}
 

### 2019-02-28 paper2018/paper.tex
compression

@@
 \begin{abstract}
- In this paper, we present a system for Natural Language Inference which uses a Haskell dynamic semantics  converter from abstract syntax trees to Coq types.
-The system makes use of the fine-grainedness of a dynamic semantics system with the powerfulness of a state-of-the-art proof assistant, like Coq.
-We evaluate the system on all sections of the FraCaS test suite, excluding section 6.
-This is the first system that does a complete run on the anaphora and ellipsis sections of the FraCaS and has the best accuracy overall when compared with previous systems.
+  We present a system for Natural Language Inference which uses a
+  dynamic semantics converter from abstract syntax trees to Coq
+  types. It combines the fine-grainedness of a dynamic semantics
+  system with the powerfulness of a state-of-the-art proof assistant,
+  like Coq. We evaluate the system on all sections of the FraCaS test
+  suite, excluding section 6. This is the first system that does a
+  complete run on the anaphora and ellipsis sections of the FraCaS. It
+  has a better overall accuracy than any previous system.
 \end{abstract}
@@
 
-\section{Background: GF, Coq, Monadic Dynamic Semantics}
+\section{Background}
 
@@
 translation preserve typing, locally:
-\begin{enumerate}
+\begin{enumerate}[noitemsep]
 \item Every GF syntactic category $C$ is mapped to a type noted $⟦C⟧$.
@@
 Therefore, we are certain that every GF syntax tree can be mapped to
-Haskell, and eventuall Coq, without error.
+Haskell, and eventually Coq, without error.
 

### 2019-02-28 paper2018/paper.tex
some more compression and fixes

@@
 In our system, we interpret gradable adjectives using a pair of a
-measure $m : objects → Z$ and and a threshold $τ : Z$, where $Z$ is
+measure $m : objects → Z$ and a threshold $τ : Z$, where $Z$ is
 treated as an abstract ordered ring by Coq. (This structure has no
@@
 approaches to gradable adjectives of
-\citet{Cresswell:1976,Kennedy:2007}.  Additionally,
-adjectival phrases, as present in the FraCaS suite, are interpreted as
+\citet{Cresswell:1976,Kennedy:2007}.  Additionally
+adjectival predicates, as present in the FraCaS suite, are interpreted as
 linear inequations in $Z$. Solving systems of such inequations is
@@
 We evaluated FraCoq against 8 sections of the FraCaS test suite, for a
-total of 259 cases. We excluded only section 7, ``temporal reference''. The reason for
-doing so is that, in our view, it contains too many examples which
-require \textit{ad-hoc} treatment, and thus makes little sense to
-include without complementing it with a more thorough suite which
-captures a more complete landscape of the phenomena that section 7
-touches.
+total of 259 cases. We excluded only section 7, ``temporal reference''.
+% The reason for
+% doing so is that, in our view, it contains too many examples which
+% require \textit{ad-hoc} treatment, and thus makes little sense to
+% include without complementing it with a more thorough suite which
+% captures a more complete landscape of the phenomena that section 7
+% touches.
 
-FraCaS intends to classify each problem as either entailment (YES),
+FraCaS classifies each problem as either entailment (YES),
 entailment of the opposite (NO) or no entailment (UNK).  In this work,
@@
 certain test case are not formed correctly. Those were already
-identified by Bill MacCartney \citet{maccartney:2007} as such (using an ``undef''
+identified by \citet{maccartney:2007} as such (using an ``undef''
 labelling), and we removed those. Second, a few test cases occur twice
-in the suite, but with two different labelling (one YES and one UNK),
+in the suite, but with two different labellings (one YES and one UNK),
 with an annotation that those labellings correspond to different

### 2019-02-28 paper2018/paper.tex
clean and compress
compress

@@
 calculus of inductive constructions (CiC), i.e.  a lambda calculus
-with dependent types. Coq is  one of the leading proof
-assistants. It has famously been instrumental in formalising a proof
-of the four-color theorem \citep{Gonthier:2008}, a proof of the odd
-order theorem \citep{Gonthier:2013}. Coq is a very powerful
-reasoning engine that makes it fit for the task of NLI, when the
-latter is formalized as theorem proving task.
-It involves a very higher order logic  and supports a number of constructions, notably dependent typing and subtyping, that are instrumental in expressing NL semantics.
+with dependent types. Coq is one of the leading proof assistants. It
+has famously been instrumental in formalising a proof of the
+four-color theorem \citep{Gonthier:2008}, a proof of the odd order
+theorem \citep{Gonthier:2013}. Coq is a very powerful reasoning engine
+that makes it fit for the task of NLI, when the latter is formalized
+as theorem proving task. It supports
+notably dependent typing and subtyping, which are instrumental in
+expressing NL semantics.
 
@@
 Our system consists of two main parts.
-\begin{enumerate}
+\begin{enumerate}[noitemsep]
 \item A converter from syntax trees to types. The syntax trees follow
@@
   semantics and comprises the bulk of our system.
-\item A number of type-theoretical combinators, that encode aspects of NL semantics which have no influence on the dynamic
-  part. Such aspects include, e.g.,  the treatment of adjectives
-  (intersective, subsective, etc.) and adverbs (veridical or not).
+\item A number of type-theoretical combinators, that encode semantical
+  aspects which have no influence on the dynamic part. Such aspects
+  include the treatment of adjectives (intersective, subsective, etc.)
+  and adverbs (veridical or not).
 \end{enumerate}
@@
 ellptic expressions are handled in the same way as anaphora.
-
+%
 For example, in \ref{ex:acd}, the argument of ``did'' is
@@
 that context.
+
 \paragraph{definites} A naive way to handle definites is an

### 2019-03-01 paper2018/paper.tex
shorten
correct and clean
concluisions

@@
 paper, we follow the approach described in
-\citet{bernardy_jolli}. More details on this will be given in the next
+\citet{bernardy_jolli}. More details are given in the next
 section.
@@
 
-\paragraph{definites} A naive way to handle definites is an
-existential (or sigma). However, if the semantics does not feature a
+\paragraph{definites} A naive way to handle definites is using an
+existential type. However, if the semantics does not feature a
 dynamic element, then the existential quantification is introduced
@@
 $∀x. person(x) → ∃y. dog(y) ∧ pet(x,y)$.
-
+%
 Instead, our take is that definites should be treated as an anaphoric
@@
 \paragraph{Adjectives}
-In our system, we interpret gradable adjectives using a pair of a
+We interpret gradable adjectives using a pair of a
 measure $m : objects → Z$ and a threshold $τ : Z$, where $Z$ is
@@
 \section{Conclusions and Future Work}
+We presented a system using  GF, Dynamic Semantics and Coq, essentially converting GF trees to Coq types using dynamic semantics.
+The system outperforms the state of the art in logical approaches when tested on the FraCaS test suite and is the only system to date that has performed a run on the ellipsis/anaphora section of the FraCaS.
+The system's precision  can form the start of a very precise NL reasoner for controlled domains.
+For the future, we plan to  extend the system to cover the remaining section of the FraCaS (tense/aspect), and also develop a more applied version to perform reasoning on controlled NL domains.
 

### 2019-03-01 paper2018/paper.tex
concluisions

@@
 
-\title{FraCaS: 80 percent complete}
+\title{A Wide-Coverage Symbolic Natural Language Inference System}
 
@@
 \section{Conclusions and Future Work}
-We presented a system using  GF, Dynamic Semantics and Coq, essentially converting GF trees to Coq types using dynamic semantics.
+We presented a system  converting GF trees to Coq types using dynamic semantics.
 The system outperforms the state of the art in logical approaches when tested on the FraCaS test suite and is the only system to date that has performed a run on the ellipsis/anaphora section of the FraCaS.
-The system's precision  can form the start of a very precise NL reasoner for controlled domains.
+The system is precise enough  to form the start of a very precise NL reasoner for controlled domains.
 For the future, we plan to  extend the system to cover the remaining section of the FraCaS (tense/aspect), and also develop a more applied version to perform reasoning on controlled NL domains.

### 2019-03-04 paper2018/paper.tex
change paper to times; add back some text

@@
 \usepackage{acl2019}
+\usepackage{times}
 \usepackage{todonotes}
@@
 provide the means to compose basic syntactic categories into more complex
-ones.  For example, consider the constructor: $AdjCN : AP → CN → CN$. This  expresses
+ones.  For example, consider the constructor: $AdjCN : AP \to CN \to CN$. This  expresses
 that one can append an adjectival phrase to a common noun and obtain
@@
 \begin{enumerate}[noitemsep]
-\item Every GF syntactic category $C$ is mapped to a type noted $⟦C⟧$.
-\item GF Functional types are mapped compositionally : $⟦A → B⟧ = ⟦A⟧ → ⟦B⟧$
-\item Every GF syntactic construction function ($f : X$) is mapped to a function $⟦f⟧$ such that $⟦f⟧ : ⟦X⟧$.
-\item GF function applications are mapped compositionally: $⟦t(u)⟧ = ⟦t⟧ (⟦u⟧)$.
+\item Every GF syntactic category $C$ is mapped to a type noted $\llbracket{}C\rrbracket{}$.
+\item GF Functional types are mapped compositionally : $\llbracket{}A \to B\rrbracket{} = \llbracket{}A\rrbracket{} \to \llbracket{}B\rrbracket{}$
+\item Every GF syntactic construction function ($f : X$) is mapped to a function $\llbracket{}f\rrbracket{}$ such that $\llbracket{}f\rrbracket{} : \llbracket{}X\rrbracket{}$.
+\item GF function applications are mapped compositionally: $\llbracket{}t(u)\rrbracket{} = \llbracket{}t\rrbracket{} (\llbracket{}u\rrbracket{})$.
 \end{enumerate}
@@
 of the interpretation would be
-$∀x. person(x) → ∃y. dog(y) ∧ pet(x,y)$.
+$\forall{} x. person(x) \to \exists{}y. dog(y) \wedge pet(x,y)$.
 %
@@
 another portion of the environment (using a monadic effect).
-%SC
-%For the
-%above example, we obtain the desired interpretation:
-%$∃y. dog(y) ∧ (∀x. person(x) → pet(x,y))$.
+For the
+above example, we obtain the desired interpretation:
+$\exists{}y. dog(y) \wedge (\forall{}x. person(x) \to pet(x,y))$.
 
@@
 this idea, we can go one level deeper in the interpretation of our
-example: ``APCOM lost θ orders. θ $\ge$ 0.  ITEL won at least θ+1
+example: ``APCOM lost $\theta$ orders. $\theta$ $\ge$ 0.  ITEL won at least $\theta$+1
 orders.''. We see here how the quantities are introduced. They are
@@
 quantity expressions.
-%SC
-%Finally, ``more'' is intepreted as ``at least
+% SC
+%Finally, ``more'' is systematically intepreted as ``at least
 %<elliptic quantity>+1''. This treatment, which we illustrated here on
@@
 We interpret gradable adjectives using a pair of a
-measure $m : objects → Z$ and a threshold $τ : Z$, where $Z$ is
+measure $m : objects \to Z$ and a threshold $\tau : Z$, where $Z$ is
 treated as an abstract ordered ring by Coq. (This structure has no
 dynamic aspect in our model, and thus is entirely handled within Coq.)
-For subsective adjectives, $τ$ will additionally depend on the class
+For subsective adjectives, $\tau$ will additionally depend on the class
 of the object in question. This structure has the benefit that
 opposite adjectives can be easily represented (measures are opposites
-$∀x. m_1 x = -m_2 x$ and thresholds do not overlap $τ_1 + τ_2 >
+$\forall{}x. m_1 x = -m_2 x$ and thresholds do not overlap $\tau_1 + \tau_2 >
 0$). Formalization aside, this idea is reminiscent of degree-based
@@
 decidable.
-% SC
-%Indeed, the tactic that Coq offers for this purpose can
-%solve all such problems in the FraCaS suite, automatically.
+Indeed, the tactic that Coq offers for this purpose can
+solve all such problems in the FraCaS suite, automatically.
 
@@
 % in general all adverbial and prepositional phrases) as VP-modifiers:
-% $Adv = VP → VP$, where $VP = object → Prop$. However, applying adverbs
+% $Adv = VP \to VP$, where $VP = object \to Prop$. However, applying adverbs
 % to verb-phrases heavily complicates the Coq proofs, because such
@@
 
-% \paragraph{error analysis}
-% Our system fails to classify correctly 28 cases out of 259.  We give
-% here a summary of the missing features which are responsible for the
-% biggest number of failures. The one biggest source of error is an
-% incomplete handling of group readings. (FraCaS
-% 013, 014, 046, 084, 111, 124, 126, 127, 137, 171, 172, 191, 193, 195, 243, 250, 333, 346). These
-% are cases where a syntactic conjunction of individuals is treated as a
-% semantic group, or where precise counting of the members of a group is
-% necessary.  Other problematic cases are where a definite plural does
-% not have a universal reading (091, 094, 095) --- these require an
-% analysis at the pragmatic level, which our system lacks. Additionally,
-% neither measure phrases (242) nor attributive comparatives (244, 245)
-% are handled.
+\paragraph{error analysis}
+Our system fails to classify correctly 28 cases out of 259.  We give
+here a summary of the missing features which are responsible for the
+biggest number of failures. The one biggest source of error is an
+incomplete handling of group readings. (FraCaS
+013, 014, 046, 084, 111, 124, 126, 127, 137, 171, 172, 191, 193, 195, 243, 250, 333, 346). These
+are cases where a syntactic conjunction of individuals is treated as a
+semantic group, or where precise counting of the members of a group is
+necessary.  Other problematic cases are where a definite plural does
+not have a universal reading (091, 094, 095) --- these require an
+analysis at the pragmatic level, which our system lacks. Additionally,
+neither measure phrases (242) nor attributive comparatives (244, 245)
+are handled.
 
 \section{Conclusions and Future Work}
-We presented a system  converting GF trees to Coq types using dynamic semantics.
-The system outperforms the state of the art in logical approaches when tested on the FraCaS test suite and is the only system to date that has performed a run on the ellipsis/anaphora section of the FraCaS.
-The system is precise enough  to form the start of a very precise NL reasoner for controlled domains.
-For the future, we plan to  extend the system to cover the remaining section of the FraCaS (tense/aspect), and also develop a more applied version to perform reasoning on controlled NL domains.
+We presented a system converting GF trees to Coq types using dynamic
+semantics. The system outperforms the state of the art in logical
+approaches when tested on the FraCaS test suite and is the only system
+to date that has performed a run on the ellipsis/anaphora section of
+the FraCaS. The system is precise enough to form the start of a very
+precise NL reasoner for controlled domains. For the future, we plan to
+extend the system to cover the remaining section of the FraCaS
+(tense/aspect), and also develop a more applied version to perform
+reasoning on controlled NL domains.
 

### 2019-03-04 paper2018/paper.tex
stergios_final

@@
 
-\paragraph{anaphora} Thanks to the above system, we can handle many
+\paragraph{Anaphora} Thanks to the above system, we can handle many
 anaphoric cases, including E-Type and Donkey pronouns. Indeed, even
@@
 
-\paragraph{ellipsis} Ellipsis is handled in essentially the same way
+\paragraph{Ellipsis} Ellipsis is handled in essentially the same way
 as anaphora. This method is made especially straightforward thanks to
@@
 
-\paragraph{definites} A naive way to handle definites is using an
+\paragraph{Definites} A naive way to handle definites is using an
 existential type. However, if the semantics does not feature a
@@
 locally. This means that the quantifier can be introduced in the wrong
-context. Consider the phrase ``everyone pets the dog''. The structure
+Context. Consider the phrase ``everyone pets the dog''. The structure
 of the interpretation would be

### 2019-05-14 BankParser.hs
small renamings

@@ -189,7 +189,6 @@ overrides (177,1,"p")= Just "s_177_1_p_NEW"
 overrides (122,4,"h")= Just "s_122_4_h_ALT"
 overrides (155,2,"p")= Just "s_155_2_p_ALT"
 overrides (086,3,"h")= Just "s_086_2_h_ALT" -- gf syntax has the wrong noun (accountant vs. lawyer)
--- overrides (099,1,"p")= Just "s_099_1_p_fixed" -- changed in FraCoq 1; but actually we had wrong interpretation for bare plurals.
 overrides (323,4,"h") = Just "s_323_4_h_NEW"
 overrides (323,1,"p") = Just "s_323_1_p_NEW"
 overrides _ = Nothing

### 2019-05-14 paper2018/paper.tex
use nodalida format
ref

@@
 \documentclass[11pt]{article}
-\usepackage{acl2019}
+\usepackage{nodalida2019}
 \usepackage{times}
@@
 \input{PaperTools/latex/newunicodedefs}
+\usepackage{hyperref}
 
@@
 `looser' or less precise compared to the definition assumed in
-platforms based in logical approaches. For example, consider the following
+platforms based in logical approaches \citep{bernardy:2019}. For example, consider the following
  example  from the SNLI dataset, predominatly used to test DL
@@
 
-\bibliography{PaperTools/bibtex/jp.bib,phorae.bib,IWCSbib.bib,NLI.bib}
+\bibliography{PaperTools/bibtex/jp,phorae,IWCSbib,NLI}
 

### 2019-07-14 paper2018/paper.tex
fit
table it is
update nodalida

@@
   types. It combines the fine-grainedness of a dynamic semantics
-  system with the powerfulness of a state-of-the-art proof assistant,
-  like Coq. We evaluate the system on all sections of the FraCaS test
+  system with the powerfulness of a state-of-the-art proof assistant. We evaluate the system on all sections of the FraCaS test
   suite, excluding section 6. This is the first system that does a
@@
 calculus of inductive constructions (CiC), i.e.  a lambda calculus
-with dependent types. Coq is one of the leading proof assistants. It
-has famously been instrumental in formalising a proof of the
-four-color theorem \citep{Gonthier:2008}, a proof of the odd order
-theorem \citep{Gonthier:2013}. Coq is a very powerful reasoning engine
+with dependent types.  Coq is a very powerful reasoning engine
 that makes it fit for the task of NLI, when the latter is formalized
@@
 \paragraph{Anaphora} Thanks to the above system, we can handle many
-anaphoric cases, including E-Type and Donkey pronouns. Indeed, even
+anaphoric cases, including E-Type and Donkey anaphora. Indeed, even
 objects which have no syntactic representation can be added to the
-environment. We follow here the general pattern outlined by
+environment. We follow here the general monadic semantics approach as outlined by
 \citet{unger:2011}. However, we use a more general scope-extension
@@
 strategy here is to leverage our dynamic semantics, revealing an
-anaphoric element of comparatives. Indeed, Consider the hypothesis of
+anaphoric element of comparatives. Indeed, consider the hypothesis of
 (FraCaS 239): ``ITEL won more orders than APCOM lost.''  We postulate
@@
 added to the environment so that, they can be referred to as elliptic
-quantity expressions.
+quantity expressions.\footnote{The degree parameter assumption is not new in the formal semantics literature \cite{Cresswell:1976,Heim:2001,Kennedy:2007,chatzikyriakidis:2017} among many others.
+The specific details and computational implementation, however, are.}
 % SC
@@
 opposite adjectives can be easily represented (measures are opposites
-$\forall{}x. m_1 x = -m_2 x$ and thresholds do not overlap $\tau_1 + \tau_2 >
+$\forall{}x. m_1 x = \lnot m_2 x$ and thresholds do not overlap $\tau_1 + \tau_2 >
 0$). Formalization aside, this idea is reminiscent of degree-based
@@
 which appeared to have been
-misclassified (as justified in supplementary material). We note that the majority of the mistaken
+misclassified. We note that the majority of the mistaken
 classifications occur in sections 3 and 4, which have not been
 previously attempted and thus, we propose, have not been properly
-scrutinized.
+scrutinized. In terms of comparison, this only has a minor effect, since our system is the first system to run sections 3 and 4. 
 % \begin{table}
@@
 particular, we test at the minimum that the automatic proof search
-built in Coq does not succeed before classifying a problem as UNK.
+built in Coq does not succeed before classifying a problem as UNK.\footnote{The other way this can be done is by introducing a timeout as \citet{Mineshima:2015} have done.}
 
@@
 \providecommand\ncases[1]{{\ensuremath{^{#1}}}}
-\begin{table}
+\begin{table}[hbt]
   \centering
@@
   \end{tabularx}
-  \caption{Overview of the accuracy of our system compared to others.
+  \caption{Accuracy of our system compared to others.
     ``Ours" refers to the approach presented in this paper. When a
@@
     system that utilizes the first-order automated theorem prover
-    \textit{nutcracker} described by \citet{bos:2008}, and ``Langpro"
+    \textit{nutcracker} \cite{bos:2008}, and ``Langpro"
     to the system presented by \citet{Abzianidze:2015}. A dash
-    indicates that no attempt was made for a given section. }
+    indicates that no attempt was made for the section. }
   \label{tab:results}
@@
 approaches in terms of coverage, with three more sections covered over
-previous state of the art. We thus cover 259 out of 337 cases (77\%),
+previous approaches. We thus cover 259 out of 337 cases (77\%),
 compared to at most 174 cases (52\%) in previous work. Additionally,
 our system performs generally the best in terms of accuracy. In
-particular, section 6 sees a large improvement in accuracy, which we
-attribute to our new analysis of comparatives, enabled by our dynamic
-semantics.
+particular, section 6 largely improves in accuracy, which we
+attribute to our dynamic semantics analysis of comparatives.
 
 \paragraph{error analysis}
-Our system fails to classify correctly 28 cases out of 259.  We give
-here a summary of the missing features which are responsible for the
-biggest number of failures. The one biggest source of error is an
+Our system fails to correctly classify  28 cases out of 259.  We give
+here a summary of the missing features which are responsible for the failures. The  biggest source of error is 
 incomplete handling of group readings. (FraCaS
@@
 semantic group, or where precise counting of the members of a group is
-necessary.  Other problematic cases are where a definite plural does
-not have a universal reading (091, 094, 095) --- these require an
-analysis at the pragmatic level, which our system lacks. Additionally,
+necessary.  Other problematic cases include definite plurals with
+no  universal readings (091, 094, 095). Additionally,
 neither measure phrases (242) nor attributive comparatives (244, 245)
@@
 semantics. The system outperforms the state of the art in logical
-approaches when tested on the FraCaS test suite and is the only system
-to date that has performed a run on the ellipsis/anaphora section of
-the FraCaS. The system is precise enough to form the start of a very
-precise NL reasoner for controlled domains. For the future, we plan to
+approaches when tested on the FraCaS  and is the only system
+to date to perform a run on the FraCaS ellipsis/anaphora section. The system is precise enough to form the start of a 
+precise NL reasoner for controlled domains. In the future, we plan to
 extend the system to cover the remaining section of the FraCaS

### 2019-07-15 paper2018/paper-long.tex
paper long version

--- 2019-07-14 paper2018/paper.tex
@@
 \documentclass[11pt]{article}
-\usepackage{nodalida2019}
+% \usepackage{nodalida2019}
+\usepackage{natbib}
 \usepackage{times}
@@
 with many linguistic phenomena \emph{at the same time}. It is the
-first systems covering the sections on ellipsis and anaphora in the
+first system covering the sections on ellipsis and anaphora in the
 FraCaS test suite and has the best coverage and accuracy on the
@@
 % This is listed below; we can let the reader figure it out.
-% \begin{itemize}
-% \item A two-phase system (rather than a one-phase as described by
-%   \citet{bernardy_type_2017}): a) using Haskell as a converter between
-%   Grammatical Framework trees and Coq types, b) using various type
-%   combinators to reach the desired NL semantics.
-% \item The first run of the Anaphora and Ellipsis sections of the
-%   FraCaS (based on work by \citet{bernardy_jolli})
-%  \item  More coverage than any previous logical approach
-%  \item State-of-the-art results in terms of accuracy compared to all
-%    previous approaches, for every section of the FraCaS
-%   \item Handing of multiple readings
-%   \item  Threshold-based interpretation of (some) adjectives plus Linear arithmetic proofs
-%   \item New handling of prepositions and adverbs compared to the
-%     earlier system in \citet{bernardy_type_2017}, as well as new
-%     handling of definitions, enabled by the dynamic semantics system.
-%   \item Improvement over the previous treatment of Plurals/Quantifiers
-%   \item Big improvement on the comparatives section enabled by
-%     handling comparatives using a threshold (BUT one different to the
-%     one used for adjectives). Again, the dynamic semantics system is
-%     crucial here.
-% \end{itemize}
+
+In sum our contributions are as follows:
+\begin{itemize}
+\item A two-phase system (rather than a one-phase as described by
+  \citet{bernardy_type_2017}): a) using Haskell as a converter between
+  Grammatical Framework trees and Coq types, b) using various type
+  combinators to reach the desired NL semantics.
+\item The first run of the Anaphora and Ellipsis sections of the
+  FraCaS (based on work by \citet{bernardy_jolli})
+ \item  More coverage than any previous logical approach
+ \item State-of-the-art results in terms of accuracy compared to all
+   previous approaches, for every section of the FraCaS
+  \item Handing of multiple readings
+  \item  Threshold-based interpretation of (some) adjectives plus Linear arithmetic proofs
+  \item New handling of prepositions and adverbs compared to the
+    earlier system in \citet{bernardy_type_2017}, as well as new
+    handling of definitions, enabled by the dynamic semantics system.
+  \item Improvement over the previous treatment of Plurals/Quantifiers
+  \item Big improvement on the comparatives section enabled by
+    handling comparatives using a threshold (BUT one different to the
+    one used for adjectives). Again, the dynamic semantics system is
+    crucial here.
+\end{itemize}
 
@@
 syntax to the concrete syntax of various natural languages.
-% These mappings can be inverted by GF, thus offering parsing from natural text into abstract syntax.
+These mappings can be inverted by GF, thus offering parsing from natural text into abstract syntax.
 However, in this project we skip the
@@
 
-% The architecture is represented schematically in Figure \ref{fig:overview}.
-% \begin{figure*}
-%   \centering
-% {
-%   \tiny
-% \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3.7cm,
-%                     semithick]
-%   \tikzstyle{every state}=[draw=black,text=black,shape=rectangle,text width=2cm]
+The architecture is represented schematically in Figure \ref{fig:overview}.
+\begin{figure*}
+  \centering
+{
+  \tiny
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3.7cm,
+                    semithick]
+  \tikzstyle{every state}=[draw=black,text=black,shape=rectangle,text width=2cm]
 
-%   \node[state]         (txt)                {text};
-%   \node[state]         (ast) [right of=txt] {AST};
-%   \node[state]         (pp)  [right of=ast] {Probabilistic program};
-%   \node[state]         (res) [right of=pp] {result in $(0,1)$};
+  \node[state]         (txt)                {text};
+  \node[state]         (ast) [right of=txt] {AST};
+  \node[state]         (pp)  [right of=ast] {Probabilistic program};
+  \node[state]         (res) [right of=pp] {result in $(0,1)$};
 
-%   \path (txt) edge              node (GF)  {GF}                  (ast)
-%         (ast) edge              node (Hask) [text width=1.5cm]{Haskell interpreter} (pp)
-%         (pp) edge               node (Coq) {Coq} (res);
+  \path (txt) edge              node (GF)  {GF}                  (ast)
+        (ast) edge              node (Hask) [text width=1.5cm]{Haskell interpreter} (pp)
+        (pp) edge               node (Coq) {Coq} (res);
 
-%   \node (Syntax) [above of=GF,node distance=1cm] {Syntax};
-%   \node (Sem) [above of=Hask,node distance=1cm] {Dynamic Semantics};
-%   \node (Hyper) [above of=Coq,node distance=1cm] {Additional Semantics$^\ast$};
+  \node (Syntax) [above of=GF,node distance=1cm] {Syntax};
+  \node (Sem) [above of=Hask,node distance=1cm] {Dynamic Semantics};
+  \node (Hyper) [above of=Coq,node distance=1cm] {Additional Semantics$^\ast$};
 
-%   \path
-%   (Syntax) edge (GF)
-%   (Sem) edge (Hask)
-%   (Hyper) edge (Coq);
-% \end{tikzpicture}
-% }
-% \caption{Phases in our system. ($\ast$) At the level of Coq, we handle
-%   the details of the adverbial (veridicality properties) and
-%   adjectival semantics (division into subsective, extentional,
-%   non-committal, etc. categories.) }
-%   \label{fig:overview}
-% \end{figure*}
+  \path
+  (Syntax) edge (GF)
+  (Sem) edge (Hask)
+  (Hyper) edge (Coq);
+\end{tikzpicture}
+}
+\caption{Phases in our system. ($\ast$) At the level of Coq, we handle
+  the details of the adverbial (veridicality properties) and
+  adjectival semantics (division into subsective, extentional,
+  non-committal, etc. categories.) }
+  \label{fig:overview}
+\end{figure*}
 
@@
 The specific details and computational implementation, however, are.}
-% SC
-%Finally, ``more'' is systematically intepreted as ``at least
-%<elliptic quantity>+1''. This treatment, which we illustrated here on
-%an example, is systematic in our implementation.
+%
+Finally, ``more'' is systematically intepreted as ``at least
+<elliptic quantity>+1''. This treatment, which we illustrated here on
+an example, is systematic in our implementation.
 
@@
 
-% \paragraph{Adverbs}
-% Another point, of minor theoretical importance but major practical
-% one, is our handling of adverbial phrases. We interpret adverbs (and
-% in general all adverbial and prepositional phrases) as VP-modifiers:
-% $Adv = VP \to VP$, where $VP = object \to Prop$. However, applying adverbs
-% to verb-phrases heavily complicates the Coq proofs, because such
-% phrases can contain quantifiers. Therefore, we instead move the
-% adverbs, so that they apply to (atomic) verbs only. Proofs can then be
-% simplify accordingly. \todo{theoretical justification in the
-%   literature?}
+\paragraph{Adverbs}
+Another point, of minor theoretical importance but major practical
+one, is our handling of adverbial phrases. We interpret adverbs (and
+in general all adverbial and prepositional phrases) as VP-modifiers:
+$Adv = VP \to VP$, where $VP = object \to Prop$. However, applying adverbs
+to verb-phrases heavily complicates the Coq proofs, because such
+phrases can contain quantifiers. Therefore, we instead move the
+adverbs, so that they apply to (atomic) verbs only. Proofs can then be
+simplify accordingly.
 
@@
 total of 259 cases. We excluded only section 7, ``temporal reference''.
-% The reason for
-% doing so is that, in our view, it contains too many examples which
-% require \textit{ad-hoc} treatment, and thus makes little sense to
-% include without complementing it with a more thorough suite which
-% captures a more complete landscape of the phenomena that section 7
-% touches.
+The reason for
+doing so is that, in our view, it contains too many examples which
+require \textit{ad-hoc} treatment, and thus makes little sense to
+include without complementing it with a more thorough suite which
+captures a more complete landscape of the phenomena that section 7
+touches.
 
@@
 scrutinized. In terms of comparison, this only has a minor effect, since our system is the first system to run sections 3 and 4. 
-% \begin{table}
-%   \centering
-%   \small
-% \begin{tabularx}{\columnwidth}{ccp{3.8cm}}
-%  test case & new class & comment \\ \hline
-%     005 & UNK & missing hypothesis: there are italian tenors \\
-%    056 &  Yes & Already identified as such by MacCartney \\
-%   069 & Unk & Mary could have used someone else's workstation \\
-%   119 & Unk & \textit{ibid.} \\
-%   181 & Yes & for the same reason as 180 \\
-%   226 & Yes &
-% \end{tabularx}
-%   \caption{Overruled FraCaS cases}
-%   \label{tab:overrulings}
-% \end{table}
+\begin{table}
+  \centering
+  \small
+\begin{tabularx}{\columnwidth}{ccp{3.8cm}}
+ test case & new class & comment \\ \hline
+    005 & UNK & missing hypothesis: there are italian tenors \\
+   056 &  Yes & Already identified as such by MacCartney \\
+  069 & Unk & Mary could have used someone else's workstation \\
+  119 & Unk & \textit{ibid.} \\
+  181 & Yes & for the same reason as 180 \\
+  226 & Yes &
+\end{tabularx}
+  \caption{Overruled FraCaS cases}
+  \label{tab:overrulings}
+\end{table}
 

### 2019-07-16 paper2018/paper.tex
minor addition

@@
 that makes it fit for the task of NLI, when the latter is formalized
-as theorem proving task. It supports
+as a theorem proving task. It supports
 notably dependent typing and subtyping, which are instrumental in
@@
 %
-For example, in \ref{ex:acd}, the argument of ``did'' is
+For example, in \ref{ex:acd} repeated below as \ref{ex:acda},  the argument of ``did'' is
 explicitly marked as an elliptic V2, which we resolve to ``speak'' in
-that context.
+that context:
+
+\begin{lingex}
+	\item
+	\label{ex:acda}
+	\textbf{P1}	Bill spoke to everyone that John did [elliptic V2].\\
+	\textbf{P2}	John spoke to Mary.\\
+	\textbf{Q} 	Did Bill speak to Mary?\\
+	\textbf{H} 	Bill spoke to Mary. \\ \textbf{Label:	Yes} [FraCas 173]
+\end{lingex}
 

### 2019-07-16 paper2018/paper.tex
merge paper.tex and paper-long.tex; add ref to phorae

@@
 object quantified over by ``every'', whose scope is extended
-accordingly.
+accordingly.  We describe the anaphora resolution system in every
+detail in a manuscript \citep{anaphora}.
 

### 2019-07-16 paper2018/paper.tex
remove duplicate reference

@@
 approaches
-\citep{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy_type_2017},
+\citep{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy:2017},
 all the way to the recent Deep Learning (DL) models for NLI. The
@@
 % \item A two-phase system (rather than a one-phase as described by
-%   \citet{bernardy_type_2017}): a) using Haskell as a converter between
+%   \citet{bernardy:2017}): a) using Haskell as a converter between
 %   Grammatical Framework trees and Coq types, b) using various type
@@
 %   \item New handling of prepositions and adverbs compared to the
-%     earlier system in \citet{bernardy_type_2017}, as well as new
+%     earlier system in \citet{bernardy:2017}, as well as new
 %     handling of definitions, enabled by the dynamic semantics system.
@@
     shown below the accuracy figure, in smaller font.  ``FraCoq''
-    refers to the work of \citet{bernardy_type_2017}. ``MINE" refers
+    refers to the work of \citet{bernardy:2017}. ``MINE" refers
     to the approach of \citet{Mineshima:2015}, ``NUT" to the CCG

### 2019-07-27 paper2018/paper.tex
fix
aff

@@
 \usepackage{hyperref}
-
+\usepackage{url}
+\aclfinalcopy % Uncomment this line for the final submission
 \title{A Wide-Coverage Symbolic Natural Language Inference System}
-
+\author{Jean-Philippe Bernardy\\
+	Department of Philosophy, Linguistics \\ and Theory of Science\\
+University of Gothenburg \\
+\hskip -0.2in	{\tt jean-philippe.bernardy@gu.se} \\\And
+	Stergios Chatzikyriakidis \\
+	Department of Philosophy, Linguistics \\ and Theory of Science\\
+University of Gothenburg \\
+\hskip 0.4in	{\tt stergios.chatzikyriakidis@gu.se} \\}
 % - Two-phase system (Dyn semantics from theorem proving)
@@
 paper, we follow the approach described in
-\citet{bernardy_jolli}. More details are given in the next
+\citet{anaphora}. More details are given in the next
 section.

### 2019-08-29 IWCS_2017/IWCSTALK1.tex
Add files via upload
IWCS2017 slides

\documentclass[11pt]{beamer}
\usepackage{beamerthemesplit}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{gb4e}
\usetheme{Boadilla}
\newcommand{\Sem}[1]{\sem{#1}}
\logo{\includegraphics{clasp.ps}} 
\renewcommand{\theenumi}{\alph{enumi}}
\input{texhacks-beamer1}
\usefoottemplate{%
\vbox{%
\tinycolouredline{structure!75}%
{\textbf{Bernardy and Chatzikyriakidis\hfill \textsc{IWCS2017, Montpellier} \hfill \insertframenumber/\inserttotalframenumber}}%
}}\title{A Type-Theoretical system for the FraCaS test suite: Grammatical Framework meets Coq
}
\author{Jean-Philippe Bernardy and Stergios Chatzikyriakidis}
\begin{document}
\date{September 20, 2017}
\frame{\titlepage}
\begin{frame}[fragile]
	\frametitle{Brief summary of the talk}

	\begin{itemize}
		\item  Present a type-theoretical framework for formal semantics leveraging two well-studied tools
		\begin{itemize}
			\item Grammatical Framework (GF, Ranta 2004, 2011)
			\item Coq

	\end{itemize}
	\item Providing a compositional resource semantics for GF

	\item Evaluation on the FraCaS test suite

	\item State-of-the-art results	\end{itemize}
\end{frame}

\begin{frame}[fragile]
	\frametitle{Structure the talk}

	\begin{itemize}
		\item Brief intro to the systems used, GF and Coq

		\item Presenting the FraCoq system

		\begin{itemize}
			\item We concentrate on the most linguistically relevant features and also the features relevant for the FraCaS

\end{itemize}

	\item Evaluation against the FraCaS test suite

	\begin{itemize}
		\item Some brief remarks about the FraCaS and NLI platforms
			\begin{itemize}

			\item Results and comparison with previous logical approaches

			\item The issue of automation

\end{itemize}\end{itemize}

						\item Conclusions and Future work

\end{itemize}
\end{frame}

\begin{frame}[fragile]
	\frametitle{Background: Grammatical Framework (GF)}

\begin{itemize}
	\item Functional programming language for writing multi-lingual grammars

	\item Involves an abstract syntax, comprised of:

\begin{itemize}
	\item  A number of syntactic
	categories

	\item A number of syntactic construction functions, which
	provide the means to compose basic syntactic categories into more complex
	ones

\begin{itemize}
	\item

	$AdjCN : AP \to CN \to CN$ (appending an adjectival phrase to a common noun and obtaining a new common noun)
\end{itemize}\end{itemize}
\item GF comes with a library of mappings from abstract
syntax to concrete
\begin{itemize}
	\item These mappings can be inverted by GF, thus offering
parsers from natural text into abstract syntax

\item   We use the parse trees constructed by Ljungl\"of (2012)
thereby avoiding any syntactic ambiguity (GF FraCaS treebank).\end{itemize}\end{itemize}
\end{frame}

\begin{frame}[fragile]
	\frametitle{Background: Type-Theoretical Semantics}

	\begin{itemize}
		\item We use the type of logics that have been traditionally dubbed as constructive

			\begin{itemize}
				\item Initiated by the work of Martin-L\"of (1971, 1984)

				\item In linguistics this types of logics go back to Ranta's seminal work (Ranta 1994) or even earlier to Sundholm (1988)
				\begin{itemize}

				\item More recent approaches can be found as well. Please see Chatzikyriakidis and Luo (2017) for a collection of papers on constructive type theories for NL semantics

\end{itemize}\end{itemize}\end{itemize}

\end{frame}

\begin{frame}[fragile]
	\frametitle{Background: Type Theoretical Semantics}

	\begin{itemize}
		\item Main features of MTTs

			\begin{itemize}
				\item  Type many-sortedness.

				\item Dependent sum and product types

				\begin{itemize}
					\item  $\Sigma$-types, often written
					$\sum_{x:A}B[x]$ and which have product types $A\times B$ as a
					special case when $B$ does not depend on $x$.

					\item Dependent product, $\Pi$-types, often written
					($\prod_{x:A}B[x]$), and which have arrow-types $A\to B$ as a
					special case

					\item They generalize universal quantification and function types and they offer type polymorphism

					\end{itemize}

				\item Proof-theoretical specification and support for effective
				reasoning.

					\begin{itemize}
						\item Most powerful proof assistants implement MTTs (e.g. Coq, Agda)

\end{itemize}\end{itemize}\end{itemize}
\end{frame}

\begin{frame}[fragile]
	\frametitle{Background: Coq}

	\begin{itemize}

		\item	Proof assistant based on the calculus of inductive
		constructions (CiC, Coquand and Paulin-Mohring 1988)

	\begin{itemize}

		\item Arguably one of the leading proof assistants
	\begin{itemize}

		\item  a proof of the four-color
theorem (Gonthier 2008)
\item  a proof of the odd order theorem (Gonthier et al. 2013)
\item  developing CompCert, a formally verified compiler for C (Leroy 2013)
\item One of the assistants used in the Univalent Foundations project (Homotopy Type Theory, The Univalent Foundations Program 2013)
\end{itemize}\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[fragile]
	\frametitle{Background: Coq}

	\begin{itemize}

		\item Important features used

		\begin{itemize}

			\item $\Pi$ types

			\begin{itemize}

				\item	in Coq: $\prod_{x:A}B[x]$ is written \texttt{forall (x:A), B} or (simply
			\texttt{A $\to$ B} when \texttt{B} does not depend on \texttt{x}) \end{itemize}
		\item 	Record types
			\begin{itemize}

				\item  Generalization of $\Sigma$-types and are encoded as (trivial) inductive types with a single
			constructor.

		\item	$\Sigma x:A. B(x)$ can be expressed as a dependent record type in Coq:

\begin{verbatim}
Record AB:Type:=mkAB{x:> A;P:B x}
\end{verbatim}
\end{itemize}\end{itemize}\end{itemize}

\end{frame}

\begin{frame}[fragile]
	\frametitle{The FraCoq system}

	\begin{itemize}

		\item We use Lj\"unglof's FraCaS treebank and take these trees to their semantic counterparts

			\item The structure of the semantic representation

		\begin{enumerate}
			\item Every GF syntactic category $C$ is mapped to a Coq $Set$, noted $\sem{C}$.
			\item GF Functional types are mapped compositionally : $\sem{A \to B}$ $=$ $\sem{A} \to \sem{B}$
			\item Every GF syntactic construction function $f : X$ is mapped to a function $\sem{f}$ such that $\sem{f} : \sem{X}$.
			\item GF function applications are mapped compositionally: $\sem{t(u)} = \sem{t} (\sem{u})$.
		\end{enumerate}

\end{itemize}
\end{frame}

\begin{frame}[fragile]
	\frametitle{The FraCoq system}

	\begin{itemize}

\item Sentences

	\begin{itemize}

		\item We interpret sentences as propositions: $\sem{S} =
		Prop$.

		\item To verify that
		$P$ entails $H$, we prove the proposition $\sem{P} \to \sem{H}$.

		\begin{verbatim}
Definition S := Prop.
		\end{verbatim}

\end{itemize}
\item Common Nouns
	\begin{itemize}

		\item Predicates over an abstract object type

			\begin{verbatim}
Parameter object : Set.
Definition CN := object->Prop.
			\end{verbatim}

\end{itemize}\end{itemize}
\end{frame}
\begin{frame}[fragile]
	\frametitle{The FraCoq system}

	\begin{itemize}

		\item Verb phrases

		\begin{itemize}

			\item Parameterize over the
			\emph{noun} of the subject (using $\Pi$ types)

			\begin{verbatim}
  Definition VP := forall (subjectClass : CN) 
  object -> Prop.
			\end{verbatim}	
\end{itemize}			
\end{itemize}
\end{frame}

\begin{frame}[fragile]
	\frametitle{The FraCoq system}

\begin{itemize}

\item Adjectives

	\begin{itemize}

		\item Functions from cn to cn (predicates to predicates)

		\begin{verbatim}
Definition A := CN -> CN.
		\end{verbatim}

\item Different classes of adjectives are captured using coercions (subtyping). All special classes of adjectives are subtypes of A.

\begin{verbatim}
Definition IntersectiveA := object -> Prop.
Definition wkIntersectiveA : IntersectiveA -> A
:= fun a cn (x:object) => a x /\ cn x.
Coercion wkIntersectiveA : IntersectiveA >-> A.
\end{verbatim}

\item Provision is made for intersective, subsective, privative and non-committal adjectives

\end{itemize}\end{itemize}
\end{frame}

\begin{frame}[fragile]
	\frametitle{The FraCoq system}

	\begin{itemize}

		\item Adverbs

		\begin{itemize}

			\item Similar method to adjectives but instead the modification is on verbal predicates

			\item The adverb cases in the FraCaS are all veridical and covariant.

			\item We define such a subclass $VeridicalAdv$ and declare it as a coercion $Adv$

				\begin{itemize}

					\item Adverbs of type $VeridicalAdv$ are also of type $Adv$

\begin{verbatim}			
Parameter on_time_Adv : VeridicalAdv .	
\end{verbatim}

\end{itemize}\end{itemize}\end{itemize}
\end{frame}

\begin{frame}[fragile]
	\frametitle{The FraCoq system}

	\begin{itemize}

		\item Noun Phrases and Predeterminers

	\begin{itemize}

		\item A clean definition of NPs as functions from predicates to truth values will not work

			\begin{itemize}

				\item	Problem with GF's abstract syntax: existence of pre-determiners which include cases like \textit{most}, \textit{all} among others and are defined as functions from NPs to NPs

				\item In general, the category includes elements that are naturally interpreted as GQs

				\item Solution: Remember the components of NPs (number, quantifier and common noun)

				\item Pre-determiners then are able to substistute the quantifier part with the appropriate quantifier

				\item This has to be done, otherwise pre-determiners introduce a dummy indefinite in these cases

\end{itemize}\end{itemize}\end{itemize}
\end{frame}

\begin{frame}[fragile]
	\frametitle{The FraCoq system}

	\begin{itemize}

		\item Generalized Quantifiers

		\begin{itemize}

			\item They turn a number
			and a common noun into a noun-phrase (which we call $NP0$).

\begin{verbatim}
	Definition Quant := Num -> CN -> NP0.
\end{verbatim}

\item Some quantifiers ignore the number and are given usual definitions (e.g. \textit{some} or \textit{all}), whereas others make essential use of number (e.g. \textit{at most})

\begin{itemize}

\item In the latter case, the function \texttt{CARD}, a context-dependent
abstract function which turns a predicate into a natural number	is used to get the correct semantics 	
\end{itemize}\end{itemize}\end{itemize}
\end{frame}

\begin{frame}[fragile]
	\frametitle{The FraCoq system}

	\begin{itemize}

		\item The definite article

	\begin{itemize}

		\item Checks for plural noun phrases

			\begin{itemize}

				\item If found, then universal quantification

				\item If not, it
				looks up the object of discourse in an abstract $environment$, which
				is a function which turns a common noun into an object

\end{itemize}\end{itemize}\end{itemize}
\begin{verbatim}
	Definition DefArt:Quant:= fun (num : Num) (P:CN)=> fun Q:VP
	=> match num with plural => (forall x, P x -> Q P x)
	 /\ Q P (environment P) /\ P (environment P) |
	_ => Q P (environment P) /\ P (environment P) end.
\end{verbatim}

\end{frame}

\begin{frame}[fragile]
	\frametitle{Evaluation: The FraCaS test suite}

	\begin{itemize}

		\item A test suite for NLI

	\begin{itemize}

		\item 346 NLI examples in the form of one or more premises followed
		by a question along with an answer to that question

\item Three potential answers

\begin{itemize}

	\item YES: The declarative sentence formed out of the question follows from the premises

	\item NO: The declarative sentence does not follow from the premises

	\item UNK: The declarative sentence neither follows nor does not follow fro the premises
\end{itemize}\end{itemize}\end{itemize}
\end{frame}

\begin{frame}[fragile]
	\frametitle{Evaluation: The FraCaS test suite}

		\begin{exe}
			\ex
			A Swede won the Nobel Prize.

			Every Swede is  Scandinavian.

			Did a Scandinavian win the Nobel prize? [Yes, FraCas 049]
		\end{exe}

		 \begin{exe}
		 	\ex
		 	No delegate finished the report on time..

		 	Did any Scandinavian delegate finish the report on time? [No, FraCaS 070]
		 \end{exe}

		 \begin{exe}
		 	\ex
		 	A Scandinavian won the Nobel Prize.

		 	Every Swede is  Scandinavian.

		 	Did a Swede win the Nobel prize? [UNK, FraCaS 065]\end{exe}

\end{frame}

\begin{frame}[fragile]
	\frametitle{Evaluation: The FraCaS test suite}

	\begin{itemize}

		\item The FraCaS has considerable weaknesses

		\begin{itemize}

			\item Small size

			\item Artificial nature of the
			examples

\end{itemize}

\item However, it covers a lot of phenomena associated with NLI

\item It is still a very good suite to test logical approaches\end{itemize}
\end{frame}

\begin{frame}[fragile]
	\frametitle{Evaluation}

	\begin{itemize}

		\item Evaluation against 5 sections of the FraCaS

		\begin{itemize}

			\item Total of 174 examples

		\item Excluded sections where a lot of context-dependency has to be taken into consideration (e.g. the section on ellipsis)

\end{itemize}

\item YES: a proof
can be constructed from the premises to the hypothesis

\item NO: a proof
of the negated hypothesis can be constructed

\item UNK: otherwise

\end{itemize}
\end{frame}

	\begin{frame}[fragile]
		\frametitle{Evaluation}

		\begin{itemize}

			\item		The following table presents the results (Ours) as well as a comparison with the approach in Mineshima et al. (MINE, 2015), Bos (Nut, 2008) and Abzianidze (Langpro, 2015)

			\begin{center}
				\begin{tabular}{rlrrrrr}
					& Section                                                                               & \# examples                                                                                                & \begin{tabular}[c]{@{}l@{}}Ours\end{tabular}                                                        & \begin{tabular}[c]{@{}l@{}}MINE \end{tabular}                                             & \begin{tabular}[c]{@{}l@{}}Nut\end{tabular}   & \begin{tabular}[c]{@{}l@{}} Langpro \end{tabular}                                                     \\ \hline
					\begin{tabular}[c]{@{}l@{}}1\\ 2\\ 3\\ 4\\5\\6\end{tabular} & \begin{tabular}[c]{@{}l@{}}Quantifiers\\ Plurals\\ Adjectives\\Comparatives\\ Attitudes\\ Total\end{tabular} & \begin{tabular}[c]{@{}l@{}}75\\ 33\\22\\ 31\\13\\174 (181)\end{tabular} & \begin{tabular}[c]{@{}l@{}}.96\\ .76 \\ .95 \\ .56    \\ .85\\ 0.83 \end{tabular} & \begin{tabular}[c]{@{}l@{}}.77\\ .67\\ .68 \\   .48      \\.77\\0.69 \end{tabular} & \begin{tabular}[c]{@{}l@{}} .53\\ .52\\ .32\\ .45\\.46\\0.50\end{tabular} & \begin{tabular}[c]{@{}l@{}} .93 (44)
						\\ .73 (24)\\ .73 (12)\\ -\\.92 (9)\\0.85\end{tabular}
				\end{tabular}
			\end{center}

\item Our approach outerperforms Mineshema et al. by 13 percentage points.

\item The approach by Abzianidze has an accuracy of 0.85 without involving the comparative section. If this section is taken out, our system's accuracy rises to 0.88

\end{itemize}

\end{frame}

	\begin{frame}[fragile]
		\frametitle{Error Analysis}

		\begin{itemize}

			\item	Improvement over earlier approaches. Still, there were a couple of difficulties

				\begin{itemize}

					\item Comparatives: cases that needed one to provide
					adequate semantics for \textit{more} but also to take care of ellipsis

			\begin{exe}
				\ex				
	\hskip 0.05in ITEL won more orders than APCOM.

	\hskip 0.05in ITEL won some orders.

	\hskip 0.05in  	Did ITEL win some orders? [Yes, FraCaS 233]

\end{exe}		
\item Definite Plurals: Universal reading was captured. Cases of existential readings were not

\begin{exe}
	\ex

\hskip 0.05in The inhabitants of Cambridge voted for a Labour MP.

\hskip 0.05in very inhabitant of Cambridge voted for a Labour MP.

\hskip 0.05in  	Did every inhabitant of Cambridge vote for a Labour MP? [UNK, FraCaS 094]

		\end{exe}

\end{itemize}\end{itemize}

\end{frame}

\begin{frame}[fragile]
	\frametitle{Automation}

	\begin{itemize}

		\item So far, our proofs are not automated

		\begin{itemize}

			\item A couple of steps (usually very few) to reach a proof

			\item Earlier approaches using Coq (e.g. Chatzikyriakidis 2014 and Mineshima et al. 2015) use Coq's tactical language LTac to define automated macros of actions

				\begin{itemize}

					\item This is not difficult to do in our case as well

					\item Just go through all the proof
					tactics or observe the tactics that are used in the proofs to create a
					macro that will automate the proofs

					\item The question remains: can that macro of tactics
					generalize outside the suite?

					\item Answer: only to a limited
					extent, i.e. when exactly the same set of tactics yields a proof

					\item For
					this reason, we have not automated proof search to obtain the results
					presented in this paper, even though this can be done easily

\end{itemize}\end{itemize}

\end{itemize}

\end{frame}

\begin{frame}[fragile]
	\frametitle{Automation}

	\begin{itemize}

		\item Automating would also make an unprincipled use of
		higher-order logic (HOL)

			\begin{itemize}

				\item No algorithm which
				can decide if a proposition has a proof or not

\begin{itemize}

	\item We
	must use heuristics both to search for proofs and to decide when to
	give up searching

\end{itemize}\end{itemize}

\item  Most problems have either
obvious proofs or obviously lack a proof (fortunately)

\begin{itemize}

	\item Due to its heuristic nature the
	proof search necessarily contains a human component

	\begin{itemize}

		\item Problematic to make a statement about the suitability of FraCoq
		outside FraCas

		\item Small dataset and lack of separation between a development and a test set does not help the situation either

		\item Related shortcoming: specialized semantics for specific lexical entries

\end{itemize}\end{itemize}\end{itemize}

\end{frame}

\begin{frame}[fragile]
	\frametitle{Future Work}

	\begin{itemize}

		\item Address the issue of
		automation

			\begin{itemize}

				\item Define a decidable fragment of
				the logic and only work within such fragment

	\begin{itemize}

		\item Possible to concisely characterize how the approach generalises

\end{itemize}

\item Train a neural network on a body of freely available proofs on the net and see whether it can generalize to automatically provide the proof tactics for the cases interested

\end{itemize}

\item Improvement at the GF level: make the abstract syntax more compatible with compositional semantics

	\begin{itemize}

		\item For example, do something about problematic syntactic categories like pre-determiners or cases where the syntax makes it impossible to recover elliptical fragments

\end{itemize}

\item Extend into the whole suite (first attempt to do anaphora using monads on the way!)

\end{itemize}

\end{frame}

\begin{frame}[fragile]
	\frametitle{Conclusions}

	\begin{itemize}

		\item We have connected two well-defined systems based on
		type-theory

		\begin{itemize}

			\item GF and Coq

			\item Providing resource semantics for GF

\end{itemize}

\item The issue of generalization remains a shortcoming

	\begin{itemize}

		\item It is possible
		to achieve very precise semantics for specific domains

\begin{itemize}

	\item Our system outerperforms previous logical systems w.r.t accuracy

\end{itemize}\end{itemize}

\item Useful in performing inference tasks on controlled natural
language domains

\item Hybrid NLI systems

\end{itemize}

\end{frame}

\begin{frame}[fragile]
	\frametitle{Conclusions}

	\begin{itemize}
		\item The system can be found here:  \href{https://github.com/GU-CLASP/FraCoq}{https://github.com/GU-CLASP/FraCoq}

\end{itemize}

\end{frame}

\end{document}

### 2019-08-29 nodalida-talk/talk.tex
nodalida talk (draft)

--- 2019-08-29 IWCS_2017/IWCSTALK1.tex
@@
 \usepackage{gb4e}
+\usepackage{tikz}
+\usepackage{enumitem}
+\usepackage{tabularx}
+\usetikzlibrary{arrows,automata}
 \usetheme{Boadilla}
+% \usetheme{Warsaw}
 \newcommand{\Sem}[1]{\sem{#1}}
-\logo{\includegraphics{clasp.ps}} 
-\renewcommand{\theenumi}{\alph{enumi}}
+\logo{\includegraphics{CLASP_logo.jpeg}} 
+% \renewcommand{\theenumi}{\alph{enumi}}
+\setitemize{label=\usebeamerfont*{itemize item}%
+  \usebeamercolor[fg]{itemize item}
+  \usebeamertemplate{itemize item}}
+\newlist{lingex}{enumerate}{3}
+\setlist[lingex,1]{parsep=0pt,itemsep=1pt,label=(\arabic*),resume=lingexcount}
+\newcommand\onelingex[1]{\begin{lingex}\item #1 \end{lingex}}
+\providecommand\ncases[1]{{\ensuremath{^{#1}}}}
+
+% \newcommand{\ex}{}
+% \newenvironment{exe}{}{}
+
 \input{texhacks-beamer1}
@@
     {\textbf{Bernardy and Chatzikyriakidis\hfill \textsc{IWCS2017, Montpellier} \hfill \insertframenumber/\inserttotalframenumber}}%
-}}\title{A Type-Theoretical system for the FraCaS test suite: Grammatical Framework meets Coq
-}
+  }}\title{FraCoq 2.0: A Wide-Coverage Symbolic Natural Language Inference System}
 \author{Jean-Philippe Bernardy and Stergios Chatzikyriakidis}
 \begin{document}
-\date{September 20, 2017}
+\date{October 2019}
 \frame{\titlepage}
-\begin{frame}[fragile]
-	\frametitle{Brief summary of the talk}
 
-	\begin{itemize}
-		\item  Present a type-theoretical framework for formal semantics leveraging two well-studied tools
-		\begin{itemize}
-			\item Grammatical Framework (GF, Ranta 2004, 2011)
-			\item Coq
-
-	\end{itemize}
-	\item Providing a compositional resource semantics for GF
-
-	\item Evaluation on the FraCaS test suite
-
-	\item State-of-the-art results	\end{itemize}
-\end{frame}
-
-\begin{frame}[fragile]
+\begin{frame}
   \frametitle{Structure the talk}
-
   \begin{itemize}
   \item Brief intro to the systems used, GF and Coq
-
-		\item Presenting the FraCoq system
-
+  \item Presenting the FraCoq (2.0!) system
     \begin{itemize}
-			\item We concentrate on the most linguistically relevant features and also the features relevant for the FraCaS
-
+    \item We concentrate on the novel features of the system (Dynamic Semantics, Anaphora
+      and Ellipsis mostly)
     \end{itemize}
-
   \item Evaluation against the FraCaS test suite
-
-	\begin{itemize}
-		\item Some brief remarks about the FraCaS and NLI platforms
-			\begin{itemize}
-
-			\item Results and comparison with previous logical approaches
-
-			\item The issue of automation
-
-\end{itemize}\end{itemize}
-
   \item Conclusions and Future work
-
   \end{itemize}
@@
 
-\begin{frame}[fragile]
-	\frametitle{Background: Grammatical Framework (GF)}
-
+\begin{frame} \frametitle{Background: Grammatical Framework (GF)}
   \begin{itemize}
-	\item Functional programming language for writing multi-lingual grammars
-
+  \item Functional programming language for writing multi-lingual
+grammars
   \item Involves an abstract syntax, comprised of:
-
     \begin{itemize}
-	\item  A number of syntactic
-	categories
-
-	\item A number of syntactic construction functions, which
-	provide the means to compose basic syntactic categories into more complex
-	ones
-
+    \item A number of syntactic categories
+    \item A number of syntactic construction functions, which provide
+the means to compose basic syntactic categories into more complex ones
       \begin{itemize}
-	\item
-
-	$AdjCN : AP \to CN \to CN$ (appending an adjectival phrase to a common noun and obtaining a new common noun)
+      \item $AdjCN : AP \to CN \to CN$ (appending an adjectival phrase to
+a common noun and obtaining a new common noun)
       \end{itemize}\end{itemize}
-\item GF comes with a library of mappings from abstract
-syntax to concrete
+  \item GF comes with a library of mappings from abstract syntax to
+concrete
     \begin{itemize}
-	\item These mappings can be inverted by GF, thus offering
-parsers from natural text into abstract syntax
+    \item These mappings can be inverted by GF, thus offering parsers
+from natural text into abstract syntax
 
     \item We use the parse trees constructed by Ljungl\"of (2012)
-thereby avoiding any syntactic ambiguity (GF FraCaS treebank).\end{itemize}\end{itemize}
+thereby avoiding any syntactic ambiguity (GF FraCaS
+treebank).\end{itemize}\end{itemize}
 \end{frame}
 
-\begin{frame}[fragile]
+\begin{frame}
   \frametitle{Background: Type-Theoretical Semantics}
@@
   \item We use the type of logics that have been traditionally dubbed as constructive
-
     \begin{itemize}
     \item Initiated by the work of Martin-L\"of (1971, 1984)
-
     \item In linguistics this types of logics go back to Ranta's seminal work (Ranta 1994) or even earlier to Sundholm (1988)
       \begin{itemize}
-
-				\item More recent approaches can be found as well. Please see Chatzikyriakidis and Luo (2017) for a collection of papers on constructive type theories for NL semantics
-
+      \item More recent approaches can be found as well. Please see
+        Chatzikyriakidis and Luo (2017) for a collection of papers on
+        constructive type theories for NL semantics
       \end{itemize}\end{itemize}\end{itemize}
-
 \end{frame}
 
-\begin{frame}[fragile]
+\begin{frame}
   \frametitle{Background: Type Theoretical Semantics}
-
   \begin{itemize}
   \item Main features of MTTs
-
     \begin{itemize}
     \item  Type many-sortedness.
-
     \item Dependent sum and product types
-
       \begin{itemize}
-					\item  $\Sigma$-types, often written
-					$\sum_{x:A}B[x]$ and which have product types $A\times B$ as a
-					special case when $B$ does not depend on $x$.
-
+      \item $\Sigma$-types, often written $\sum_{x:A}B[x]$ and which
+        have product types $A\times B$ as a special case when $B$ does
+        not depend on $x$.
       \item Dependent product, $\Pi$-types, often written
@@
         special case
-
-					\item They generalize universal quantification and function types and they offer type polymorphism
-
+      \item They generalize universal quantification and function
+        types and they offer type polymorphism
       \end{itemize}
-
     \item Proof-theoretical specification and support for effective
       reasoning.
-
       \begin{itemize}
       \item Most powerful proof assistants implement MTTs (e.g. Coq, Agda)
-
       \end{itemize}\end{itemize}\end{itemize}
@@
 
-\begin{frame}[fragile]
+\begin{frame}
   \frametitle{Background: Coq}
-
   \begin{itemize}
-
   \item	Proof assistant based on the calculus of inductive
     constructions (CiC, Coquand and Paulin-Mohring 1988)
-
     \begin{itemize}
-
     \item Arguably one of the leading proof assistants
       \begin{itemize}
-
       \item  a proof of the four-color
@@
 
-\begin{frame}[fragile]
+\begin{frame}
   \frametitle{Background: Coq}
-
   \begin{itemize}
-
   \item Important features used
-
     \begin{itemize}
-
     \item $\Pi$ types 
-
       \begin{itemize}
-
       \item	in Coq: $\prod_{x:A}B[x]$ is written \texttt{forall (x:A), B} or (simply
@@
       \begin{itemize}
-
       \item  Generalization of $\Sigma$-types and are encoded as (trivial) inductive types with a single
         constructor.
-
       \item	$\Sigma x:A. B(x)$ can be expressed as a dependent record type in Coq:
-
-\begin{verbatim}
-Record AB:Type:=mkAB{x:> A;P:B x}
-\end{verbatim}
+        \texttt{Record AB:Type:=mkAB\{x:> A;P:B x\}}
       \end{itemize}\end{itemize}\end{itemize}
-
 \end{frame}
 
-\begin{frame}[fragile]
-	\frametitle{The FraCoq system}
+\begin{frame}{Overview of the system}
+{
+\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3.8cm, semithick,scale=0.4, every node/.style={scale=0.8}]
+  \tikzstyle{every state}=[draw=black,text=black,shape=rectangle,text width=2cm]
 
+  \node[state]         (txt)                {text};
+  \node[state]         (ast) [right of=txt] {AST};
+  \node[state]         (pp)  [right of=ast] {Program};
+  \node[state]         (res) [right of=pp] {result in $(0,1)$};
+
+  \path (txt) edge              node (GF)  {GF}                  (ast)
+        (ast) edge              node (Hask) [text width=1.5cm]{Haskell} (pp)
+        (pp) edge               node (Coq) {Coq} (res);
+
+  \node (Syntax) [above of=GF,node distance=1cm] {Syntax};
+  \node (Sem) [above of=Hask,node distance=1cm] {Dynamic Semantics};
+  \node (Hyper) [above of=Coq,node distance=1cm] {Additional Semantics$^\ast$};
+
+  \path
+  (Syntax) edge (GF)
+  (Sem) edge (Hask)
+  (Hyper) edge (Coq);
+\end{tikzpicture}
+}
   \begin{itemize}
-
-		\item We use Lj\"unglof's FraCaS treebank and take these trees to their semantic counterparts
-
-			\item The structure of the semantic representation
-
-		\begin{enumerate}
-			\item Every GF syntactic category $C$ is mapped to a Coq $Set$, noted $\sem{C}$.
-			\item GF Functional types are mapped compositionally : $\sem{A \to B}$ $=$ $\sem{A} \to \sem{B}$
-			\item Every GF syntactic construction function $f : X$ is mapped to a function $\sem{f}$ such that $\sem{f} : \sem{X}$.
-			\item GF function applications are mapped compositionally: $\sem{t(u)} = \sem{t} (\sem{u})$.
-		\end{enumerate}
-
+  \item We use Lj\"unglof's FraCaS treebank and take these trees to
+    their semantic counterparts
+  \item GF constructions are translated to Haskell programs (types are
+    respected). Dynamic semantics are handled at this level.
+  \item The programs output Coq propositions and types. At the level
+    of Coq, we add several kinds of linguistically relevant
+    information (veridicality properties; adjectival semantics such as
+    division into subsective, extentional, non-committal, etc.)
   \end{itemize}
@@
 
-\begin{frame}[fragile]
-	\frametitle{The FraCoq system}
-
+\begin{frame}
+  \frametitle{FraCoq: Dynamic Semantics}
   \begin{itemize}
-
-\item Sentences
-
+  \item Monadic structure \cite{Shan:2002}
+  \item Two kind of effects
     \begin{itemize}
-
-		\item We interpret sentences as propositions: $\sem{S} =
-		Prop$.
-
-		\item To verify that
-		$P$ entails $H$, we prove the proposition $\sem{P} \to \sem{H}$.
-
-		\begin{verbatim}
-Definition S := Prop.
-		\end{verbatim}
-
-\end{itemize}
-\item Common Nouns
-	\begin{itemize}
-
-		\item Predicates over an abstract object type
-
-			\begin{verbatim}
-Parameter object : Set.
-Definition CN := object->Prop.
-			\end{verbatim}
-
-\end{itemize}\end{itemize}
-\end{frame}
-\begin{frame}[fragile]
-	\frametitle{The FraCoq system}
-
-	\begin{itemize}
-
-		\item Verb phrases
-
-		\begin{itemize}
-
-			\item Parameterize over the
-			\emph{noun} of the subject (using $\Pi$ types)
-
-			\begin{verbatim}
-  Definition VP := forall (subjectClass : CN) 
-  object -> Prop.
-			\end{verbatim}	
+    \item Manage a state of referents for anaphoric expressions (NPs,
+      but also other linguistic categories)
+    \item Non-determinism. (Linguistic expressions can have several
+      interpretations.)
     \end{itemize}
@@
 
-\begin{frame}[fragile]
-	\frametitle{The FraCoq system}
-
+\begin{frame}
+  \frametitle{The FraCoq system: basic idea}
   \begin{itemize}
-
-\item Adjectives
-
+  \item Sentences are interpreted as propositions (with effect). 
+  \item VP and CN are interpreted as predicates (also with effect).
+  \item Adjectives: Functions from CN to CN (again: effect!)
     \begin{itemize}
-
-		\item Functions from cn to cn (predicates to predicates)
-
-		\begin{verbatim}
-Definition A := CN -> CN.
-		\end{verbatim}
-
     \item Different classes of adjectives are captured using coercions (subtyping). All special classes of adjectives are subtypes of A.
-
-\begin{verbatim}
-Definition IntersectiveA := object -> Prop.
-Definition wkIntersectiveA : IntersectiveA -> A
-:= fun a cn (x:object) => a x /\ cn x.
-Coercion wkIntersectiveA : IntersectiveA >-> A.
-\end{verbatim}
-
     \item Provision is made for intersective, subsective, privative and non-committal adjectives
-
-\end{itemize}\end{itemize}
+    \end{itemize}
+  \item (Generalized) Quantifiers. Similar to Montague semantics, but
+    we also take into account quantities using an abstract function
+    which can count the cardinalities of sets. (plus: effects, see below.)
+  \end{itemize}
 \end{frame}
 
-\begin{frame}[fragile]
-	\frametitle{The FraCoq system}
-
+\begin{frame}
+  \frametitle{Anaphora}
   \begin{itemize}
-
-		\item Adverbs
-
+  \item Wide-coverage: E-Type pronouns and Donkey anaphora, etc.
+  \item New scope-extension mechanism
+    \begin{lingex}
+    \item
+      \label{ex:anaphora-scope}
+      \textbf{P1}	Every committee has a chairman.\\
+      \textbf{P2}	He is appointed its members.\\
+      \textbf{H} 	Every committee has a chairman appointed by members of the committee. \\ \textbf{Label:	YES} [FraCaS 122]
+    \end{lingex}
     \begin{itemize}
-
-			\item Similar method to adjectives but instead the modification is on verbal predicates
-
-			\item The adverb cases in the FraCaS are all veridical and covariant.
-
-			\item We define such a subclass $VeridicalAdv$ and declare it as a coercion $Adv$
-
-				\begin{itemize}
-
-					\item Adverbs of type $VeridicalAdv$ are also of type $Adv$
-
-\begin{verbatim}			
-Parameter on_time_Adv : VeridicalAdv .	
-\end{verbatim}
-
-\end{itemize}\end{itemize}\end{itemize}
+    \item ``Every'' introduces a referent in the context.
+    \item ``He'' find this referent.
+    \item Normally, the referent would be out of scope, but we
+      liberally extend scopes until we obtain a valid logical formula.
+    \end{itemize}
+  \end{itemize}
 \end{frame}
 
-\begin{frame}[fragile]
-	\frametitle{The FraCoq system}
-
+\begin{frame}
+  \frametitle{Ellipsis}
   \begin{itemize}
-
-		\item Noun Phrases and Predeterminers
-
-	\begin{itemize}
-
-		\item A clean definition of NPs as functions from predicates to truth values will not work
-
-			\begin{itemize}
-
-				\item	Problem with GF's abstract syntax: existence of pre-determiners which include cases like \textit{most}, \textit{all} among others and are defined as functions from NPs to NPs
-
-				\item In general, the category includes elements that are naturally interpreted as GQs
-
-				\item Solution: Remember the components of NPs (number, quantifier and common noun)
-
-				\item Pre-determiners then are able to substistute the quantifier part with the appropriate quantifier
-
-				\item This has to be done, otherwise pre-determiners introduce a dummy indefinite in these cases
-
-\end{itemize}\end{itemize}\end{itemize}
+  \item Similar to anaphora
+  \item
+    \begin{lingex}
+    \item
+      \label{ex:acda}
+      \textbf{P1}	Bill spoke to everyone that John did [elliptic V2].\\
+      \textbf{P2}	John spoke to Mary.\\
+      \textbf{Q} 	Did Bill speak to Mary?\\
+      \textbf{H} 	Bill spoke to Mary. \\ \textbf{Label:	Yes} [FraCas 173]
+    \end{lingex}
+  \item the verb ``spoke'' is introduced as a referent in the context when its encountered
+  \item the argument of ``did'' is explicitly marked as an elliptic V2 by GF.
+  \item we resolve it to ``speak'' by loading it from the environment.
+  \end{itemize}
 \end{frame}
 
-\begin{frame}[fragile]
-	\frametitle{The FraCoq system}
-
+\begin{frame}{Phrasal comparatives}
   \begin{itemize}
-
-		\item Generalized Quantifiers
-
+  \item Handled using the dynamic semantics
+  \item Example:
     \begin{itemize}
-
-			\item They turn a number
-			and a common noun into a noun-phrase (which we call $NP0$).
-
-\begin{verbatim}
-	Definition Quant := Num -> CN -> NP0.
-\end{verbatim}
-
-\item Some quantifiers ignore the number and are given usual definitions (e.g. \textit{some} or \textit{all}), whereas others make essential use of number (e.g. \textit{at most})
-
-\begin{itemize}
-
-\item In the latter case, the function \texttt{CARD}, a context-dependent
-abstract function which turns a predicate into a natural number	is used to get the correct semantics 	
-\end{itemize}\end{itemize}\end{itemize}
+    \item ``ITEL won more orders than APCOM lost.''
+    \item becomes: ``APCOM lost some orders. ITEL won more
+      orders [than some elliptic quantity].''
+    \item Where is the elliptic quantity introduced?
+    \item ``APCOM lost $\theta$ orders.  ITEL won at least $\theta$+1 orders.''
+    \end{itemize}
+  \end{itemize}
+  (Simplified, see the paper for details.)
 \end{frame}
 
-\begin{frame}[fragile]
-	\frametitle{The FraCoq system}
-
+\begin{frame}{Gradable adjectives}
+  Example: tall.
   \begin{itemize}
-
-		\item The definite article
-
-	\begin{itemize}
-
-		\item Checks for plural noun phrases
-
-			\begin{itemize}
-
-				\item If found, then universal quantification
-
-				\item If not, it
-				looks up the object of discourse in an abstract $environment$, which
-				is a function which turns a common noun into an object
-
-\end{itemize}\end{itemize}\end{itemize}
-\begin{verbatim}
-	Definition DefArt:Quant:= fun (num : Num) (P:CN)=> fun Q:VP
-	=> match num with plural => (forall x, P x -> Q P x)
-	 /\ Q P (environment P) /\ P (environment P) |
-	_ => Q P (environment P) /\ P (environment P) end.
-\end{verbatim}
-
+  \item Interpreted as a pair of a measure $tallness : objects \to Z$, and a threshold $minTall : Z$.
+    \item An individual $x$ is considered $tall$ iff $tallness\, x > minTall$.
+  \end{itemize}
 \end{frame}
 
-\begin{frame}[fragile]
+\begin{frame}
   \frametitle{Evaluation: The FraCaS test suite}
-
   \begin{itemize}
-
   \item A test suite for NLI
-
     \begin{itemize}
-
-		\item 346 NLI examples in the form of one or more premises followed
-		by a question along with an answer to that question
-
+    \item 346 NLI examples in the form of one or more premises
+      followed by a question along with an answer to that question
     \item Three potential answers
-
       \begin{itemize}
-
       \item YES: The declarative sentence formed out of the question follows from the premises
-
       \item NO: The declarative sentence does not follow from the premises
-
       \item UNK: The declarative sentence neither follows nor does not follow fro the premises
@@
 
-\begin{frame}[fragile]
+\begin{frame}
   \frametitle{Evaluation: The FraCaS test suite}
@@
 
-\begin{frame}[fragile]
-	\frametitle{Evaluation: The FraCaS test suite}
-
+\begin{frame}
+  \frametitle{Evaluation}
   \begin{itemize}
-
-		\item The FraCaS has considerable weaknesses
-
+  \item Evaluation against 8 sections of the FraCaS
     \begin{itemize}
-
-			\item Small size
-
-			\item Artificial nature of the
-			examples
-
+    \item Total of 259 examples
+    \item Only exclude the ``temporal-reference'' section.
     \end{itemize}
-
-\item However, it covers a lot of phenomena associated with NLI
-
-\item It is still a very good suite to test logical approaches\end{itemize}
-\end{frame}
-
-\begin{frame}[fragile]
-	\frametitle{Evaluation}
-
-	\begin{itemize}
-
-		\item Evaluation against 5 sections of the FraCaS
-
-		\begin{itemize}
-
-			\item Total of 174 examples
-
-		\item Excluded sections where a lot of context-dependency has to be taken into consideration (e.g. the section on ellipsis)
-
-\end{itemize}
-
-\item YES: a proof
-can be constructed from the premises to the hypothesis
-
-\item NO: a proof
-of the negated hypothesis can be constructed
-
+  \item YES: a proof can be constructed from the premises to the hypothesis
+  \item NO: a proof of the negated hypothesis can be constructed
   \item UNK: otherwise
-
   \end{itemize}
@@
 
-	\begin{frame}[fragile]
+\begin{frame}
   \frametitle{Evaluation}
+  \small
+  \begin{tabularx}{9cm}{Xr@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r}
+Section      & {\kern -2em} \#cases & FC2     & FC1 & MINE & Nut  & Langpro  \\ \hline
+Quantifiers  & 75          & .96  & .96    & .77  & .53  & .93  \\
+             &             & \ncases{74}     &        &      &      &     \ncases{44} \\
+Plurals      & 33          & .82      & .76    & .67  & .52  & .73 \\
+      &          &       &     &   &   & \ncases{24} \\
+Anaphora     & 28          & .86      &   -    & -    & -    &  -       \\
+Ellipsis     & 52          & .87      &   -    & -    & -    &  -       \\
+Adjectives   & 22          & .95 & .95    & .68  & .32  & .73 \\
+             &          &  \ncases{20} &     &   &   &  \ncases{12} \\
+Comparatives {\kern -1em}& 31          & .87      & .56    & .48  & .45  &  -       \\
+Temporal     & 75          &  -       &   -    &   -  &  -   &  -       \\
+Verbs        & 8           & .75      &   -    & -    & -    &  -       \\
+Attitudes    & 13          & .92      & .85    & .77  & .46  & .92  \\ 
+    &          &       &     &   &   & \ncases {9}  \\ \hline
+Total        & 337         & .89      & .83    & .69  & .50  & .85  \\
+             &             & \ncases{259}    & \ncases{174}  & \ncases{174}& \ncases{174}& \ncases{89}
+  \end{tabularx}
 
   \begin{itemize}
-
-			\item		The following table presents the results (Ours) as well as a comparison with the approach in Mineshima et al. (MINE, 2015), Bos (Nut, 2008) and Abzianidze (Langpro, 2015)
-
-			\begin{center}
-				\begin{tabular}{rlrrrrr}
-					& Section                                                                               & \# examples                                                                                                & \begin{tabular}[c]{@{}l@{}}Ours\end{tabular}                                                        & \begin{tabular}[c]{@{}l@{}}MINE \end{tabular}                                             & \begin{tabular}[c]{@{}l@{}}Nut\end{tabular}   & \begin{tabular}[c]{@{}l@{}} Langpro \end{tabular}                                                     \\ \hline
-					\begin{tabular}[c]{@{}l@{}}1\\ 2\\ 3\\ 4\\5\\6\end{tabular} & \begin{tabular}[c]{@{}l@{}}Quantifiers\\ Plurals\\ Adjectives\\Comparatives\\ Attitudes\\ Total\end{tabular} & \begin{tabular}[c]{@{}l@{}}75\\ 33\\22\\ 31\\13\\174 (181)\end{tabular} & \begin{tabular}[c]{@{}l@{}}.96\\ .76 \\ .95 \\ .56    \\ .85\\ 0.83 \end{tabular} & \begin{tabular}[c]{@{}l@{}}.77\\ .67\\ .68 \\   .48      \\.77\\0.69 \end{tabular} & \begin{tabular}[c]{@{}l@{}} .53\\ .52\\ .32\\ .45\\.46\\0.50\end{tabular} & \begin{tabular}[c]{@{}l@{}} .93 (44)
-						\\ .73 (24)\\ .73 (12)\\ -\\.92 (9)\\0.85\end{tabular}
-				\end{tabular}
-			\end{center}
-
-\item Our approach outerperforms Mineshema et al. by 13 percentage points.
-
-\item The approach by Abzianidze has an accuracy of 0.85 without involving the comparative section. If this section is taken out, our system's accuracy rises to 0.88
-
+  \item ``FC2" refers to the approach presented in this paper.
+  % \item When a system does not handle the nominal number of test cases (shown in the second column), the actual number of test cases attempted is shown below the accuracy figure, in smaller font.
+  % \item
+    ``FC1'' refers to \cite{bernardy:2017}.
+  % \item
+    ``MINE" refers to \cite{Mineshima:2015},
+  % \item
+    ``NUT" to the CCG+\textit{nutcracker} \cite{bos:2008},
+  % \item
+    ``Langpro" to \cite{Abzianidze:2015}.
+  % \item A dash indicates that no attempt was made for the section.
   \end{itemize}
@@
 
-	\begin{frame}[fragile]
+\begin{frame}
   \frametitle{Error Analysis}
@@
   \begin{itemize}
-
-			\item	Improvement over earlier approaches. Still, there were a couple of difficulties
-
+  \item	Improvement over earlier approaches. Still, there were a couple of difficulties, notably:
     \begin{itemize}
-
-					\item Comparatives: cases that needed one to provide
-					adequate semantics for \textit{more} but also to take care of ellipsis
-
+    \item Definite Plurals: Universal reading was captured. Cases of existential readings were not
       \begin{exe}
 	\ex	
-	\hskip 0.05in ITEL won more orders than APCOM.
-
-	\hskip 0.05in ITEL won some orders.
-
-	\hskip 0.05in  	Did ITEL win some orders? [Yes, FraCaS 233]
-
+        \hskip 0.05in The inhabitants of Cambridge voted for a Labour MP.
+        \hskip 0.05in very inhabitant of Cambridge voted for a Labour MP.
+        \hskip 0.05in  	Did every inhabitant of Cambridge vote for a Labour MP? [UNK, FraCaS 094]			
       \end{exe}
-\item Definite Plurals: Universal reading was captured. Cases of existential readings were not
-
-\begin{exe}
-	\ex
-
-\hskip 0.05in The inhabitants of Cambridge voted for a Labour MP.
-
-\hskip 0.05in very inhabitant of Cambridge voted for a Labour MP.
-
-\hskip 0.05in  	Did every inhabitant of Cambridge vote for a Labour MP? [UNK, FraCaS 094]
-
-		\end{exe}
-
+    \item A syntactic conjunction of individuals is treated as a
+      semantic group, or where precise counting of the members of a
+      group is necessary.
     \end{itemize}\end{itemize}
@@
 
-\begin{frame}[fragile]
-	\frametitle{Automation}
-
+\begin{frame}
+  \frametitle{Future Work}
   \begin{itemize}
-
-		\item So far, our proofs are not automated
-
-		\begin{itemize}
-
-			\item A couple of steps (usually very few) to reach a proof
-
-			\item Earlier approaches using Coq (e.g. Chatzikyriakidis 2014 and Mineshima et al. 2015) use Coq's tactical language LTac to define automated macros of actions
-
-				\begin{itemize}
-
-					\item This is not difficult to do in our case as well
-
-					\item Just go through all the proof
-					tactics or observe the tactics that are used in the proofs to create a
-					macro that will automate the proofs
-
-					\item The question remains: can that macro of tactics
-					generalize outside the suite?
-
-					\item Answer: only to a limited
-					extent, i.e. when exactly the same set of tactics yields a proof
-
-					\item For
-					this reason, we have not automated proof search to obtain the results
-					presented in this paper, even though this can be done easily
-
-\end{itemize}\end{itemize}
-
+    \item Improve support for group readings.
+    % \item Improvement at the GF level: make the abstract syntax more compatible with compositional semantics
+    \item Address the last section of FraCaS test suite (it contains
+      many \textit{adhoc} cases, and much preliminary work to do to
+      define general approach.)
   \end{itemize}
-
 \end{frame}
 
-\begin{frame}[fragile]
-	\frametitle{Automation}
-
+\begin{frame}
+  \frametitle{Conclusions}
   \begin{itemize}
-
-		\item Automating would also make an unprincipled use of
-		higher-order logic (HOL)
-
-			\begin{itemize}
-
-				\item No algorithm which
-				can decide if a proposition has a proof or not
-
-\begin{itemize}
-
-	\item We
-	must use heuristics both to search for proofs and to decide when to
-	give up searching
-
-\end{itemize}\end{itemize}
-
-\item  Most problems have either
-obvious proofs or obviously lack a proof (fortunately)
-
-\begin{itemize}
-
-	\item Due to its heuristic nature the
-	proof search necessarily contains a human component
-
-	\begin{itemize}
-
-		\item Problematic to make a statement about the suitability of FraCoq
-		outside FraCas
-
-		\item Small dataset and lack of separation between a development and a test set does not help the situation either
-
-		\item Related shortcoming: specialized semantics for specific lexical entries
-
-\end{itemize}\end{itemize}\end{itemize}
-
-\end{frame}
-
-\begin{frame}[fragile]
-	\frametitle{Future Work}
-
-	\begin{itemize}
-
-		\item Address the issue of
-		automation
-
-			\begin{itemize}
-
-				\item Define a decidable fragment of
-				the logic and only work within such fragment
-
-	\begin{itemize}
-
-		\item Possible to concisely characterize how the approach generalises
-
-\end{itemize}
-
-\item Train a neural network on a body of freely available proofs on the net and see whether it can generalize to automatically provide the proof tactics for the cases interested
-
-\end{itemize}
-
-\item Improvement at the GF level: make the abstract syntax more compatible with compositional semantics
-
-	\begin{itemize}
-
-		\item For example, do something about problematic syntactic categories like pre-determiners or cases where the syntax makes it impossible to recover elliptical fragments
-
-\end{itemize}
-
-\item Extend into the whole suite (first attempt to do anaphora using monads on the way!)
-
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}[fragile]
-	\frametitle{Conclusions}
-
-	\begin{itemize}
-
   \item We have connected two well-defined systems based on
     type-theory
-
     \begin{itemize}
-
-			\item GF and Coq
-
+    \item Combines the strengths of 3 systems: GF (Parsing), Haskell
+      (dynamic semantics), and Coq (powerful reasoning mechanisms).
     \item Providing resource semantics for GF
-
     \end{itemize}
-
-\item The issue of generalization remains a shortcoming
-
-	\begin{itemize}
-
-		\item It is possible
-		to achieve very precise semantics for specific domains
-
-\begin{itemize}
-
-	\item Our system outerperforms previous logical systems w.r.t accuracy
-
-\end{itemize}\end{itemize}
 
@@
   \item Hybrid NLI systems
-
+  \item The system can be found here:  \href{https://github.com/GU-CLASP/FraCoq}{https://github.com/GU-CLASP/FraCoq}
   \end{itemize}
-
 \end{frame}
 
-\begin{frame}[fragile]
-	\frametitle{Conclusions}
+\bibliography{jp,phorae,IWCSbib,NLI}
 
-	\begin{itemize}
-		\item The system can be found here:  \href{https://github.com/GU-CLASP/FraCoq}{https://github.com/GU-CLASP/FraCoq}
-
-\end{itemize}
-
-\end{frame}
+\bibliographystyle{plain}
 

### 2019-09-30 nodalida-talk/talk.tex
crucify gb4e
oops
corrections
additions

@@
 \usepackage{beamerthemesplit}
+\usepackage{graphicx}
 \usepackage{latexsym}
@@
 \usepackage{amssymb}
-\usepackage{gb4e}
 \usepackage{tikz}
@@
 \newcommand{\Sem}[1]{\sem{#1}}
-\logo{\includegraphics{CLASP_logo.jpeg}} 
-% \renewcommand{\theenumi}{\alph{enumi}}
+%\logo{\includegraphics[height=0.7cm]{CLASP_logo.jpeg}} 
+\logo{\includegraphics{clasp.ps}} 
 \setitemize{label=\usebeamerfont*{itemize item}%
@@
 \newcommand\onelingex[1]{\begin{lingex}\item #1 \end{lingex}}
+\newcommand{\nologo}{\setbeamertemplate{logo}{}} % command to set the logo to nothing
 \providecommand\ncases[1]{{\ensuremath{^{#1}}}}
 
-% \newcommand{\ex}{}
-% \newenvironment{exe}{}{}
+\newcommand{\ex}{\item}
+\newenvironment{exe}{\begin{lingex}}{\end{lingex}}
 
@@
   \node[state]         (pp)  [right of=ast] {Program};
-  \node[state]         (res) [right of=pp] {result in $(0,1)$};
+  \node[state]         (res) [right of=pp] {Result};
 
@@
     information (veridicality properties; adjectival semantics such as
-    division into subsective, extentional, non-committal, etc.)
+    division into subsective, extensional, non-committal, etc.)
   \end{itemize}
@@
 \begin{frame}
-  \frametitle{FraCoq: Dynamic Semantics}
+  \frametitle{FraCoq 2.0: Dynamic Semantics}
   \begin{itemize}
@@
 \begin{frame}
-  \frametitle{The FraCoq system: basic idea}
+  \frametitle{The FraCoq 2.0 system: basic idea}
   \begin{itemize}
@@
       \textbf{P1}	Every committee has a chairman.\\
-      \textbf{P2}	He is appointed its members.\\
+      \textbf{P2}	He is appointed by its members.\\
       \textbf{H} 	Every committee has a chairman appointed by members of the committee. \\ \textbf{Label:	YES} [FraCaS 122]
@@
     \item ``Every'' introduces a referent in the context.
-    \item ``He'' find this referent.
+    \item ``He'' finds this referent.
     \item Normally, the referent would be out of scope, but we
@@
       \textbf{P2}	John spoke to Mary.\\
-      \textbf{Q} 	Did Bill speak to Mary?\\
       \textbf{H} 	Bill spoke to Mary. \\ \textbf{Label:	Yes} [FraCas 173]
     \end{lingex}
-  \item the verb ``spoke'' is introduced as a referent in the context when its encountered
-  \item the argument of ``did'' is explicitly marked as an elliptic V2 by GF.
-  \item we resolve it to ``speak'' by loading it from the environment.
+  \item The verb ``spoke'' is introduced as a referent in the context when its encountered
+  \item The argument of ``did'' is explicitly marked as an elliptic V2 by GF.
+  \item We resolve it to ``speak'' by loading it from the environment.
   \end{itemize}
@@
   \end{itemize}
-  (Simplified, see the paper for details.)
+  (Simplified, see the paper for details)
 \end{frame}
@@
 
+{\nologo
 \begin{frame}
@@
 
-\end{frame}
+\end{frame}		}
 
@@
     \item Definite Plurals: Universal reading was captured. Cases of existential readings were not
-      \begin{exe}
-	\ex	
-        \hskip 0.05in The inhabitants of Cambridge voted for a Labour MP.
-        \hskip 0.05in very inhabitant of Cambridge voted for a Labour MP.
-        \hskip 0.05in  	Did every inhabitant of Cambridge vote for a Labour MP? [UNK, FraCaS 094]			
-      \end{exe}
+
+   \begin{lingex}
+  	\item
+  	\label{ex:acda}
+  	\textbf{P1}	The inhabitants of Cambridge voted for a Labour MP.\\
+  	\textbf{P2}	Every inhabitant of Cambridge voted for a Labour MP.\\
+  	\textbf{H} Every inhabitant of Cambridge voted for a Labour MP.	\\ \textbf{Label:	UNK} [FraCas 094]
+  \end{lingex}
     \item A syntactic conjunction of individuals is treated as a

### 2019-09-30 nodalida-talk/talk.tex
additions

@@
   	\textbf{P1}	The inhabitants of Cambridge voted for a Labour MP.\\
-  	\textbf{P2}	Every inhabitant of Cambridge voted for a Labour MP.\\
   	\textbf{H} Every inhabitant of Cambridge voted for a Labour MP.	\\ \textbf{Label:	UNK} [FraCas 094]
   \end{lingex}
+
+    \end{itemize}\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Error Analysis}
+
+\begin{itemize}
+	\item Group readings
+
+	\begin{itemize}
+
 		\item A syntactic conjunction of individuals is treated as a
@@
 	group is necessary.
-    \end{itemize}\end{itemize}
+
+	\end{itemize}
+
+		\begin{lingex}
+			\item
+			\label{ex:acda}
+			\textbf{P1}		Smith signed one contract.\\
+						\textbf{P2}		Jones signed another contract.\
+			\textbf{H} Smith and Jones signed two contracts.\\ \textbf{Label:	YES, on a collective reading of the sentence} [FraCas 111]
+		\end{lingex}
+
+\end{itemize}
 
@@
   \begin{itemize}
-    \item Improve support for group readings.
+    \item Improve support for group readings
     % \item Improvement at the GF level: make the abstract syntax more compatible with compositional semantics
-    \item Address the last section of FraCaS test suite (it contains
-      many \textit{adhoc} cases, and much preliminary work to do to
-      define general approach.)
-  \end{itemize}
+    \item Address the  section of  the FraCaS test suite  left, i.e. section 7: Temporal reference (tense and aspect)
+
+     \begin{itemize}
+    	\item  Contains
+      many \textit{adhoc} cases, and much preliminary work is needed in order to
+      define general approach
+  \end{itemize}  \end{itemize}
 \end{frame}
@@
   \begin{itemize}
-  \item We have connected two well-defined systems based on
+  \item We have connected three well-defined systems based on
     type-theory
@@
 
-  \item Hybrid NLI systems
+  \item Hybrid NLI systems?
   \item The system can be found here:  \href{https://github.com/GU-CLASP/FraCoq}{https://github.com/GU-CLASP/FraCoq}

### 2020-01-30 paper2020/paper.tex
[temporal] paper stub

\documentclass[a4paper,11pt]{article}
\usepackage[hyperref]{naaclhlt2019}
\usepackage{natbib}

\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

%\newcommand\BibTeX{B{\sc ib}\TeX}

%\usepackage[backend=bibtex,citestyle=authoryear,style=authoryear,natbib]{biblatex}
% new style not compatible with biblatex
%\bibliography{../bbcl_lccm2018/bbcl_lccm2018.bib,iwcs2019}

\usepackage[verbose]{newunicodechar}

\title{Bayesian Inference Semantics: A Modelling System and  A Test Suite}
\author{Jean-Philippe Bernardy \qquad Stergios Chatzikyriakidis \\
  University of Gothenburg\\
  {\tt firstname.lastname@@gu.se} \\}

\date{}

\begin{document}
\maketitle

\section{Introduction}
Stergios
\section{Background}
\paragraph{GF}
If needed
\paragraph{Coq}
If needed
\paragraph{Temporal Semantics}
Stergios

\section{Our Compositional Semantics}
JP

\subsection{Out treatment of time}
JP
\subsection{...}
If needed
\section{Results and evaluation}
JP
\section{Conclusions and Future Work}
Stergios
\section*{Acknowledgements}

The research reported in this paper was supported by grant 2014-39 from the
Swedish Research Council, which funds the Centre for Linguistic Theory and
Studies in Probability (CLASP) in the Department of Philosophy, Linguistics,
and Theory of Science at the University of Gothenburg. We are grateful to
our colleagues in CLASP for helpful discussion of some of the ideas presented
here. We also thank three anonymous reviewers for their useful comments on an
earlier draft of the paper.

\bibliographystyle{acl_natbib}

\end{document}

### 2020-02-05 paper2020/paper.tex
stub analysis

@@
 \section{Our Compositional Semantics}
-JP
 
-\subsection{Out treatment of time}
-JP
-\subsection{...}
-If needed
+- See our pervious papers, (adjectives, adverbs, nouns, verbs, anaphora, etc.)
+
+- Here we simply ignore these aspects and it is safe for the reader to
+assume a montegovian semantics and assume that anaphoroi are resolved
+as they should be.
+
+\section{Our Treatment of time}
+
+Our basic approach is to extend predicates with two time parameters,
+one corresponding with the starting time of the action and one with
+its stopping time.
+
+walk(john,t0,t1)
+
+\paragraph{time context}
+
+The compositional semantics carry a current time reference (itself an
+interval). This current interval propagates through the compositional
+interpraetation. By default every interpretation passes the temporal
+context to its components --- but some may act on it.
+
+\paragraph{tenses and temporal adverbs}
+
+Tenses are represented syntactically (in GF!) as an attribute of the
+clause. In our representation we deal only with present and past
+tenses --- we find that FraCas does not exercise (if a variant of the
+past is used, the additional information is also carried by adverbs,
+in a more specific way.)
+
+If the temporal context was modified by an adverb (or an averbial
+clause, such as ``before James swam''), then the tense is ignored.
+%
+Otherwise, if the tense is the present, then we use a simple
+$(now,now)$ interval, with $now$ being a constant.
+%
+If the tense is the past, then we locally quantify over a time
+interval $(t_0,t_1)$, such that $t_1 < now$.
+
+\paragraph{time references and aspectual classes}
+
+A common theme in the testsuite is reference to previous occurences of an event
+
+P1	Smith left.
+P2	Jones left.
+P3	Smith left before Jones left.
+Q 	Did Jones leave after Smith left?
+
+We have to arrange that the two occurences of ``Jones left'', in P1
+and P3 refer to the same time intervals.
+
+For this purpose we postulate the unicity of action for certain timed propositions:
+
+P-unicity : P t1 t2 → P t3 t4 → (t1 = t3) ∧ (t2 = t4)
+
+P-unicity holds only if P is an \emph{achievement} or an \emph{actitivty}.
+
+(The difference between actitivty and achievement is that for the
+latter time intervals can be assumed to be of nil duration. However
+this appears to play little role, if any, in our analysis.)
+
+A contrario, if P is a \emph{stative}, then we get a time-interval subsumption property:
+
+P-subsumption : [t3,t4] ⊆ [t1,t2] -> P t1 t2 -> P t3 t4
+
+This principle is used for example below:
+
+P1	Smith arrived in Paris on the 5th of May, 1995.
+P2	Today is the 15th of May, 1995.
+P3	She is still in Paris.
+Q 	Was Smith in Paris on the 7th of May, 1995?
+
+\paragraph{(un)repeatable achievements}
+Using unicity of action, we can refute 279:
+
+279:
+P1	Smith wrote a novel in 1991.
+Q 	Did Smith write it in 1992?
+H 	Smith wrote it in 1992.
+
+∃x. novel(x)
+∀t<NOW.  ∧ write(smith,x,t,t) [the scope for the existential is extended according to our previous analysis of anaphora]
+
+However, the testsuite instructs that we should not be able to refute
+280, with the justification that ``wrote a novel'' is a repeatable
+accomplishment:
+
+280:
+P1	Smith wrote a novel in 1991.
+Q 	Did Smith write a novel in 1992?
+H 	Smith wrote a novel in 1992.
+
+∀t<NOW. ∃x. novel(x) ∧ write(smith,x,t,t)
+
+Our analysis does not need to separate these as special cases. Indeed,
+even if $write(smith,x)$ is an activity and thus subject to unicity of
+action, in (280), $x$ is quantified existentially; we have two
+different actions: $write(smith,x) t1 t2$ and $write(smith,y) t3 t4$,
+and thus we can't deduce equality of the interval $t1,t2$ and $t3,t4$;
+in turn the hypothesis can't be refuted.
+
+\paragraph{interval adverbs, point adverbs, universal adverbs}
+Time adverbs, such as "Since", "always", "never", "in two years", "by
+11 am", etc. introduce *local* quantification over a timespan.
+
+time interval:
+⟦``in 1992, s''⟧(_) = ∃t1,t2.  [t1,t2] ⊆ 1992, ⟦s⟧(t1,t2)
+
+Universal:
+⟦``always s''⟧(_) = ∀t1,t2.  [t1,t2] ⊆ 1992, ⟦s⟧(t1,t2)
+
+relative time:
+⟦``in two years, s''⟧(t0,tx) = ∃t2.  ⟦s⟧(t0+2years,t2)
+
+\paragraph{action-modification verbs}
+
+``start to'', ``finish'', etc.
+
+Lexical semantics transform the interval (see start_VVTiming in Coq code)
+
 \section{Results and evaluation}
 JP
+\paragraph{interaction with negation and group readings}
+
 \section{Conclusions and Future Work}

### 2020-02-05 paper2020/paper.tex
draft result table

@@
 \usepackage{natbib}
+\usepackage{tabularx}
 
@@
 
-\title{Bayesian Inference Semantics: A Modelling System and  A Test Suite}
+\title{FraCaS: Temporal Analysis}
 \author{Jean-Philippe Bernardy \qquad Stergios Chatzikyriakidis \\
@@
 time interval:
-⟦``in 1992, s''⟧(_) = ∃t1,t2.  [t1,t2] ⊆ 1992, ⟦s⟧(t1,t2)
+⟦``in 1992, s''⟧(...) = ∃t1,t2.  [t1,t2] ⊆ 1992, ⟦s⟧(t1,t2)
 
 Universal:
-⟦``always s''⟧(_) = ∀t1,t2.  [t1,t2] ⊆ 1992, ⟦s⟧(t1,t2)
+⟦``always s''⟧(...) = ∀t1,t2.  [t1,t2] ⊆ 1992, ⟦s⟧(t1,t2)
 
@@
 
-Lexical semantics transform the interval (see start_VVTiming in Coq code)
+example:
+
+Definition finishVV : VVT := fun tvp => fun x => fun t0 t1 =>
+exists t0', exists t1', t0 <= t1' <= t1 /\ tvp t0' t1' x.
 
 \section{Results and evaluation}
-JP
+\providecommand\ncases[1]{{\ensuremath{^{#1}}}}
+\begin{table}[hbt]
+  \centering
+  \small
+\begin{tabularx}{\columnwidth}{Xr@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r}
+Section      & {\kern -2em} \#FraCaS
+                          & This        & FC2         & FC & MINE & Nut  & LP  \\ \hline
+Quantifiers  & 75         & .91        & .96         & .96    & .77  & .53  & .93  \\
+             &            & \ncases{74}& \ncases{74} &        &      &      &     \ncases{44} \\
+Plurals      & 33         & .64        & .82         & .76    & .67  & .52  & .73 \\
+             &            &            &             &     &   &   & \ncases{24} \\
+Anaphora     & 28         & .79        & .86         &   -    & -    & -    &  -       \\
+Ellipsis     & 52         & .81        & .87         &   -    & -    & -    &  -       \\
+Adjectives   & 22         & .95        & .95         & .95    & .68  & .32  & .73 \\
+             &            &  \ncases{20}&  \ncases{20}&     &   &   &  \ncases{12} \\
+Comparatives & 31         & .64        & .87         & .56    & .48  & .45  &  -       \\
+Temporal     & 75         &  -         &  -          &   -    &   -  &  -   &  -       \\
+Verbs        & 8          & .71        & .75         &   -    & -    & -    &  -       \\
+Attitudes    & 13         & .85        & .92         & .85    & .77  & .46  & .92  \\ 
+             &            &            &             &        &      &      & \ncases {9}  \\ \hline
+Total        & 337        & .78        & .89         & .83    & .69  & .50  & .85  \\
+             &            & \ncases{329}& \ncases{259}& \ncases{174}  & \ncases{174}& \ncases{174}& \ncases{89}
+  \end{tabularx}
+  \caption{Accuracy of our system compared to others.
+    ``Ours" refers to the approach presented in this paper. When a
+    system does not handle the nominal number of test cases (shown in
+    the second column), the actual number of test cases attempted is
+    shown below the accuracy figure, in smaller font.  ``FraCoq''
+    refers to the work of \citet{bernardy:2017}. ``MINE" refers
+    to the approach of \citet{Mineshima:2015}, ``NUT" to the CCG
+    system that utilizes the first-order automated theorem prover
+    \textit{nutcracker} \cite{bos:2008}, and ``Langpro"
+    to the system presented by \citet{Abzianidze:2015}. A dash
+    indicates that no attempt was made for the section. }
+  \label{tab:results}
+\end{table}
 \paragraph{interaction with negation and group readings}
-
 \section{Conclusions and Future Work}

### 2020-02-05 ResultParser.hs

@@ -87,7 +87,9 @@ doIt args = do
   mapM_ print (parseWords Nothing ws)
   mapM_ print consolidated
   mapM_ (showScores consolidated) (zip3 [1..] sectionStarts (tail sectionStarts))
-  (showScores' consolidated) (0,\x -> x < 251 || x >= 326)
+  (showScores' consolidated) (0,const True {- excludeTemporal -})
+
+excludeTemporal = \x -> x < 251 || x >= 326
 
 main :: IO ()
 main = do

### 2020-02-06 paper2020/paper.tex
expand

@@
 one corresponding with the starting time of the action and one with
-its stopping time.
+its stopping time. For example, if John walked between $t_0$ and
+$t_1$, we would have: $walk(john,t0,t1)$. From now on we will call an
+interval of time points $[t_0,t_1]$ a timespan. Every timespan
+$[t_0,t_1]$ has the property $t_0 ≤ t_1$ (it starts no later than it
+stops).
 
-walk(john,t0,t1)
+\paragraph{temporal context}
 
-\paragraph{time context}
+We adjust the montegovian semantics so that every category takes a
+timespan as an additional parameter. This timespan serves as a time
+reference for semantics, and we call it a \emph{temporal context}. This current interval propagates through the compositional
+interpretation down to atoms. By default, every interpretation passes
+the temporal context down to its components without changing
+it. However some key elements will act on it on nontrivial ways, which
+we proceed to detail below.
 
-The compositional semantics carry a current time reference (itself an
-interval). This current interval propagates through the compositional
-interpraetation. By default every interpretation passes the temporal
-context to its components --- but some may act on it.
+\paragraph{tenses}
 
-\paragraph{tenses and temporal adverbs}
+The prime non-trivial manipulation of timespans are tenses. In our
+syntax, inherited from GF, tenses are represented syntactically as an
+attribute of the clause. TODO{show example}
 
-Tenses are represented syntactically (in GF!) as an attribute of the
-clause. In our representation we deal only with present and past
-tenses --- we find that FraCas does not exercise (if a variant of the
-past is used, the additional information is also carried by adverbs,
-in a more specific way.)
+In our representation we deal only with present and simple past
+tenses.  Indeed we find that FraCas does not exercise additional
+specific tenses (if a more complicated tense is used, the additional
+information is also carried by adverbs or adverbial phrases, in a more
+specific way.)
 
-If the temporal context was modified by an adverb (or an averbial
-clause, such as ``before James swam''), then the tense is ignored.
-%
-Otherwise, if the tense is the present, then we use a simple
-$(now,now)$ interval, with $now$ being a constant.
-%
+TODO: optional time context.
+
+The procedure to handle tense annotations is as follows:
+\begin{itemize}
+\item If the temporal context was modified by an adverb (or an
+  averbial clause, such as ``before James swam''), then the tense does
+  not create a new interval, but it may constrain it. Typically, a
+  past tense adds the constraint that the temporal context ends before
+  the constant $now$, which is a constant representing the current
+  time.
+\item Otherwise, if the tense is the present, then the temporal
+  context becomes the simple $(now,now)$ interval.
+\item
   If the tense is the past, then we locally quantify over a time
-interval $(t_0,t_1)$, such that $t_1 < now$.
+  interval $[t_0,t_1]$, such that $t_1 < now$.
+\end{itemize}
+
+\paragraph{temporal adverbs}
+
+The other single most important source of interesting timespans are
+adverbs. Most of the temporal adverbs fall in either of the following
+categories:
+
+\begin{enumerate}
+\item[exact interval adverbs] For such adverbs, an exact interval is
+  provided. Typical such adverbs specify in fact a single point in
+  time (so the start and the end of the interval coincide).
+
+  ⟦``at 5 pm, s''⟧(...) = ⟦s⟧(5pm,5pm)
+
+\item[existential interval] The majority of temporal adverbs
+  existentially quantify over a timespan. Examples include ``since
+  1991'', ``in 1996'', ``for two years'', etc. The common theme is to
+  introduce the interval and then restrict its bounds or its duration
+  in someway. Sometimes the restriction is an equality (``for exactly
+  two hours'').
+
+  ⟦``in 1992, s''⟧(...) = ∃t1,t2.  [t1,t2] ⊆ 1992, ⟦s⟧(t1,t2)
+
+\item[universal interval] A few adverbs introduce intervals via a
+  universal quantification (sometimes with a constraint). Examples
+  include: ``always'', ``never''.
+
+  If there is no explicit time context, then ``always'' has no
+  constraint on the interval, otherwise the quantified interval must
+  be included in it:
+  ⟦``always s''⟧(t0,t1) = ∀t0',t1'. [t0',t1'] ⊆ [t0,t1], ⟦s⟧(t0',t1')
+
+  % A particularly interesting case found in FraCas is ``every month'':
+
+  % ⟦``every month s''⟧(t0,t1) = ∀t. [t,t+1month] ⊆ [t0,t1], ⟦s⟧(t,t+1month)
+\end{enumerate}
 
@@
 
-A common theme in the testsuite is reference to previous occurences of an event
+A common theme in the testsuite is to refer to previous occurences of
+an event:
 
-P1	Smith left.
-P2	Jones left.
-P3	Smith left before Jones left.
-Q 	Did Jones leave after Smith left?
+(262)
+P1	Smith left after Jones left.
+P2	Jones left after Anderson left.
+Q 	Did Smith leave after Anderson left?
+H 	Smith left after Anderson left.
 
-We have to arrange that the two occurences of ``Jones left'', in P1
-and P3 refer to the same time intervals.
+To be able to conclude as the testsuite expects, we have to make sure
+that the two occurences of ``Jones left'', in P1 and P3 refer to the
+same time intervals.
 
@@
 
+Unicity of action appears to be a non-logical principle. Indeed, it is
+quite possible that ``Jones left'' several times. However, it seems
+that this principle is never contradicted by the testsuite. As such,
+even though unicity of action is only a pragmatic rule, it can be
+taken as a valid \emph{by default}: it is only when we have a
+sufficiently constrained situation that one should reject it. Consider
+the following discourse:
+
+(1) Smith left at 1pm.
+(2) Smith went to its appointment with the lawyer.
+(3) Smith came back at 2pm
+(4) Smith left at 4pm.
+
+One would normally not say that there is contradiction. However if (1)
+and (4) are presented in isolation, a contradiction can be flagged.
+
 \paragraph{(un)repeatable achievements}
-Using unicity of action, we can refute 279:
+The principle of using unicity of action interacts well with the usual interpretation of quantifiers (and anaphora).
+Indeed, using it, we can refute (279):
 
@@
 Q 	Did Smith write it in 1992?
-H 	Smith wrote it in 1992.
 
-∃x. novel(x)
-∀t<NOW.  ∧ write(smith,x,t,t) [the scope for the existential is extended according to our previous analysis of anaphora]
+Indeed, following our account, the above inference problem is be
+interperted as ∃x. novel(x) ∀t<NOW.  ∧ write(smith,x,t,t) [the scope
+for the existential is extended according to our previous analysis of
+anaphora]
 
@@
 
-\paragraph{interval adverbs, point adverbs, universal adverbs}
-Time adverbs, such as "Since", "always", "never", "in two years", "by
-11 am", etc. introduce *local* quantification over a timespan.
-
-time interval:
-⟦``in 1992, s''⟧(...) = ∃t1,t2.  [t1,t2] ⊆ 1992, ⟦s⟧(t1,t2)
-
-Universal:
-⟦``always s''⟧(...) = ∀t1,t2.  [t1,t2] ⊆ 1992, ⟦s⟧(t1,t2)
-
-relative time:
-⟦``in two years, s''⟧(t0,tx) = ∃t2.  ⟦s⟧(t0+2years,t2)
-
 \paragraph{action-modification verbs}
@@
   \small
-\begin{tabularx}{\columnwidth}{Xr@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r}
+\begin{tabularx}{\columnwidth}{Xr@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r}
 Section      & {\kern -2em} \#FraCaS
@@
 Comparatives & 31         & .64        & .87         & .56    & .48  & .45  &  -       \\
-Temporal     & 75         &  -         &  -          &   -    &   -  &  -   &  -       \\
-Verbs        & 8          & .71        & .75         &   -    & -    & -    &  -       \\
+Temporal     & 75         & .71        &  -          &   -    &   -  &  -   &  -       \\
+Verbs        & 8          & .75        & .75         &   -    & -    & -    &  -       \\
 Attitudes    & 13         & .85        & .92         & .85    & .77  & .46  & .92  \\ 

### 2020-02-07 paper2020/paper.tex
complete description of temporal treatment

@@
 We adjust the montegovian semantics so that every category takes a
-timespan as an additional parameter. This timespan serves as a time
-reference for semantics, and we call it a \emph{temporal context}. This current interval propagates through the compositional
-interpretation down to atoms. By default, every interpretation passes
-the temporal context down to its components without changing
-it. However some key elements will act on it on nontrivial ways, which
-we proceed to detail below.
+\emph{optional} timespan as an additional parameter. This timespan
+serves as a time reference for semantics, and we call it a
+\emph{temporal context}. This context propagates through the
+compositional interpretation down to atoms. By default, every
+interpretation passes the temporal context down to its components
+without changing it. However some key elements will act on it on
+nontrivial ways, which we proceed to detail below.
+
+The timespan in the context is optional because, in certain
+situations, the semantics is different depending on whether a timespan
+has been specified externally or not, as we explain below.
 
@@
 
-In our representation we deal only with present and simple past
+In our semantics we deal only with present and simple past
 tenses.  Indeed we find that FraCas does not exercise additional
@@
 specific way.)
-
-TODO: optional time context.
 
@@
   past tense adds the constraint that the temporal context ends before
-  the constant $now$, which is a constant representing the current
+  the timepoint $now$, which is a logical constant representing the current
   time.
@@
 \begin{enumerate}
-\item[exact interval adverbs] For such adverbs, an exact interval is
+\item[exact adverbs] For such adverbs, an exact interval is
   provided. Typical such adverbs specify in fact a single point in
@@
 
-\item[existential interval] The majority of temporal adverbs
+\item[existentially quantifying] The majority of temporal adverbs
   existentially quantify over a timespan. Examples include ``since
@@
 
-\item[universal interval] A few adverbs introduce intervals via a
+\item[universally quantifying] A few adverbs introduce intervals via a
   universal quantification (sometimes with a constraint). Examples
@@
 
-A common theme in the testsuite is to refer to previous occurences of
-an event:
+A common theme in the testsuite is the reference to previous
+occurences of an event:
 
@@
 To be able to conclude as the testsuite expects, we have to make sure
-that the two occurences of ``Jones left'', in P1 and P3 refer to the
+that the two occurences of ``Jones left'' (in P1 and P3) refer to the
 same time intervals.
 
-For this purpose we postulate the unicity of action for certain timed propositions:
+For this purpose we postulate the unicity of action for certain timed
+propositions:
 
@@
 
-Indeed, following our account, the above inference problem is be
-interperted as ∃x. novel(x) ∀t<NOW.  ∧ write(smith,x,t,t) [the scope
+Indeed, following our account, the above (contradictory) inference problem is to be
+interperted as
+∃x. novel(x) ∧
+∃t1,t2. [t1,t2] ⊆ 1991 ∧ write(smith,x,t1,t2) ∧
+∃t3,t4. [t3,t4] ⊆ 1992 ∧ write(smith,x,t3,t4)
+⟶ ⊥
+
+Thanks to unicity of action we find [t1,t2] = [t3,t4], and due to 1991
+and 1992 being disjoint we obtain contradiction.
+
+[the scope
 for the existential is extended according to our previous analysis of
@@
 
-However, the testsuite instructs that we should not be able to refute
+However, the testsuite instructs that we should \emph{not} be able to refute
 280, with the justification that ``wrote a novel'' is a repeatable
@@
 
-∀t<NOW. ∃x. novel(x) ∧ write(smith,x,t,t)
+Here our interpretation is:
 
-Our analysis does not need to separate these as special cases. Indeed,
-even if $write(smith,x)$ is an activity and thus subject to unicity of
-action, in (280), $x$ is quantified existentially; we have two
-different actions: $write(smith,x) t1 t2$ and $write(smith,y) t3 t4$,
-and thus we can't deduce equality of the interval $t1,t2$ and $t3,t4$;
-in turn the hypothesis can't be refuted.
+∃x. novel(x) ∧
+∃t1,t2. [t1,t2] ⊆ 1991 ∧ write(smith,x,t1,t2) ∧
+∃y. novel(y) ∧
+∃t3,t4. [t3,t4] ⊆ 1992 ∧ write(smith,x,t3,t4)
+⟶ ⊥
+
+Our analysis does not need to treat this last case specially. Indeed,
+even if $write(smith,x,.,.)$ is an activity and thus subject to
+unicity of action, in (280), $x$ is quantified existentially; we have
+two \emph{different} actions: $write(smith,x,t1,t2)$ and
+$write(smith,y,t3,t4)$, and thus we can't deduce equality of the
+interval $t1,t2$ and $t3,t4$; in turn the hypothesis can't be refuted.
 
@@
 
-``start to'', ``finish'', etc.
+The final class of lexeme carrying a temporal-dependent semantics are
+verbs taking a proposition as argument, like ``finish'', ``start'',
+etc. These verbs modify the temporal in non-trivial ways. Consider for
+example ``finish to ...''. The timespan of the argument of finish
+should end within the timespan of the finishing action:
 
-example:
-
-Definition finishVV : VVT := fun tvp => fun x => fun t0 t1 =>
-exists t0', exists t1', t0 <= t1' <= t1 /\ tvp t0' t1' x.
+⟦finish to s⟧(t0,t1) = ∃(t0',t1'). t1' ∈ [t0,t1] ∧ s(t0',t1')
 

### 2020-02-07 Dynamic.hs
simplification

@@ -202,7 +202,6 @@ data Env = Env {vpEnv :: VPEnv
                ,envDefinites :: [(Exp,Object)] -- map from CN to pure objects
                ,envMissing :: [(Exp,Var)] -- definites that we could not find. A map from CN to missing variables
                ,envPluralizingQuantifier :: Bool
-               ,envFacts :: [Prop]
                ,freshVars :: [String]
                }
          -- deriving Show
@@ -409,7 +408,6 @@ assumed = Env {vp2Env = return $ \x y -> (mkRel2 "assumedV2" x y)
                ,envDefinites = []
                ,envMissing = []
                ,envPluralizingQuantifier = False
-               ,envFacts = []
                ,freshVars = allVars}
 
 assumedReadEnv :: ReadEnv
@@ -533,26 +531,6 @@ forceTime tMeta cl = noExtraObjs (useTime tMeta cl)
 useTime :: TimeSpan -> Cl' -> S'
 useTime t s ExtraArgs{..} = s ExtraArgs{extraTime = ForceTime t,..}
 
--- | Look in envFacts for the time at which s' happened.
--- That is: Find the times t when Prop(t) happened, looking up true facts in environment.
-referenceTimesFor :: Cl' -> Dynamic [TimeSpan]
-referenceTimesFor s' = do
-  tMeta0 <- getFresh
-  tMeta1 <- getFresh
-  facts <- gets envFacts
-  let e = forceTime (Var tMeta0, Var tMeta1) s'
-  return $ map mkSpan $ catMaybes $ map (solveThe [tMeta0,tMeta1] e) facts
-  where mkSpan [t0,t1] = (t0,t1)
-
-referenceTimeFor :: Cl' -> Dynamic TimeSpan
-referenceTimeFor s = do
-  ts <- referenceTimesFor s
-  case ts of
-    [] -> do
-      facts <- gets envFacts
-      error ("referenceTimesFor: no time for " ++ show (forceTime (Var "τ₀", Var "τ₁") s) ++ "\n facts:\n" ++ intercalate "\n" (map show facts))
-    (t:_) -> return t
-
 -- | S' shall use the given time constraint
 usingTime :: Temporal -> S' -> S'
 usingTime e s' ExtraArgs{..} = s' ExtraArgs{extraTime = e, ..} 
@@ -570,10 +548,6 @@ tempToArgs = pairToList . temporalToLogic
[…]

### 2020-02-07 MS.hs

@@ -479,8 +479,8 @@ withAtLeastDuration delta = do -- with duration meaning, as in FraCas 278
   return (\s extraObjs ->
            let (p,actualSpan) = s extraObjs {extraTime = tSpan}
            in (quantTime Exists t1 true $
-               quantTime Exists t2 ((Con "IS_INTERVAL" `apps` [intervalStart (actualSpan),intervalStart (extraTime extraObjs)])
-                                ∧
+               quantTime Exists t2 ((Con "IS_INTERVAL" `apps` [intervalStart (actualSpan),intervalStart (extraTime extraObjs)]) ∧
+                                    -- (Con "IS_INTERVAL" `apps` [intervalEnd (actualSpan),Con "NOW"]) ∧
                                 -- it may finish later, see Fracas 304
                                 -- we constraint the returned time, so that we have a stronger interpretation of achievements (see Fracas 306)
                                  (Con "IS_INTERVAL" `apps` [Con "Plus" `apps` [Var t1, delta], Var t2])) $
@@ -498,6 +498,9 @@ withExactDuration delta = do -- as in FraCas 278
 
 intervalStart UnspecifiedTime = Con "NOW"
 intervalStart (ExactTime (a,b)) = a
+
+intervalEnd UnspecifiedTime = Con "NOW"
+intervalEnd (ExactTime (a,b)) = a
 
 inInterval :: Temporal -> Temporal -> Exp
 UnspecifiedTime `inInterval` _ = true
@@ -1071,15 +1074,7 @@ predVP np vp = withClause $ do
   tense <- asks envTense
   p'' <- case tense of
     t | t `elem` [Past, PresentPerfect, PastPerfect] -> do
-      ts <- referenceTimesFor p' -- (1)
-      -- fixme: this should happen at a lower level (lexical
-      -- item). But we do not have dynamic access to arguments at that
-      -- level at the moment so this will do temporarily.
-      -- Why? Because events could refer to occurences inside a quantifier:
-      -- Example "every boy climbed and fell after they climbed." (ATOM)
-      case ts of
-         [] -> -- not a reference to a previous event.
-           do -- Allocate own time.
+              -- Allocate own time.
               t1 <- getFresh
               t2 <- getFresh
               return $ \e@ExtraArgs{..} ->
@@ -1092,9 +1087,7 @@ predVP np vp = withClause $ do
                       -- be able to override it, to search for it when we reach
[…]

### 2020-03-10 paper2020/paper.tex
revision

@@
 
-- See our pervious papers, (adjectives, adverbs, nouns, verbs, anaphora, etc.)
+We are relying on a largely standard montegovian semantics. In brief,
+sentences are interpreted as propositions, verbs and noun-phrases as
+predicates.  We use type-raising for noun-phrases to support
+quantifiers.
 
-- Here we simply ignore these aspects and it is safe for the reader to
-assume a montegovian semantics and assume that anaphoroi are resolved
-as they should be.
+We support additionally various constructinos and phenomena present in
+the testsuite, including adjectives, adverbs, nouns, verbs, anaphora,
+etc. We direct the reader to our previous work for details \ifanon
+(citations anonymized) \else \citet{} \fi, but the particular
+treatment of such phenomena is largely independent from our treatment
+of time.
+%
+Here we simply ignore these aspects beyond the fact that they
+are handled correctly in most of the case of the FraCaS testsuite.  In
+particular, non-temporal anaphoroi are resolved as they should be.
 
@@
 
-Our basic approach is to extend predicates with two time parameters,
-one corresponding with the starting time of the action and one with
-its stopping time. For example, if John walked between $t_0$ and
-$t_1$, we would have: $walk(john,t0,t1)$. From now on we will call an
-interval of time points $[t_0,t_1]$ a timespan. Every timespan
-$[t_0,t_1]$ has the property $t_0 ≤ t_1$ (it starts no later than it
-stops).
+In montegovian semantics, (intransitive) verbs are one-place
+predicates; in types, they are functions from entities to propositions
+($e \to t$).  Our basic approach is to generalise verb to take two
+additional time parameters, one corresponding with the starting time
+of the action and one corresponding to its stopping time. For example,
+if John walked between $t_0$ and $t_1$, we would have:
+$walk(john,t0,t1)$. From now on we will call an interval of time
+points $[t_0,t_1]$ a timespan. Every timespan $[t_0,t_1]$ has the
+property $t_0 ≤ t_1$ (it starts no later than it stops). We are thus
+using a simple newtonian model of time, corresponding to layman
+intuition.
+
+In principle, common nouns and adjectives should undergo the same
+procedure. For simplicity we will however only consider verbs from now
+on. (In fact in our implementation we chose not to modify nouns nor
+adjectives to limit the increase in complexity of the formulas.)
 
@@
 
-We adjust the montegovian semantics so that every category takes a
+We adjust the montegovian semantics so that the interpretation of
+every category (propositions, verb phrases, etc.) takes an
 \emph{optional} timespan as an additional parameter. This timespan
@@
 \emph{temporal context}. This context propagates through the
-compositional interpretation down to atoms. By default, every
-interpretation passes the temporal context down to its components
-without changing it. However some key elements will act on it on
-nontrivial ways, which we proceed to detail below.
+compositional interpretation down to lexical items with atomic
+representation (verbs). By default, every interpretation passes the
+temporal context down to its components without changing it. However
+some key elements will act on it on nontrivial ways, which we proceed
+to detail below.
 
@@
 
-The prime non-trivial manipulation of timespans are tenses. In our
-syntax, inherited from GF, tenses are represented syntactically as an
-attribute of the clause. TODO{show example}
+The principal non-trivial manipulation of timespans are tense
+markers. In our syntax, inherited from GF, tenses are represented
+syntactically as an attribute of the clause.
+
+Example:
+
+A scandinavian won the nobel prize.
+
+useCl past pPos (predVP (detCN (detQuant indefArt numSg) (useN (lexemeN "scandinavian_N"))) (complSlash (slashV2a (lexemeV2 "win_V2")) (detCN (detQuant indefArt numSg) (useN (lexemeN "nobel_prize_N")))))
 

### 2020-03-10 paper2020/paper.tex
texification

@@
+\newif\ifanon\anontrue
+
 \documentclass[a4paper,11pt]{article}
+\usepackage[utf8]{inputenc}
+
 \usepackage[hyperref]{naaclhlt2019}
@@
 \usepackage{tabularx}
+\usepackage{amsmath}
+\usepackage{stmaryrd}
+\usepackage{todonotes}
+\usepackage{enumitem}
+\newlist{lingex}{enumerate}{3}
+\setlist[lingex,1]{parsep=0pt,itemsep=1pt,label=(\arabic*),resume=lingexcount}
+\newcommand\onelingex[1]{\begin{lingex}\item #1 \end{lingex}}
+
+\newlist{subex}{enumerate}{3}
+\setlist[subex,1]{parsep=0pt,itemsep=1pt,label=\textbf{P\arabic*}}
+\newcommand\hyp{\item[\textbf{H}]}
+\newcommand\fracasex[2]{\begin{lingex}\item[(#1)] \begin{subex} #2 \end{subex} \end{lingex} }
+\newcommand\jp[1]{\todo[backgroundcolor=blue!10]{JP: #1}}
+
+\newcommand\constant[1]{\mathsf{#1}}
 
@@
 \usepackage[verbose]{newunicodechar}
+\input{../PaperTools/latex/newunicodedefs}
 
@@
 if John walked between $t_0$ and $t_1$, we would have:
-$walk(john,t0,t1)$. From now on we will call an interval of time
+$walk(john,t_0,t_1)$. From now on we will call an interval of time
 points $[t_0,t_1]$ a timespan. Every timespan $[t_0,t_1]$ has the
@@
 
-The timespan in the context is optional because, in certain
-situations, the semantics is different depending on whether a timespan
-has been specified externally or not, as we explain below.
+\newcommand\nospan{\ensuremath{─}}
+\newcommand\anyspan{\ensuremath{\_}}
+The timespan in the context is
+optional because, in certain situations, the semantics is different
+depending on whether a timespan has been specified externally or not,
+as we explain below. A non-present timespan will be represented as
+\nospan{}. If a semantic function does not depend on a timespan, we will write \anyspan{} instead
 
@@
 
-useCl past pPos (predVP (detCN (detQuant indefArt numSg) (useN (lexemeN "scandinavian_N"))) (complSlash (slashV2a (lexemeV2 "win_V2")) (detCN (detQuant indefArt numSg) (useN (lexemeN "nobel_prize_N")))))
+useCl past pPos (predVP (detCN (detQuant indefArt numSg) scandinavian\_CN) (complSlash (slashV2a win\_V2) (detCN (detQuant indefArt numSg) nobel\_prize\_CN)))
 
 In our semantics we deal only with present and simple past
-tenses.  Indeed we find that FraCas does not exercise additional
+tenses. \jp{Is this enough for theory?} Indeed we find that FraCas does not exercise additional
 specific tenses (if a more complicated tense is used, the additional
@@
 
-  ⟦``at 5 pm, s''⟧(...) = ⟦s⟧(5pm,5pm)
+  \[⟦\text{at 5 pm, \(s\)}⟧(...) = ⟦s⟧(5pm,5pm)\]
 
@@
 
-  ⟦``in 1992, s''⟧(...) = ∃t1,t2.  [t1,t2] ⊆ 1992, ⟦s⟧(t1,t2)
+  \[⟦\text{in 1992, \(s\)}⟧(...) = ∃t_1,t_2.  [t_1,t_2] ⊆ 1992, ⟦s⟧(t_1,t_2)\]
 
@@
   be included in it:
-  ⟦``always s''⟧(t0,t1) = ∀t0',t1'. [t0',t1'] ⊆ [t0,t1], ⟦s⟧(t0',t1')
+  \[⟦\text{always \(s\)}⟧(t_0,t_1) = ∀t_0',t_1'. [t_0',t_1'] ⊆ [t_0,t_1], ⟦s⟧(t_0',t_1')\]
 
@@
 
-  % ⟦``every month s''⟧(t0,t1) = ∀t. [t,t+1month] ⊆ [t0,t1], ⟦s⟧(t,t+1month)
+  % ⟦``every month s''⟧(t_0,t_1) = ∀t. [t,t+1month] ⊆ [t_0,t_1], ⟦s⟧(t,t+1month)
 \end{enumerate}
@@
 
-(262)
-P1	Smith left after Jones left.
-P2	Jones left after Anderson left.
-Q 	Did Smith leave after Anderson left?
-H 	Smith left after Anderson left.
+\fracasex{262}{
+\item	Smith left after Jones left.
+\item	Jones left after Anderson left.
+\hyp 	Did Smith leave after Anderson left?
+}
 
@@
 
-P-unicity : P t1 t2 → P t3 t4 → (t1 = t3) ∧ (t2 = t4)
+\[\constant{unicity}_P : P t_1 t_2 → P t_3 t_4 → (t_1 = t_3) ∧ (t_2 = t_4)\]
 
@@
 
-P-subsumption : [t3,t4] ⊆ [t1,t2] -> P t1 t2 -> P t3 t4
+\[\constant {subsumption}_P : [t_3,t_4] ⊆ [t_1,t_2] → P t_1 t_2 → P t_3 t_4\]
 
@@
 
-P1	Smith arrived in Paris on the 5th of May, 1995.
-P2	Today is the 15th of May, 1995.
-P3	She is still in Paris.
-Q 	Was Smith in Paris on the 7th of May, 1995?
+\fracasex{314}
+{\item	Smith arrived in Paris on the 5th of May, 1995.
+\item	Today is the 15th of May, 1995.
+\item	She is still in Paris.
+\hyp 	Was Smith in Paris on the 7th of May, 1995? 
+}
 
@@
 
-(1) Smith left at 1pm.
-(2) Smith went to its appointment with the lawyer.
-(3) Smith came back at 2pm
-(4) Smith left at 4pm.
-
+\begin{enumerate}[parsep=0pt,itemsep=1pt,label=(\arabic*)]
+\item Smith left at 1pm.
+\item Smith went to its appointment with the lawyer.
+\item Smith came back at 2pm
+\item Smith left at 4pm.
+\end{enumerate}
 One would normally not say that there is contradiction. However if (1)
@@
 
-279:
-P1	Smith wrote a novel in 1991.
-Q 	Did Smith write it in 1992?
-
+\fracasex{279}{
+\item	Smith wrote a novel in 1991.
+\hyp 	Did Smith write it in 1992?
+}
 Indeed, following our account, the above (contradictory) inference problem is to be
 interperted as
-∃x. novel(x) ∧
-∃t1,t2. [t1,t2] ⊆ 1991 ∧ write(smith,x,t1,t2) ∧
-∃t3,t4. [t3,t4] ⊆ 1992 ∧ write(smith,x,t3,t4)
-⟶ ⊥
 
-Thanks to unicity of action we find [t1,t2] = [t3,t4], and due to 1991
-and 1992 being disjoint we obtain contradiction.
+\[\begin{array}{l}
+∃x. novel(x) ∧ \\
+∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2) ∧ \\
+∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,x,t_3,t_4) \\
+⟶ ⊥ \\
+\end{array}
+\]
+
+Thanks to unicity of action we find \([t_1,t_2] = [t_3,t_4]\), and due
+to the years 1991 and 1992 being disjoint we obtain contradiction.
 
@@
 However, the testsuite instructs that we should \emph{not} be able to refute
-280, with the justification that ``wrote a novel'' is a repeatable
+(280), with the justification that ``wrote a novel'' is a repeatable
 accomplishment:
 
-280:
-P1	Smith wrote a novel in 1991.
-Q 	Did Smith write a novel in 1992?
-H 	Smith wrote a novel in 1992.
-
+\fracasex{280}{
+\item	Smith wrote a novel in 1991.
+\item 	Did Smith write a novel in 1992?
+\hyp 	Smith wrote a novel in 1992.
+}
 Here our interpretation is:
 
-∃x. novel(x) ∧
-∃t1,t2. [t1,t2] ⊆ 1991 ∧ write(smith,x,t1,t2) ∧
-∃y. novel(y) ∧
-∃t3,t4. [t3,t4] ⊆ 1992 ∧ write(smith,x,t3,t4)
-⟶ ⊥
-
+\[\begin{array}{l}
+∃x. novel(x) ∧ \\
+∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2) ∧ \\
+∃y. novel(y) ∧ \\
+∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,x,t_3,t_4) \\
+⟶ ⊥ \\
+\end{array}
+\]
 Our analysis does not need to treat this last case specially. Indeed,
@@
 unicity of action, in (280), $x$ is quantified existentially; we have
-two \emph{different} actions: $write(smith,x,t1,t2)$ and
-$write(smith,y,t3,t4)$, and thus we can't deduce equality of the
-interval $t1,t2$ and $t3,t4$; in turn the hypothesis can't be refuted.
+two \emph{different} actions: $write(smith,x,t_1,t_2)$ and
+$write(smith,y,t_3,t_4)$, and thus we can't deduce equality of the
+interval $t_1,t_2$ and $t_3,t_4$; in turn the hypothesis can't be refuted.
 
@@
 
-⟦finish to s⟧(t0,t1) = ∃(t0',t1'). t1' ∈ [t0,t1] ∧ s(t0',t1')
+\[⟦\text{finish to \(s\)}⟧(t_0,t_1) = ∃(t_0',t_1'). t_1' ∈ [t_0,t_1] ∧ ⟦s⟧(t_0',t_1')\]
 

### 2020-03-11 paper2020/paper.tex
formatting tweaks
tweak
revise

@@
 \newlist{subex}{enumerate}{3}
-\setlist[subex,1]{parsep=0pt,itemsep=1pt,label=\textbf{P\arabic*}}
-\newcommand\hyp{\item[\textbf{H}]}
+\setlist[subex,1]{parsep=0pt,itemsep=1pt,font=\bfseries,label=P\arabic*}
+\newcommand\hyp{\item[H]}
 \newcommand\fracasex[2]{\begin{lingex}\item[(#1)] \begin{subex} #2 \end{subex} \end{lingex} }
@@
 
-\begin{enumerate}
+\begin{enumerate}[align=left,font=\itshape]
 \item[exact adverbs] For such adverbs, an exact interval is
@@
 
-  \[⟦\text{at 5 pm, \(s\)}⟧(...) = ⟦s⟧(5pm,5pm)\]
+  \[⟦\text{at 5 pm, \(s\)}⟧(\anyspan) = ⟦s⟧(5pm,5pm)\]
 
@@
 
-  \[⟦\text{in 1992, \(s\)}⟧(...) = ∃t_1,t_2.  [t_1,t_2] ⊆ 1992, ⟦s⟧(t_1,t_2)\]
-
+  \begin{multline*}
+  ⟦\text{in 1992, \(s\)}⟧(\anyspan) = \\ ∃t_1,t_2.  [t_1,t_2] ⊆ 1992, ⟦s⟧(t_1,t_2)
+\end{multline*}
+In the FraCaS test suite, we normally do not find several
+time-modifying adverbs modifying a single verb. Indeed, sentences such
+as ``in 1992, in 1991 john wrote a novel'' are infelicitous. This
+justifies ignoring the input timespan in the above interpretation --
+we're in particular not interested in modeling felicity with our
+semantics, only giving an accurate semantics when the input is
+felicitous.
 \item[universally quantifying] A few adverbs introduce intervals via a
   universal quantification (sometimes with a constraint). Examples
-  include: ``always'', ``never''.
+  include ``always'' and ``never''.
 
@@
   be included in it:
-  \[⟦\text{always \(s\)}⟧(t_0,t_1) = ∀t_0',t_1'. [t_0',t_1'] ⊆ [t_0,t_1], ⟦s⟧(t_0',t_1')\]
+  \begin{multline*}
+  ⟦\text{always \(s\)}⟧(t_0,t_1) = \\ ∀t_0',t_1'. [t_0',t_1'] ⊆ [t_0,t_1], ⟦s⟧(t_0',t_1')
+\end{multline*}
+Note that, here, we do use the input interval, allowing to interpret
+phrases such as ``In 1994, Itel was always on time.''.
 
-  % A particularly interesting case found in FraCas is ``every month'':
-
-  % ⟦``every month s''⟧(t_0,t_1) = ∀t. [t,t+1month] ⊆ [t_0,t_1], ⟦s⟧(t,t+1month)
+% A specifically interesting case found in FraCas is the adverbial phrase ``every month'', which we modeled
+% \begin{equation}
+%   ⟦\text{every month s}⟧(t_0,t_1) = ∀t. [t,t+1month] ⊆ [t_0,t_1], ⟦s⟧(t,t+1month)
+% \end{equation}
 \end{enumerate}
@@
 
-To be able to conclude as the testsuite expects, we have to make sure
-that the two occurences of ``Jones left'' (in P1 and P3) refer to the
+To be able to conclude that there is entailment, as the testsuite expects, we have to make sure
+that the two occurences of ``Jones left'' (in \textbf{P1} and \textbf{P2}) refer to the
 same time intervals.
 
-For this purpose we postulate the unicity of action for certain timed
+For this purpose we postulate \emph{unicity of action} for certain time-dependent
 propositions:
 
-\[\constant{unicity}_P : P t_1 t_2 → P t_3 t_4 → (t_1 = t_3) ∧ (t_2 = t_4)\]
+\[\constant{unicity}_P : P (t_1,t_2) → P (t_3,t_4) → (t_1 = t_3) ∧ (t_2 = t_4)\]
 
-P-unicity holds only if P is an \emph{achievement} or an \emph{actitivty}.
+Unicity of action holds only if the aspectual class of the proposition
+$P$ is \emph{achievement} or \emph{actitivty}.\todo{reference in literature}
 
@@
 
-A contrario, if P is a \emph{stative}, then we get a time-interval subsumption property:
-
-\[\constant {subsumption}_P : [t_3,t_4] ⊆ [t_1,t_2] → P t_1 t_2 → P t_3 t_4\]
-
-This principle is used for example below:
-
-\fracasex{314}
-{\item	Smith arrived in Paris on the 5th of May, 1995.
-\item	Today is the 15th of May, 1995.
-\item	She is still in Paris.
-\hyp 	Was Smith in Paris on the 7th of May, 1995? 
-}
-
 Unicity of action appears to be a non-logical principle. Indeed, it is
@@
 even though unicity of action is only a pragmatic rule, it can be
-taken as a valid \emph{by default}: it is only when we have a
+taken as a valid one \emph{by default}: it is only when we have a
 sufficiently constrained situation that one should reject it. Consider
@@
 
+\paragraph{Statives}
+\textit{A contrario}, if \(P\) is \emph{stative}, then we get a time-interval subsumption property:
+
+\begin{multline*}
+\constant {subsumption}_P :\\ [t_3,t_4] ⊆ [t_1,t_2] → P (t_1,t_2) → P (t_3,t_4)
+\end{multline*}
+This principle is used to reason about problem (314), below:
+
+\fracasex{314}
+{\item	Smith arrived in Paris on the 5th of May, 1995.
+\item	Today is the 15th of May, 1995.
+\item	She is still in Paris.
+\hyp 	Was Smith in Paris on the 7th of May, 1995? 
+}
+
+Indeed, from \textbf{P3} we get that Smith was in paris between May
+5th and May 15th. Because being in Paris is stative, we also get that
+Smith was in Paris in any subinterval.
+
 \paragraph{(un)repeatable achievements}
 The principle of using unicity of action interacts well with the usual interpretation of quantifiers (and anaphora).
-Indeed, using it, we can refute (279):
+Indeed, using it, we can refute problem (279), as expected in the testsuite:
 
@@
 \]
+Note here that the scope for the existential is extended beyond the
+scope of \textbf{P1}. This extension can work according to
+\citet{unger:2011}, and our implemented analysis of anaphora\ifanon (citation anonymized) \else\citep{bernardy_computational_2018,bernardy_wide_2019}\fi.
 
-Thanks to unicity of action we find \([t_1,t_2] = [t_3,t_4]\), and due
+Thanks to the unicity of action of $write(smith,x,...)$ we find \([t_1,t_2] = [t_3,t_4]\), and due
 to the years 1991 and 1992 being disjoint we obtain contradiction.
 
-[the scope
-for the existential is extended according to our previous analysis of
-anaphora]
-
 However, the testsuite instructs that we should \emph{not} be able to refute
-(280), with the justification that ``wrote a novel'' is a repeatable
+problem (280), with the justification that ``wrote a novel'' is a repeatable
 accomplishment:
@@
 ∃y. novel(y) ∧ \\
-∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,x,t_3,t_4) \\
+∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,y,t_3,t_4) \\
 ⟶ ⊥ \\
@@
 two \emph{different} actions: $write(smith,x,t_1,t_2)$ and
-$write(smith,y,t_3,t_4)$, and thus we can't deduce equality of the
-interval $t_1,t_2$ and $t_3,t_4$; in turn the hypothesis can't be refuted.
+$write(smith,y,t_3,t_4)$, and thus we cannot deduce equality of the
+intervals $t_1,t_2$ and $t_3,t_4$. In turn, the hypothesis cannot be refuted.
 
@@
 
-The final class of lexeme carrying a temporal-dependent semantics are
+The final class of lexemes carrying a temporal-dependent semantics are
 verbs taking a proposition as argument, like ``finish'', ``start'',
-etc. These verbs modify the temporal in non-trivial ways. Consider for
-example ``finish to ...''. The timespan of the argument of finish
-should end within the timespan of the finishing action:
+etc. These verbs modify the temporal context in non-trivial
+ways. Consider for example ``finish to ...''. The timespan of the
+argument of ``finish'' should end within the timespan of the finishing
+action:
 
-\[⟦\text{finish to \(s\)}⟧(t_0,t_1) = ∃(t_0',t_1'). t_1' ∈ [t_0,t_1] ∧ ⟦s⟧(t_0',t_1')\]
+\begin{multline*}
+⟦\text{finish to \(s\)}⟧(t_0,t_1) = \\ ∃(t_0',t_1'). t_1' ∈ [t_0,t_1] ∧ ⟦s⟧(t_0',t_1')
+\end{multline*}
 
@@
 our colleagues in CLASP for helpful discussion of some of the ideas presented
-here. We also thank three anonymous reviewers for their useful comments on an
+here. We also thank anonymous reviewers for their useful comments on an
 earlier draft of the paper.

### 2020-04-02 paper2020/paper.tex
revision

@@
 \usepackage{tabularx}
+\usepackage{times}
 \usepackage{amsmath}
@@
 
-We are relying on a largely standard montegovian semantics. In brief,
-sentences are interpreted as propositions, verbs and noun-phrases as
-predicates.  We use type-raising for noun-phrases to support
-quantifiers.
+Our interpretation is cast withing the framework of a largely standard
+montegovian semantics \cite{todo}. In brief, sentences are interpreted as
+propositions, verbs and noun-phrases as predicates.  We use
+type-raising of noun-phrases, to support quantifiers \cite{PTQ}.
 
@@
 etc. We direct the reader to our previous work for details \ifanon
-(citations anonymized) \else \citet{} \fi, but the particular
+(citations anonymized) \else \citet{bernardy_type_2017,bernardy_wide_2019} \fi, but the particular
 treatment of such phenomena is largely independent from our treatment
@@
 predicates; in types, they are functions from entities to propositions
-($e \to t$).  Our basic approach is to generalise verb to take two
+($e \to t$).  Our basic approach is to generalise the interpretation of 
+verbs, so that it takes two
 additional time parameters, one corresponding with the starting time
-of the action and one corresponding to its stopping time. For example,
+of the action and one corresponding to its stopping time ($(time × time × e) \to t$). For example,
 if John walked between $t_0$ and $t_1$, we would have:
@@
 points $[t_0,t_1]$ a timespan. Every timespan $[t_0,t_1]$ has the
-property $t_0 ≤ t_1$ (it starts no later than it stops). We are thus
-using a simple newtonian model of time, corresponding to layman
-intuition.
+property $t_0 ≤ t_1$ (it starts no later than it stops). (We are thus
+using a simple newtonian model of time, corresponding to a layman
+intuition.)
 
@@
 We adjust the montegovian semantics so that the interpretation of
-every category (propositions, verb phrases, etc.) takes an
-\emph{optional} timespan as an additional parameter. This timespan
-serves as a time reference for semantics, and we call it a
-\emph{temporal context}. This context propagates through the
-compositional interpretation down to lexical items with atomic
-representation (verbs). By default, every interpretation passes the
-temporal context down to its components without changing it. However
-some key elements will act on it on nontrivial ways, which we proceed
-to detail below.
+every category (propositions, verb phrases, etc.) takes a
+\emph{temporal context} as an additional parameter, serves as a time
+reference for the interpretation of all time-dependent semantics
+within the phrase.  This context propagates through the compositional
+interpretation down to lexical items with atomic representation
+(verbs). By default, every interpretation passes the temporal context
+down to its components without changing it. However some key elements
+will act on it on nontrivial ways, which we proceed to detail below.
 
+This temporal context is an \emph{optional} timespan. That is, it can
+be an timespan or an explicitly unspecified context.
 \newcommand\nospan{\ensuremath{─}}
-\newcommand\anyspan{\ensuremath{\_}}
-The timespan in the context is
+\newcommand\anyspan{\ensuremath{\_}} The timespan in the context is
 optional because, in certain situations, the semantics is different
@@
 as we explain below. A non-present timespan will be represented as
-\nospan{}. If a semantic function does not depend on a timespan, we will write \anyspan{} instead
+\nospan{}. If a semantic function does not depend on the temporal
+context at all, we will write \anyspan{} instead.
 
@@
 
-The principal non-trivial manipulation of timespans are tense
+\newcommand\varid[1]{\mathnormal{#1}}
+%
+The principal non-trivial manipulator of timespans are tense
 markers. In our syntax, inherited from GF, tenses are represented
-syntactically as an attribute of the clause.
+syntactically as an attribute of the clause. An illustration of a
+past-tense clause and its interpretation follows. Notice in particular
+the $past$ argument to $useCl$.
 
-Example:
-
-A scandinavian won the nobel prize.
-
-useCl past pPos (predVP (detCN (detQuant indefArt numSg) scandinavian\_CN) (complSlash (slashV2a win\_V2) (detCN (detQuant indefArt numSg) nobel\_prize\_CN)))
-
+\onelingex{A scandinavian won the nobel prize.\label{ex:onelingex}
+\\
+{\small\begin{multline*}
+\varid{useCl}\, \varid{past}\, \varid{pPos}\\ (\varid{predVP}\, (\varid{detCN}\, (\varid{detQuant}\, \varid{indefArt}\, \varid{numSg})\\ \varid{scandinavian\_CN})\\ (\varid{complSlash}\, (\varid{slashV2a}\, \varid{win\_V2})\\ (\varid{detCN}\, (\varid{detQuant}\, \varid{indefArt}\, \varid{numSg})\\ \varid{nobel\_prize\_CN})))
+\end{multline*}
+}}
 In our semantics we deal only with present and simple past
-tenses. \jp{Is this enough for theory?} Indeed we find that FraCas does not exercise additional
-specific tenses (if a more complicated tense is used, the additional
-information is also carried by adverbs or adverbial phrases, in a more
-specific way.)
+tenses. \jp{Is this enough for theory?} Indeed we find that FraCas
+does not exercise additional specific tenses (if a more complicated
+tense is used, the additional information is also carried by adverbs
+or adverbial phrases, in a more specific way.)
 
@@
 \begin{itemize}
-\item If the temporal context was modified by an adverb (or an
+\item If the tense is the past, and the temporal context is
+  unspecified, then we locally quantify over a time interval
+  $[t_0,t_1]$, such that $t_1 < now$, where $now$ is a logical
+  constant representing the current timepoint. The temporal context
+  then becomes this interval.
+\item If the tense is the present and the temporal context is
+  unspecified, then the temporal context becomes the simple
+  $(now,now)$ interval.
+\item If the temporal context is specified (for example due to the presence of an adverb or an
   averbial clause, such as ``before James swam''), then the tense does
@@
   past tense adds the constraint that the temporal context ends before
-  the timepoint $now$, which is a logical constant representing the current
-  time.
-\item Otherwise, if the tense is the present, then the temporal
-  context becomes the simple $(now,now)$ interval.
-\item
-  If the tense is the past, then we locally quantify over a time
-  interval $[t_0,t_1]$, such that $t_1 < now$.
+  the timepoint $now$.
 \end{itemize}
@@
 \begin{enumerate}[align=left,font=\itshape]
-\item[exact adverbs] For such adverbs, an exact interval is
+\item[exact] For such adverbs, an exact interval is
   provided. Typical such adverbs specify in fact a single point in
@@
   introduce the interval and then restrict its bounds or its duration
-  in someway. Sometimes the restriction is an equality (``for exactly
-  two hours'').
+  in some way. Sometimes the restriction is an equality, as in ``for
+  exactly two hours''.
 
@@
 \end{multline*}
-Note that, here, we do use the input interval, allowing to interpret
-phrases such as ``In 1994, Itel was always on time.''.
+Note that here we \emph{do} use the input interval, resulting in a correct interpretation for
+phrases such as ``In 1994, Itel was always on time.'' .
 
@@
 the following discourse:
-
 \begin{enumerate}[parsep=0pt,itemsep=1pt,label=(\arabic*)]
@@
 \item Smith went to its appointment with the lawyer.
-\item Smith came back at 2pm
 \item Smith left at 4pm.
 \end{enumerate}
-One would normally not say that there is contradiction. However if (1)
-and (4) are presented in isolation, a contradiction can be flagged.
+One would normally not say that there is contradiction. However if the
+middle sentence were not present, a contradiction should be
+flagged. We leave such discourse analysis as future work, and simply
+apply unicity of action everywhere.
 
@@
 5th and May 15th. Because being in Paris is stative, we also get that
-Smith was in Paris in any subinterval.
+Smith was in Paris in any subinterval. Contrary to unicity of action,
+subsumption is always valid.
 
 \paragraph{(un)repeatable achievements}
-The principle of using unicity of action interacts well with the usual interpretation of quantifiers (and anaphora).
-Indeed, using it, we can refute problem (279), as expected in the testsuite:
+The principle of using unicity of action interacts well with the usual
+interpretation of quantifiers (and anaphora).  Indeed, using it, we
+can refute problem (279), as expected by the testsuite:
 
@@
 
-Thanks to the unicity of action of $write(smith,x,...)$ we find \([t_1,t_2] = [t_3,t_4]\), and due
-to the years 1991 and 1992 being disjoint we obtain contradiction.
+Thanks to the unicity of action of $write(smith,x,...)$ we find
+\([t_1,t_2] = [t_3,t_4]\), and due to the years 1991 and 1992 being
+disjoint we obtain contradiction.
 
@@
 two \emph{different} actions: $write(smith,x,t_1,t_2)$ and
-$write(smith,y,t_3,t_4)$, and thus we cannot deduce equality of the
+$write(smith,y,t_3,t_4)$, because $x \neq y$, and thus we cannot deduce equality of the
 intervals $t_1,t_2$ and $t_3,t_4$. In turn, the hypothesis cannot be refuted.
@@
 
+\ifanon
+Anonymized
+\else
 The research reported in this paper was supported by grant 2014-39 from the
@@
 earlier draft of the paper.
-
+\fi
 \bibliographystyle{acl_natbib}

### 2020-04-02 paper2020/paper.tex
error analysis

@@
 \paragraph{interaction with negation and group readings}
+One reason explaining the lower performance of our systems on some
+sections of the testsuite is that our interpretation of time interacts
+imperfectly with negation and group readings.
+
+The issue is that the quantification introduced by the temporal marker
+is outside the negation. For example, the premiss of (038) is
+interpreted as ``there exists a time in the past when every delegate
+did not finish the report'' --- which is incorrect in this context.
+\fracasex{038}{
+\item	No delegate finished the report.
+\hyp 	Some delegate finished the report on time.
+}
+%
+We stress that such interaction between quantifiers and negation
+already occurs without temporal analysis: one must make a difference
+between narrow scoping and wide scoping \cite{todo}. Unfortunately,
+the montegovian system over which we base our implementaiton does not
+support this phenomena. (Indeed the test suite does not appear to
+handle this case). We propose that the situation becomes
+quantitatively more difficult to handle with the added complication of
+time-quantification, but not fundamentally different.
+%
+A similar issue occurs with group readings, which are interpreted as
+conjunctions. Here, the problem is to solve the order of scoping
+between this conjunction and time-interval quantifiers.
+
 \section{Conclusions and Future Work}

### 2020-04-23 paper2020/paper.tex
wibble
failed attempt to include bibliography
notes

@@
   University of Gothenburg\\
-  {\tt firstname.lastname@@gu.se} \\}
+  {\tt firstname.lastname@gu.se} \\}
 
@@
 Stergios
+
+We do not care about classifying sentences as felicitous/infelicitous
+or grammatical/ungrammatical, rather we assume that the sentence is
+correct and care about giving it an appropriate interpretation as a
+logical formula.  Indeed, our evaluation uses entailment properties
+only.
+
 \section{Background}
@@
 Stergios
+
+We use notions of ``stative'', ``achievement'', ``activity''. Our
+definition do not seem to match perfectly the definitions found in the
+literature; however we think they are good definitions, given that
+they generally afford the correct inferences for the testsuite. (We
+assume the testsuite is correct.)
 
@@
 (The difference between actitivty and achievement is that for the
-latter time intervals can be assumed to be of nil duration. However
+latter time intervals can be assumed to be of nil\footnote{TODO: needs qualification} duration. However
 this appears to play little role, if any, in our analysis.)
@@
 Indeed, from \textbf{P3} we get that Smith was in paris between May
-5th and May 15th. Because being in Paris is stative, we also get that
+5th and May 15th. Because "being in Paris" is stative, we also get that
 Smith was in Paris in any subinterval. Contrary to unicity of action,
@@
 
+% \bibliography{../PaperTools/bibtex/jp}
+
 \end{document}

### 2020-05-04 paper2020/paper.tex
minor comments

@@
 definition do not seem to match perfectly the definitions found in the
-literature; however we think they are good definitions, given that
+literature; In general, we use the term achievement to include accomplishment as well.
+Insofar as the test suite is concerned, we find that these two categories can be collapsed into a single one.
+As such,  we think it is  a good classification, given that
 they generally afford the correct inferences for the testsuite. (We
@@
 (The difference between actitivty and achievement is that for the
-latter time intervals can be assumed to be of nil\footnote{TODO: needs qualification} duration. However
-this appears to play little role, if any, in our analysis.)
+latter time intervals can be assumed to be of nil duration. In reality, this is an oversimplification as achievements are usually of short duration. but not nil. However, 
+this appears to play little role, if any, in our analysis, and given the task-driven nature of it as well.)
 
@@
 Our analysis does not need to treat this last case specially. Indeed,
-even if $write(smith,x,.,.)$ is an activity and thus subject to
+even if $write(smith,x,.,.)$ is an activity \todo{This should be an accomplishment and thus, achievement in our classification} and thus subject to
 unicity of action, in (280), $x$ is quantified existentially; we have

### 2020-05-18 paper2020/paper.tex
intro

@@
 Stergios
+The semantics of tense and aspect has been a long standing issue in the study of formal semantics since the early days of Montague Grammar and a number of different ideas have been put forth on the table to deal with them throughout the years \citet{dowty:2012,prior:2003,steedman:2000,higginbotham:2009,fernando:2015}.
+The semantics of tense and aspect has been also central in the study of Natural Language Inference (NLI).
+The various datasets for NLI that have been proposed by the years contain examples that have some implicit or explicit reliance to inferences related to tense and aspect.
+One of the early datasets used to test logical approaches, the FraCaS test suite \citet{cooper:1996} contains a whole section of the dataset only for temporal and aspectual inference.
+This part of the FraCaS has been notoriously difficult to tackle.
+In fact, all the computational systems ran on the FraCaS so far skip this section, and usually also the anaphora and ellipsis sections.The exception with respect to the ellipsis and anaphora sections is the system presented by \citet{bernardy_type_2017,bernardy_wide_2019} where a whole run of the FraCaS excluding the temporal section has been attempted.
+In this paper, we take up the challenge of providing a computationally viable account of tense and aspect to deal with the notorious section 8 of the FraCaS test suite.
+This account is not meant to be a theoretically full account of tense and aspect, but an account that is driven by the need to cover the test suite in a way that is general enough to capture the test suite examples.
+The account is evaluated on the entailment properties of various temporal and aspectual examples, as given by the test suite.
+As such, we are not getting into the discussion of how tense and aspect might affect grammaticality or infelicitousness of various sentences.
+We take the sentences of the FraCaS to be syntactically correct and the entailment patterns they are annotated with valid.
 
-We do not care about classifying sentences as felicitous/infelicitous
-or grammatical/ungrammatical, rather we assume that the sentence is
-correct and care about giving it an appropriate interpretation as a
-logical formula.  Indeed, our evaluation uses entailment properties
-only.
+The paper is structured as follows: in chapter 2, we give a brief background of the system we are going to be using, saying a few words about its main individual components: the Grammatical Framework that is used as the syntactic parser, the Coq proof assistant which does all the reasoning and some brief remarks on temporal semantics \textbf{(JP: Anything to say on Dynamic Semantics?)}.
+In section 3, we discuss the main aspects of the compositional semantics of our system \textbf{This could go in the background actually}, and in section 4, we give our main treatment of time, using various examples from the FraCaS in exemplifying its effectiveness.
+In section 5, we have a brief look at how our system performs with respect to the temporal section of the FraCaS, but we also ran the system across the whole of the FraCaS, this being the first time that a complete run through the whole suite has been made.
+In section 6, we conclude and discuss avenues for future work.
+
+%We do not care about classifying sentences as felicitous/infelicitous
+%or grammatical/ungrammatical, rather we assume that the sentence is
+%correct and care about giving it an appropriate interpretation as a
+%logical formula.  Indeed, our evaluation uses entailment properties
+%only.
 

### 2020-05-19 paper2020/paper.tex
rework background

@@
 \section{Background}
+
 \paragraph{GF}
-If needed
+We work on a syntactic representation of the FraCaS test suite in GF
+abstract syntax~\citep{Ljunglof:2012}. Thus, we skip the
+parsing phase and avoid any syntactic ambiguity.
+
+For the purpose of this paper, the important point is that GF the
+syntax aims at a balance of sufficient abstraction to provide a
+semantically-relevant structure, but at the same time it embeds
+sufficiently many syntactic features to be able to reconstruct
+unambiguously natural-language text. Example \label{ex:onelingex}
+shows an example abstract syntax tree and its realisation in English.
+
 \paragraph{Coq}
 If needed
+
+\paragraph{Montegovian semantics}
+\todo{integrate to background}
+
+Our interpretation is cast withing the framework of a largely standard
+montegovian semantics \cite{todo}. In brief, sentences are interpreted as
+propositions, verbs and noun-phrases as predicates.  We use
+type-raising of noun-phrases, to support quantifiers~\cite{PTQ}.
+
+We support additionally the basic constructions and phenomena present
+in the testsuite, including adjectives, adverbs, nouns, verbs,
+anaphora, etc. We direct the reader to our previous work for details
+\ifanon (citations anonymized) \else
+\citet{bernardy_type_2017} \fi, but the particular
+treatment of such phenomena is essentially independent from our
+treatment of time:
+in this paper we simply ignore these aspects beyond the fact that they
+are handled correctly in most of the cases of the FraCaS testsuite.
+
+\paragraph{Dynamic semantics}
+We additionally assume support for (non-temporal) anaphora, using a
+monadic-based dynamic semantics \cite{TODO}.  Our particular
+implementation is not perfect \ifanon (citations anonymized) \else
+\citet{ bernardy_wide_2019} \fi but non-temporal anaphoroi in the
+testsuite are generally resolved as they should be.
+
+As for other basic phenomena, there is not much interaction between
+our treatment of time and non-temporal anaphora. One critical
+exception is discussed in \cref{sec:repeatable-achievements}.
+
 \paragraph{Temporal Semantics}
@@
 assume the testsuite is correct.)
-
-\section{Our Compositional Semantics}
-
-Our interpretation is cast withing the framework of a largely standard
-montegovian semantics \cite{todo}. In brief, sentences are interpreted as
-propositions, verbs and noun-phrases as predicates.  We use
-type-raising of noun-phrases, to support quantifiers \cite{PTQ}.
-
-We support additionally various constructinos and phenomena present in
-the testsuite, including adjectives, adverbs, nouns, verbs, anaphora,
-etc. We direct the reader to our previous work for details \ifanon
-(citations anonymized) \else \citet{bernardy_type_2017,bernardy_wide_2019} \fi, but the particular
-treatment of such phenomena is largely independent from our treatment
-of time.
-%
-Here we simply ignore these aspects beyond the fact that they
-are handled correctly in most of the case of the FraCaS testsuite.  In
-particular, non-temporal anaphoroi are resolved as they should be.
 
@@
 \paragraph{(un)repeatable achievements}
+\label{sec:repeatable-achievements}
 The principle of using unicity of action interacts well with the usual
@@
 \section{Results and evaluation}
+\todo{How this system is new; comparison with other systems; first system that does the whole fracas}
 \providecommand\ncases[1]{{\ensuremath{^{#1}}}}

### 2020-05-19 paper2020/paper.tex
eval intro

@@
 \section{Results and evaluation}
-\todo{How this system is new; comparison with other systems; first system that does the whole fracas}
+Our target is the FraCaS testsuite, which aims at covering a wide
+range of common natural-language phenomena. The testsuite is structured in 9 sections:
+
+The suite is structured according to the semantic phenomena involved
+in the inference process for each example, and contains nine sections:
+Quantifiers, Plurals, Anaphora, Ellipsis, Adjectives, Comparatives,
+Temporal, Verbs and Attitudes.  The system described here focuses on
+the Temporal section. However, it also supports the other eight
+sections.  To our knowledge this is the first system which attempt to
+target the temporal section. But in fact, our system is even provides
+support for all the other section. A couple of decades after its
+formulation, we propose a first attempt at covering the whole suite.
+As such, there it is no other system to compare our system with, in
+all aspects.  We can however compare with systems which target parts
+of the FraCaS testsuite, as shown in \cref{tab:results}.
+
 \providecommand\ncases[1]{{\ensuremath{^{#1}}}}

### 2020-05-21 paper2020/paper.tex
fix compilation
minor

@@
 \usepackage{enumitem}
+\usepackage{cleveref}
 \newlist{lingex}{enumerate}{3}
@@
 tense is used, the additional information is also carried by adverbs
-or adverbial phrases, in a more specific way.)
+or adverbial phrases, in a more specific way.) It is not difficult to see how other tenses can be captured under the same general framework.
 

### 2020-05-26 paper2020/paper.tex
Coq section

@@
 \paragraph{Coq}
-If needed
+ Coq is an interactive theorem prover (proof assistant). It is based on the
+calculus of coinductive constructions (CiC), i.e.  a lambda calculus
+with dependent types and coinduction (we do not use any coinduction in this paper).  Coq is a very powerful reasoning engine
+that makes it fit for implementing natural language semantics in it but also, on a more practical side, for the the task of NLI, when the latter is formalized
+as a theorem proving task. Coq supports, as already mentioned,
+ dependent typing, but also subtyping, which are instrumental in
+expressing NL semantics.
 

### 2020-05-26 paper2020/paper.tex
minor

@@
 
-\paragraph{Montegovian semantics}
+\paragraph{Montagovian semantics}
 \todo{integrate to background}
@@
 Our interpretation is cast withing the framework of a largely standard
-montegovian semantics \cite{todo}. In brief, sentences are interpreted as
+montagovian semantics \cite{todo}. In brief, sentences are interpreted as
 propositions, verbs and noun-phrases as predicates.  We use
@@
 
-In montegovian semantics, (intransitive) verbs are one-place
+In montagovian semantics, (intransitive) verbs are one-place
 predicates; in types, they are functions from entities to propositions
@@
 
-We adjust the montegovian semantics so that the interpretation of
+We adjust the montagovian semantics so that the interpretation of
 every category (propositions, verb phrases, etc.) takes a
@@
 between narrow scoping and wide scoping \cite{todo}. Unfortunately,
-the montegovian system over which we base our implementaiton does not
+the montagovian system over which we base our implementaiton does not
 support this phenomena. (Indeed the test suite does not appear to

### 2020-05-26 paper2020/paper.tex
minor

@@
 
-Our interpretation is cast withing the framework of a largely standard
+Our interpretation is cast within the framework of a largely standard
 montagovian semantics \cite{todo}. In brief, sentences are interpreted as
@@
 support this phenomena. (Indeed the test suite does not appear to
-handle this case). We propose that the situation becomes
+handle this case).
+\todo{You mean the system?
+It seems unclear what you are talking and in what way a test suite is not handling an example.
+It does not exist or?} We propose that the situation becomes
 quantitatively more difficult to handle with the added complication of

### 2020-05-26 paper2020/paper.tex
minor

@@
 The various datasets for NLI that have been proposed by the years contain examples that have some implicit or explicit reliance to inferences related to tense and aspect.
-One of the early datasets used to test logical approaches, the FraCaS test suite \citet{cooper:1996} contains a whole section of the dataset only for temporal and aspectual inference.
+One of the early datasets used to test logical approaches, the FraCaS test suite \citet{cooper:1996} contains a whole section of the dataset only for temporal and aspectual inference (section 7).
 This part of the FraCaS has been notoriously difficult to tackle.
@@
 \item Smith left at 1pm.
-\item Smith went to its appointment with the lawyer.
+\item Smith went to his appointment with the lawyer.
 \item Smith left at 4pm.
@@
 \item	She is still in Paris.
-\hyp 	Was Smith in Paris on the 7th of May, 1995?
+\hyp 	Smith was in Paris on the 7th of May, 1995.
 }
@@
 \item	Smith wrote a novel in 1991.
-\hyp 	Did Smith write it in 1992?
+\hyp 	 Smith wrote it in 1992.
 }
@@
 \hyp 	Smith wrote a novel in 1992.
-}
+}\todo{something wrong with P2 here, is there a premise there at all?}
 Here our interpretation is:
@@
 % \bibliography{../PaperTools/bibtex/jp}
-
+% Things we cannot handle (I think): John was writing a book in 1991 -> John wrote a book in 1991 (I think we predict entailment here).
+This is because there is no distinction with progressive and non progressive aspect.
+We did not do it, I guess because it was needed, but just for future reference.
+There are similar examples playing around with LEXICAL aspect (aspect grammatically encoded) that we could look at in future work
+% Accomplishments usually involve a culmination point in the action which is necessary for the event to be completed.
+This is not necessary for activities.
+John stopped writing a book not-> John wrote a book (but does imply that John was writing a book), but, John stopped running -> John run and John was runnning
 \end{document}

### 2020-05-29 paper2020/paper.tex
fix some todos

@@
 \paragraph{Montagovian semantics}
-\todo{integrate to background}
 
 Our interpretation is cast within the framework of a largely standard
-montagovian semantics \cite{todo}. In brief, sentences are interpreted as
+montagovian semantics. In brief, sentences are interpreted as
 propositions, verbs and noun-phrases as predicates.  We use
-type-raising of noun-phrases, to support quantifiers~\cite{PTQ}.
+type-raising of noun-phrases, to support quantifiers~\citep{montague_proper_1974}.
 
@@
 We additionally assume support for (non-temporal) anaphora, using a
-monadic-based dynamic semantics \cite{TODO}.  Our particular
+monadic-based dynamic semantics \cite{unger_dynamic_2011,charlow_monadic_2015,charlow_monadic_2017}.  Our particular
 implementation is not perfect \ifanon (citations anonymized) \else
-\citet{ bernardy_wide_2019} \fi but non-temporal anaphoroi in the
+\citet{bernardy_wide_2019} \fi but non-temporal anaphoroi in the
 testsuite are generally resolved as they should be.
@@
 \item	Smith wrote a novel in 1991.
-\item 	Did Smith write a novel in 1992?
 \hyp 	Smith wrote a novel in 1992.
-}\todo{something wrong with P2 here, is there a premise there at all?}
+}
 Here our interpretation is:
@@
 between narrow scoping and wide scoping \cite{todo}. Unfortunately,
-the montagovian system over which we base our implementaiton does not
-support this phenomena. (Indeed the test suite does not appear to
-handle this case).
-\todo{You mean the system?
-It seems unclear what you are talking and in what way a test suite is not handling an example.
-It does not exist or?} We propose that the situation becomes
+the montagovian system over which we base our implementation does not
+support this phenomena. (In fact the FraCaS test suite does not appear to
+test if the difference between narrow and wide scoping is handled properly.)
+We propose that the situation becomes
 quantitatively more difficult to handle with the added complication of

### 2020-06-03 paper2020/paper.tex
intro revision

@@
 \section{Introduction}
-Stergios
-The semantics of tense and aspect has been a long standing issue in the study of formal semantics since the early days of Montague Grammar and a number of different ideas have been put forth on the table to deal with them throughout the years \citet{dowty:2012,prior:2003,steedman:2000,higginbotham:2009,fernando:2015}.
-The semantics of tense and aspect has been also central in the study of Natural Language Inference (NLI).
-The various datasets for NLI that have been proposed by the years contain examples that have some implicit or explicit reliance to inferences related to tense and aspect.
-One of the early datasets used to test logical approaches, the FraCaS test suite \citet{cooper:1996} contains a whole section of the dataset only for temporal and aspectual inference (section 7).
-This part of the FraCaS has been notoriously difficult to tackle.
-In fact, all the computational systems ran on the FraCaS so far skip this section, and usually also the anaphora and ellipsis sections.The exception with respect to the ellipsis and anaphora sections is the system presented by \citet{bernardy_type_2017,bernardy_wide_2019} where a whole run of the FraCaS excluding the temporal section has been attempted.
-In this paper, we take up the challenge of providing a computationally viable account of tense and aspect to deal with the notorious section 8 of the FraCaS test suite.
-This account is not meant to be a theoretically full account of tense and aspect, but an account that is driven by the need to cover the test suite in a way that is general enough to capture the test suite examples.
-The account is evaluated on the entailment properties of various temporal and aspectual examples, as given by the test suite.
-As such, we are not getting into the discussion of how tense and aspect might affect grammaticality or infelicitousness of various sentences.
-We take the sentences of the FraCaS to be syntactically correct and the entailment patterns they are annotated with valid.
+\label{sec:introduction}
 
-The paper is structured as follows: in chapter 2, we give a brief background of the system we are going to be using, saying a few words about its main individual components: the Grammatical Framework that is used as the syntactic parser, the Coq proof assistant which does all the reasoning and some brief remarks on temporal semantics \textbf{(JP: Anything to say on Dynamic Semantics?)}.
-In section 3, we discuss the main aspects of the compositional semantics of our system \textbf{This could go in the background actually}, and in section 4, we give our main treatment of time, using various examples from the FraCaS in exemplifying its effectiveness.
-In section 5, we have a brief look at how our system performs with respect to the temporal section of the FraCaS, but we also ran the system across the whole of the FraCaS, this being the first time that a complete run through the whole suite has been made.
-In section 6, we conclude and discuss avenues for future work.
+The semantics of tense and aspect has been a long standing issue in
+the study of formal semantics since the early days of Montague Grammar
+and a number of different ideas have been put forth on the table to
+deal with them throughout the
+years~\citep{dowty:2012,prior:2003,steedman:2000,higginbotham:2009,fernando:2015}. The
+semantics of tense and aspect has been also central in the study of
+Natural Language Inference (NLI)\todo{citation?}. The various datasets
+for NLI that have been proposed by the years contain examples that
+have some implicit or explicit reliance on inferences related to tense
+and aspect. One of the early datasets used to test logical approaches,
+the FraCaS test suite~\citep{cooper:1996} contains a whole section
+dedicated to temporal and aspectual inference (section 7). This part
+of the FraCaS test suite has been notoriously difficult to tackle. In
+fact, all the computational systems ran on the FraCaS test suite so
+far skip this section. Additionally, they also skip the anaphora
+and ellipsis sections, with the exception of the system presented by
+\citet{bernardy_type_2017,bernardy_wide_2019}, which offers a complete whole run of the
+FraCaS test suite, excluding the temporal section. In this
+paper, we take up the challenge of providing a computationally viable
+account of tense and aspect to deal with the notorious section 8\todo{or 7? 8 was not discussed previously} of
+the FraCaS test suite. Our account is not meant to be a theoretically
+full account of tense and aspect, but rather an account that is driven by the
+need to cover the test suite in a way that is general enough to
+capture the test suite examples. The account is evaluated on the
+entailment properties of various temporal and aspectual examples, as
+given by the test suite. As such, we are not getting into the
+discussion of how tense and aspect might affect grammaticality or
+infelicitousness of various sentences. We take the sentences of the
+FraCaS suite to be syntactically correct and the entailment patterns they
+are annotated with valid.
+
+The paper is structured as follows: in \cref{sec:background}, we give
+a brief summary of the frameworks whose various subsystems rely on. In
+particular, the Grammatical Framework is used to construct the
+syntactic parser; the Coq proof assistant checks all the reasoning; a
+monad-based dynamic semantics deals with references (anaphora).  We
+also provide some brief remarks on temporal semantics. In section 3,
+we discuss the main aspects of the compositional semantics of our
+system \textbf{This could go in the background actually}, and in
+section 4, we give our main treatment of time, using various examples
+from the FraCaS suite in exemplifying its effectiveness. In section 5,
+we take a brief look at how our system performs with respect to the
+FraCaS suite. We ran the system across the whole suite, this being the
+first time that a complete run through the whole suite has been made,
+but we look in particular at the performance on the temporal
+section. In section 6, we conclude and discuss avenues for future
+work.
 
@@
 \section{Background}
+\label{sec:background}
 

### 2020-06-03 paper2020/paper.tex
changes

--- 2020-05-29 paper2020/paper.tex
@@
 \section{Conclusions and Future Work}
-Stergios
+We have presented a first attempt for a computational approach dealing with the temporal section of the FraCaS test suite.
+This is the first time a system is ran against this section, at the same time, this concludes our work in handling the whole of the FraCaS test suite using a combination of a number of well-studied tools: type theory, Grammatical Framework, Monadic Dynamic Semantics and proof assistant technology (Coq).
 \section*{Acknowledgements}

### 2020-06-03 paper2020/paper.tex
conclusions

--- 2020-06-03 paper2020/paper.tex
@@
 \section{Introduction}
-\label{sec:introduction}
-
-The semantics of tense and aspect has been a long standing issue in
-the study of formal semantics since the early days of Montague Grammar
-and a number of different ideas have been put forth on the table to
-deal with them throughout the
-years~\citep{dowty:2012,prior:2003,steedman:2000,higginbotham:2009,fernando:2015}. The
-semantics of tense and aspect has been also central in the study of
-Natural Language Inference (NLI)\todo{citation?}. The various datasets
-for NLI that have been proposed by the years contain examples that
-have some implicit or explicit reliance on inferences related to tense
-and aspect. One of the early datasets used to test logical approaches,
-the FraCaS test suite~\citep{cooper:1996} contains a whole section
-dedicated to temporal and aspectual inference (section 7). This part
-of the FraCaS test suite has been notoriously difficult to tackle. In
-fact, all the computational systems ran on the FraCaS test suite so
-far skip this section. Additionally, they also skip the anaphora
-and ellipsis sections, with the exception of the system presented by
-\citet{bernardy_type_2017,bernardy_wide_2019}, which offers a complete whole run of the
-FraCaS test suite, excluding the temporal section. In this
-paper, we take up the challenge of providing a computationally viable
-account of tense and aspect to deal with the notorious section 8\todo{or 7? 8 was not discussed previously} of
-the FraCaS test suite. Our account is not meant to be a theoretically
-full account of tense and aspect, but rather an account that is driven by the
-need to cover the test suite in a way that is general enough to
-capture the test suite examples. The account is evaluated on the
-entailment properties of various temporal and aspectual examples, as
-given by the test suite. As such, we are not getting into the
-discussion of how tense and aspect might affect grammaticality or
-infelicitousness of various sentences. We take the sentences of the
-FraCaS suite to be syntactically correct and the entailment patterns they
-are annotated with valid.
-
-The paper is structured as follows: in \cref{sec:background}, we give
-a brief summary of the frameworks whose various subsystems rely on. In
-particular, the Grammatical Framework is used to construct the
-syntactic parser; the Coq proof assistant checks all the reasoning; a
-monad-based dynamic semantics deals with references (anaphora).  We
-also provide some brief remarks on temporal semantics. In section 3,
-we discuss the main aspects of the compositional semantics of our
-system \textbf{This could go in the background actually}, and in
-section 4, we give our main treatment of time, using various examples
-from the FraCaS suite in exemplifying its effectiveness. In section 5,
-we take a brief look at how our system performs with respect to the
-FraCaS suite. We ran the system across the whole suite, this being the
-first time that a complete run through the whole suite has been made,
-but we look in particular at the performance on the temporal
-section. In section 6, we conclude and discuss avenues for future
-work.
+Stergios
+The semantics of tense and aspect has been a long standing issue in the study of formal semantics since the early days of Montague Grammar and a number of different ideas have been put forth on the table to deal with them throughout the years \citet{dowty:2012,prior:2003,steedman:2000,higginbotham:2009,fernando:2015}.
+The semantics of tense and aspect has been also central in the study of Natural Language Inference (NLI)\ The various datasets for NLI that have been proposed by the years contain examples that have some implicit or explicit reliance ton nferences related to tense and aspect.
+One of the early datasets used to test logical approaches, the FraCaS test suite ~\citetp{cooper:196} contins a whole section of the dataset only for temporal and aspectual inference (section 7).
+This part of the FraCaS test suite has been  difficult to tackle.
+In fact, all the computational systems ran on the FraCaS so far skip this section, and usuall.
+Additionally, they ora and ellipsis sns.The exception with respect to the ellipsis and anaphora sections is, with the exception of the system presented by \cibernardy_wide_2019}, which offeres a complete whole run of the FraCaS test uite, excluding the temporion as been atIn this paper, we take upge of providing a computationally viable account of tense and aspect to deal with the notorious section 8 of the FraCaS test suite.
+ThisOur account is not meant to be a theoretically full account of tense and aspect, but rather an acnt that is driven by the need to cover the test suite in a way that is general enough to captutest suite examples.
+The account is evaluated on the entailment properties of various temporal and aspectual examples, as given by the test suite.
+As such, we are not getting into the discussion of how tense and aspect might affect grammaticality or infelicitousness of various sentences.
+We take the sentences of the FraCaS suite to be syntactically correct and the entailment patterns they are annotated with valid.
+paper is structured as follows: in chapter 2\cref{sec:background}, we give a brief background of the system wsummary of the frameworks whose vare going to be using, saying a few words about its main individual components: the Grammatical Framework that is used asto construct the syntactic parser,; the Coq proof assistant which doechecks all the reasoning; and some brief remarks on temporal semantics \textbf{(JP: Anything to say on Dynamic Semantics?)}.
+In section 3, we discuss the main aspects of the compositional semantics of our system \textbf{This could go in the background actually}, and in section 4, we give our main treatment of time, using various examples from the FraCaS suite in exemplifying its effectiveness.
+In section 5, we havtake a brief look at how our system performs with respect to the temporal section of the FraCaS, but we also ran the system across the whole of the FraCaSsuite, this being the first time that a complete run through the whole suite has been made,
+. In section 6, we conclude and discuss avenues for future work.
 
@@
 \section{Background}
-\label{sec:background}
 
@@
 \section{Conclusions and Future Work}
-Stergios
+We have presented a first attempt for a computational approach dealing with the temporal section of the FraCaS test suite.
+This is the first time a system is ran against this section, at the same time, this concludes our work in handling the whole of the FraCaS test suite using a combination of a number of well-studied tools: type theory, parsing using the Grammatical Framework (GF), Monadic Dynamic Semantics and proof assistant technology (Coq).
+It achieves an accuracy of 0.71 on the Temporal Section and 0.78 overall.
+One of the things to be looked at is fixing the issues associated with parts of the test suite that ``broke'' so to say when the temporal analysis was introduced.
+Some of these have been already mentioned: interaction of the temporal variables with group readings and negation.
+
+Another issue to be looking at is the codification of lexical aspect.
+In what we have been doing so far, we did not distinguish between the present continuous and the present or the simple past and the past continuous and their interaction with non-lexical aspect.
+The main reason for this  is that in the test suite, even though there are examples of inferences based on the either lexical or non-lexical aspect, no example where an interaction between the two exists.
+For example, one such case is the one shown below and exemplifies the difference between activities and accomplishments with respect to the different inferences they give when used  in the past continuous:
+
+\fracasex{extra1}{
+	\item	John was writing a book. 
+	\hyp 	John wrote a book (UNK).
+}
+
+\fracasex{extra2}{
+	\item	John was running. 
+	\hyp 	John ran (YES).
+}
+
+\fracasex{extra1}{
+	\item	John wrote a book. 
+	\hyp 	John was writing a book (YES).
+}
+
+\fracasex{extra2}{
+	\item	John ran. 
+	\hyp 	John was running (YES?).
+}
+
 \section*{Acknowledgements}

### 2020-06-03 Dynamic.hs
very simple support for progressive

@@ -111,6 +111,7 @@ instance Monoid Temporal where
   mempty = UnspecifiedTime
 
 data ExtraArgs = ExtraArgs { extraPreps :: [(Var,Object)] -- prepositions
+                           , extraProgressive :: Bool
                            , extraAdvs :: (Object -> Prop) -> Object -> Prop -- adverbs
                            , extraCompClass :: Optional CN''
                            , extraTime :: Temporal
@@ -450,7 +451,7 @@ type Pron = NP
 -- Extra objects and S'
 
 emptyObjs :: ExtraArgs
-emptyObjs = ExtraArgs {extraPreps = [], extraAdvs = id, extraTime = mempty, extraCompClass = mempty}
+emptyObjs = ExtraArgs {extraProgressive = False, extraPreps = [], extraAdvs = id, extraTime = mempty, extraCompClass = mempty}
 
 noExtraObjs :: S' -> Prop
 noExtraObjs f = fst $ f emptyObjs
@@ -480,7 +481,7 @@ verbAspect "live_Vin" = Activity
 verbAspect "run_V2" = Activity
 verbAspect "use_VVTiming" = Activity
 verbAspect "discover_V2" = Activity -- treated as such in FraCas 291
-verbAspect "_BE_" = Activity -- mostly used for "it is now date"
+verbAspect "_BE_" = Activity -- mostly used for "it is now <DATE>"
 verbAspect "finish_VVTiming" = Activity -- Because we can finish "within" an interval, see Coq code.
 verbAspect _ = Achievement
 
@@ -496,7 +497,8 @@ appArgs isTimed nm objs@(_:_) (ExtraArgs {..}) = (extraAdvs (app (pAdverbs time'
             ExactTime (_,t0) -> ExactTime (t0,t0)
             ForceTime tspan -> ForceTime tspan
             UnspecifiedTime -> UnspecifiedTime
-        nmPrep = nm ++ concatMap fst prepositions
+        nmPrep = prog ++ nm ++ concatMap fst prepositions
+        prog = if extraProgressive then "PROG_" else ""
         indirectObjects = init objs
         subject = last objs
         cleanedPrepositions = sortBy (compare `on` fst) $ nubBy ((==) `on` fst) extraPreps
[…]

### 2020-06-03 MS.hs

@@ -935,7 +935,9 @@ complVQ :: VQ -> QS -> VP
 complVQ = id
 
 progrVPa :: VP -> VP
-progrVPa = id -- ignoring tense
+progrVPa vp = do
+  vp' <- vp
+  return (\x -> progressively (vp' x))
 
 complVV :: VV -> VP -> VP
 complVV vv vp = vv <*> vp

### 2020-06-04 paper2020/paper.tex
salvage previous commit

--- 2020-06-03 paper2020/paper.tex
@@
 We have presented a first attempt for a computational approach dealing with the temporal section of the FraCaS test suite.
-This is the first time a system is ran against this section, at the same time, this concludes our work in handling the whole of the FraCaS test suite using a combination of a number of well-studied tools: type theory, Grammatical Framework, Monadic Dynamic Semantics and proof assistant technology (Coq).
+This is the first time a system is ran against this section, at the same time, this concludes our work in handling the whole of the FraCaS test suite using a combination of a number of well-studied tools: type theory, parsing using the Grammatical Framework (GF), Monadic Dynamic Semantics and proof assistant technology (Coq).
+It achieves an accuracy of 0.71 on the Temporal Section and 0.78 overall.
+One of the things to be looked at is fixing the issues associated with parts of the test suite that ``broke'' so to say when the temporal analysis was introduced.
+Some of these have been already mentioned: interaction of the temporal variables with group readings and negation.
+
+Another issue to be looking at is the codification of lexical aspect.
+In what we have been doing so far, we did not distinguish between the present continuous and the present or the simple past and the past continuous and their interaction with non-lexical aspect.
+The main reason for this  is that in the test suite, even though there are examples of inferences based on the either lexical or non-lexical aspect, no example where an interaction between the two exists.
+For example, one such case is the one shown below and exemplifies the difference between activities and accomplishments with respect to the different inferences they give when used  in the past continuous:
+
+\fracasex{extra1}{
+	\item	John was writing a book. 
+	\hyp 	John wrote a book (UNK).
+}
+
+\fracasex{extra2}{
+	\item	John was running. 
+	\hyp 	John ran (YES).
+}
+
+\fracasex{extra1}{
+	\item	John wrote a book. 
+	\hyp 	John was writing a book (YES).
+}
+
+\fracasex{extra2}{
+	\item	John ran. 
+	\hyp 	John was running (YES?).
+}
+
 \section*{Acknowledgements}

### 2020-06-04 paper2020/paper.tex
revise and notes in background sec.

@@
 \paragraph{GF}
-We work on a syntactic representation of the FraCaS test suite in GF
-abstract syntax~\citep{Ljunglof:2012}. Thus, we skip the
-parsing phase and avoid any syntactic ambiguity.
+The grammatical framework (GF, \citet{Ranta:GF}) is a powerful parser
+generator for natural languages, based on type-theoretical abstract grammars.
+This work relies a syntactic representation of the FraCaS test suite in GF
+abstract syntax~\citep{Ljunglof:2012}. Thus, we skip the parsing phase
+and avoid any syntactic ambiguity.
 
@@
 sufficiently many syntactic features to be able to reconstruct
-unambiguously natural-language text. Example \label{ex:onelingex}
+unambiguously natural-language text. \cref{ex:onelingex}
 shows an example abstract syntax tree and its realisation in English.
@@
 \paragraph{Coq}
- Coq is an interactive theorem prover (proof assistant). It is based on the
-calculus of coinductive constructions (CiC), i.e.  a lambda calculus
-with dependent types and coinduction (we do not use any coinduction in this paper).  Coq is a very powerful reasoning engine
-that makes it fit for implementing natural language semantics in it but also, on a more practical side, for the the task of NLI, when the latter is formalized
-as a theorem proving task. Coq supports, as already mentioned,
- dependent typing, but also subtyping, which are instrumental in
-expressing NL semantics.
+Coq is an interactive theorem prover (proof assistant). It is based on
+the calculus of coinductive constructions~\citep{werner_une_1994} We
+do not use any coinduction (or even induction) in this paper, relying
+on the pure lambda-calculus inner core of Coq.  Coq is a very
+powerful reasoning engine that makes it fit for implementing natural
+language semantics in it but also, on a more practical side, for the
+the task of NLI, when the latter is formalized as a theorem proving
+task. Coq supports, as already mentioned, dependent typing, but also
+subtyping, which are instrumental in expressing NL semantics.
 
@@
 We additionally assume support for (non-temporal) anaphora, using a
-monadic-based dynamic semantics \cite{unger_dynamic_2011,charlow_monadic_2015,charlow_monadic_2017}.  Our particular
-implementation is not perfect \ifanon (citations anonymized) \else
-\citet{bernardy_wide_2019} \fi but non-temporal anaphoroi in the
-testsuite are generally resolved as they should be.
+monadic-based dynamic semantics
+\cite{unger_dynamic_2011,charlow_monadic_2015,charlow_monadic_2017}.
+Our particular implementation is not perfect \ifanon (citations
+anonymized) \else \citet{bernardy_wide_2019} \fi but non-temporal
+anaphoroi in the testsuite are generally resolved as they should be.
 
@@
 \paragraph{Temporal Semantics}
-Stergios
+\todo{More general notion of temporal semantics? Should this be moved somewhere else?}
 
@@
 definition do not seem to match perfectly the definitions found in the
-literature; In general, we use the term achievement to include accomplishment as well.
-Insofar as the test suite is concerned, we find that these two categories can be collapsed into a single one.
-As such,  we think it is  a good classification, given that
-they generally afford the correct inferences for the testsuite. (We
-assume the testsuite is correct.)
+literature; In general, we use the term achievement to include
+accomplishment as well. Insofar as the test suite is concerned, we
+find that these two categories can be collapsed into a single one.  As
+such, we think it is a good classification, given that they generally
+afford the correct inferences for the testsuite. (We assume the
+testsuite is correct.)
 

### 2020-06-04 paper2020/paper.tex
wibble

@@
 sufficiently many syntactic features to be able to reconstruct
-unambiguously natural-language text. \cref{ex:onelingex}
+unambiguously natural-language text. \cref{ex:nobel}
 shows an example abstract syntax tree and its realisation in English.
@@
 \paragraph{Temporal Semantics}
-\todo{More general notion of temporal semantics? Should this be moved somewhere else?}
+\jp{More general intro to temporal semantics? Should this be moved
+  somewhere else?}
 
@@
 additional time parameters, one corresponding with the starting time
-of the action and one corresponding to its stopping time ($(time × time × e) \to t$). For example,
+of the action and one corresponding to its stopping time ($(e × time × time) \to t$). For example,
 if John walked between $t_0$ and $t_1$, we would have:
@@
 using a simple newtonian model of time, corresponding to a layman
-intuition.)
+intuition of a linear constant flow of time.)
 
@@
 every category (propositions, verb phrases, etc.) takes a
-\emph{temporal context} as an additional parameter, serves as a time
+\emph{temporal context} as an additional parameter, which serves as a time
 reference for the interpretation of all time-dependent semantics
@@
 be an timespan or an explicitly unspecified context.
-\newcommand\nospan{\ensuremath{─}}
-\newcommand\anyspan{\ensuremath{\_}} The timespan in the context is
+\newcommand\nospan{\ensuremath{-}}
+\newcommand\anyspan{\ensuremath{\_}}\todo{better notation} The timespan in the context is
 optional because, in certain situations, the semantics is different
@@
 
-\onelingex{A scandinavian won the nobel prize.\label{ex:onelingex}
+\onelingex{A scandinavian won the nobel prize.\label{ex:nobel}
 \\

### 2020-06-09 paper2020/paper.tex
formatting
blurb on progressive

@@
 
-\paragraph{time references and aspectual classes}
+\paragraph{Time references and aspectual classes}
 
@@
 
+\paragraph{Progressive aspect}
+
+We treat verbs in the progressive form as different semantically. For
+example, ``John was writing a book'' is encoded as
+$∃t1, t2. t1<= t2, t2 <= now, PROG_write(John,book,t1,t2)$, while
+``John wrote a book'' is encoded as
+$∃t1, t2. t1<= t2, t2 <= now, write(John,book,t1,t2)$. This is
+necessary, because $PROG_write$ is subject to
+$\constant{subsumption}$: if John is writing in interval $[t1,t2]$
+then he is writing in any subinterval. However, the same cannot be
+said of $write$: the end-points of the interval indicates the time
+needed to complete the achievement. (For example, John wrote a book in
+1993 does not entail John wrote a book in March 1993.) Thus, $write$
+is on the contrary subject to $\constant{unicity}$. Finally, we also
+have $write(x,y,t1,t2) -> PROG_write(x,y,t1,t2)$. That is, the
+achievement variant implies the stative variant, in the same interval.
+Consequently we get the entailment from ``John wrote a book'' to
+``John was writing a book''.
+
+Finally we note that not all verbs are subject to the
+stative/achievement distinction induced by the progressive. For
+example, the phrases ``John ran'' and ``John is running'' appear to be
+logically equivalent, for our purposes\todo{For the FraCaS suite?}.
+
 \section{Results and evaluation}
@@
 \section{Conclusions and Future Work}
-We have presented a first attempt for a computational approach dealing with the temporal section of the FraCaS test suite.
-This is the first time a system is ran against this section, at the same time, this concludes our work in handling the whole of the FraCaS test suite using a combination of a number of well-studied tools: type theory, parsing using the Grammatical Framework (GF), Monadic Dynamic Semantics and proof assistant technology (Coq).
-It achieves an accuracy of 0.71 on the Temporal Section and 0.78 overall.
-One of the things to be looked at is fixing the issues associated with parts of the test suite that ``broke'' so to say when the temporal analysis was introduced.
-Some of these have been already mentioned: interaction of the temporal variables with group readings and negation.
+We have presented a first attempt for a computational approach dealing
+with the temporal section of the FraCaS test suite. This is the first
+time a system is ran against this section, at the same time, this
+concludes our work in handling the whole of the FraCaS test suite
+using a combination of a number of well-studied tools: type theory,
+parsing using the Grammatical Framework (GF), Monadic Dynamic
+Semantics and proof assistant technology (Coq). It achieves an
+accuracy of 0.71 on the Temporal Section and 0.78 overall. One of the
+things to be looked at is fixing the issues associated with parts of
+the test suite that ``broke'' so to say when the temporal analysis was
+introduced. Some of these have been already mentioned: interaction of
+the temporal variables with group readings and negation.
 
-Another issue to be looking at is the codification of lexical aspect.
-In what we have been doing so far, we did not distinguish between the present continuous and the present or the simple past and the past continuous and their interaction with non-lexical aspect.
-The main reason for this  is that in the test suite, even though there are examples of inferences based on the either lexical or non-lexical aspect, no example where an interaction between the two exists.
-For example, one such case is the one shown below and exemplifies the difference between activities and accomplishments with respect to the different inferences they give when used  in the past continuous:
+Another issue to be looking at is the codification of lexical
+aspect.\todo{update} In what we have been doing so far, we did not distinguish
+between the present continuous and the present or the simple past and
+the past continuous and their interaction with non-lexical aspect.
+The main reason for this is that in the test suite, even though there
+are examples of inferences based on the either lexical or non-lexical
+aspect, no example where an interaction between the two exists. For
+example, one such case is the one shown below and exemplifies the
+difference between activities and accomplishments with respect to the
+different inferences they give when used in the past continuous:
 
@@
 % \bibliography{../PaperTools/bibtex/jp}
-% Things we cannot handle (I think): John was writing a book in 1991 -> John wrote a book in 1991 (I think we predict entailment here).
-This is because there is no distinction with progressive and non progressive aspect.
-We did not do it, I guess because it was needed, but just for future reference.
-There are similar examples playing around with LEXICAL aspect (aspect grammatically encoded) that we could look at in future work
-% Accomplishments usually involve a culmination point in the action which is necessary for the event to be completed.
-This is not necessary for activities.
-John stopped writing a book not-> John wrote a book (but does imply that John was writing a book), but, John stopped running -> John run and John was runnning
+
+% Things we cannot handle (I think): John was writing a book in 1991
+% -> John wrote a book in 1991 (I think we predict entailment
+% here). This is because there is no distinction with progressive and
+% non progressive aspect. We did not do it, I guess because it was
+% needed, but just for future reference. There are similar examples
+% playing around with LEXICAL aspect (aspect grammatically encoded)
+% that we could look at in future work
+
+% Accomplishments usually involve a culmination point in the action
+% which is necessary for the event to be completed. This is not
+% necessary for activities. John stopped writing a book not-> John
+% wrote a book (but does imply that John was writing a book), but,
+% John stopped running -> John run and John was runnning
+
 \end{document}

### 2020-06-09 paper2020/paper.tex
typography

@@
 example, ``John was writing a book'' is encoded as
-$∃t1, t2. t1<= t2, t2 <= now, PROG_write(John,book,t1,t2)$, while
+$∃t1, t2. t1≤ t2, t2 ≤ now, PROG\_write(John,book,t1,t2)$, while
 ``John wrote a book'' is encoded as
-$∃t1, t2. t1<= t2, t2 <= now, write(John,book,t1,t2)$. This is
-necessary, because $PROG_write$ is subject to
+$∃t1, t2. t1≤ t2, t2 ≤ now, write(John,book,t1,t2)$. This is
+necessary, because $PROG\_write$ is subject to
 $\constant{subsumption}$: if John is writing in interval $[t1,t2]$
@@
 is on the contrary subject to $\constant{unicity}$. Finally, we also
-have $write(x,y,t1,t2) -> PROG_write(x,y,t1,t2)$. That is, the
+have $write(x,y,t1,t2) → PROG\_write(x,y,t1,t2)$. That is, the
 achievement variant implies the stative variant, in the same interval.
@@
 % Things we cannot handle (I think): John was writing a book in 1991
-% -> John wrote a book in 1991 (I think we predict entailment
+% → John wrote a book in 1991 (I think we predict entailment
 % here). This is because there is no distinction with progressive and
@@
 % which is necessary for the event to be completed. This is not
-% necessary for activities. John stopped writing a book not-> John
+% necessary for activities. John stopped writing a book not→ John
 % wrote a book (but does imply that John was writing a book), but,
-% John stopped running -> John run and John was runnning
+% John stopped running → John run and John was runnning
 

### 2020-06-09 paper2020/paper.tex
wibble

@@
 example, ``John was writing a book'' is encoded as
-$∃t1, t2. t1≤ t2, t2 ≤ now, PROG\_write(John,book,t1,t2)$, while
-``John wrote a book'' is encoded as
-$∃t1, t2. t1≤ t2, t2 ≤ now, write(John,book,t1,t2)$. This is
+$∃(t_1, t_2). t_1≤ t_2, t_2 ≤ now, PROG\_write(John,book,t_1,t_2)$,
+while ``John wrote a book'' is encoded as
+$∃(t_1, t_2). t_1≤ t_2, t_2 ≤ now, write(John,book,t_1,t_2)$. This is
 necessary, because $PROG\_write$ is subject to
-$\constant{subsumption}$: if John is writing in interval $[t1,t2]$
+$\constant{subsumption}$: if John is writing in interval $[t_1,t_2]$
 then he is writing in any subinterval. However, the same cannot be
 said of $write$: the end-points of the interval indicates the time
-needed to complete the achievement. (For example, John wrote a book in
-1993 does not entail John wrote a book in March 1993.) Thus, $write$
-is on the contrary subject to $\constant{unicity}$. Finally, we also
-have $write(x,y,t1,t2) → PROG\_write(x,y,t1,t2)$. That is, the
+needed to complete the achievement. (For example, ``John wrote a book
+in 1993'' neither entails ``John wrote a book in January 1993'' nor
+``John wrote a book in December 1993''.) That is, $write$ is on the
+contrary subject to $\constant{unicity}$. Finally, we also have
+$write(x,y,t_1,t_2) → PROG\_write(x,y,t_1,t_2)$. That is, the
 achievement variant implies the stative variant, in the same interval.

### 2020-06-09 paper2020/paper.tex
bib fixes
renumber sections

@@
 monad-based dynamic semantics deals with references (anaphora).  We
-also provide some brief remarks on temporal semantics. In section 3,
-we discuss the main aspects of the compositional semantics of our
-system \textbf{This could go in the background actually}, and in
-section 4, we give our main treatment of time, using various examples
-from the FraCaS suite in exemplifying its effectiveness. In section 5,
+also provide some brief remarks on temporal semantics. In \cref{sec:our-semantics},
+we discuss the main aspects of the compositional semantics of our, using various examples
+from the FraCaS suite in exemplifying its effectiveness. In \cref{sec:results},
 we take a brief look at how our system performs with respect to the
@@
 but we look in particular at the performance on the temporal
-section. In section 6, we conclude and discuss avenues for future
+section. In \cref{sec:conclusions}, we conclude and discuss avenues for future
 work.
@@
 \section{Our Treatment of time}
+\label{sec:our-semantics}
 
@@
 \section{Results and evaluation}
+\label{sec:results}
 Our target is the FraCaS testsuite, which aims at covering a wide
@@
 \section{Conclusions and Future Work}
+\label{sec:conclusion}
 We have presented a first attempt for a computational approach dealing
@@
 
-% \bibliography{../PaperTools/bibtex/jp}
+\bibliography{../PaperTools/bibtex/jp,paper2020}
 

### 2020-06-09 paper2020/paper.tex
intro

@@
 years~\citep{dowty:2012,prior:2003,steedman:2000,higginbotham:2009,fernando:2015}. The
-semantics of tense and aspect has been also central in the study of
+semantics of tense and aspect has been also considered in the study of
 Natural Language Inference (NLI)\todo{citation?}. The various datasets
@@
 paper, we take up the challenge of providing a computationally viable
-account of tense and aspect to deal with the notorious section 8\todo{or 7? 8 was not discussed previously} of
+account of tense and aspect to deal with the notorious section 7\todo{or 7? 8 was not discussed previously} of
 the FraCaS test suite. Our account is not meant to be a theoretically

### 2020-06-09 paper2020/paper.tex
handle various todos

@@
 \paragraph{Temporal Semantics}
-\jp{More general intro to temporal semantics? Should this be moved
-  somewhere else?}
 
-We use notions of ``stative'', ``achievement'', ``activity''. Our
-definition do not seem to match perfectly the definitions found in the
-literature; In general, we use the term achievement to include
-accomplishment as well. Insofar as the test suite is concerned, we
-find that these two categories can be collapsed into a single one.  As
-such, we think it is a good classification, given that they generally
-afford the correct inferences for the testsuite. (We assume the
-testsuite is correct.)
+In this paper we borrow several notions from classical temporal
+semantics such as ``stative'', ``achievement'', ``activity'', etc.,
+even though our definitions do not perfectly match the classical
+ones. We explain our precise meaning forthese terms in the body of the
+paper. Nevertheless, we refer the reader to \citet{steedman:2000} for
+an extensive review of temporal semantics.
+
+For the \textit{cognoscenti}, we can already point out some
+differences in terminology: we use the term activity as a general term
+which encompasses the three classical notions of activites,
+achievements and accomplishments. Indeed, insofar as the test suite is
+concerned, we find that these three categories can be collapsed into a
+single one.  As such, we think it is a good classification, given that
+they generally afford the correct inferences for the testsuite. (In
+this paper, we always assume that the FraCaS testsuite is correct.)
 
@@
 }}
-In our semantics we deal only with present and simple past
-tenses. \jp{Is this enough for theory?} Indeed we find that FraCas
+In our semantics we deal only with present and past
+tenses (simple and continuous). Indeed we find that FraCas
 does not exercise additional specific tenses (if a more complicated
@@
 
-The procedure to handle tense annotations is as follows:
+Even though we will discuss a refinement to handle the past continuous
+at the end of this section, the procedure to handle tense annotations
+is as follows:
 \begin{itemize}
@@
 Unicity of action holds only if the aspectual class of the proposition
-$P$ is \emph{achievement} or \emph{actitivty}.\todo{reference in literature}
+$P$ is \emph{actitivty}~\citep{steedman:2000} (which, for our
+purposes, includes \emph{achievements} and \emph{accomplishments}).
 
-(The difference between actitivty and achievement is that for the
-latter time intervals can be assumed to be of nil duration. In reality, this is an oversimplification as achievements are usually of short duration. but not nil. However, 
-this appears to play little role, if any, in our analysis, and given the task-driven nature of it as well.)
+(The difference between actitivty and accomplishments on the one and
+achievement on the other hand is that for the latter, time intervals
+can be assumed to be of nil duration. In reality, this is an
+oversimplification as achievements are usually of short duration, but
+not nil. However, this appears to play little role, if any, in our
+analysis, and given the its task-driven nature as well.)
 
@@
 Our analysis does not need to treat this last case specially. Indeed,
-even if $write(smith,x,.,.)$ is an activity \todo{This should be an accomplishment and thus, achievement in our classification} and thus subject to
+even if $write(smith,x,.,.)$ is an activity and thus subject to
 unicity of action, in (280), $x$ is quantified existentially; we have
 two \emph{different} actions: $write(smith,x,t_1,t_2)$ and
-$write(smith,y,t_3,t_4)$, because $x \neq y$, and thus we cannot deduce equality of the
-intervals $t_1,t_2$ and $t_3,t_4$. In turn, the hypothesis cannot be refuted.
+$write(smith,y,t_3,t_4)$, because $x \neq y$, and thus we cannot
+deduce equality of the intervals $t_1,t_2$ and $t_3,t_4$. In turn, the
+hypothesis cannot be refuted.
 
@@
 example, the phrases ``John ran'' and ``John is running'' appear to be
-logically equivalent, for our purposes\todo{For the FraCaS suite?}.
+logically equivalent, for entailment purposes.
 

### 2020-06-09 paper2020/paper.tex
blip
fix a bunch of citations

@@
 semantics of tense and aspect has been also considered in the study of
-Natural Language Inference (NLI)\todo{citation?}. The various datasets
+Natural Language Inference (NLI). The various datasets
 for NLI that have been proposed by the years contain examples that
@@
 and ellipsis sections, with the exception of the system presented by
-\citet{bernardy_type_2017,bernardy_wide_2019}, which offers a complete whole run of the
+\citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019}, which offers a complete whole run of the
 FraCaS test suite, excluding the temporal section. In this
 paper, we take up the challenge of providing a computationally viable
-account of tense and aspect to deal with the notorious section 7\todo{or 7? 8 was not discussed previously} of
+account of tense and aspect to deal with the notorious section 7 of
 the FraCaS test suite. Our account is not meant to be a theoretically
@@
 \paragraph{GF}
-The grammatical framework (GF, \citet{Ranta:GF}) is a powerful parser
+The grammatical framework (GF, \citet{ranta_grammatical_2004}) is a powerful parser
 generator for natural languages, based on type-theoretical abstract grammars.
@@
 \ifanon (citations anonymized) \else
-\citet{bernardy_type_2017} \fi, but the particular
+\citet{bernardy_type-theoretical_2017} \fi, but the particular
 treatment of such phenomena is essentially independent from our
@@
 monadic-based dynamic semantics
-\cite{unger_dynamic_2011,charlow_monadic_2015,charlow_monadic_2017}.
+\cite{unger_dynamic_2011,charlow_monadic_2015,charlow_modular_2017}.
 Our particular implementation is not perfect \ifanon (citations
-anonymized) \else \citet{bernardy_wide_2019} \fi but non-temporal
+anonymized) \else \citet{bernardy_wide-coverage_2019} \fi but non-temporal
 anaphoroi in the testsuite are generally resolved as they should be.
@@
 \section{Our Treatment of time}
+\todo{JP: add a completely worked out example}
 \label{sec:our-semantics}
@@
 \newcommand\nospan{\ensuremath{-}}
-\newcommand\anyspan{\ensuremath{\_}}\todo{better notation} The timespan in the context is
+\newcommand\anyspan{\ensuremath{\mathbf{?}}}\todo{better notation} The timespan in the context is
 optional because, in certain situations, the semantics is different
@@
 
-\[\constant{unicity}_P : P (t_1,t_2) → P (t_3,t_4) → (t_1 = t_3) ∧ (t_2 = t_4)\]
+\begin{multline*}
+\constant{unicity}_P : P (t_1,t_2) → P (t_3,t_4) → \\ (t_1 = t_3) ∧ (t_2 = t_4)
+\end{multline*}
 
@@
     shown below the accuracy figure, in smaller font.  ``FraCoq''
-    refers to the work of \citet{bernardy:2017}. ``MINE" refers
+    refers to the work of \citet{bernardy_type-theoretical_2017}. ``MINE" refers
     to the approach of \citet{Mineshima:2015}, ``NUT" to the CCG
@@
 already occurs without temporal analysis: one must make a difference
-between narrow scoping and wide scoping \cite{todo}. Unfortunately,
+between narrow scoping and wide scoping \cite{todo}\todo{Stergios?}. Unfortunately,
 the montagovian system over which we base our implementation does not

### 2020-06-09 paper2020/paper.tex
fix more citations

@@
 scope of \textbf{P1}. This extension can work according to
-\citet{unger:2011}, and our implemented analysis of anaphora\ifanon (citation anonymized) \else\citep{bernardy_computational_2018,bernardy_wide_2019}\fi.
+\citet{unger_dynamic_2011}, and our implemented analysis of anaphora\ifanon (citation anonymized) \else\citep{bernardy_computational_2018,bernardy_wide_2019}\fi.
 
@@
     \textit{nutcracker} \cite{bos:2008}, and ``Langpro"
-    to the system presented by \citet{Abzianidze:2015}. A dash
+    to the system presented by \citet{abzianidze_tableau_2015}. A dash
     indicates that no attempt was made for the section. }
@@
 
-\bibliography{../PaperTools/bibtex/jp,paper2020}
+\bibliography{../PaperTools/bibtex/jp,paper2020,../paper2018/NLI.bib}
 

### 2020-06-09 paper2020/paper.tex
wibble
import example

@@
 \section{Our Treatment of time}
-\todo{JP: add a completely worked out example}
 \label{sec:our-semantics}
@@
 \[\begin{array}{l}
-∃x. novel(x) ∧ \\
+∀x. novel(x) ∧ \\
 ∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2) ∧ \\
@@
 Note here that the scope for the existential is extended beyond the
-scope of \textbf{P1}. This extension can work according to
-\citet{unger_dynamic_2011}, and our implemented analysis of anaphora\ifanon (citation anonymized) \else\citep{bernardy_computational_2018,bernardy_wide_2019}\fi.
+scope of \textbf{P1}, and its polarity switched (to universal). This extension can work according to
+\citet{unger_dynamic_2011}, and our implemented analysis of
+anaphora\ifanon (citation anonymized)
+\else\citep{bernardy_computational_2018,bernardy_wide-coverage_2019}\fi.
 
@@
 \[\begin{array}{l}
-∃x. novel(x) ∧ \\
-∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2) ∧ \\
-∃y. novel(y) ∧ \\
-∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,y,t_3,t_4) \\
+(∃x. novel(x) ∧ \\
+∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2)) ∧ \\
+(∃y. novel(y) ∧ \\
+∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,y,t_3,t_4)) \\
 ⟶ ⊥ \\
@@
 logically equivalent, for entailment purposes.
+
+\section{Worked out example}
+To give a sense of the additional details necessary to deal with the
+precision demanded by a proof-assistant such as Coq we show how
+problem (279) is worked out fully.
+
+Input from \citet{Ljunglof:2012}:
+
+\begin{verbatim}
+s_279_1_p :: Phr
+s_279_1_p=((sentence) ((useCl) ((past)) ((pPos)) ((predVP) ((usePN) ((lexemePN "smith_PN"))) ((advVP) ((complSlash) ((slashV2a) ((lexemeV2 "write_V2"))) ((detCN) ((detQuant) ((indefArt)) ((numSg))) ((useN) ((lexemeN "novel_N"))))) ((lexemeAdv "in_1991_Adv"))))))
+s_279_3_h :: Phr
+s_279_3_h=((sentence) ((useCl) ((past)) ((pPos)) ((predVP) ((usePN) ((lexemePN "smith_PN"))) ((advVP) ((complSlash) ((slashV2a) ((lexemeV2 "write_V2"))) ((usePron) ((it_Pron)))) ((lexemeAdv "in_1992_Adv"))))))
+\end{verbatim}
+Interpretation by our dynamic and temporal semantics:
+
+\begin{verbatim}
+Definition Problem279aFalse:= (FORALL (fun a=>novel_N a) (fun a=>(exists (b: Time), ((exists (c: Time), (IS_INTERVAL Date_19910101 b /\ IS_INTERVAL c Date_19911231 /\ IS_INTERVAL b c /\ appTime b c (write_V2 a) (PN2object smith_PN))))) -> Not (exists (f: Time), ((exists (g: Time), (IS_INTERVAL Date_19920101 f /\ IS_INTERVAL g Date_19921231 /\ IS_INTERVAL f g /\ appTime f g (write_V2 a) (PN2object smith_PN))))))).
+\end{verbatim}
+Unfolding definitions by Coq:
+
+\begin{verbatim}
+  forall x : object,
+  novel_N x ->
+  (exists b c : Z,
+     Date_19910101 <= b /\ c <= Date_19911231 /\ b <= c /\ write_V2 x SMITH b c) ->
+  (exists f g : Z,
+     Date_19920101 <= f /\ g <= Date_19921231 /\ f <= g /\ write_V2 x SMITH f g) ->
+  False
+\end{verbatim}
+Proof:
+
+\begin{verbatim}
+Theorem  problem279 : Problem279aFalse.
+cbv.
+intros novel isSmithsNovel P1 H.
+destruct P1 as [t0 [t1 [ct1 [ct2 [ct3 P1]]]]].
+destruct H as [u0 [u1 [cu1 [cu2 [cu3 H]]]]].
+specialize writeUnique with (x := novel)(y := SMITH) as A.
+unfold UniqueActivity in A.
+specialize (A _ _ _ _ P1 H) as B.
+lia.
+Qed.
+\end{verbatim}
 

### 2020-06-09 paper2020/paper.tex
some cleanup in example

@@
 
+{\small
 \begin{verbatim}
-s_279_1_p :: Phr
-s_279_1_p=((sentence) ((useCl) ((past)) ((pPos)) ((predVP) ((usePN) ((lexemePN "smith_PN"))) ((advVP) ((complSlash) ((slashV2a) ((lexemeV2 "write_V2"))) ((detCN) ((detQuant) ((indefArt)) ((numSg))) ((useN) ((lexemeN "novel_N"))))) ((lexemeAdv "in_1991_Adv"))))))
-s_279_3_h :: Phr
-s_279_3_h=((sentence) ((useCl) ((past)) ((pPos)) ((predVP) ((usePN) ((lexemePN "smith_PN"))) ((advVP) ((complSlash) ((slashV2a) ((lexemeV2 "write_V2"))) ((usePron) ((it_Pron)))) ((lexemeAdv "in_1992_Adv"))))))
+s_279_1_p=
+ sentence (useCl past pPos
+  (predVP (usePN (lexemePN "smith_PN"))
+          (advVP (complSlash (slashV2a (lexemeV2 "write_V2")))
+                     (detCN (detQuant indefArt numSg) (useN (lexemeN "novel_N"))))
+                     ((lexemeAdv "in_1991_Adv")))))
+s_279_3_h=sentence (useCl past pPos (predVP (usePN (lexemePN "smith_PN")) (advVP (complSlash (slashV2a (lexemeV2 "write_V2")) (usePron it_Pron)) ((lexemeAdv "in_1992_Adv")))))
 \end{verbatim}
@@
 \begin{verbatim}
-Definition Problem279aFalse:= (FORALL (fun a=>novel_N a) (fun a=>(exists (b: Time), ((exists (c: Time), (IS_INTERVAL Date_19910101 b /\ IS_INTERVAL c Date_19911231 /\ IS_INTERVAL b c /\ appTime b c (write_V2 a) (PN2object smith_PN))))) -> Not (exists (f: Time), ((exists (g: Time), (IS_INTERVAL Date_19920101 f /\ IS_INTERVAL g Date_19921231 /\ IS_INTERVAL f g /\ appTime f g (write_V2 a) (PN2object smith_PN))))))).
+FORALL (fun a=>novel_N a)
+(fun a=>(exists (b: Time),
+((exists (c: Time),
+(IS_INTERVAL Date_19910101 b /\
+ IS_INTERVAL c Date_19911231 /\
+ IS_INTERVAL b c /\
+ appTime b c (write_V2 a) (PN2object smith_PN))))) ->
+Not (exists (f: Time),
+((exists (g: Time),
+ (IS_INTERVAL Date_19920101 f /\
+  IS_INTERVAL g Date_19921231 /\
+  IS_INTERVAL f g /\
+  appTime f g (write_V2 a) (PN2object smith_PN)))))).
 \end{verbatim}
@@
 (exists b c : Z,
-     Date_19910101 <= b /\ c <= Date_19911231 /\ b <= c /\ write_V2 x SMITH b c) ->
+   Date_19910101 <= b /\
+   c <= Date_19911231 /\
+   b <= c /\ write_V2 x SMITH b c) ->
 (exists f g : Z,
-     Date_19920101 <= f /\ g <= Date_19921231 /\ f <= g /\ write_V2 x SMITH f g) ->
+   Date_19920101 <= f /\
+   g <= Date_19921231 /\
+   f <= g /\ write_V2 x SMITH f g) ->
 False
@@
 \end{verbatim}
+}
 

### 2020-06-09 paper2020/paper.tex
example

@@
 
-\[\begin{array}{l}
+\begin{equation}
+  \label{eq:smith-novel-1}
+\begin{array}{l}
 ∀x. novel(x) ∧ \\
@@
 \end{array}
-\]
+\end{equation}
 Note here that the scope for the existential is extended beyond the
@@
 s_279_1_p=
- sentence (useCl past pPos
-  (predVP (usePN (lexemePN "smith_PN"))
-          (advVP (complSlash (slashV2a (lexemeV2 "write_V2")))
-                     (detCN (detQuant indefArt numSg) (useN (lexemeN "novel_N"))))
-                     ((lexemeAdv "in_1991_Adv")))))
-s_279_3_h=sentence (useCl past pPos (predVP (usePN (lexemePN "smith_PN")) (advVP (complSlash (slashV2a (lexemeV2 "write_V2")) (usePron it_Pron)) ((lexemeAdv "in_1992_Adv")))))
+sentence 
+(useCl past pPos
+ (predVP
+   (usePN (lexemePN "smith_PN"))
+   (advVP
+     (complSlash
+      (slashV2a (lexemeV2 "write_V2")))
+     (detCN (detQuant indefArt numSg)
+       (useN (lexemeN "novel_N"))))
+       (lexemeAdv "in_1991_Adv")))
+s_279_3_h=
+sentence
+(useCl past pPos
+ (predVP (usePN
+   (lexemePN "smith_PN"))
+   (advVP
+     (complSlash
+      (slashV2a (lexemeV2 "write_V2"))
+      (usePron it_Pron))
+     (lexemeAdv "in_1992_Adv"))))
 \end{verbatim}
@@
  IS_INTERVAL b c /\
- appTime b c (write_V2 a) (PN2object smith_PN))))) ->
+ appTime b c (write_V2 a)
+   (PN2object smith_PN))))) ->
 Not (exists (f: Time),
@@
   IS_INTERVAL f g /\
-  appTime f g (write_V2 a) (PN2object smith_PN)))))).
+  appTime f g (write_V2 a) 
+    (PN2object smith_PN)))))).
 \end{verbatim}
-Unfolding definitions by Coq:
 
+Unfolding definitions at the Coq level:
 \begin{verbatim}
@@
 \end{verbatim}
+(Very close to \cref{eq:smith-novel-1})
+
 Proof:
-
 \begin{verbatim}
@@
 intros novel isSmithsNovel P1 H.
-destruct P1 as [t0 [t1 [ct1 [ct2 [ct3 P1]]]]].
-destruct H as [u0 [u1 [cu1 [cu2 [cu3 H]]]]].
-specialize writeUnique with (x := novel)(y := SMITH) as A.
+destruct P1 as
+   [t0 [t1 [ct1 [ct2 [ct3 P1]]]]].
+destruct H as
+   [u0 [u1 [cu1 [cu2 [cu3 H]]]]].
+specialize writeUnique
+  with (x := novel)(y := SMITH) as A.
 unfold UniqueActivity in A.
@@
 }
+
+Critical: UniqueActivity; lia does arithmetic work.
 

### 2020-06-10 paper2020/paper.tex
example: commentary

@@
 precision demanded by a proof-assistant such as Coq we show how
-problem (279) is worked out fully.
+problem (279) is worked out in full details.
 
-Input from \citet{Ljunglof:2012}:
+We start with input trees in GF format, given by
+\citet{Ljunglof:2012}. They can be rendered as follows:
 
@@
 \end{verbatim}
-Interpretation by our dynamic and temporal semantics:
 
+  Of particular note is the use of the pronoun ``it'', and the fact
+  that adverbials expressions such that ``in 1992'' are lexicalized.
+  We also follow the GF convention to postfix lexical items with the
+  name of their category.
+
+Our dynamic and temporal semantics gives the following interpretation
+for \verb!s_279_1_p! implies \verb!s_279_3_h!.
 \begin{verbatim}
@@
 \end{verbatim}
+In the above, one should remark the top-level quantification over the
+novel (as explained in \cref{sec:our-semantics}), the quantification
+over time intervals as individual timepoints, and the use of custom
+operators for several constructions (FORALL, Not, IS_INTERVAL,
+appTime). This use of custom operators is useful for several
+generalisations (for example, we have quantifiers such as MOST in
+addition to FORALL.)
 
-Unfolding definitions at the Coq level:
+Unfolding the definitions for these operators yield the following
+proposition:
 \begin{verbatim}
@@
 \end{verbatim}
-(Very close to \cref{eq:smith-novel-1})
+This is very close to our idealised representation of the problem
+\cref{eq:smith-novel-1}. One difference is the use of abstract Coq
+integers for timepoints. Using a discrete time allows us to use
+predifined Coq tactics. The discrete nature of integers does not
+interfere with the reasoning.
 
-Proof:
+Finally, we can show a Coq proof for the above proposition:
 \begin{verbatim}
@@
 
-Critical: UniqueActivity; lia does arithmetic work.
+The intros and destruct tactics serve bookkeeping purposes. The
+critical part is the use of the \texttt{writeUnique} axiom, which
+witnesses the aspectual class of the predicate \texttt{write\_V2}.
+The proof is completed by the use of the \texttt{lia} tactic, which is
+embeds a decision procedure for linear arithmetic
+problems. Fortunately, \texttt{lia} can take care of all the problems
+which arise in the FraCaS testsuite.
 

### 2020-06-11 paper2020/paper.tex
fix compilation

@@
 over time intervals as individual timepoints, and the use of custom
-operators for several constructions (FORALL, Not, IS_INTERVAL,
-appTime). This use of custom operators is useful for several
-generalisations (for example, we have quantifiers such as MOST in
-addition to FORALL.)
+operators for several constructions (\verb!FORALL!, \verb!Not!, \verb!IS_INTERVAL!,
+\verb!appTime!). This use of custom operators is useful for several
+generalisations (for example, we have quantifiers such as \verb!MOST! in
+addition to \verb!FORALL!.)
 
@@
 
-\bibliography{../PaperTools/bibtex/jp,paper2020,../paper2018/NLI.bib}
+\bibliography{../PaperTools/bibtex/jp,paper2020,../paper2018/NLI}
 

### 2020-06-22 paper2020/paper.tex
minor

@@
 
-\paragraph{Montagovian semantics}
+\paragraph{Montagovian Semantics}
 
@@
 
-\paragraph{Dynamic semantics}
+\paragraph{Dynamic Semantics}
 We additionally assume support for (non-temporal) anaphora, using a
@@
 
-\section{Our Treatment of time}
+\section{Our Treatment of Time}
 \label{sec:our-semantics}
@@
 
-\paragraph{temporal context}
+\paragraph{Temporal Context}
 
@@
 
-\paragraph{tenses}
+\paragraph{Tenses}
 
@@
 
-\paragraph{temporal adverbs}
+\paragraph{Temporal Adverbs}
 
@@
 
-\paragraph{(un)repeatable achievements}
+\paragraph{(Un)repeatable Achievements}
 \label{sec:repeatable-achievements}
@@
 
-\paragraph{action-modification verbs}
+\paragraph{Action-modification Verbs}
 
@@
 
-\paragraph{Progressive aspect}
+\paragraph{Progressive Aspect}
 
@@
 
-\section{Results and evaluation}
+\section{Results and Evaluation}
 \label{sec:results}
@@
 We stress that such interaction between quantifiers and negation
-already occurs without temporal analysis: one must make a difference
-between narrow scoping and wide scoping \cite{todo}\todo{Stergios?}. Unfortunately,
-the montagovian system over which we base our implementation does not
-support this phenomena. (In fact the FraCaS test suite does not appear to
+already occurs without any temporal analysis: one must make a difference
+between narrow  and wide scope.\footnote{The literature on wide and narrow scope readings of quantifiers and quantifiers and negation is vast.
+We direct the reader to \cite{steedman:2012} for a comprehensive and thorough discussion of the relevant issues.} Unfortunately,
+the Montagovian system over which we base our implementation does not
+support these phenomena. (In fact the FraCaS test suite does not appear to
 test if the difference between narrow and wide scoping is handled properly.)

### 2020-06-22 paper2020/paper.tex
conclusions

@@
 
-Another issue to be looking at is the codification of lexical
-aspect.\todo{update} In what we have been doing so far, we did not distinguish
-between the present continuous and the present or the simple past and
-the past continuous and their interaction with non-lexical aspect.
-The main reason for this is that in the test suite, even though there
-are examples of inferences based on the either lexical or non-lexical
-aspect, no example where an interaction between the two exists. For
-example, one such case is the one shown below and exemplifies the
-difference between activities and accomplishments with respect to the
-different inferences they give when used in the past continuous:
+Another thing we would be willing to try is an extension of the FraCaS test suite that will include more fine-grained examples of lexical and grammatical aspect and  the interaction between those two, for example cases where one needs to actually distinguish between achievements and accomplishments on the basis of their inferential properties:
 
 \fracasex{extra1}{
-	\item	John was writing a book. 
-	\hyp 	John wrote a book (UNK).
+	\item	John found his keys.  
+	\hyp 	John was finding his keys  (UNK).
 }
@@
 \fracasex{extra2}{
-	\item	John was running. 
-	\hyp 	John ran (YES).
-}
-
-\fracasex{extra1}{
 	\item	John wrote a book.  
@@
 
-\fracasex{extra2}{
-	\item	John ran.
-	\hyp 	John was running (YES?).
+In the first of the two examples involving an achievement verb, the inference is UNK, since there is no guarantee that the action is non-instanteous.
+To the contrary, for accomplishment verbs, the inference follows.
+
+Further cases to be included in an extended FraCaS future suite involve examples where the interaction between different tenses needs to be captured:
+%Another issue to be looking at is the codification of lexical
+%aspect.\todo{update} In what we have been doing so far, we did not distinguish
+%between the present continuous and the present or the simple past and
+%the past continuous and their interaction with non-lexical aspect.
+%The main reason for this is that in the test suite, even though there
+%are examples of inferences based on the either lexical or non-lexical
+%aspect, no example where an interaction between the two exists. For
+%example, one such case is the one shown below and exemplifies the
+%difference between activities and accomplishments with respect to the
+%different inferences they give when used in the past continuous:
+
+\fracasex{extra3}{
+	\item	When the phone rang, John had entered the house.  
+	\hyp 	John entered the house before the phone rang (YES).
 }

### 2020-06-23 paper2020/paper.tex
revise intro

@@
 and a number of different ideas have been put forth on the table to
-deal with them throughout the
-years~\citep{dowty:2012,prior:2003,steedman:2000,higginbotham:2009,fernando:2015}. The
-semantics of tense and aspect has been also considered in the study of
+deal with them throughout the years. Recent proposals include the
+works of the following authors:
+\citet{dowty:2012,prior:2003,steedman:2000,higginbotham:2009,fernando:2015}.
+%
+The
+semantics of tense and aspect have been also considered in the study of
 Natural Language Inference (NLI). The various datasets
@@
 the FraCaS test suite~\citep{cooper:1996} contains a whole section
-dedicated to temporal and aspectual inference (section 7). This part
+dedicated to temporal and aspectual inference (section 7 of the dataset). This part
 of the FraCaS test suite has been notoriously difficult to tackle. In
-fact, all the computational systems ran on the FraCaS test suite so
-far skip this section. Additionally, they also skip the anaphora
-and ellipsis sections, with the exception of the system presented by
-\citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019}, which offers a complete whole run of the
-FraCaS test suite, excluding the temporal section. In this
-paper, we take up the challenge of providing a computationally viable
-account of tense and aspect to deal with the notorious section 7 of
-the FraCaS test suite. Our account is not meant to be a theoretically
-full account of tense and aspect, but rather an account that is driven by the
-need to cover the test suite in a way that is general enough to
-capture the test suite examples. The account is evaluated on the
-entailment properties of various temporal and aspectual examples, as
-given by the test suite. As such, we are not getting into the
-discussion of how tense and aspect might affect grammaticality or
-infelicitousness of various sentences. We take the sentences of the
-FraCaS suite to be syntactically correct and the entailment patterns they
-are annotated with valid.
+fact, so far, no computational system has been capable to deal with it: when
+authors report accuracy over the FraCaS test suite they
+skip this section.
+%
+In fact, they also often skip the anaphora and ellipsis sections, the
+exception being the system presented by
+\citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019},
+which include support for anaphora and ellipsis but still omit the
+temporal section.
+%
+In this paper, we take up the challenge of providing a
+computationally viable account of tense and aspect to deal with the
+notoriously difficult section 7 of the FraCaS test suite. Our account is not meant
+to be a theoretically extensive account of tense and aspect, but rather an
+account that is driven by the need to cover the test suite in a way
+that is general enough to capture the test suite examples, \emph{while still covering the rest of the FraCaS test suite}.
+%
+The account is evaluated on the entailment properties of various
+temporal and aspectual examples, as given by the test suite. As such,
+we are not getting into the discussion of how tense and aspect might
+affect grammaticality or infelicitousness of various sentences.
+We assume that the sentences of the FraCaS suite are syntactically and
+semantically correct, and strive to produce accurate logical
+representations given that assumption. We further assume that the
+entailment annotations of various problems are valid, and we use those
+to evaluate the correctness of the logical representations of
+sentences.
 
 The paper is structured as follows: in \cref{sec:background}, we give
-a brief summary of the frameworks whose various subsystems rely on. In
-particular, the Grammatical Framework is used to construct the
-syntactic parser; the Coq proof assistant checks all the reasoning; a
-monad-based dynamic semantics deals with references (anaphora).  We
-also provide some brief remarks on temporal semantics. In \cref{sec:our-semantics},
-we discuss the main aspects of the compositional semantics of our, using various examples
-from the FraCaS suite in exemplifying its effectiveness. In \cref{sec:results},
-we take a brief look at how our system performs with respect to the
-FraCaS suite. We ran the system across the whole suite, this being the
-first time that a complete run through the whole suite has been made,
-but we look in particular at the performance on the temporal
-section. In \cref{sec:conclusions}, we conclude and discuss avenues for future
+a brief summary of the computational frameworks whose various
+subsystems rely on. In particular, the Grammatical Framework is used
+to construct the syntactic parser, the Coq proof assistant checks all
+the reasoning and a monad-based dynamic semantics deals with
+Montague-style semantics, and references (anaphora).  We also provide
+some brief remarks on temporal semantics. In \cref{sec:our-semantics},
+we discuss the main aspects of the compositional semantics of our
+system, using various examples from the FraCaS suite to illustrate its
+effectiveness. In \cref{sec:results}, we evaluate how our system
+performs with respect to the FraCaS suite. We ran the system across
+the whole suite: our system is thus the first which is capable of
+handling the complete FraCaS test suite.  Yet, we are interested in
+particular in the performance on the temporal section. In
+\cref{sec:conclusions}, we conclude and discuss avenues for future
 work.
-
-%We do not care about classifying sentences as felicitous/infelicitous
-%or grammatical/ungrammatical, rather we assume that the sentence is
-%correct and care about giving it an appropriate interpretation as a
-%logical formula.  Indeed, our evaluation uses entailment properties
-%only.
 

### 2020-06-23 paper2020/paper.tex
revisions: background

@@
 generator for natural languages, based on type-theoretical abstract grammars.
-This work relies a syntactic representation of the FraCaS test suite in GF
-abstract syntax~\citep{Ljunglof:2012}. Thus, we skip the parsing phase
+The present work leverages a syntactic representation of the FraCaS test suite in GF
+abstract syntax~\citep{Ljunglof:2012}. Thanks to this, we skip the parsing phase
 and avoid any syntactic ambiguity.
 
-For the purpose of this paper, the important point is that GF the
+For the purpose of this paper, the important point is that the GF
 syntax aims at a balance of sufficient abstraction to provide a
@@
 sufficiently many syntactic features to be able to reconstruct
-unambiguously natural-language text. \cref{ex:nobel}
-shows an example abstract syntax tree and its realisation in English.
+natural-language text. That is, the parse trees generally satisfy the
+homomorphism requirement of \citet{Montague:EFL70,Montague1974}, and
+we can focus on the translation of syntactic trees to logical forms.
+\cref{ex:nobel} shows an example abstract syntax tree and its
+realisation in English.
 
@@
 do not use any coinduction (or even induction) in this paper, relying
-on the pure lambda-calculus inner core of Coq.  Coq is a very
-powerful reasoning engine that makes it fit for implementing natural
-language semantics in it but also, on a more practical side, for the
-the task of NLI, when the latter is formalized as a theorem proving
-task. Coq supports, as already mentioned, dependent typing, but also
-subtyping, which are instrumental in expressing NL semantics.
+on the pure lambda-calculus inner core of Coq.  Coq is a very powerful
+reasoning engine that makes it fit for implementing natural language
+semantics.  Coq supports, as already mentioned, dependent typing, but
+also subtyping. Both concepts are instrumental in expressing NL
+semantics\citet{TODO: chatzikyriakidis and luo}.  Besides, on a more
+practical side, it works well for the the task of NLI, when the latter
+is formalized as a theorem proving task: its many tactics mean that
+many tasks in theorem proving are trivialized. In particular, all
+problems of time-intervals inclusion, which occur in every temporal
+problems, are solved with Coq's linear arithmetic tactic.
 
@@
 We additionally assume support for (non-temporal) anaphora, using a
-monadic-based dynamic semantics
-\cite{unger_dynamic_2011,charlow_monadic_2015,charlow_modular_2017}.
-Our particular implementation is not perfect \ifanon (citations
-anonymized) \else \citet{bernardy_wide-coverage_2019} \fi but non-temporal
-anaphoroi in the testsuite are generally resolved as they should be.
+monadic-based dynamic semantics, generally following the state of the
+art in this matter
+\citep{unger_dynamic_2011,charlow_monadic_2015,charlow_modular_2017}.
+Our particular implementation does not handle every case \ifanon
+(citations anonymized) \else \citet{bernardy_wide-coverage_2019} \fi
+but non-temporal anaphoroi in the testsuite are generally resolved as
+they should be: on the whole accuracy is not affected significantly by
+issues in this subsystem.
 
-As for other basic phenomena, there is not much interaction between
-our treatment of time and non-temporal anaphora. One critical
-exception is discussed in \cref{sec:repeatable-achievements}.
+As it is the case for other basic phenomena, there is not much
+interaction between our treatment of time and non-temporal
+anaphora. One critical exception is discussed in
+\cref{sec:repeatable-achievements}.
 
@@
 even though our definitions do not perfectly match the classical
-ones. We explain our precise meaning forthese terms in the body of the
+ones. We explain our precise meaning for these terms in the body of the
 paper. Nevertheless, we refer the reader to \citet{steedman:2000} for
-an extensive review of temporal semantics.
+an extensive review of formal temporal semantics.
 
@@
 concerned, we find that these three categories can be collapsed into a
-single one.  As such, we think it is a good classification, given that
-they generally afford the correct inferences for the testsuite. (In
-this paper, we always assume that the FraCaS testsuite is correct.)
+single one (they are subject to \cref{eq:unicity}).  As such, we think
+it is a good classification, given that it generally affords the
+correct inferences for the testsuite. (In this paper, we always assume
+that the problems in the FraCaS testsuite are correctly annotated.)
 
@@
   \constant{unicity}_P : P (t_1,t_2) → P (t_3,t_4) → \\ (t_1 = t_3) ∧ (t_2 = t_4)
+  \label{eq:unicity}
 \end{multline*}

### 2020-06-23 paper2020/paper.tex
revise sem. sec

@@
 verbs, so that it takes two
-additional time parameters, one corresponding with the starting time
+additional time parameters, one corresponding to the starting time
 of the action and one corresponding to its stopping time ($(e × time × time) \to t$). For example,
@@
 procedure. For simplicity we will however only consider verbs from now
-on. (In fact in our implementation we chose not to modify nouns nor
-adjectives to limit the increase in complexity of the formulas.)
+on. (In fact, even in our implementation we chose not to extend nouns
+nor adjectives with timespan parameters. This choice limits the
+increase in complexity of the formulas compared to non-temporal
+semantics, at the expense of inaccuracy for a couple of problems in
+the FraCaS test suite: problems 271 and 272 use a an adjective as a
+copula which is subject to temporal reasoning.)
 
@@
 This temporal context is an \emph{optional} timespan. That is, it can
-be an timespan or an explicitly unspecified context.
+be a timespan or an explicitly unspecified context.
 \newcommand\nospan{\ensuremath{-}}
@@
 %
-The principal non-trivial manipulator of timespans are tense
+The principal non-trivial manipulators of timespans are tense
 markers. In our syntax, inherited from GF, tenses are represented
-syntactically as an attribute of the clause. An illustration of a
-past-tense clause and its interpretation follows. Notice in particular
-the $past$ argument to $useCl$.
+syntactically as an attribute of clauses. An illustration of a
+past-tense clause and its interpretation follows in
+\cref{ex:nobel}. Notice in particular the $\varid{past}$ argument to
+the $\varid{useCl}$ constructor.
 
@@
 }}
-In our semantics we deal only with present and past
-tenses (simple and continuous). Indeed we find that FraCas
-does not exercise additional specific tenses (if a more complicated
-tense is used, the additional information is also carried by adverbs
-or adverbial phrases, in a more specific way.) It is not difficult to see how other tenses can be captured under the same general framework.
+%
+In our semantics we deal only with present and past tenses (simple and
+continuous). Indeed we find that FraCas does not exercise additional
+specific tenses. (When a more complicated tense is used, the
+additional information is also carried by adverbs or adverbial
+phrases, in a more specific way). While we believe that many other
+tenses can be captured under the same general framework, we leave a
+detailed study to further work.
 
-Even though we will discuss a refinement to handle the past continuous
+Even though we discuss a refinement to handle the past continuous
 at the end of this section, the procedure to handle tense annotations
@@
 \item[exact] For such adverbs, an exact interval is
-  provided. Typical such adverbs specify in fact a single point in
+  provided. In fact, such adverbs typically specify a single point in
   time (so the start and the end of the interval coincide).
@@
 In the FraCaS test suite, we normally do not find several
-time-modifying adverbs modifying a single verb. Indeed, sentences such
+time-modifying adverbs modifying a single verb phrase. Indeed, sentences such
 as ``in 1992, in 1991 john wrote a novel'' are infelicitous. This
 justifies ignoring the input timespan in the above interpretation --
-we're in particular not interested in modeling felicity with our
+we are in particular not interested in modeling felicity with our
 semantics, only giving an accurate semantics when the input is
@@
 same time intervals.
-
+%
 For this purpose we postulate \emph{unicity of action} for certain time-dependent
@@
 
-\begin{multline*}
+\begin{multline}
   \constant{unicity}_P : P (t_1,t_2) → P (t_3,t_4) → \\ (t_1 = t_3) ∧ (t_2 = t_4)
   \label{eq:unicity}
-\end{multline*}
+\end{multline}
 
@@
 
-(The difference between actitivty and accomplishments on the one and
+(The difference between actitivty and accomplishments on the one hand and
 achievement on the other hand is that for the latter, time intervals
@@
 oversimplification as achievements are usually of short duration, but
-not nil. However, this appears to play little role, if any, in our
-analysis, and given the its task-driven nature as well.)
+not nil. However, this plays little role in our
+analysis. As far as we can tell the FraCaS test suite does exercise temporal semantics to
+such a level of precision.)
 
@@
 flagged. We leave such discourse analysis as future work, and simply
-apply unicity of action everywhere.
+apply unicity of action everywhere: it is valid uniformly in the
+FraCaS test suite for activity aspect classes.
 
@@
 Indeed, from \textbf{P3} we get that Smith was in paris between May
-5th and May 15th. Because "being in Paris" is stative, we also get that
+5th and May 15th. Because ``being in Paris'' is stative, we also get that
 Smith was in Paris in any subinterval. Contrary to unicity of action,
 subsumption is always valid.
+
+\paragraph{Class-modifying adverbs}
+It should be noted that some adverbs can locally disable the
+application of $\constant {subsumption}$. For example, problem 299
+features the sentence ``Smith lived in Birmingham for exactly a
+year''.  Even though ``live'' is normally stative, one can no longer
+apply subsumption in the context of ``exactly a year'' --- this can be
+done by propagating another context flag in the montegovian semantics
+(in addition to the temporal context).
 
@@
 
-We treat verbs in the progressive form as different semantically. For
-example, ``John was writing a book'' is encoded as
+We treat verbs in the progressive form as different semantically from
+the non-progressive form. For example, ``John was writing a book'' is
+encoded as
 $∃(t_1, t_2). t_1≤ t_2, t_2 ≤ now, PROG\_write(John,book,t_1,t_2)$,
@@
 necessary, because $PROG\_write$ is subject to
-$\constant{subsumption}$: if John is writing in interval $[t_1,t_2]$
-then he is writing in any subinterval. However, the same cannot be
-said of $write$: the end-points of the interval indicates the time
-needed to complete the achievement. (For example, ``John wrote a book
-in 1993'' neither entails ``John wrote a book in January 1993'' nor
-``John wrote a book in December 1993''.) That is, $write$ is on the
-contrary subject to $\constant{unicity}$. Finally, we also have
-$write(x,y,t_1,t_2) → PROG\_write(x,y,t_1,t_2)$. That is, the
-achievement variant implies the stative variant, in the same interval.
-Consequently we get the entailment from ``John wrote a book'' to
-``John was writing a book''.
+$\constant{subsumption}$: if John is writing in the interval
+$[t_1,t_2]$ then he is writing in any subinterval of
+$[t_1,t_2]$. However, the same cannot be said of $write$: the
+end-points of the interval indicate the time needed to complete the
+achievement. (For example, ``John wrote a book in 1993'' neither
+entails ``John wrote a book in January 1993'' nor ``John wrote a book
+in December 1993''.) In fact, \emph{write}, in the non-progressive
+from, is on the contrary subject to $\constant{unicity}$. Finally, we
+also have $write(x,y,t_1,t_2) → PROG\_write(x,y,t_1,t_2)$. That is,
+the achievement (or \emph{activity} in our terminology) variant implies the stative variant, in the
+same interval.  Consequently we get the entailment from ``John wrote a
+book'' to ``John was writing a book''.
 
@@
 \label{sec:conclusion}
+\todo{summary of advances}
+
 We have presented a first attempt for a computational approach dealing

### 2020-06-23 paper2020/paper.tex
typography

@@
 
-{\small
+\begin{small}
 \begin{verbatim}
@@
 \end{verbatim}
+\end{small}
 
@@
 for \verb!s_279_1_p! implies \verb!s_279_3_h!.
+{\small
 \begin{verbatim}
@@
 \end{verbatim}
+}
 In the above, one should remark the top-level quantification over the
@@
 proposition:
+{\small
 \begin{verbatim}
@@
 \end{verbatim}
-This is very close to our idealised representation of the problem
+}This is very close to our idealised representation of the problem
 \cref{eq:smith-novel-1}. One difference is the use of abstract Coq
@@
 Finally, we can show a Coq proof for the above proposition:
+{\small
 \begin{verbatim}
@@
 Our target is the FraCaS testsuite, which aims at covering a wide
-range of common natural-language phenomena. The testsuite is structured in 9 sections:
-
+range of common natural-language phenomena. 
+%
 The suite is structured according to the semantic phenomena involved

### 2020-06-23 paper2020/paper.tex
revise conclusion

@@
 \paragraph{interaction with negation and group readings}
-One reason explaining the lower performance of our systems on some
+One reason explaining the lower performance of our system on some
 sections of the testsuite is that our interpretation of time interacts
@@
 A similar issue occurs with group readings, which are interpreted as
-conjunctions. Here, the problem is to solve the order of scoping
-between this conjunction and time-interval quantifiers.
+conjunctions. Here, the problem is an improper resolution of scoping
+order, between this conjunction and time-interval quantifiers.
 
@@
 \label{sec:conclusion}
-\todo{summary of advances}
 
 We have presented a first attempt for a computational approach dealing
-with the temporal section of the FraCaS test suite. This is the first
-time a system is ran against this section, at the same time, this
-concludes our work in handling the whole of the FraCaS test suite
-using a combination of a number of well-studied tools: type theory,
-parsing using the Grammatical Framework (GF), Monadic Dynamic
+with the temporal section of the FraCaS test suite. To do this, we
+have provided a new, simplified taxonomy of aspectual classes for verb
+phrases, guided by the applicability of the unicity of action and
+temporal subsumption properties.
+
+This paper concludes our work in handling the whole of the FraCaS test
+suite using a combination of a number of well-studied tools: type
+theory, parsing using the Grammatical Framework (GF), Monadic Dynamic
 Semantics and proof assistant technology (Coq). It achieves an
@@
 things to be looked at is fixing the issues associated with parts of
-the test suite that ``broke'' so to say when the temporal analysis was
+the test suite that ``broke'' when the temporal analysis was
 introduced. Some of these have been already mentioned: interaction of
@@
 
-Another thing we would be willing to try is an extension of the FraCaS test suite that will include more fine-grained examples of lexical and grammatical aspect and  the interaction between those two, for example cases where one needs to actually distinguish between achievements and accomplishments on the basis of their inferential properties:
+Another thing we are willing to try is an extension of the FraCaS
+test suite that will include more fine-grained examples of lexical and
+grammatical aspect and the interaction between those two, for example
+cases where one needs to actually distinguish between achievements and
+accomplishments on the basis of their inferential properties:
 
@@
 
-In the first of the two examples involving an achievement verb, the inference is UNK, since there is no guarantee that the action is non-instanteous.
-To the contrary, for accomplishment verbs, the inference follows.
+In the first of the two examples involving an achievement verb, the
+inference is UNK, since there is no guarantee that the action is
+non-instanteous. To the contrary, for accomplishment verbs, the
+inference follows.
 
-Further cases to be included in an extended FraCaS future suite involve examples where the interaction between different tenses needs to be captured:
+Further cases to be included in an extended FraCaS future suite
+involve examples where the interaction between different tenses needs
+to be captured:
 %Another issue to be looking at is the codification of lexical

### 2020-06-23 paper2020/paper.tex
fix some citations

@@
 works of the following authors:
-\citet{dowty:2012,prior:2003,steedman:2000,higginbotham:2009,fernando:2015}.
+\citet{dowty:2012,prior:2003,steedman_productions_2000,higginbotham:2009,fernando:2015}.
 %
@@
 natural-language text. That is, the parse trees generally satisfy the
-homomorphism requirement of \citet{Montague:EFL70,Montague1974}, and
+homomorphism requirement of \citet{montague_english_1970,montague_proper_1974}, and
 we can focus on the translation of syntactic trees to logical forms.
@@
 Unicity of action holds only if the aspectual class of the proposition
-$P$ is \emph{actitivty}~\citep{steedman:2000} (which, for our
+$P$ is \emph{actitivty}~\citep{steedman_productions_2000} (which, for our
 purposes, includes \emph{achievements} and \emph{accomplishments}).

### 2020-06-26 paper2020/paper.tex
typography

@@
 we can focus on the translation of syntactic trees to logical forms.
-\cref{ex:nobel} shows an example abstract syntax tree and its
+Example \ref{ex:nobel} shows an example abstract syntax tree and its
 realisation in English.
@@
 \newcommand\nospan{\ensuremath{-}}
-\newcommand\anyspan{\ensuremath{\mathbf{?}}}\todo{better notation} The timespan in the context is
+\newcommand\anyspan{\ensuremath{\mathbf{*}}} The timespan in the context is
 optional because, in certain situations, the semantics is different
@@
 past-tense clause and its interpretation follows in
-\cref{ex:nobel}. Notice in particular the $\varid{past}$ argument to
+Example \ref{ex:nobel}. Notice in particular the $\varid{past}$ argument to
 the $\varid{useCl}$ constructor.

### 2020-06-26 paper2020/paper.tex
spellchecking and abstract

@@
 \maketitle
+
+\begin{abstract}
+  In this paper, we propose an implementation of temporal semantics
+  which is suitable for inference problems. This implementation
+  translates syntax trees to logical formulas, suitable for
+  consumption by the Coq proof assistant. We support several phenomena
+  including: temporal references, temporal adverbs, aspectual classes
+  and progressives.  We apply these semantics to temporal problems of
+  the FraCaS testsuite and obtain an accuracy of 71 percent.
+\end{abstract}
 
@@
 Example \ref{ex:nobel} shows an example abstract syntax tree and its
-realisation in English.
+realization in English.
 
@@
 Coq is an interactive theorem prover (proof assistant). It is based on
-the calculus of coinductive constructions~\citep{werner_une_1994} We
-do not use any coinduction (or even induction) in this paper, relying
+the calculus of co-inductive constructions~\citep{werner_une_1994} We
+do not use any co-induction (or even induction) in this paper, relying
 on the pure lambda-calculus inner core of Coq.  Coq is a very powerful
@@
 property $t_0 ≤ t_1$ (it starts no later than it stops). (We are thus
-using a simple newtonian model of time, corresponding to a layman
+using a simple Newtonian model of time, corresponding to a layman
 intuition of a linear constant flow of time.)
@@
 \item If the temporal context is specified (for example due to the presence of an adverb or an
-  averbial clause, such as ``before James swam''), then the tense does
+  adverbial clause, such as ``before James swam''), then the tense does
   not create a new interval, but it may constrain it. Typically, a
@@
 A common theme in the testsuite is the reference to previous
-occurences of an event:
+occurrences of an event:
 
@@
 To be able to conclude that there is entailment, as the testsuite expects, we have to make sure
-that the two occurences of ``Jones left'' (in \textbf{P1} and \textbf{P2}) refer to the
+that the two occurrences of ``Jones left'' (in \textbf{P1} and \textbf{P2}) refer to the
 same time intervals.
@@
 Unicity of action holds only if the aspectual class of the proposition
-$P$ is \emph{actitivty}~\citep{steedman_productions_2000} (which, for our
+$P$ is \emph{activity}~\citep{steedman_productions_2000} (which, for our
 purposes, includes \emph{achievements} and \emph{accomplishments}).
 
-(The difference between actitivty and accomplishments on the one hand and
+(The difference between activity and accomplishments on the one hand and
 achievement on the other hand is that for the latter, time intervals
@@
 
-Indeed, from \textbf{P3} we get that Smith was in paris between May
+Indeed, from \textbf{P3} we get that Smith was in Paris between May
 5th and May 15th. Because ``being in Paris'' is stative, we also get that
-Smith was in Paris in any subinterval. Contrary to unicity of action,
+Smith was in Paris in any sub-interval. Contrary to unicity of action,
 subsumption is always valid.
@@
 apply subsumption in the context of ``exactly a year'' --- this can be
-done by propagating another context flag in the montegovian semantics
+done by propagating another context flag in the montagovian semantics
 (in addition to the temporal context).
@@
 Indeed, following our account, the above (contradictory) inference problem is to be
-interperted as
+interpreted as
 
@@
 $\constant{subsumption}$: if John is writing in the interval
-$[t_1,t_2]$ then he is writing in any subinterval of
+$[t_1,t_2]$ then he is writing in any sub-interval of
 $[t_1,t_2]$. However, the same cannot be said of $write$: the
@@
 \verb!appTime!). This use of custom operators is useful for several
-generalisations (for example, we have quantifiers such as \verb!MOST! in
+generalizations (for example, we have quantifiers such as \verb!MOST! in
 addition to \verb!FORALL!.)
@@
 \end{verbatim}
-}This is very close to our idealised representation of the problem
+}This is very close to our idealized representation of the problem
 \cref{eq:smith-novel-1}. One difference is the use of abstract Coq
 integers for timepoints. Using a discrete time allows us to use
-predifined Coq tactics. The discrete nature of integers does not
+predefined Coq tactics. The discrete nature of integers does not
 interfere with the reasoning.
@@
 inference is UNK, since there is no guarantee that the action is
-non-instanteous. To the contrary, for accomplishment verbs, the
+non-instantaneous. To the contrary, for accomplishment verbs, the
 inference follows.
@@
 
-\section*{Acknowledgements}
+\section*{Acknowledgments}
 
@@
 \end{document}
+
+% LocalWords:  dowty steedman higginbotham fernando NLI FraCaS GF une
+% LocalWords:  aspectual dataset anaphora bernardy grammaticality NL
+% LocalWords:  infelicitousness ranta Ljunglof montague english TODO
+% LocalWords:  prover chatzikyriakidis Coq's anonymized monadic unger
+% LocalWords:  charlow anaphoroi stative activites timespan timespans
+% LocalWords:  useCl scandinavian nobel FraCas timepoint Itel unicity
+% LocalWords:  Statives contrario th Un lexemes PROG lexicalized lia
+% LocalWords:  postfix timepoints FORALL appTime writeUnique FC CCG
+% LocalWords:  FraCoq Mineshima Langpro abzianidze premiss UNK Centre

### 2020-09-30 paper2020/paper.tex
list todos

@@
 dedicated to temporal and aspectual inference (section 7 of the dataset). This part
-of the FraCaS test suite has been notoriously difficult to tackle. In
-fact, so far, no computational system has been capable to deal with it: when
+of the FraCaS test suite has been difficult to tackle. That is,
+so far, no computational system has been capable to deal with it: when
 authors report accuracy over the FraCaS test suite they

### 2020-10-01 paper2020/paper.tex
repair citation
kill todo and minor fixes

@@
 \usepackage{enumitem}
-\usepackage{cleveref}
+\usepackage[capitalize]{cleveref}
 \newlist{lingex}{enumerate}{3}
@@
 \citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019},
-which include support for anaphora and ellipsis but still omit the
+which includes support for anaphora and ellipsis but still omit the
 temporal section.
@@
 \paragraph{GF}
-The grammatical framework (GF, \citet{ranta_grammatical_2004}) is a powerful parser
-generator for natural languages, based on type-theoretical abstract grammars.
-The present work leverages a syntactic representation of the FraCaS test suite in GF
-abstract syntax~\citep{Ljunglof:2012}. Thanks to this, we skip the parsing phase
-and avoid any syntactic ambiguity.
+The grammatical framework (GF, \citet{ranta_grammatical_2004}) is a
+powerful parser generator for natural languages, based on
+type-theoretical abstract grammars.  The present work leverages a
+syntactic representation of the FraCaS test suite in GF abstract
+syntax~\citep{Ljunglof:2012}. Thanks to this, we skip the parsing
+phase and avoid any syntactic ambiguity.
 
-For the purpose of this paper, the important point is that the GF
-syntax aims at a balance of sufficient abstraction to provide a
+For the purpose of this paper, the important feature of GF
+syntax is that it aims at a balance of sufficient abstraction to provide a
 semantically-relevant structure, but at the same time it embeds
@@
 we can focus on the translation of syntactic trees to logical forms.
+%
+Consequently, the system presented
+here does not aim at textual natural language understanding, but rather
+provides a testable, systematic formal semantics of temporal phenomena.
 Example \ref{ex:nobel} shows an example abstract syntax tree and its
@@
 also subtyping. Both concepts are instrumental in expressing NL
-semantics\citet{TODO: chatzikyriakidis and luo}.  Besides, on a more
+semantics\citet{chatzikyriakidis_natural_2014}.  Besides, on a more
 practical side, it works well for the the task of NLI, when the latter

### 2020-10-01 paper2020/paper.tex
deal with review 1

@@
 
-\section{Background}
+\paragraph{Temporal Semantics}
+\todo{Fold this in some other section}
+In this paper we borrow several notions from classical temporal
+semantics such as ``stative'', ``achievement'', ``activity'', etc.,
+even though our definitions do not perfectly match the classical
+ones. We explain our precise meaning for these terms in the body of the
+paper. Nevertheless, we refer the reader to \citet{steedman:2000} for
+an extensive review of formal temporal semantics.
+
+For the \textit{cognoscenti}, we can already point out some
+differences in terminology: we use the term activity as a general term
+which encompasses the three classical notions of activites,
+achievements and accomplishments. Indeed, insofar as the test suite is
+concerned, we find that these three categories can be collapsed into a
+single one (they are subject to \cref{eq:unicity}).  As such, we think
+it is a good classification, given that it generally affords the
+correct inferences for the testsuite. (In this paper, we always assume
+that the problems in the FraCaS testsuite are correctly annotated.)
+
+\section{Temporal-Semantics in a Logic-based NLI System}
 \label{sec:background}
 
+Our temporal analysis places itself in the context of a complete NLI
+system -- which is why we can test it on the FraCaS suite. In this
+section we give a brief overview of the phases of the system,
+referring the reader to published work for details.
+
 \paragraph{GF}
-The grammatical framework (GF, \citet{ranta_grammatical_2004}) is a
+The first phase of the system, parsing, is taken care of by the 
+Grammatical Framework (GF, \citet{ranta_grammatical_2004}), which is a
 powerful parser generator for natural languages, based on
@@
 
-\paragraph{Coq}
-Coq is an interactive theorem prover (proof assistant). It is based on
+\paragraph{Dynamic Semantics}
+Parse trees are then processed by a dynamic semantic component.
+Its role is essentially to support  (non-temporal) anaphora, using a
+monadic-based dynamic semantics, generally following the state of the
+art in this matter
+\citep{unger_dynamic_2011,charlow_monadic_2015,charlow_modular_2017}.
+Our particular implementation does not handle every case \ifanon
+(citations anonymized) \else \citet{bernardy_computational_2020,bernardy_wide-coverage_2019} \fi
+but non-temporal anaphoroi in the testsuite are generally resolved as
+they should be: on the whole accuracy is not affected significantly by
+issues in this subsystem.
+
+As it is the case for other basic phenomena, there is not much
+interaction between our treatment of time and non-temporal
+anaphora. One critical exception is discussed in
+\cref{sec:repeatable-achievements}.
+
+\paragraph{Montagovian Semantics}
+Non-withstanding special support for anaphora, the core of the translation of syntax trees
+to logical form follows a largely standard
+montagovian semantics. In brief, sentences are interpreted as
+propositions, verbs and noun-phrases as predicates.  We use
+type-raising of noun-phrases, to support
+quantifiers~\citep{montague_proper_1974}.
+
+We support additionally the basic constructions and phenomena present
+in the testsuite, including adjectives, adverbs, nouns, verbs,
+anaphora, etc. The method is outlined by
+\citet{montague_english_1970,montague_proper_1973}, but we direct the
+reader to our previous work for details \ifanon (citations anonymized)
+\else \citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019} \fi, but the particular
+treatment of such phenomena is essentially independent from our
+treatment of time: in this paper we simply ignore these aspects beyond
+the fact that they are handled correctly in most of the cases of the
+FraCaS testsuite.
+
+\paragraph{Inference using Coq}
+Logical forms are then fed to the Coq
+Coq interactive theorem prover (proof assistant). Coq is based on
 the calculus of co-inductive constructions~\citep{werner_une_1994} We
@@
 problems, are solved with Coq's linear arithmetic tactic.
-
-\paragraph{Montagovian Semantics}
-
-Our interpretation is cast within the framework of a largely standard
-montagovian semantics. In brief, sentences are interpreted as
-propositions, verbs and noun-phrases as predicates.  We use
-type-raising of noun-phrases, to support quantifiers~\citep{montague_proper_1974}.
-
-We support additionally the basic constructions and phenomena present
-in the testsuite, including adjectives, adverbs, nouns, verbs,
-anaphora, etc. We direct the reader to our previous work for details
-\ifanon (citations anonymized) \else
-\citet{bernardy_type-theoretical_2017} \fi, but the particular
-treatment of such phenomena is essentially independent from our
-treatment of time:
-in this paper we simply ignore these aspects beyond the fact that they
-are handled correctly in most of the cases of the FraCaS testsuite.
-
-\paragraph{Dynamic Semantics}
-We additionally assume support for (non-temporal) anaphora, using a
-monadic-based dynamic semantics, generally following the state of the
-art in this matter
-\citep{unger_dynamic_2011,charlow_monadic_2015,charlow_modular_2017}.
-Our particular implementation does not handle every case \ifanon
-(citations anonymized) \else \citet{bernardy_wide-coverage_2019} \fi
-but non-temporal anaphoroi in the testsuite are generally resolved as
-they should be: on the whole accuracy is not affected significantly by
-issues in this subsystem.
-
-As it is the case for other basic phenomena, there is not much
-interaction between our treatment of time and non-temporal
-anaphora. One critical exception is discussed in
-\cref{sec:repeatable-achievements}.
-
-\paragraph{Temporal Semantics}
-
-In this paper we borrow several notions from classical temporal
-semantics such as ``stative'', ``achievement'', ``activity'', etc.,
-even though our definitions do not perfectly match the classical
-ones. We explain our precise meaning for these terms in the body of the
-paper. Nevertheless, we refer the reader to \citet{steedman:2000} for
-an extensive review of formal temporal semantics.
-
-For the \textit{cognoscenti}, we can already point out some
-differences in terminology: we use the term activity as a general term
-which encompasses the three classical notions of activites,
-achievements and accomplishments. Indeed, insofar as the test suite is
-concerned, we find that these three categories can be collapsed into a
-single one (they are subject to \cref{eq:unicity}).  As such, we think
-it is a good classification, given that it generally affords the
-correct inferences for the testsuite. (In this paper, we always assume
-that the problems in the FraCaS testsuite are correctly annotated.)
 
@@
 reference for the interpretation of all time-dependent semantics
-within the phrase.  This context propagates through the compositional
+within the phrase. (While some categories do not need this temporal context, we pass it everywhere for consistency.)  This context propagates through the compositional
 interpretation down to lexical items with atomic representation
@@
   in some way. Sometimes the restriction is an equality, as in ``for
-  exactly two hours''.
+  exactly two hours''. In the following example we show the inclusion
+  constraint, for ``in 1992''.
 
@@
 The principle of using unicity of action interacts well with the usual
-interpretation of quantifiers (and anaphora).  Indeed, using it, we
+interpretation of existential quantifiers (and anaphora).  Indeed, using it, we
 can refute problem (279), as expected by the testsuite:
@@
 stative/achievement distinction induced by the progressive. For
-example, the phrases ``John ran'' and ``John is running'' appear to be
+example, the phrases ``John ran'' and ``John was running'' appear to be
 logically equivalent, for entailment purposes.
@@
 generalizations (for example, we have quantifiers such as \verb!MOST! in
-addition to \verb!FORALL!.)
+addition to \verb!FORALL! --- see \ifanon (citation anonymized)
+\else \citet{bernardy_type-theoretical_2017} \fi)
 

### 2020-10-01 paper2020/paper.tex
deal with reviewer 2 and 3

@@
 work.
-
-\paragraph{Temporal Semantics}
-\todo{Fold this in some other section}
-In this paper we borrow several notions from classical temporal
-semantics such as ``stative'', ``achievement'', ``activity'', etc.,
-even though our definitions do not perfectly match the classical
-ones. We explain our precise meaning for these terms in the body of the
-paper. Nevertheless, we refer the reader to \citet{steedman:2000} for
-an extensive review of formal temporal semantics.
-
-For the \textit{cognoscenti}, we can already point out some
-differences in terminology: we use the term activity as a general term
-which encompasses the three classical notions of activites,
-achievements and accomplishments. Indeed, insofar as the test suite is
-concerned, we find that these three categories can be collapsed into a
-single one (they are subject to \cref{eq:unicity}).  As such, we think
-it is a good classification, given that it generally affords the
-correct inferences for the testsuite. (In this paper, we always assume
-that the problems in the FraCaS testsuite are correctly annotated.)
 
@@
 
+\paragraph{Aside: aspectual classes in the literature}
+In this paper we borrow several notions from classical temporal
+semantics such as ``stative'', ``achievement'', ``activity'', etc.,
+even though our definitions do not perfectly match the classical
+ones. We explain our precise meaning for these terms in the body of the
+paper. Nevertheless, we refer the reader to \citet{steedman:2000} for
+an extensive review of formal temporal semantics.
+
+For the \textit{cognoscenti}, we can already point out some
+differences in terminology: we use the term activity as a general term
+which encompasses the three classical notions of activites,
+achievements and accomplishments. Indeed, insofar as the test suite is
+concerned, we find that these three categories can be collapsed into a
+single one (they are subject to \cref{eq:unicity}).  As such, we think
+it is a good classification, given that it generally affords the
+correct inferences for the testsuite. (In this paper, we always assume
+that the problems in the FraCaS testsuite are correctly annotated.)
+
 \paragraph{Time references and aspectual classes}
@@
 such a level of precision.)
+
+Unicity of action plays the role of event coreference in
+(neo-)davidsonian account \todo{citation}. It is also a fine-grained
+principle, allowing coreference to take into account certain arguments
+when referencing. As we detail below, taking arguments into account
+yields is critical to handle repeatability of achievements.
 
@@
 
-Thanks to the unicity of action of $write(smith,x,...)$ we find
-\([t_1,t_2] = [t_3,t_4]\), and due to the years 1991 and 1992 being
-disjoint we obtain contradiction.
+Thanks to the unicity of action of $write(smith,x,...)$ (the subject
+and direct object are fixed) we find \([t_1,t_2] = [t_3,t_4]\), and
+due to the years 1991 and 1992 being disjoint we obtain contradiction.
 
@@
 
-  Of particular note is the use of the pronoun ``it'', and the fact
-  that adverbials expressions such that ``in 1992'' are lexicalized.
-  We also follow the GF convention to postfix lexical items with the
-  name of their category.
+Of particular note is the use of the pronoun ``it'', and the fact that
+adverbials expressions such that ``in 1992'' are lexicalized.  We also
+follow the GF convention to postfix lexical items with the name of
+their category. Most of the other categories follow usual naming
+conventions. We remind the reader that ``slash'' categories are used
+to swap the order of arguments (compared to non-slashed categories of
+similar names).
 
@@
   \caption{Accuracy of our system compared to others.
-    ``Ours" refers to the approach presented in this paper. When a
+    ``This" refers to the approach presented in this paper. When a
     system does not handle the nominal number of test cases (shown in
     the second column), the actual number of test cases attempted is
-    shown below the accuracy figure, in smaller font.  ``FraCoq''
-    refers to the work of \citet{bernardy_type-theoretical_2017}. ``MINE" refers
+    shown below the accuracy figure, in smaller font.  ``FC''
+    refers to the work of \citet{bernardy_type-theoretical_2017}, and ``FC2'' its followup~\citep{bernardy_wide-coverage_2019}. ``MINE" refers
     to the approach of \citet{Mineshima:2015}, ``NUT" to the CCG
@@
 phrases, guided by the applicability of the unicity of action and
-temporal subsumption properties.
+temporal subsumption properties. While part of this simplification is
+accidental (conflation of activity and accomplishment), we find that
+other parts (the automatic distinction between repeatable and
+unrepeatable achievements) constitute theoretical improvements.
 

### 2020-10-04 paper2020/paper.tex
minor fixes

@@
 of the FraCaS test suite has been difficult to tackle. That is,
-so far, no computational system has been capable to deal with it: when
+so far, no computational system has been capable to deal with it in its entirety: when
 authors report accuracy over the FraCaS test suite they
@@
 which includes support for anaphora and ellipsis but still omit the
-temporal section.
+temporal section.\footnote{\citet{maccartney:2007} have made a run against the whole test suite, but upon a closer look, this is not really the case, as they do not deal with multi-premise cases.
+For example, only 36/75 cases in the temporal section are treated.
+The general accuracy of the system is .59 and .61 for the temporal section.
+Our system, as shown \cref{tab:results}, presents considerable improvements and coverage over this system.}
 %
@@
 computationally viable account of tense and aspect to deal with the
-notoriously difficult section 7 of the FraCaS test suite. Our account is not meant
+  section 7 of the FraCaS test suite. Our account is not meant
 to be a theoretically extensive account of tense and aspect, but rather an
@@
 %
+
 The account is evaluated on the entailment properties of various
@@
 reasoning engine that makes it fit for implementing natural language
-semantics.  Coq supports, as already mentioned, dependent typing, but
-also subtyping. Both concepts are instrumental in expressing NL
+semantics.  Coq also supports  dependent typing and 
+ subtyping. Both concepts are instrumental in expressing NL
 semantics\citet{chatzikyriakidis_natural_2014}.  Besides, on a more
@@
 the Temporal section. However, it also supports the other eight
-sections.  To our knowledge this is the first system which attempt to
-target the temporal section. But in fact, our system is even provides
-support for all the other section. A couple of decades after its
+sections.  To our knowledge this is the first system which attempts to
+target the temporal section in full. But in fact, our system even provides
+support for all the other sections. Thus, a couple of decades after its
 formulation, we propose a first attempt at covering the whole suite.

### 2020-10-06 paper2020/paper.tex
wibble
kill some more todos and clarify others

@@
 which includes support for anaphora and ellipsis but still omit the
-temporal section.\footnote{\citet{maccartney:2007} have made a run against the whole test suite, but upon a closer look, this is not really the case, as they do not deal with multi-premise cases.
-For example, only 36/75 cases in the temporal section are treated.
-The general accuracy of the system is .59 and .61 for the temporal section.
-Our system, as shown \cref{tab:results}, presents considerable improvements and coverage over this system.}
+temporal section.\jp{check todo in footnote}\footnote{One can consider that \citet{maccartney:2007} have made a run against the whole test suite.
+However, they do not deal with multi-premise cases.
+Consequently only 36/75 cases in the temporal section are treated{TODO: jp: treated or attempted?
+I suppose attempted (otherwise .61 accuracy makes no sense)}.
+The general accuracy of the system is .59, and .61 for the temporal section.
+Our system, as shown \cref{tab:results}, presents considerable improvements in coverage and accuracy over that of \citeauthor{maccartney:2007}.}
 %
@@
 particular in the performance on the temporal section. In
-\cref{sec:conclusions}, we conclude and discuss avenues for future
+\cref{sec:conclusion}, we conclude and discuss avenues for future
 work.
@@
 
-This paper concludes our work in handling the whole of the FraCaS test
-suite using a combination of a number of well-studied tools: type
+\cite{bernardy_wide-coverage_2019} presented a
+logical system for handling 8 of the 9 sections of the FRACAS test
+suite, but excluded section 7, suggesting that it requires many
+examples that require ad hoc treatment. Here, we take up this
+challenge and show that a system similar to theirs can be extended to
+cover the remaining section of the test suite.
+%
+We use the same combination of a number of well-studied tools: type
 theory, parsing using the Grammatical Framework (GF), Monadic Dynamic
-Semantics and proof assistant technology (Coq). It achieves an
+Semantics and proof assistant technology (Coq). The system achieves an
 accuracy of 0.71 on the Temporal Section and 0.78 overall.  One of the

### 2020-10-09 paper2020/paper.tex
clarify todos
address comparision with Androutsopoulos1998time

@@
 Unicity of action plays the role of event coreference in
-(neo-)davidsonian account \todo{citation}. It is also a fine-grained
+(neo-)davidsonian account \todo{stergios: citation}. It is also a fine-grained
 principle, allowing coreference to take into account certain arguments
@@
 Note here that the scope for the existential is extended beyond the
-scope of \textbf{P1}, and its polarity switched (to universal). This extension can work according to
+scope of \textbf{P1}, and its polarity switched (to universal). This extension can follow the account of
 \citet{unger_dynamic_2011}, and our implemented analysis of
@@
 
+Outside of inference, formal interpretation of tense is found in
+natural-language interface to databases. Of note is the work of
+\citet{androutsopoulos1998time}, which handles many of the time-aware
+adverbial clauses that we address. However, we cover many more
+interaction logical aspects of inference, such as coreference via
+unity of action and interaction with quantifiers.
+
 \cite{bernardy_wide-coverage_2019} presented a
@@
 Semantics and proof assistant technology (Coq). The system achieves an
-accuracy of 0.71 on the Temporal Section and 0.78 overall.  One of the
+accuracy of 0.71 on the Temporal Section and 0.78 overall.
+One of the
 things to be looked at is fixing the issues associated with parts of

### 2020-10-27 paper2020/paper.tex
minor additions

@@
 Unicity of action plays the role of event coreference in
-(neo-)davidsonian account \todo{stergios: citation}. It is also a fine-grained
+(neo-)Davidsonian accounts \citep{parsons1990events}. It is also a fine-grained
 principle, allowing coreference to take into account certain arguments
@@
 same interval.  Consequently we get the entailment from ``John wrote a
-book'' to ``John was writing a book''.
+book'' to ``John was writing a book''.\footnote{Note that this is a different interpretation of the usual accounts of the progressive in the literature which take a progressive sentence  to be  true at an interval $t_0$
+ iff there is an interval $t_1$ such that $t_0$ is a non final
+subinterval of $t_1$ and the progressiver sentence is true at $t_1$ \citep{ogihara2007tense}. In terms of implementation, the solution we provide here seems to be simpler.}
 

### 2020-10-27 paper2020/paper.tex
minor additions

--- 2020-10-09 paper2020/paper.tex
@@
 same interval.  Consequently we get the entailment from ``John wrote a
-book'' to ``John was writing a book''.
+book'' to ``John was writing a book''.\footnote{Note that this is a different interpretation of the usual accounts of the progressive in the literature which take a progressive sentence  to be  true at an interval $t_0$
+ iff there is an interval $t_1$ such that $t_0$ is a non final
+subinterval of $t_1$ and the progressive sentence is true at $t_1$ \citep{ogihara2007tense}. In terms of implementation, the solution we provide here seems to be simpler.}
 

### 2020-10-28 MS.hs
Correct handling of negation in sec. 1

@@ -1072,26 +1072,25 @@ predVP np vp = withClause $ do
   MkNP pre n q cn gender <- np
   np' <- evalQuant pre q n (cn,gender) Subject
   vp' <- vp
-  let p' = np' vp'
   tense <- asks envTense
-  p'' <- case tense of
+  modifier <- case tense of
     t | t `elem` [Past, PresentPerfect, PastPerfect] -> do
               -- Allocate own time.
               t1 <- getFresh
               t2 <- getFresh
-              return $ \e@ExtraArgs{..} ->
+              return $ \prop e@ExtraArgs{..} ->
                   case extraTime of
                     UnspecifiedTime -> modP (quantTime Exists t2 (Con "PAST" `app` Var t2) .
                                              quantTime Exists t1 (Con "IS_INTERVAL" `apps` [Var t1,Var t2]))
-                                            (usingTime (ExactTime $ (Var t1,Var t2)) p' e)
+                                            (usingTime (ExactTime $ (Var t1,Var t2)) prop e)
                       -- (b) We do not have a specified time
                       -- Use own time. This time MUST be overridable by (2), otherwise we'll never
                       -- be able to override it, to search for it when we reach
                       -- point (1) at a later occurence of the same event.
-                    _ -> p' e -- (a) We have a specified time already, e.g coming from an adverbial phrase. Change nothing
-    _ -> return p' -- no specific time info, leave as such. This is important because the time may come from an adverbial phrase.
+                    _ -> prop e -- (a) We have a specified time already, e.g coming from an adverbial phrase. Change nothing
+    _ -> return id -- no specific time info, leave as such. This is important because the time may come from an adverbial phrase.
   modify (pushS (predVP np vp))
-  return $ usingCompClass cn p''
+  return $ usingCompClass cn ((np' (modifier . vp')))
 
 questCl :: Cl -> Cl
 questCl = id

### 2020-10-28 paper2020/paper.tex
finish checking examples and update paper

@@
                           & This        & FC2         & FC & MINE & Nut  & LP  \\ \hline
-Quantifiers  & 75         & .91        & .96         & .96    & .77  & .53  & .93  \\
+Quantifiers  & 75         & .93        & .96         & .96    & .77  & .53  & .93  \\
              &            & \ncases{74}& \ncases{74} &        &      &      &     \ncases{44} \\
-Plurals      & 33         & .64        & .82         & .76    & .67  & .52  & .73 \\
+Plurals      & 33         & .76        & .82         & .76    & .67  & .52  & .73 \\
              &            &            &             &     &   &   & \ncases{24} \\
@@
              &            & \ncases{20}&  \ncases{20}&     &   &   &  \ncases{12} \\
-Comparatives & 31         & .64        & .87         & .56    & .48  & .45  &  -       \\
+Comparatives & 31         & .65        & .87         & .56    & .48  & .45  &  -       \\
 Temporal     & 75         & .71        &  -          &   -    &   -  &  -   &  -       \\
@@
              &            &            &             &        &      &      & \ncases {9}  \\ \hline
-Total        & 337        & .78        & .89         & .83    & .69  & .50  & .85  \\
+Total        & 337        & .80        & .89         & .83    & .69  & .50  & .85  \\
              &            & \ncases{329}& \ncases{259}& \ncases{174}  & \ncases{174}& \ncases{174}& \ncases{89}
@@
 \end{table}
-\paragraph{interaction with negation and group readings}
+\paragraph{interaction with group readings}
 One reason explaining the lower performance of our system on some
 sections of the testsuite is that our interpretation of time interacts
-imperfectly with negation and group readings.
+imperfectly with group readings.
 
-The issue is that the quantification introduced by the temporal marker
-is outside the negation. For example, the premiss of (038) is
-interpreted as ``there exists a time in the past when every delegate
-did not finish the report'' --- which is incorrect in this context.
-\fracasex{038}{
-\item	No delegate finished the report.
-\hyp 	Some delegate finished the report on time.
-}
-%
-We stress that such interaction between quantifiers and negation
-already occurs without any temporal analysis: one must make a difference
+We stress that group readings is already flawed in the system which we base our analysis on.
+The issue is related to the difference
 between narrow  and wide scope.\footnote{The literature on wide and narrow scope readings of quantifiers and quantifiers and negation is vast.
-We direct the reader to \cite{steedman:2012} for a comprehensive and thorough discussion of the relevant issues.} Unfortunately,
-the Montagovian system over which we base our implementation does not
-support these phenomena. (In fact the FraCaS test suite does not appear to
+We direct the reader to \cite{steedman:2012} for a comprehensive and thorough discussion of the relevant issues.} (In fact the FraCaS test suite does not appear to
 test if the difference between narrow and wide scoping is handled properly.)
@@
 time-quantification, but not fundamentally different.
-%
-A similar issue occurs with group readings, which are interpreted as
-conjunctions. Here, the problem is an improper resolution of scoping
-order, between this conjunction and time-interval quantifiers.
 

### 2020-10-28 paper2020/paper.tex
update numbers in paper

@@
              &            & \ncases{74}& \ncases{74} &        &      &      &     \ncases{44} \\
-Plurals      & 33         & .76        & .82         & .76    & .67  & .52  & .73 \\
+Plurals      & 33         & .79        & .82         & .76    & .67  & .52  & .73 \\
              &            &            &             &     &   &   & \ncases{24} \\
@@
              &            &            &             &        &      &      & \ncases {9}  \\ \hline
-Total        & 337        & .80        & .89         & .83    & .69  & .50  & .85  \\
+Total        & 337        & .81        & .89         & .83    & .69  & .50  & .85  \\
              &            & \ncases{329}& \ncases{259}& \ncases{174}  & \ncases{174}& \ncases{174}& \ncases{89}

### 2020-10-28 paper2020/paper.tex
more correct error analysis

@@
 interaction between our treatment of time and non-temporal
-anaphora. One critical exception is discussed in
-\cref{sec:repeatable-achievements}.
+anaphora. Critical exceptions are discussed in
+\cref{sec:repeatable-achievements} and \cref{sec:interaction-with-anaphora}.
 
@@
 \end{table}
-\paragraph{interaction with group readings}
+\paragraph{interaction with anaphora}
+\label{sec:interaction-with-anaphora}
 One reason explaining the lower performance of our system on some
 sections of the testsuite is that our interpretation of time interacts
-imperfectly with group readings.
+imperfectly with anaphora and ellipsis. Consider the following example:
 
-We stress that group readings is already flawed in the system which we base our analysis on.
-The issue is related to the difference
-between narrow  and wide scope.\footnote{The literature on wide and narrow scope readings of quantifiers and quantifiers and negation is vast.
-We direct the reader to \cite{steedman:2012} for a comprehensive and thorough discussion of the relevant issues.} (In fact the FraCaS test suite does not appear to
-test if the difference between narrow and wide scoping is handled properly.)
-We propose that the situation becomes
-quantitatively more difficult to handle with the added complication of
-time-quantification, but not fundamentally different.
+\fracasex{232}
+{\item	ITEL won more orders than APCOM did.
+\item	APCOM won ten orders.
+\hyp 	ITEL won at least eleven orders.
+}
+
+In the first premiss, our system essentially resolves the ellipsis to
+get the following reading: ``ITEL won X orders and APCOM won Y orders
+and X > Y.''. One would need each of the verb phrases ``won X orders''
+and ``won Y orders'' to introduce their own timespans with existential
+quantifiers. However, the organisation of our system is such that the
+existentials are introduced before the ellipsis is
+expanded. Consequently we get a wrong interpretation and the inference
+cannot be made.
 
@@
 introduced. Some of these have been already mentioned: interaction of
-the temporal variables with group readings and negation.
+the temporal variables with anaphora.
 

### 2020-11-08 paper2020/paper.tex
minor

@@
   and progressives.  We apply these semantics to temporal problems of
-  the FraCaS testsuite and obtain an accuracy of 71 percent.
+  the FraCaS testsuite and obtain an accuracy of 81 percent.
 \end{abstract}
@@
 However, they do not deal with multi-premise cases.
-Consequently only 36/75 cases in the temporal section are treated{TODO: jp: treated or attempted?
-I suppose attempted (otherwise .61 accuracy makes no sense)}.
+Consequently only 36/75 cases in the temporal section are attempted.
 The general accuracy of the system is .59, and .61 for the temporal section.

### 2020-11-09 paper2020/paper.tex
minor

@@
  subtyping. Both concepts are instrumental in expressing NL
-semantics\citet{chatzikyriakidis_natural_2014}.  Besides, on a more
+semantics~\citep{chatzikyriakidis_natural_2014}.  Besides, on a more
 practical side, it works well for the the task of NLI, when the latter
@@
 \end{table}
-\paragraph{interaction with anaphora}
+\paragraph{Interaction with anaphora}
 \label{sec:interaction-with-anaphora}
@@
 
-In the first premiss, our system essentially resolves the ellipsis to
+In the first premise, our system essentially resolves the ellipsis to
 get the following reading: ``ITEL won X orders and APCOM won Y orders
@@
 
-Outside of inference, formal interpretation of tense is found in
-natural-language interface to databases. Of note is the work of
+Besides inference, formal interpretation of tense is found in
+natural-language interfaces to databases. Of note is the work of
 \citet{androutsopoulos1998time}, which handles many of the time-aware
 adverbial clauses that we address. However, we cover many more
-interaction logical aspects of inference, such as coreference via
+ logical aspects of inference, such as coreference via
 unity of action and interaction with quantifiers.
@@
 suite, but excluded section 7, suggesting that it requires many
-examples that require ad hoc treatment. Here, we take up this
+examples that need an ad hoc treatment. Here, we take up this
 challenge and show that a system similar to theirs can be extended to
@@
 test suite that will include more fine-grained examples of lexical and
-grammatical aspect and the interaction between those two, for example
+grammatical aspect, as well as the interaction between those two, for example
 cases where one needs to actually distinguish between achievements and

### 2020-11-09 paper2020/paper.tex
done

@@
 which includes support for anaphora and ellipsis but still omit the
-temporal section.\jp{check todo in footnote}\footnote{One can consider that \citet{maccartney:2007} have made a run against the whole test suite.
+temporal section.\footnote{One can consider that \citet{maccartney:2007} have made a run against the whole test suite.
 However, they do not deal with multi-premise cases.

### 2020-11-09 paper2020/paper.tex
update formatting

@@
 
-\documentclass[a4paper,11pt]{article}
-\usepackage[utf8]{inputenc}
+\documentclass[a4paper,twoside]{article}
 
-\usepackage[hyperref]{naaclhlt2019}
+\usepackage{subcaption}
+% \usepackage{calc}
 \usepackage{natbib}
@@
 \usepackage{times}
+\usepackage{amssymb}
+\usepackage{amstext}
 \usepackage{amsmath}
+\usepackage{amsthm}
 \usepackage{stmaryrd}
@@
 \usepackage{enumitem}
+\usepackage{multicol}
+\usepackage{url}
 \usepackage[capitalize]{cleveref}
+\usepackage[verbose]{newunicodechar}
+\input{../PaperTools/latex/newunicodedefs}
+\usepackage{SCITEPRESS}     % Please add other packages that you may need BEFORE the SCITEPRESS.sty package.
+
 \newlist{lingex}{enumerate}{3}
@@
 
-\aclfinalcopy % Uncomment this line for the final submission
-%\def\aclpaperid{***} %  Enter the acl Paper ID here
-
-%\setlength\titlebox{5cm}
-% You can expand the titlebox if you need extra space
-% to show all the authors. Please do not make the titlebox
-% smaller than 5cm (the original size); we will check this
-% in the camera-ready version and ask you to change it back.
-
-%\newcommand\BibTeX{B{\sc ib}\TeX}
-
-%\usepackage[backend=bibtex,citestyle=authoryear,style=authoryear,natbib]{biblatex}
-% new style not compatible with biblatex
-%\bibliography{../bbcl_lccm2018/bbcl_lccm2018.bib,iwcs2019}
-
-\usepackage[verbose]{newunicodechar}
-\input{../PaperTools/latex/newunicodedefs}
+\begin{document}
 
 \title{FraCaS: Temporal Analysis}
-\author{Jean-Philippe Bernardy \qquad Stergios Chatzikyriakidis \\
-  University of Gothenburg\\
-  {\tt firstname.lastname@gu.se} \\}
+\author{\authorname{Jean-Philippe Bernardy} \authorname{Stergios Chatzikyriakidis}
+  \affiliation{University of Gothenburg}
+  \email{firstname.lastname@gu.se}}
 
-\date{}
+\keywords{Temporal Semantics, Type-Theory}
 
-\begin{document}
-\maketitle
-
-\begin{abstract}
-  In this paper, we propose an implementation of temporal semantics
+\abstract{In this paper, we propose an implementation of temporal semantics
   which is suitable for inference problems. This implementation
@@
   the FraCaS testsuite and obtain an accuracy of 81 percent.
-\end{abstract}
+}
+
+\onecolumn \maketitle \normalsize \setcounter{footnote}{0} \vfill
 
@@
 \fi
-\bibliographystyle{acl_natbib}
+\bibliographystyle{apalike}
 

### 2020-11-09 paper2020/paper.tex
extend discussion on progressive

@@
 while ``John wrote a book'' is encoded as
-$∃(t_1, t_2). t_1≤ t_2, t_2 ≤ now, write(John,book,t_1,t_2)$. This is
-necessary, because $PROG\_write$ is subject to
-$\constant{subsumption}$: if John is writing in the interval
+$∃(t_1, t_2). t_1≤ t_2, t_2 ≤ now, write(John,book,t_1,t_2)$. This distinction is
+necessary because in our analysis the progressive form ($PROG\_write$) is subject to
+$\constant{subsumption}$. That is, if John is writing in the interval
 $[t_1,t_2]$ then he is writing in any sub-interval of
-$[t_1,t_2]$. However, the same cannot be said of $write$: the
+$[t_1,t_2]$. This interpretation corresponds to the idea the the action takes place continuously over the whole interval.
+However, the same cannot be said of the non-continuous form ($write$): the
 end-points of the interval indicate the time needed to complete the
@@
 entails ``John wrote a book in January 1993'' nor ``John wrote a book
-in December 1993''.) In fact, \emph{write}, in the non-progressive
-from, is on the contrary subject to $\constant{unicity}$. Finally, we
+in December 1993''.) (In fact, \emph{write}, in the non-progressive
+from, is on the contrary subject to $\constant{unicity}$.) Finally, we
 also have $write(x,y,t_1,t_2) → PROG\_write(x,y,t_1,t_2)$. That is,
-the achievement (or \emph{activity} in our terminology) variant implies the stative variant, in the
+the achievement (or \emph{activity} in our terminology) variant implies the stative variant, for the
 same interval.  Consequently we get the entailment from ``John wrote a
-book'' to ``John was writing a book''.\footnote{Note that this is a different interpretation of the usual accounts of the progressive in the literature which take a progressive sentence  to be  true at an interval $t_0$
- iff there is an interval $t_1$ such that $t_0$ is a non final
-subinterval of $t_1$ and the progressive sentence is true at $t_1$ \citep{ogihara2007tense}. In terms of implementation, the solution we provide here seems to be simpler.}
+book in 1993'' to ``John was writing a book in 1993'', but not the other way around.
 
-Finally we note that not all verbs are subject to the
+We note however that this interpretation differs only slightly from
+the usual accounts of the progressive in the
+literature. \Citet{ogihara2007tense} summarizes the position of
+\citet{bennett1978toward} as follows: a progressive sentence is true
+at an interval $[t_0,t_1]$ iff there is an interval $[t'_0,t'_1]$ such
+that $[t_0,t_1]$ is a non-final subinterval of $[t'_0,t'_1]$ and the
+progressive sentence is true at $[t'_0,t'_1]$. This is very similar to
+our approach ($\constant{subsumption}$ for the progressive form only),
+but there is a difference regarding final intervals. Yet in our view
+this difference is hard to justify: we cannot see why ``John was
+writing a book in 1993'' entails that he was writing it January,
+February, etc. but not in December.
+
+\Citet{ogihara2007tense} argues that this simple account of the
+progressive fails to reject a sentence such as ``Lee is resembling
+Terri.'' while ``Lee is walking'' is acceptable. We argue instead that
+the latter should be rejected for pragmatic reasons. Indeed, when a
+predicate holds for a very long interval, one typically uses the
+simple present tense in English. Therefore the continuous form
+pragmatically implies that the predicate holds for a limited interval.
+But, without further context, the predicate ``resemble Terri'' does
+not vary over time (while ``walk'' generally does). Therefore the
+continuous form ``Lee is resembling Terri'' is confusing: one implies
+a limited interval, but the semantics of resembling normally yield an unlimited interval.
+%
+Because we do not account for pragmatics, we prefer to retain the
+simplest account based on $\constant{subsumption}$.
+
+Finally we stress that not all verbs are subject to the
 stative/achievement distinction induced by the progressive. For
-example, the phrases ``John ran'' and ``John was running'' appear to be
-logically equivalent, for entailment purposes.
+example, the phrases ``John ran'' and ``John was running'' appear to
+be logically equivalent, for entailment purposes.
 
@@
 
-Another thing we are willing to try is an extension of the FraCaS
+Another thing that we are willing to try is an extension of the FraCaS
 test suite that will include more fine-grained examples of lexical and
@@
 to be captured:
+
 %Another issue to be looking at is the codification of lexical
@@
 
-% Things we cannot handle (I think): John was writing a book in 1991
-% → John wrote a book in 1991 (I think we predict entailment
-% here). This is because there is no distinction with progressive and
-% non progressive aspect. We did not do it, I guess because it was
-% needed, but just for future reference. There are similar examples
-% playing around with LEXICAL aspect (aspect grammatically encoded)
-% that we could look at in future work
-
-% Accomplishments usually involve a culmination point in the action
-% which is necessary for the event to be completed. This is not
-% necessary for activities. John stopped writing a book not→ John
-% wrote a book (but does imply that John was writing a book), but,
-% John stopped running → John run and John was runnning
-
 \end{document}

### 2020-11-09 paper2020/paper.tex
clarify FraCaS extension status

@@
 
-Another thing that we are willing to try is an extension of the FraCaS
-test suite that will include more fine-grained examples of lexical and
+Another extension of this work is to reflect more temporal semantic
+inference properties in an extended test suite. Indeed, there as properties which are
+not captured in the FraCaS test suite, such as fine-grained examples of lexical and
 grammatical aspect, as well as the interaction between those two, for example
@@
 to be captured:
-
-%Another issue to be looking at is the codification of lexical
-%aspect.\todo{update} In what we have been doing so far, we did not distinguish
-%between the present continuous and the present or the simple past and
-%the past continuous and their interaction with non-lexical aspect.
-%The main reason for this is that in the test suite, even though there
-%are examples of inferences based on the either lexical or non-lexical
-%aspect, no example where an interaction between the two exists. For
-%example, one such case is the one shown below and exemplifies the
-%difference between activities and accomplishments with respect to the
-%different inferences they give when used in the past continuous:
 

### 2020-11-09 paper2020/table.tex
draft table of examples

% Proposing a table of examples to include

\subsection{Correctly handled examples}

\fracasex{252}{\answer yes
\item	Since 1992 ITEL has been in Birmingham.
\item	It is now 1996.
\hyp    Itel was in Birmingham in 1993.

} \fracasex{253	}{\answer unknown
\item	ITEL has developed a new editor since 1992.
\item	It is now 1996.
\hyp    ITEL developed a new editor in 1993.

} \fracasex{254	}{\answer unknown
\item	ITEL has expanded since 1992.
\item	It is now 1996.
\hyp    ITEL expanded in 1993.

} \fracasex{258	}{\answer no
\item	In March 1993 APCOM founded ITEL.
\hyp    ITEL existed in 1992.

} \fracasex{260	}{\answer yes
\item	Yesterday APCOM signed the contract.
\item	Today is Saturday, July 14th.
\hyp    APCOM signed the contract Friday, 13th.

} \fracasex{261	}{\answer yes
\item	Smith left before Jones left.
\item	Jones left before Anderson left.
\hyp    Smith left before Anderson left.

} \fracasex{262	}{\answer yes
\item	Smith left after Jones left.
\item	Jones left after Anderson left.
\hyp    Smith left after Anderson left.

} \fracasex{264	}{\answer yes
\item	Smith left.
\item	Jones left.
\item	Smith left before Jones left.
\hyp    Jones left after Smith left.
} \fracasex{265	}{\answer yes
\item	Smith left.
\item	Jones left.
\item	Smith left after Jones left.
\hyp    Jones left before Smith left.

} \fracasex{266	}{\answer yes
\item	Smith left.
\item	Jones left.
\item	Jones left before Smith left.
\hyp    Smith left after Jones left.
} \fracasex{267	}{\answer yes
\item	Jones revised the contract.
\item	Smith revised the contract.
\item	Jones revised the contract before Smith did.
\hyp    Smith revised the contract after Jones did.
} \fracasex{268	}{\answer yes
\item	Jones revised the contract.
\item	Smith revised the contract.
\item	Jones revised the contract after Smith did.
\hyp    Smith revised the contract before Jones did.

} \fracasex{269	}{\answer unknown
\item	Smith swam.
\item	Jones swam.
\item	Smith swam before Jones swam.
\hyp    Jones swam after Smith swam.

 However we do get

} \fracasex{270	}{\answer yes
\item	Smith swam to the shore.
\item	Jones swam to the shore.
\item	Smith swam to the shore before Jones swam to the shore.
\hyp    Jones swam to the shore after Smith swam to the shore.

} \fracasex{273	}{\answer unknown
\item	Smith was writing a report.
\item	Jones was writing a report.
\item	Smith was writing a report before Jones was writing a report.
\hyp    Jones was writing a report after Smith was writing a report.

} \fracasex{274	}{\answer unknown
\item	Smith was writing a report.
\item	Jones was writing a report.
\item	Smith was writing a report after Jones was writing a report.
\hyp    Jones was writing a report before Smith was writing a report.

} \fracasex{277	}{\answer unknown
\item	Smith lived in Birmingham in 1991.
\hyp    Smith lived in Birmingham in 1992.

} \fracasex{279	}{\answer no
\item	Smith wrote a novel in 1991.
\hyp    Smith wrote it in 1992.

} \fracasex{280	}{\answer unknown
\item	Smith wrote a novel in 1991.
\hyp    Smith wrote a novel in 1992.

} \fracasex{281	}{\answer unknown
\item	Smith was running a business in 1991.
\hyp    Smith was running it in 1992.

} \fracasex{282	}{\answer no
\item	Smith discovered a new species in 1991.
\hyp    Smith discovered it in 1992.

} \fracasex{283	}{\answer unknown
\item	Smith discovered a new species in 1991.
\hyp    Smith discovered a new species in 1992.

} \fracasex{284	}{\answer yes
\item	Smith wrote a report in two hours.
\item	Smith started writing the report at 8 am.
\hyp    Smith had finished writing the report by 11 am.

} \fracasex{287	}{\answer unknown
\item	Smith wrote a report in two hours.
\hyp    Smith wrote a report in one hour.
} \fracasex{288	}{\answer yes
\item	Smith wrote a report in two hours.
\hyp    Smith wrote a report.
} \fracasex{289	}{\answer no
\item	Smith discovered a new species in two hours.
\hyp    Smith spent two hours discovering the new species.
Why 	Achievements are typically (more or less) instantaneous
} \fracasex{290	}{\answer yes
\item	Smith discovered a new species in two hours.
\hyp    Smith discovered a new species.
} \fracasex{291	}{\answer yes **
\item	Smith discovered many new species in two hours.
\hyp    Smith spent two hours discovering new species.
A 	?Yes
Why 	Repeated achievement can last two hours.
} \fracasex{292	}{\answer unknown
\item	Smith was running his own business in two years.
\hyp    Smith spent two years running his own business.
Why 	Premise refers to time taken to inception of activity, not duration of activity.
} \fracasex{293	}{\answer unknown
\item	Smith was running his own business in two years.
\hyp    Smith spent more than two years running his own business.
Why 	Cf. similar inference for accomplishment, (286)
} \fracasex{295	}{\answer unknown
\item	In two years Smith owned a chain of businesses.
\hyp    Smith owned a chain of business for two years.
Why 	States behave like activities.
Note 	Sic: the original did have "chain of business" in the question.
} \fracasex{297	}{\answer yes
\item	In two years Smith owned a chain of businesses.
\hyp    Smith owned a chain of business.
Note 	Sic: the original did have "chain of business" in the question.
} \fracasex{298	}{\answer yes
\item	Smith lived in Birmingham for two years.
\hyp    Smith lived in Birmingham for a year.
Why 	State
} \fracasex{299	}{\answer no
\item	Smith lived in Birmingham for two years.
\hyp    Smith lived in Birmingham for exactly a year.
} \fracasex{300	}{\answer yes
\item	Smith lived in Birmingham for two years.
\hyp    Smith lived in Birmingham.
} \fracasex{301	}{\answer yes
\item	Smith ran his own business for two years.
\hyp    Smith ran his own business for a year.
Why 	Activity
} \fracasex{303	}{\answer yes
\item	Smith wrote a report for two hours.
\hyp    Smith wrote a report for an hour.
Why 	Accomplishment
} \fracasex{304	}{\answer unknown
\item	Smith wrote a report for two hours.
\hyp    Smith wrote a report.
Why 	He may not have finished it

} \fracasex{307	}{\answer yes
\item	In 1994 ITEL sent a progress report every month.
\hyp    ITEL sent a progress report in July 1994.

} \fracasex{311	}{\answer yes
\item	Smith had left the house at a quarter past five.
\item	Then she took a taxi to the station.
\hyp    Smith left the house before she took a taxi to the station.
} \fracasex{312	}{\answer yes
\item	ITEL always delivers reports late.
\item	In 1993 ITEL delivered reports.
\hyp    ITEL delivered reports late in 1993.
Note 	Sic: the original did have "delivered" in the question.
} \fracasex{313	}{\answer no
\item	ITEL never delivers reports late.
\item	In 1993 ITEL delivered reports.
\hyp    ITEL delivered reports late in 1993.
} \fracasex{314	}{\answer yes
\item	Smith arrived in Paris on the 5th of May, 1995.
\item	Today is the 15th of May, 1995.
\item	She is still in Paris.
\hyp    Smith was in Paris on the 7th of May, 1995.

} \fracasex{320	}{\answer yes
\item	When Jones got his job at the CIA, he knew that he would never be allowed to write his memoirs.
\hyp    It is the case that Jones is not and will never be allowed to write his memoirs.

\subsection{Incorrectly handled examples}
{

} \fracasex{259	}{\answer yes
\item	The conference started on July 4th, 1994.
\item	It lasted 2 days.
\hyp    The conference was over on July 8th, 1994.

} \fracasex{263	}{\answer unknown
\item	Smith was present after Jones left.
\item	Jones left after Anderson was present.
\hyp    Smith was present after Anderson was present.

} \fracasex{278	}{\answer no
\item	Smith wrote his first novel in 1991.
\hyp    Smith wrote his first novel in 1992.

} \fracasex{285	}{\answer unknown
\item	Smith wrote a report in two hours.
\hyp    Smith spent two hours writing the report.

} \fracasex{286	}{\answer no
\item	Smith wrote a report in two hours.
\hyp    Smith spent more than two hours writing the report.

} \fracasex{294	}{\answer yes
\item	Smith was running his own business in two years.
\hyp    Smith ran his own business.

} \fracasex{296	}{\answer unknown
\item	In two years Smith owned a chain of businesses.
\hyp    Smith owned a chain of business for more than two years.

} \fracasex{302	}{\answer yes
\item	Smith ran his own business for two years.
\hyp    Smith ran his own business.

} \fracasex{306	}{\answer yes
\item	Smith discovered new species for two years.
\hyp    Smith discovered new species.

} \fracasex{315	}{\answer yes
\item	When Smith arrived in Katmandu she had been travelling for three days.
\hyp    Smith had been travelling the day before she arrived in Katmandu.

} \fracasex{316	}{\answer yes
\item	Jones graduated in March and has been employed ever since.
\item	Jones has been unemployed in the past.
\hyp    Jones was unemployed at some time before he graduated.

} \fracasex{317	}{\answer yes
\item	Every representative has read this report.
\item	No two representatives have read it at the same time.
\item	No representative took less than half a day to read the report.
\item	There are sixteen representatives.
\hyp    It took the representatives more than a week to read the report.

} \fracasex{318	}{\answer no
\item	While Jones was updating the program, Mary came in and told him about the board meeting.
\item	She finished before he did.
\hyp    Mary's story lasted as long as Jones's updating the program.

} \fracasex{321	}{\answer yes
\item	Smith has been to Florence twice in the past.
\item	Smith will go to Florence twice in the coming year.
\hyp    Two years from now Smith will have been to Florence at least four times.

} \fracasex{323	}{\answer yes
\item	No one gambling seriously stops until he is broke.
\item	No one can gamble when he is broke.

\hyp    Everyone who starts gambling seriously stops the moment he is broke.
}

### 2020-11-09 paper2020/table.tex
list failure reasons

@@
 \hyp    Jones swam after Smith swam.
-
- However we do get
 
@@
 \hyp    It is the case that Jones is not and will never be allowed to write his memoirs.
+}
 
 \subsection{Incorrectly handled examples}
+
+We list below a number of selected examples from the temporal section
+of the FraCaS testsuite where our semantics fails to list the correct
+inference. We list the reason why it fails under the item \textbf{F}.
+
 {
@@
 \hyp    The conference was over on July 8th, 1994.
-
+\failsf Incorrect lexical semantics for ``last $n$ days''.
 } \fracasex{263	}{\answer unknown
@@
 \hyp    Smith was present after Anderson was present.
+\failsf Copula do not get timespans
 
@@
 \hyp    Smith wrote his first novel in 1992.
+\failsf Unicity of ordinals not implemented
 
@@
 \hyp    Smith spent two hours writing the report.
+\failsf Temporal semanticts of ``spend'' incorrect
 
@@
 \hyp    Smith spent more than two hours writing the report.
+\failsf Temporal semanticts of ``spend'' incorrect
 
@@
 \hyp    Smith ran his own business.
+\failsf Parsing error for ``in two years''
 
@@
 \hyp    Smith owned a chain of business for more than two years.
+\failsf Parsing error for ``in two years''
 
@@
 \hyp    Smith had been travelling the day before she arrived in Katmandu.
+\failsf Incorrect idiomatic semantics for ``the day before''
 
@@
 \hyp    Jones was unemployed at some time before he graduated.
+\failsf Incorrect idiomatic semantics
 
@@
 \hyp    It took the representatives more than a week to read the report.
+\failsf Counting semantics not implemented
 
@@
 \hyp    Mary's story lasted as long as Jones's updating the program.
+\failsf Incorrect interaction with anaphora
 
@@
 \hyp    Two years from now Smith will have been to Florence at least four times.
+\failsf Counting semantics not implemented
 
@@
 \hyp    Everyone who starts gambling seriously stops the moment he is broke.
+\failsf Idiomatic semantics of ``the moment'' not supported
 }

### 2020-11-09 paper2020/paper.tex
draft table of examples
list failure reasons
fix 2 more cases

@@
 \newcommand\hyp{\item[H]}
+\newcommand\answer{\item[A]}
+\newcommand\failsf{\item[F]}
 \newcommand\fracasex[2]{\begin{lingex}\item[(#1)] \begin{subex} #2 \end{subex} \end{lingex} }
@@
 Comparatives & 31         & .65        & .87         & .56    & .48  & .45  &  -       \\
-Temporal     & 75         & .71        &  -          &   -    &   -  &  -   &  -       \\
+Temporal     & 75         & .73        &  -          &   -    &   -  &  -   &  -       \\
 Verbs        & 8          & .75        & .75         &   -    & -    & -    &  -       \\
@@
 
+\section{Examples}
+\input{table}
+
 \section*{Acknowledgments}

### 2020-11-09 Dynamic.hs

@@ -94,14 +94,15 @@ onS' f s extra = first f (s extra)
   -- f (p eos)
 
 type TimeSpan = (Exp,Exp)
-data Temporal = ForceTime TimeSpan | ExactTime TimeSpan | UnspecifiedTime {-  TenseTime Temporal-} deriving Show
+data Temporal = -- ForceTime TimeSpan |
+                ExactTime TimeSpan | UnspecifiedTime {-  TenseTime Temporal-} deriving Show
 
 now :: Temporal
 now = ExactTime (Con "NOW", Con "NOW")
 
 instance Semigroup Temporal where
-  ForceTime t <> _ = ForceTime t
-  _ <> ForceTime t = ForceTime t
+  -- ForceTime t <> _ = ForceTime t
+  -- _ <> ForceTime t = ForceTime t
   UnspecifiedTime <> x = x
   x <> UnspecifiedTime = x
   -- TenseTime _ <> x = x -- time specification given by tense, this is overridden by specific times.
@@ -495,7 +496,7 @@ appArgs isTimed nm objs@(_:_) (ExtraArgs {..}) = (extraAdvs (app (pAdverbs time'
           Activity -> extraTime
           Achievement -> case extraTime of
             ExactTime (_,t0) -> ExactTime (t0,t0)
-            ForceTime tspan -> ForceTime tspan
+            -- ForceTime tspan -> ForceTime tspan
             UnspecifiedTime -> UnspecifiedTime
         nmPrep = prog ++ nm ++ concatMap fst prepositions
         prog = if extraProgressive then "PROG_" else ""
@@ -527,14 +528,14 @@ sentenceApplyAdv adv s' ExtraArgs{..} = s' ExtraArgs {extraAdvs = adv . extraAdv
 --------------------------
 -- Time
 
-forceTime :: TimeSpan -> Cl' -> Prop
-forceTime tMeta cl = noExtraObjs (useTime tMeta cl)
+-- forceTime :: TimeSpan -> Cl' -> Prop
+-- forceTime tMeta cl = noExtraObjs (useTime tMeta cl)
 -- HACK: setting the time is at the "UseCl" level, which has to set a
 -- time. But we need to override it from the level above (S), so we
 -- use the "ForceTime" hack.
 
-useTime :: TimeSpan -> Cl' -> S'
[…]

### 2020-11-09 MS.hs

@@ -480,7 +480,7 @@ withAtLeastDuration delta = do -- with duration meaning, as in FraCas 278
            let (p,actualSpan) = s extraObjs {extraTime = tSpan}
            in (quantTime Exists t1 true $
                quantTime Exists t2 ((Con "IS_INTERVAL" `apps` [intervalStart (actualSpan),intervalStart (extraTime extraObjs)]) ∧
-                                    -- (Con "IS_INTERVAL" `apps` [intervalEnd (actualSpan),Con "NOW"]) ∧
+                                    (Con "IS_INTERVAL" `apps` [Var t2,intervalEnd (extraTime extraObjs)]) ∧
                                 -- it may finish later, see Fracas 304
                                 -- we constraint the returned time, so that we have a stronger interpretation of achievements (see Fracas 306)
                                  (Con "IS_INTERVAL" `apps` [Con "Plus" `apps` [Var t1, delta], Var t2])) $
@@ -1087,7 +1087,8 @@ predVP np vp = withClause $ do
                       -- Use own time. This time MUST be overridable by (2), otherwise we'll never
                       -- be able to override it, to search for it when we reach
                       -- point (1) at a later occurence of the same event.
-                    _ -> prop e -- (a) We have a specified time already, e.g coming from an adverbial phrase. Change nothing
+                    _ -> prop e -- ExactTime (t1,t2) -> first (∧ (Con "PAST" `app` t2)) (prop e) 
+                      -- (a) We have a specified time already, e.g coming from an adverbial phrase. Only fix it to the past.
     _ -> return id -- no specific time info, leave as such. This is important because the time may come from an adverbial phrase.
   modify (pushS (predVP np vp))
   return $ usingCompClass cn ((np' (modifier . vp')))

### 2020-11-11 paper2020/paper.tex
improve listing of examples

@@
 
-\section{Examples}
-\input{table}
+\begin{figure*}
+\begin{subfigure}{0.5\textwidth}
+
+\fracasex{252}{\answer yes
+\item	Since 1992 ITEL has been in Birmingham.
+\item	It is now 1996.
+\hyp    Itel was in Birmingham in 1993.
+} \fracasex{260	}{\answer yes
+\item	Yesterday APCOM signed the contract.
+\item	Today is Saturday, July 14th.
+\hyp    APCOM signed the contract Friday, 13th.
+
+} \fracasex{262	}{\answer yes
+\item	Smith left after Jones left.
+\item	Jones left after Anderson left.
+\hyp    Smith left after Anderson left.
+
+} \fracasex{312	}{\answer yes
+\item	ITEL always delivers reports late.
+\item	In 1993 ITEL delivered reports.
+\hyp    ITEL delivered reports late in 1993.
+
+} \fracasex{320	}{\answer yes
+\item	When Jones got his job at the CIA, he knew that he would never be allowed to write his memoirs.
+\hyp    It is the case that Jones is not and will never be allowed to write his memoirs.
+}
+
+  \caption{Selected correctly handled examples}
+  \label{fig:examples}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+{
+} \fracasex{259	}{\answer yes
+\item	The conference started on July 4th, 1994.
+\item	It lasted 2 days.
+\hyp    The conference was over on July 8th, 1994.
+\failsf Incorrect lexical semantics for ``last $n$ days''.
+} \fracasex{294	}{\answer yes
+\item	Smith was running his own business in two years.
+\hyp    Smith ran his own business.
+\failsf Parsing error for ``in two years''
+} \fracasex{317	}{\answer yes
+\item	Every representative has read this report.
+\item	No two representatives have read it at the same time.
+\item	No representative took less than half a day to read the report.
+\item	There are sixteen representatives.
+\hyp    It took the representatives more than a week to read the report.
+\failsf Counting semantics not implemented
+% } \fracasex{318	}{\answer no
+% \item	While Jones was updating the program, Mary came in and told him about the board meeting.
+% \item	She finished before he did.
+% \hyp    Mary's story lasted as long as Jones's updating the program.
+% \failsf Incorrect interaction with anaphora
+}
+  \caption{Selected incorrectly handled examples.
+  We list below a number of selected examples from the temporal section
+of the FraCaS testsuite where our semantics fails to list the correct
+inference. We list the reason why it fails under the item \textbf{F}.
+}
+  \label{fig:examples}
+\end{subfigure}
+\end{figure*}
+% \section{Examples}
+% \input{table}
 

### 2020-11-11 paper2020/paper.tex
typography
update accuracy numbers in text

@@
 
+\begin{figure*}
+\begin{subfigure}{0.5\textwidth}
+
+\fracasex{252}{\answer yes
+\item	Since 1992 ITEL has been in Birmingham.
+\item	It is now 1996.
+\hyp    Itel was in Birmingham in 1993.
+} \fracasex{260	}{\answer yes
+\item	Yesterday APCOM signed the contract.
+\item	Today is Saturday, July 14th.
+\hyp    APCOM signed the contract Friday, 13th.
+
+} \fracasex{262	}{\answer yes
+\item	Smith left after Jones left.
+\item	Jones left after Anderson left.
+\hyp    Smith left after Anderson left.
+
+} \fracasex{312	}{\answer yes
+\item	ITEL always delivers reports late.
+\item	In 1993 ITEL delivered reports.
+\hyp    ITEL delivered reports late in 1993.
+
+} \fracasex{320	}{\answer yes
+\item	When Jones got his job at the CIA, he knew that he would never be allowed to write his memoirs.
+\hyp    It is the case that Jones is not and will never be allowed to write his memoirs.
+}
+
+  \caption{Selected correctly handled examples}
+  \label{fig:correct-examples}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+{
+} \fracasex{259	}{\answer yes
+\item	The conference started on July 4th, 1994.
+\item	It lasted 2 days.
+\hyp    The conference was over on July 8th, 1994.
+\failsf Incorrect lexical semantics for ``last $n$ days''.
+} \fracasex{294	}{\answer yes
+\item	Smith was running his own business in two years.
+\hyp    Smith ran his own business.
+\failsf Parsing error for ``in two years''
+} \fracasex{317	}{\answer yes
+\item	Every representative has read this report.
+\item	No two representatives have read it at the same time.
+\item	No representative took less than half a day to read the report.
+\item	There are sixteen representatives.
+\hyp    It took the representatives more than a week to read the report.
+\failsf Counting semantics not implemented
+% } \fracasex{318	}{\answer no
+% \item	While Jones was updating the program, Mary came in and told him about the board meeting.
+% \item	She finished before he did.
+% \hyp    Mary's story lasted as long as Jones's updating the program.
+% \failsf Incorrect interaction with anaphora
+}
+  \caption{Selected incorrectly handled examples.
+  We list below a number of selected examples from the temporal section
+of the FraCaS testsuite where our semantics fails to list the correct
+inference. We list the reason why it fails under the item \textbf{F}.
+}
+  \label{fig:incorrect-examples}
+\end{subfigure}
+% \caption{Selected examples.}
+\end{figure*}
+
 \section{Results and Evaluation}
@@
 \end{table}
+Additionally, to provide the reader with a better sense of what the
+complete system is capable and not capable of, we list a number of
+correctly and incorrectly handled examples in \cref{fig:correct-examples} and \cref{fig:incorrect-examples}.
 \paragraph{Interaction with anaphora}
@@
 Semantics and proof assistant technology (Coq). The system achieves an
-accuracy of 0.71 on the Temporal Section and 0.78 overall.
+accuracy of 0.73 on the Temporal Section and 0.81 overall.
 One of the
@@
 
-\begin{figure*}
-\begin{subfigure}{0.5\textwidth}
-
-\fracasex{252}{\answer yes
-\item	Since 1992 ITEL has been in Birmingham.
-\item	It is now 1996.
-\hyp    Itel was in Birmingham in 1993.
-} \fracasex{260	}{\answer yes
-\item	Yesterday APCOM signed the contract.
-\item	Today is Saturday, July 14th.
-\hyp    APCOM signed the contract Friday, 13th.
-
-} \fracasex{262	}{\answer yes
-\item	Smith left after Jones left.
-\item	Jones left after Anderson left.
-\hyp    Smith left after Anderson left.
-
-} \fracasex{312	}{\answer yes
-\item	ITEL always delivers reports late.
-\item	In 1993 ITEL delivered reports.
-\hyp    ITEL delivered reports late in 1993.
-
-} \fracasex{320	}{\answer yes
-\item	When Jones got his job at the CIA, he knew that he would never be allowed to write his memoirs.
-\hyp    It is the case that Jones is not and will never be allowed to write his memoirs.
-}
-
-  \caption{Selected correctly handled examples}
-  \label{fig:examples}
-\end{subfigure}
-\begin{subfigure}{0.5\textwidth}
-{
-} \fracasex{259	}{\answer yes
-\item	The conference started on July 4th, 1994.
-\item	It lasted 2 days.
-\hyp    The conference was over on July 8th, 1994.
-\failsf Incorrect lexical semantics for ``last $n$ days''.
-} \fracasex{294	}{\answer yes
-\item	Smith was running his own business in two years.
-\hyp    Smith ran his own business.
-\failsf Parsing error for ``in two years''
-} \fracasex{317	}{\answer yes
-\item	Every representative has read this report.
-\item	No two representatives have read it at the same time.
-\item	No representative took less than half a day to read the report.
-\item	There are sixteen representatives.
-\hyp    It took the representatives more than a week to read the report.
-\failsf Counting semantics not implemented
-% } \fracasex{318	}{\answer no
-% \item	While Jones was updating the program, Mary came in and told him about the board meeting.
-% \item	She finished before he did.
-% \hyp    Mary's story lasted as long as Jones's updating the program.
-% \failsf Incorrect interaction with anaphora
-}
-  \caption{Selected incorrectly handled examples.
-  We list below a number of selected examples from the temporal section
-of the FraCaS testsuite where our semantics fails to list the correct
-inference. We list the reason why it fails under the item \textbf{F}.
-}
-  \label{fig:examples}
-\end{subfigure}
-\end{figure*}
-% \section{Examples}
-% \input{table}
-
 \section*{Acknowledgments}

### 2020-11-11 paper2020/paper.tex
final tweak and spelling

@@
   including: temporal references, temporal adverbs, aspectual classes
-  and progressives.  We apply these semantics to temporal problems of
-  the FraCaS testsuite and obtain an accuracy of 81 percent.
+  and progressives.  We apply these semantics to the complete FraCaS testsuite.
+  We obtain an accuracy of 81 percent overall and 73 percent for problems explicitly marked
+  as related to temporal reference.
 }
@@
 Example \ref{ex:nobel} shows an example abstract syntax tree and its
-realization in English.
+realisation in English.
 
@@
 practical side, it works well for the the task of NLI, when the latter
-is formalized as a theorem proving task: its many tactics mean that
-many tasks in theorem proving are trivialized. In particular, all
+is formalised as a theorem proving task: its many tactics mean that
+many tasks in theorem proving are trivialised. In particular, all
 problems of time-intervals inclusion, which occur in every temporal
@@
 justifies ignoring the input timespan in the above interpretation --
-we are in particular not interested in modeling felicity with our
+we are in particular not interested in modelling felicity with our
 semantics, only giving an accurate semantics when the input is
@@
 the usual accounts of the progressive in the
-literature. \Citet{ogihara2007tense} summarizes the position of
+literature. \Citet{ogihara2007tense} summarises the position of
 \citet{bennett1978toward} as follows: a progressive sentence is true
@@
 Because we do not account for pragmatics, we prefer to retain the
-simplest account based on $\constant{subsumption}$.
+simplest account based on the subinterval property (which we call
+$\constant{subsumption}$ here).
 
@@
 \verb!appTime!). This use of custom operators is useful for several
-generalizations (for example, we have quantifiers such as \verb!MOST! in
+generalisations (for example, we have quantifiers such as \verb!MOST! in
 addition to \verb!FORALL! --- see \ifanon (citation anonymized)
@@
 \end{verbatim}
-}This is very close to our idealized representation of the problem
+}This is very close to our idealised representation of the problem
 \cref{eq:smith-novel-1}. One difference is the use of abstract Coq
@@
     to the approach of \citet{Mineshima:2015}, ``NUT" to the CCG
-    system that utilizes the first-order automated theorem prover
+    system that utilises the first-order automated theorem prover
     \textit{nutcracker} \cite{bos:2008}, and ``Langpro"
@@
 suite, but excluded section 7, suggesting that it requires many
-examples that need an ad hoc treatment. Here, we take up this
+examples that need an \textit{ad hoc} treatment. Here, we take up this
 challenge and show that a system similar to theirs can be extended to
@@
 
-\section*{Acknowledgments}
+\section*{Acknowledgements}
 
@@
 
+% Local Variables:
+% ispell-local-dictionary: "british"
+% End:
+
 % LocalWords:  dowty steedman higginbotham fernando NLI FraCaS GF une
@@
 % LocalWords:  infelicitousness ranta Ljunglof montague english TODO
-% LocalWords:  prover chatzikyriakidis Coq's anonymized monadic unger
+% LocalWords:  prover chatzikyriakidis Coq's  monadic unger
 % LocalWords:  charlow anaphoroi stative activites timespan timespans
@@
 % LocalWords:  FraCoq Mineshima Langpro abzianidze premiss UNK Centre
+% LocalWords:  Stergios testsuite datasets maccartney compositional
+% LocalWords:  Montagovian montagovian coreference neo Davidsonian
+% LocalWords:  ogihara bennett iff subinterval APCOM existentials hoc
+% LocalWords:  androutsopoulos

### 2020-11-11 paper2020/paper.tex
anon
typography

@@
 \title{FraCaS: Temporal Analysis}
-\author{\authorname{Jean-Philippe Bernardy} \authorname{Stergios Chatzikyriakidis}
+\author{\ifanon Anonymous for review\else
+  \authorname{Jean-Philippe Bernardy} \authorname{Stergios Chatzikyriakidis}
   \affiliation{University of Gothenburg}
-  \email{firstname.lastname@gu.se}}
+  \email{firstname.lastname@gu.se}\fi}
 
@@
 In the first premise, our system essentially resolves the ellipsis to
-get the following reading: ``ITEL won X orders and APCOM won Y orders
-and X > Y.''. One would need each of the verb phrases ``won X orders''
-and ``won Y orders'' to introduce their own timespans with existential
+get the following reading: ``ITEL won \(X\) orders and APCOM won \(Y\) orders
+and $X > Y$.''. One would need each of the verb phrases ``won \(X\) orders''
+and ``won $Y$ orders'' to introduce their own timespans with existential
 quantifiers. However, the organisation of our system is such that the

### 2020-12-19 paper2020/paper.tex
minor_reiew

@@
 \title{FraCaS: Temporal Analysis}
-\author{\ifanon Anonymous for review\else
-  \authorname{Jean-Philippe Bernardy} \authorname{Stergios Chatzikyriakidis}
-  \affiliation{University of Gothenburg}
-  \email{firstname.lastname@gu.se}\fi}
+\author{\authorname{Jean-Philippe Bernardy} \authorname{Stergios Chatzikyriakidis}
+  \affiliation{CLASP, Deparment of Philosophy, Linguistics and Theory of Science, University of Gothenburg}
+  \email{firstname.lastname@gu.se}}
 
-\keywords{Temporal Semantics, Type-Theory}
+\keywords{Temporal Semantics, Type-Theory, Natural Language Inference, Formal Semantics, FraCaS Test Suite, Coq}
 
@@
 \begin{figure*}
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}{0.47\textwidth}
 
@@
 \end{subfigure}
-\begin{subfigure}{0.5\textwidth}
+\hfill
+\begin{subfigure}{0.47\textwidth}
 {
@@
 \providecommand\ncases[1]{{\ensuremath{^{#1}}}}
-\begin{table}[hbt]
+\begin{table}[t]
   \centering
@@
 
-\ifanon
-Anonymized
-\else
+%\ifanon
+%Anonymized
+%\else
 The research reported in this paper was supported by grant 2014-39 from the
@@
 earlier draft of the paper.
-\fi
+%\fi
 \bibliographystyle{apalike}

### 2020-12-19 paper2020/paper_2021.tex
IWCS_format

--- 2020-12-19 paper2020/paper.tex
@@
-\newif\ifanon\anontrue
-
-\documentclass[a4paper,twoside]{article}
+\documentclass[11pt,a4paper]{article}
+\usepackage[hyperref]{acl2020}
+\usepackage{times}
+\usepackage{latexsym}
+\renewcommand{\UrlFont}{\ttfamily\small}
 
@@
 \input{../PaperTools/latex/newunicodedefs}
-\usepackage{SCITEPRESS}     % Please add other packages that you may need BEFORE the SCITEPRESS.sty package.
+% \usepackage{SCITEPRESS}     % Please add other packages that you may need BEFORE the SCITEPRESS.sty package.
 
@@
 
+%\aclfinalcopy % Uncomment this line for the final submission
+%\def\aclpaperid{***} %  Enter the acl Paper ID here
+
+%\setlength\titlebox{5cm}
+% You can expand the titlebox if you need extra space
+% to show all the authors. Please do not make the titlebox
+% smaller than 5cm (the original size); we will check this
+% in the camera-ready version and ask you to change it back.
+
+\newcommand\BibTeX{B\textsc{ib}\TeX}
+
+\title{Temporal Semantics to Go: A Complete Run of the FraCaS Test Suite}
+
+\author{First Author \\
+	Affiliation / Address line 1 \\
+	Affiliation / Address line 2 \\
+	Affiliation / Address line 3 \\
+	\texttt{email@domain} \\\And
+	Second Author \\
+	Affiliation / Address line 1 \\
+	Affiliation / Address line 2 \\
+	Affiliation / Address line 3 \\
+	\texttt{email@domain} \\}
+
+\date{}
+
 \begin{document}
+	\maketitle
 
-\title{FraCaS: Temporal Analysis}
-\author{\authorname{Jean-Philippe Bernardy} \authorname{Stergios Chatzikyriakidis}
-  \affiliation{CLASP, Deparment of Philosophy, Linguistics and Theory of Science, University of Gothenburg}
-  \email{firstname.lastname@gu.se}}
+% \keywords{Temporal Semantics, Type-Theory}
 
-\keywords{Temporal Semantics, Type-Theory, Natural Language Inference, Formal Semantics, FraCaS Test Suite, Coq}
+\begin{abstract}In this paper, we propose an implementation of temporal semantics that
+  translates syntax trees to logical formulas, suitable for
+ consumption by the Coq proof assistant. The analysis  supports a wide range of phenomena
+ including: temporal references, temporal adverbs, aspectual classes
+ and progressives.
+The new semantics are built on top of a previous system handling all sections of the FraCaS test suite except the temporal reference section, and we  obtain an accuracy of 81 percent overall and 73 percent for the problems explicitly marked
+as related to temporal reference. To the best of our knowledge, this is the best performance of a logical system on the whole of the FraCaS. 
+\end{abstract}
 
-\abstract{In this paper, we propose an implementation of temporal semantics
-  which is suitable for inference problems. This implementation
-  translates syntax trees to logical formulas, suitable for
-  consumption by the Coq proof assistant. We support several phenomena
-  including: temporal references, temporal adverbs, aspectual classes
-  and progressives.  We apply these semantics to the complete FraCaS testsuite.
-  We obtain an accuracy of 81 percent overall and 73 percent for problems explicitly marked
-  as related to temporal reference.
-}
-
-\onecolumn \maketitle \normalsize \setcounter{footnote}{0} \vfill
+% \onecolumn \maketitle \normalsize \setcounter{footnote}{0} \vfill
 
@@
 \citep{unger_dynamic_2011,charlow_monadic_2015,charlow_modular_2017}.
-Our particular implementation does not handle every case \ifanon
-(citations anonymized) \else \citet{bernardy_computational_2020,bernardy_wide-coverage_2019} \fi
+Our particular implementation does not handle every case 
+(citations anonymized) \citet{bernardy_computational_2020,bernardy_wide-coverage_2019} \fi
 but non-temporal anaphoroi in the testsuite are generally resolved as
@@
 \citet{montague_english_1970,montague_proper_1973}, but we direct the
-reader to our previous work for details \ifanon (citations anonymized)
-\else \citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019} \fi, but the particular
+reader to our previous work for details  (citations anonymized)
+\citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019}, but the particular
 treatment of such phenomena is essentially independent from our
@@
 \citet{unger_dynamic_2011}, and our implemented analysis of
-anaphora\ifanon (citation anonymized)
-\else\citep{bernardy_computational_2018,bernardy_wide-coverage_2019}\fi.
+anaphora (citation anonymized)
+\citep{bernardy_computational_2018,bernardy_wide-coverage_2019}.
 
@@
 generalisations (for example, we have quantifiers such as \verb!MOST! in
-addition to \verb!FORALL! --- see \ifanon (citation anonymized)
-\else \citet{bernardy_type-theoretical_2017} \fi)
+addition to \verb!FORALL! --- see (citation anonymized)
+ \citet{bernardy_type-theoretical_2017} )
 
@@
 which arise in the FraCaS testsuite.
-
+\iffalse
 \begin{figure*}
@@
 \end{figure*}
-
+\fi
 \section{Results and Evaluation}
@@
 \providecommand\ncases[1]{{\ensuremath{^{#1}}}}
-\begin{table}[t]
+\begin{table}[hbt]
   \centering
@@
 \end{table}
-Additionally, to provide the reader with a better sense of what the
-complete system is capable and not capable of, we list a number of
-correctly and incorrectly handled examples in \cref{fig:correct-examples} and \cref{fig:incorrect-examples}.
+%Additionally, to provide the reader with a better sense of what the
+%complete system is capable and not capable of, we list a number of
+%correctly and incorrectly handled examples in \cref{fig:correct-examples} and \cref{fig:incorrect-examples}.
 \paragraph{Interaction with anaphora}
@@
 with the temporal section of the FraCaS test suite. To do this, we
-have provided a new, simplified taxonomy of aspectual classes for verb
+have provided a simplified taxonomy of aspectual classes for verb
 phrases, guided by the applicability of the unicity of action and
@@
 \cite{bernardy_wide-coverage_2019} presented a
-logical system for handling 8 of the 9 sections of the FRACAS test
+logical system for handling 8 of the 9 sections of the FraCaS test
 suite, but excluded section 7, suggesting that it requires many
-examples that need an \textit{ad hoc} treatment. Here, we take up this
-challenge and show that a system similar to theirs can be extended to
-cover the remaining section of the test suite.
+examples that need an \textit{ad hoc} treatment. Here, we took up this
+challenge and have shown that a system similar to theirs can be extended to
+cover the remaining section of the test suite, without considerably decreasing the performance of the 
+rest of the sections.
+This is indeed a common problem with logical approaches, namely the fact that one can have theoretically motivated implementations of individual phenomena, e.g.
+anaphora, ellipsis, quantifiers, temporal reference etc., but when one tries to put all these together into a unified system, this proves to be a daunting task.
+We believe that this paper presents an exception, and provides a system that can deal with all these different semantic phenomena under a unified system with very good results.
 %
-We use the same combination of a number of well-studied tools: type
+We use the same combination of a number of well-studied tools as \cite{bernardy_wide-coverage_2019} : type
 theory, parsing using the Grammatical Framework (GF), Monadic Dynamic
@@
 
-\section*{Acknowledgements}
-
-%\ifanon
-%Anonymized
-%\else
+\iffalse\section*{Acknowledgements}
 The research reported in this paper was supported by grant 2014-39 from the
@@
 here. We also thank anonymous reviewers for their useful comments on an
-earlier draft of the paper.
-%\fi
-\bibliographystyle{apalike}
+earlier draft of the paper.\fi
+
+\bibliographystyle{acl_natbib}
 

### 2021-02-01 paper2020/acl2021.tex
add acl style

%
% File acl2021.tex
%
%% Based on the style files for EMNLP 2020, which were
%% Based on the style files for ACL 2020, which were
%% Based on the style files for ACL 2018, NAACL 2018/19, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2021}
\usepackage{times}
\usepackage{latexsym}
\renewcommand{\UrlFont}{\ttfamily\small}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B\textsc{ib}\TeX}

\title{Instructions for IWCS 2021 Proceedings}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  \texttt{email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  \texttt{email@domain} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
This document contains the instructions for preparing a manuscript for the proceedings of ACL-IJCNLP 2021.
The document itself conforms to its own specifications, and is therefore an example of what your manuscript should look like.
These instructions should be used for both papers submitted for review and for final versions of accepted papers.
Authors are asked to conform to all the directions reported in this document.
\end{abstract}

\section{Credits}

This document has been adapted by Roberto Navigli
from the instructions for earlier ACL, NAACL and EMNLP proceedings, including those for 
EMNLP 2020 by Yulan He,
ACL 2020 by Steven Bethard, Ryan Cotterrell and Rui Yan, 
ACL 2019 by Douwe Kiela and Ivan Vuli\'{c},
NAACL 2019 by Stephanie Lukin and Alla Roskovskaya, 
ACL 2018 by Shay Cohen, Kevin Gimpel, and Wei Lu, 
NAACL 2018 by Margaret Michell and Stephanie Lukin,
2017/2018 (NA)ACL bibtex suggestions from Jason Eisner,
ACL 2017 by Dan Gildea and Min-Yen Kan, 
NAACL 2017 by Margaret Mitchell, 
ACL 2012 by Maggie Li and Michael White, 
ACL 2010 by Jing-Shing Chang and Philipp Koehn, 
ACL 2008 by Johanna D. Moore, Simone Teufel, James Allan, and Sadaoki Furui, 
ACL 2005 by Hwee Tou Ng and Kemal Oflazer, 
ACL 2002 by Eugene Charniak and Dekang Lin, 
and earlier ACL and EACL formats written by several people, including
John Chen, Henry S. Thompson and Donald Walker.
Additional elements were taken from the formatting instructions of the \emph{International Joint Conference on Artificial Intelligence} and the \emph{Conference on Computer Vision and Pattern Recognition}.

\section{Introduction}

The following instructions are directed to authors of papers submitted to ACL-IJCNLP 2021 or accepted for publication in its proceedings.
All authors are required to adhere to these specifications.
Authors are required to provide a Portable Document Format (PDF) version of their papers.
\textbf{The proceedings are designed for printing on A4 paper.}

\section{Electronically-available resources}

ACL provides this description and accompanying style files at
\begin{quote}
\url{https://2021.aclweb.org/files/acl-ijcnlp2021-templates.zip}
\end{quote}
We strongly recommend the use of these style files, which have been appropriately tailored for the ACL-IJCNLP 2021 proceedings. You can also start from the Overleaf template linked from the conference website.

\paragraph{\LaTeX-specific details:}
The templates include the \LaTeX2e{} source (\texttt{\small acl2021.tex}),
the \LaTeX2e{} style file used to format it (\texttt{\small acl2021.sty}),
an ACL bibliography style (\texttt{\small acl\_natbib.bst}),
an example bibliography (\texttt{\small acl2021.bib}),
and the bibliography for the ACL Anthology (\texttt{\small anthology.bib}).

\section{Length of Submission}
\label{sec:length}

The conference accepts submissions of long papers and short papers. Long papers may consist of up to eight (8) pages of content plus unlimited pages for references and ethics/broader impact statement. Short papers may consist of up to four (4) pages of content, plus unlimited pages for references and ethics/broader impact statement. Upon acceptance, final versions of both long and short papers will be given one additional content page to address the reviewers' comments.

For both long and short papers, all illustrations and tables that are part of the main text must be accommodated within these page limits, observing the formatting instructions given in the present document. 
\textbf{Papers that do not conform to the specified length and formatting requirements are subject to be rejected without review.}

The conference encourages the submission of additional material that is relevant to the reviewers but not an integral part of the paper.
There are two such types of material: appendices, which can be read, and non-readable supplementary materials, often data or code. 
Additional material must be submitted as separate files, and must adhere to the same anonymity guidelines as the main paper. Appendices should be no longer than 4 pages.
The paper must be self-contained: it is optional for reviewers to look at the supplementary material.
The paper should not refer, for further detail, to documents, code or data resources that are not available to the reviewers.
Refer to Section~\ref{sec:supplementary} for further information.

Workshop chairs may have different rules for allowed length and whether supplemental material is welcome.
As always, the respective call for papers is the authoritative source.

\section{Anonymity}
As reviewing will be double-blind, papers submitted for review should not include any author information (such as names or affiliations). Furthermore, self-references that reveal the author's identity, \emph{e.g.},
\begin{quote}
We previously showed \citep{Gusfield:97} \ldots
\end{quote}
should be avoided. Instead, use citations such as 
\begin{quote}
\citet{Gusfield:97} previously showed\ldots
\end{quote}
Please do not use anonymous citations and do not include acknowledgements.
\textbf{Papers that do not conform to these requirements may be rejected without review.}

Any preliminary non-archival versions of submitted papers should be listed in the submission form but not in the review version of the paper.
Reviewers are generally aware that authors may present preliminary versions of their work in other venues, but will not be provided the list of previous presentations from the submission form.

Please do not include github links that reveal the authors' identities in your submission. If you feel it is important to include your source code, you can zip the code and submit it to softconf.

Once a paper has been accepted to the conference, the camera-ready version of the paper should include the author's names and affiliations, and is allowed to use self-references and provide the related github link.

\paragraph{\LaTeX-specific details:}
For an anonymized submission, ensure that {\small\verb|\aclfinalcopy|} at the top of this document is commented out, and that you have filled in the paper ID number (assigned during the submission process on softconf) where {\small\verb|***|} appears in the {\small\verb|\def\aclpaperid{***}|} definition at the top of this document.
For a camera-ready submission, ensure that {\small\verb|\aclfinalcopy|} at the top of this document is not commented out.

\section{Multiple Submission Policy}

ACL-IJCNLP 2021 will not consider any paper that is under review in a journal or another conference at the time of submission, and submitted papers must not be submitted elsewhere during the ACL-IJCNLP 2021 review period. This policy covers all refereed and archival conferences and workshops (e.g., NAACL). The only exception is that a paper can be dual-submitted to both ACL-IJCNLP 2021 and an ACL-IJCNLP workshop.

In addition, we will not consider any paper that overlaps significantly in content or results with papers that will be (or have been) published elsewhere. 
Authors submitting more than one paper to ACL-IJCNLP 2021 must ensure that their submissions do not overlap significantly ($>25$\%) with each other in content or results.

\section{Formatting Instructions}

Manuscripts must be in two-column format.
Exceptions to the two-column format include the title, authors' names and complete addresses, which must be centered at the top of the first page, and any full-width figures or tables (see the guidelines in Section~\ref{ssec:title-authors}).
\textbf{Type single-spaced.}
Start all pages directly under the top margin.
The manuscript should be printed single-sided and its length should not exceed the maximum page limit described in Section~\ref{sec:length}.
Pages should be numbered in the version submitted for review, but \textbf{pages should not be numbered in the camera-ready version}.

\paragraph{\LaTeX-specific details:}
The style files will generate page numbers when {\small\verb|\aclfinalcopy|} is commented out, and remove them otherwise.

\subsection{File Format}
\label{sect:pdf}

For the production of the electronic manuscript you must use Adobe's Portable Document Format (PDF).
Please make sure that your PDF file includes all the necessary fonts (especially tree diagrams, symbols, and fonts with Asian characters).
When you print or create the PDF file, there is usually an option in your printer setup to include none, all or just non-standard fonts.
Please make sure that you select the option of including ALL the fonts.
\textbf{Before sending it, test your PDF by printing it from a computer different from the one where it was created.}
Moreover, some word processors may generate very large PDF files, where each page is rendered as an image.
Such images may reproduce poorly.
In this case, try alternative ways to obtain the PDF.
One way on some systems is to install a driver for a postscript printer, send your document to the printer specifying ``Output to a file'', then convert the file to PDF.

It is of utmost importance to specify the \textbf{A4 format} (21 cm x 29.7 cm) when formatting the paper.
Print-outs of the PDF file on A4 paper should be identical to the hardcopy version.
If you cannot meet the above requirements about the production of your electronic submission, please contact the publication chairs as soon as possible.

\paragraph{\LaTeX-specific details:}
PDF files are usually produced from \LaTeX{} using the \texttt{\small pdflatex} command.
If your version of \LaTeX{} produces Postscript files, \texttt{\small ps2pdf} or \texttt{\small dvipdf} can convert these to PDF.
To ensure A4 format in \LaTeX, use the command {\small\verb|\special{papersize=210mm,297mm}|}
in the \LaTeX{} preamble (below the {\small\verb|\usepackage|} commands) and use \texttt{\small dvipdf} and/or \texttt{\small pdflatex}; or specify \texttt{\small -t a4} when working with \texttt{\small dvips}.

\subsection{Layout}
\label{ssec:layout}

Format manuscripts two columns to a page, in the manner these
instructions are formatted.
The exact dimensions for a page on A4 paper are:

\begin{itemize}
\item Left and right margins: 2.5 cm
\item Top margin: 2.5 cm
\item Bottom margin: 2.5 cm
\item Column width: 7.7 cm
\item Column height: 24.7 cm
\item Gap between columns: 0.6 cm
\end{itemize}

\noindent Papers should not be submitted on any other paper size.
If you cannot meet the above requirements about the production of your electronic submission, please contact the publication chairs above as soon as possible.

\subsection{Fonts}

For reasons of uniformity, Adobe's \textbf{Times Roman} font should be used.
If Times Roman is unavailable, you may use Times New Roman or \textbf{Computer Modern Roman}.

Table~\ref{font-table} specifies what font sizes and styles must be used for each type of text in the manuscript.

\begin{table}
\centering
\begin{tabular}{lrl}
\hline \textbf{Type of Text} & \textbf{Font Size} & \textbf{Style} \\ \hline
paper title & 15 pt & bold \\
author names & 12 pt & bold \\
author affiliation & 12 pt & \\
the word ``Abstract'' & 12 pt & bold \\
section titles & 12 pt & bold \\
subsection titles & 11 pt & bold \\
document text & 11 pt  &\\
captions & 10 pt & \\
abstract text & 10 pt & \\
bibliography & 10 pt & \\
footnotes & 9 pt & \\
\hline
\end{tabular}
\caption{\label{font-table} Font guide. }
\end{table}

\paragraph{\LaTeX-specific details:}
To use Times Roman in \LaTeX2e{}, put the following in the preamble:
\begin{quote}
\small
\begin{verbatim}
\usepackage{times}
\usepackage{latexsym}
\end{verbatim}
\end{quote}

\subsection{Ruler}
A printed ruler (line numbers in the left and right margins of the article) should be presented in the version submitted for review, so that reviewers may comment on particular lines in the paper without circumlocution.
The presence or absence of the ruler should not change the appearance of any other content on the page.
The camera ready copy should not contain a ruler.

\paragraph{Reviewers:}
note that the ruler measurements may not align well with lines in the paper -- this turns out to be very difficult to do well when the paper contains many figures and equations, and, when done, looks ugly.
In most cases one would expect that the approximate location will be adequate, although you can also use fractional references (\emph{e.g.}, this line ends at mark $295.5$).

\paragraph{\LaTeX-specific details:}
The style files will generate the ruler when {\small\verb|\aclfinalcopy|} is commented out, and remove it otherwise.

\subsection{Title and Authors}
\label{ssec:title-authors}

Center the title, author's name(s) and affiliation(s) across both columns.
Do not use footnotes for affiliations.
Place the title centered at the top of the first page, in a 15-point bold font.
Long titles should be typed on two lines without a blank line intervening.
Put the title 2.5 cm from the top of the page, followed by a blank line, then the author's names(s), and the affiliation on the following line.
Do not use only initials for given names (middle initials are allowed).
Do not format surnames in all capitals (\emph{e.g.}, use ``Mitchell'' not ``MITCHELL'').
Do not format title and section headings in all capitals except for proper names (such as ``BLEU'') that are
conventionally in all capitals.
The affiliation should contain the author's complete address, and if possible, an electronic mail address.

The title, author names and addresses should be completely identical to those entered to the electronical paper submission website in order to maintain the consistency of author information among all publications of the conference.
If they are different, the publication chairs may resolve the difference without consulting with you; so it is in your own interest to double-check that the information is consistent.

Start the body of the first page 7.5 cm from the top of the page.
\textbf{Even in the anonymous version of the paper, you should maintain space for names and addresses so that they will fit in the final (accepted) version.}

\subsection{Abstract}
Use two-column format when you begin the abstract.
Type the abstract at the beginning of the first column.
The width of the abstract text should be smaller than the
width of the columns for the text in the body of the paper by 0.6 cm on each side.
Center the word \textbf{Abstract} in a 12 point bold font above the body of the abstract.
The abstract should be a concise summary of the general thesis and conclusions of the paper.
It should be no longer than 200 words.
The abstract text should be in 10 point font.

\subsection{Text}
Begin typing the main body of the text immediately after the abstract, observing the two-column format as shown in the present document.

Indent 0.4 cm when starting a new paragraph.

\subsection{Sections}

Format section and subsection headings in the style shown on the present document.
Use numbered sections (Arabic numerals) to facilitate cross references.
Number subsections with the section number and the subsection number separated by a dot, in Arabic numerals.

\subsection{Footnotes}
Put footnotes at the bottom of the page and use 9 point font.
They may be numbered or referred to by asterisks or other symbols.\footnote{This is how a footnote should appear.}
Footnotes should be separated from the text by a line.\footnote{Note the line separating the footnotes from the text.}

\subsection{Graphics}

Place figures, tables, and photographs in the paper near where they are first discussed, rather than at the end, if possible.
Wide illustrations may run across both columns.
Color is allowed, but adhere to Section~\ref{ssec:accessibility}'s guidelines on accessibility.

\paragraph{Captions:}
Provide a caption for every illustration; number each one sequentially in the form:
``Figure 1. Caption of the Figure.''
``Table 1. Caption of the Table.''
Type the captions of the figures and tables below the body, using 10 point text.
Captions should be placed below illustrations.
Captions that are one line are centered (see Table~\ref{font-table}).
Captions longer than one line are left-aligned (see Table~\ref{tab:accents}).

\begin{table}
\centering
\begin{tabular}{lc}
\hline
\textbf{Command} & \textbf{Output}\\
\hline
\verb|{\"a}| & {\"a} \\
\verb|{\^e}| & {\^e} \\
\verb|{\`i}| & {\`i} \\ 
\verb|{\.I}| & {\.I} \\ 
\verb|{\o}| & {\o} \\
\verb|{\'u}| & {\'u}  \\ 
\verb|{\aa}| & {\aa}  \\\hline
\end{tabular}
\begin{tabular}{lc}
\hline
\textbf{Command} & \textbf{Output}\\
\hline
\verb|{\c c}| & {\c c} \\ 
\verb|{\u g}| & {\u g} \\ 
\verb|{\l}| & {\l} \\ 
\verb|{\~n}| & {\~n} \\ 
\verb|{\H o}| & {\H o} \\ 
\verb|{\v r}| & {\v r} \\ 
\verb|{\ss}| & {\ss} \\
\hline
\end{tabular}
\caption{Example commands for accented characters, to be used in, \emph{e.g.}, \BibTeX\ names.}\label{tab:accents}
\end{table}

\paragraph{\LaTeX-specific details:}
The style files are compatible with the caption and subcaption packages; do not add optional arguments.
\textbf{Do not override the default caption sizes.}

\subsection{Hyperlinks}
Within-document and external hyperlinks are indicated with Dark Blue text, Color Hex \#000099.

\subsection{Citations}
Citations within the text appear in parentheses as~\citep{Gusfield:97} or, if the author's name appears in the text itself, as \citet{Gusfield:97}.
Append lowercase letters to the year in cases of ambiguities.  
Treat double authors as in~\citep{Aho:72}, but write as in~\citep{Chandra:81} when more than two authors are involved. Collapse multiple citations as in~\citep{Gusfield:97,Aho:72}.

Refrain from using full citations as sentence constituents.
Instead of
\begin{quote}
  ``\citep{Gusfield:97} showed that ...''
\end{quote}
write
\begin{quote}
``\citet{Gusfield:97} showed that ...''
\end{quote}

\begin{table*}
\centering
\begin{tabular}{lll}
\hline
\textbf{Output} & \textbf{natbib command} & \textbf{Old ACL-style command}\\
\hline
\citep{Gusfield:97} & \small\verb|\citep| & \small\verb|\cite| \\
\citealp{Gusfield:97} & \small\verb|\citealp| & no equivalent \\
\citet{Gusfield:97} & \small\verb|\citet| & \small\verb|\newcite| \\
\citeyearpar{Gusfield:97} & \small\verb|\citeyearpar| & \small\verb|\shortcite| \\
\hline
\end{tabular}
\caption{\label{citation-guide}
Citation commands supported by the style file.
The style is based on the natbib package and supports all natbib citation commands.
It also supports commands defined in previous ACL style files for compatibility.
}
\end{table*}

\paragraph{\LaTeX-specific details:}
Table~\ref{citation-guide} shows the syntax supported by the style files.
We encourage you to use the natbib styles.
You can use the command {\small\verb|\citet|} (cite in text) to get ``author (year)'' citations as in \citet{Gusfield:97}.
You can use the command {\small\verb|\citep|} (cite in parentheses) to get ``(author, year)'' citations as in \citep{Gusfield:97}.
You can use the command {\small\verb|\citealp|} (alternative cite without  parentheses) to get ``author year'' citations (which is useful for  using citations within parentheses, as in \citealp{Gusfield:97}).

\subsection{References}
Gather the full set of references together under the heading \textbf{References}; place the section before any Appendices. 
Arrange the references alphabetically by first author, rather than by order of occurrence in the text.

Provide as complete a citation as possible, using a consistent format, such as the one for \emph{Computational Linguistics\/} or the one in the  \emph{Publication Manual of the American 
Psychological Association\/}~\citep{APA:83}.
Use full names for authors, not just initials.

Submissions should accurately reference prior and related work, including code and data.
If a piece of prior work appeared in multiple venues, the version that appeared in a refereed, archival venue should be referenced.
If multiple versions of a piece of prior work exist, the one used by the authors should be referenced.
Authors should not rely on automated citation indices to provide accurate references for prior and related work.

The following text cites various types of articles so that the references section of the present document will include them.
\begin{itemize}
\item Example article in journal: \citep{Ando2005}.
\item Example article in proceedings, with location: \citep{borschinger-johnson-2011-particle}.
\item Example article in proceedings, without location: \citep{andrew2007scalable}.
\item Example arxiv paper: \citep{rasooli-tetrault-2015}. 
\end{itemize}

\paragraph{\LaTeX-specific details:}
The \LaTeX{} and Bib\TeX{} style files provided roughly follow the American Psychological Association format.
If your own bib file is named \texttt{\small acl2021.bib}, then placing the following before any appendices in your \LaTeX{}  file will generate the references section for you:
\begin{quote}\small
\verb|\bibliographystyle{acl_natbib}|\\
\verb|\bibliography{acl2021}|
\end{quote}

You can obtain the complete ACL Anthology as a Bib\TeX\ file from \url{https://aclweb.org/anthology/anthology.bib.gz}.
To include both the anthology and your own bib file, use the following instead of the above.
\begin{quote}\small
\verb|\bibliographystyle{acl_natbib}|\\
\verb|\bibliography{anthology,acl2021}|
\end{quote}

\subsection{Digital Object Identifiers}
As part of our work to make ACL materials more widely used and cited outside of our discipline, ACL has registered as a CrossRef member, as a registrant of Digital Object Identifiers (DOIs), the standard for registering permanent URNs for referencing scholarly materials.

All camera-ready references are required to contain the appropriate DOIs (or, as a second resort, the hyperlinked ACL Anthology Identifier) to all cited works.
Appropriate records should be found for most materials in the current ACL Anthology at \url{http://aclanthology.info/}.
As examples, we cite \citep{goodman-etal-2016-noise} to show you how papers with a DOI will appear in the bibliography.
We cite \citep{harper-2014-learning} to show how papers without a DOI but with an ACL Anthology Identifier will appear in the bibliography.

\paragraph{\LaTeX-specific details:}
Please ensure that you use Bib\TeX\ records that contain DOI or URLs for any of the ACL materials that you reference.
If the Bib\TeX{} file contains DOI fields, the paper title in the references section will appear as a hyperlink to the DOI, using the hyperref \LaTeX{} package.

\section{Supplementary Materials}
\label{sec:supplementary}

Supplementary material may report preprocessing decisions, model parameters, and other details necessary for the replication of the experiments reported in the paper.
Seemingly small preprocessing decisions can sometimes make a large difference in performance, so it is crucial to record such decisions to precisely characterize state-of-the-art methods.

Nonetheless, supplementary material should be supplementary (rather than central) to the paper.
\textbf{Submissions that misuse the supplementary material may be rejected without review.}
Supplementary material may include explanations or details of proofs or derivations that do not fit into the paper, lists of
features or feature templates, sample inputs and outputs for a system, hyperparameter values, pseudo-code or source code, and data.

The paper should not rely on the supplementary material: while the paper may refer to and cite the supplementary material and the supplementary material will be available to the reviewers, they will not be asked to review the supplementary material.

Supplementary material can be uploaded as up to three separate files: a single appendix file in PDF format, a single .tgz or .zip archive containing software, and/or a single .tgz or .zip archive containing data.

\subsection{Appendices}
\label{sec:appendix}
Appendices are material that can be read, and include lemmas, formulas, proofs, and tables that are not critical to the reading and understanding of the paper. 
In the submitted version, appendices should be uploaded as a separate file. \textbf{Submissions which include appendices in the main paper will be rejected without review.} In the camera ready version, instead, appendices should directly follow the text and the references.

\paragraph{\LaTeX-specific details:}
In your camera ready version use {\small\verb|\appendix|} before any appendix section to switch the section numbering over to letters.

\subsection{Software and Data}
\label{sec:software_and_data}
Submissions may include software and data used in the work and described in the paper.
Any accompanying software and/or data should include licenses and documentation of research review as appropriate.

\section{Accessibility}
\label{ssec:accessibility}

In an effort to accommodate people who are color-blind (as well as those printing to paper), grayscale readability is strongly encouraged.
Color is not forbidden, but authors should ensure that tables and figures do not rely solely on color to convey critical distinctions.
A simple criterion:
All curves and points in your figures should be clearly distinguishable without color.

\section{Translation of non-English Terms}

It is also advised to supplement non-English characters and terms with appropriate transliterations and/or translations since not all readers understand all such characters and terms.
Inline transliteration or translation can be represented in the order of:
\begin{center}
\begin{tabular}{c}
original-form \\
transliteration \\
``translation''
\end{tabular}
\end{center}

\section{\LaTeX{} Compilation Issues}
You may encounter the following error during compilation: 
\begin{quote}
{\small\verb|\pdfendlink|} ended up in different nesting level than {\small\verb|\pdfstartlink|}.
\end{quote}
This happens when \texttt{\small pdflatex} is used and a citation splits across a page boundary.
To fix this, the style file contains a patch consisting of two lines:
(1) {\small\verb|\RequirePackage{etoolbox}|} (line 455 in \texttt{\small acl2021.sty}), and
(2) A long line below (line 456 in \texttt{\small acl2021.sty}).

If you still encounter compilation issues even with the patch enabled, disable the patch by commenting the two lines, and then disable the \texttt{\small hyperref} package by loading the style file with the \texttt{\small nohyperref} option:

\noindent
{\small\verb|\usepackage[nohyperref]{acl2021}|}

\noindent
Then recompile, find the problematic citation, and rewrite the sentence containing the citation. (See, {\em e.g.}, \url{http://tug.org/errors.html})

\section*{Acknowledgments}

The acknowledgments should go immediately before the references. Do not number the acknowledgments section.
\textbf{Do not include this section when submitting your paper for review.}

\bibliographystyle{acl_natbib}
\bibliography{anthology,acl2021}

%\appendix

\end{document}

### 2021-02-01 paper2020/paper.tex
minor_reiew
remove scitepress
typography
Use IWCS style files

--- 2020-11-11 paper2020/paper.tex
@@
 
-\documentclass[a4paper,twoside]{article}
+\documentclass[11pt,a4paper]{article}
+\usepackage[hyperref]{acl2021}
+\usepackage{times}
+\usepackage{latexsym}
+\renewcommand{\UrlFont}{\ttfamily\small}
+
+% This is not strictly necessary, and may be commented out,
+% but it will improve the layout of the manuscript,
+% and will typically save some space.
+\usepackage{microtype}
+
+%\aclfinalcopy % Uncomment this line for the final submission
+%\def\aclpaperid{***} %  Enter the acl Paper ID here
+
+%\setlength\titlebox{5cm}
+% You can expand the titlebox if you need extra space
+% to show all the authors. Please do not make the titlebox
+% smaller than 5cm (the original size); we will check this
+% in the camera-ready version and ask you to change it back.
 
@@
 \title{FraCaS: Temporal Analysis}
+% \author{\ifanon Anonymous for review\else
+%   \authorname{Jean-Philippe Bernardy} \authorname{Stergios Chatzikyriakidis}
+%   \affiliation{University of Gothenburg}
+%   \email{firstname.lastname@gu.se}\fi}
+
 \author{\ifanon Anonymous for review\else
-  \authorname{Jean-Philippe Bernardy} \authorname{Stergios Chatzikyriakidis}
-  \affiliation{University of Gothenburg}
-  \email{firstname.lastname@gu.se}\fi}
+  {Jean-Philippe Bernardy}\footnote{University of Gothenburg} ~~ {Stergios Chatzikyriakidis}\footnote{University of Gothenburg}
+}
 
-\keywords{Temporal Semantics, Type-Theory}
+\maketitle
 
-\abstract{In this paper, we propose an implementation of temporal semantics
+% \keywords{Temporal Semantics, Type-Theory}
+
+\begin{abstract}
+In this paper, we propose an implementation of temporal semantics
   which is suitable for inference problems. This implementation
@@
   as related to temporal reference.
-}
+\end{abstract}
 
-\onecolumn \maketitle \normalsize \setcounter{footnote}{0} \vfill
+% \onecolumn \maketitle \normalsize \setcounter{footnote}{0} \vfill
 
@@
 \onelingex{A scandinavian won the nobel prize.\label{ex:nobel}
-\\
 {\small\begin{multline*}
@@
   constraint, for ``in 1992''.
-
+%
   \begin{multline*}
@@
 propositions:
-
+%
 \begin{multline}
@@
 \textit{A contrario}, if \(P\) is \emph{stative}, then we get a time-interval subsumption property:
-
 \begin{multline*}
@@
 action:
-
 \begin{multline*}
@@
 
-\begin{figure*}
-\begin{subfigure}{0.5\textwidth}
-
-\fracasex{252}{\answer yes
-\item	Since 1992 ITEL has been in Birmingham.
-\item	It is now 1996.
-\hyp    Itel was in Birmingham in 1993.
-} \fracasex{260	}{\answer yes
-\item	Yesterday APCOM signed the contract.
-\item	Today is Saturday, July 14th.
-\hyp    APCOM signed the contract Friday, 13th.
-
-} \fracasex{262	}{\answer yes
-\item	Smith left after Jones left.
-\item	Jones left after Anderson left.
-\hyp    Smith left after Anderson left.
-
-} \fracasex{312	}{\answer yes
-\item	ITEL always delivers reports late.
-\item	In 1993 ITEL delivered reports.
-\hyp    ITEL delivered reports late in 1993.
-
-} \fracasex{320	}{\answer yes
-\item	When Jones got his job at the CIA, he knew that he would never be allowed to write his memoirs.
-\hyp    It is the case that Jones is not and will never be allowed to write his memoirs.
-}
-
-  \caption{Selected correctly handled examples}
-  \label{fig:correct-examples}
-\end{subfigure}
-\begin{subfigure}{0.5\textwidth}
-{
-} \fracasex{259	}{\answer yes
-\item	The conference started on July 4th, 1994.
-\item	It lasted 2 days.
-\hyp    The conference was over on July 8th, 1994.
-\failsf Incorrect lexical semantics for ``last $n$ days''.
-} \fracasex{294	}{\answer yes
-\item	Smith was running his own business in two years.
-\hyp    Smith ran his own business.
-\failsf Parsing error for ``in two years''
-} \fracasex{317	}{\answer yes
-\item	Every representative has read this report.
-\item	No two representatives have read it at the same time.
-\item	No representative took less than half a day to read the report.
-\item	There are sixteen representatives.
-\hyp    It took the representatives more than a week to read the report.
-\failsf Counting semantics not implemented
-% } \fracasex{318	}{\answer no
-% \item	While Jones was updating the program, Mary came in and told him about the board meeting.
-% \item	She finished before he did.
-% \hyp    Mary's story lasted as long as Jones's updating the program.
-% \failsf Incorrect interaction with anaphora
-}
-  \caption{Selected incorrectly handled examples.
-  We list below a number of selected examples from the temporal section
-of the FraCaS testsuite where our semantics fails to list the correct
-inference. We list the reason why it fails under the item \textbf{F}.
-}
-  \label{fig:incorrect-examples}
-\end{subfigure}
-% \caption{Selected examples.}
-\end{figure*}
-
 \section{Results and Evaluation}
@@
 
+\ifanon
+
+\else
 \section*{Acknowledgements}
-
-\ifanon
-Anonymized
-\else
 The research reported in this paper was supported by grant 2014-39 from the

### 2021-02-01 paper2020/selected-examples.tex
Use IWCS style files

--- 2020-12-19 paper2020/paper.tex
@@
-\newif\ifanon\anontrue
-
-\documentclass[a4paper,twoside]{article}
-
-\usepackage{subcaption}
-% \usepackage{calc}
-\usepackage{natbib}
-\usepackage{tabularx}
-\usepackage{times}
-\usepackage{amssymb}
-\usepackage{amstext}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{stmaryrd}
-\usepackage{todonotes}
-\usepackage{enumitem}
-\usepackage{multicol}
-\usepackage{url}
-\usepackage[capitalize]{cleveref}
-\usepackage[verbose]{newunicodechar}
-\input{../PaperTools/latex/newunicodedefs}
-\usepackage{SCITEPRESS}     % Please add other packages that you may need BEFORE the SCITEPRESS.sty package.
-
-\newlist{lingex}{enumerate}{3}
-\setlist[lingex,1]{parsep=0pt,itemsep=1pt,label=(\arabic*),resume=lingexcount}
-\newcommand\onelingex[1]{\begin{lingex}\item #1 \end{lingex}}
-
-\newlist{subex}{enumerate}{3}
-\setlist[subex,1]{parsep=0pt,itemsep=1pt,font=\bfseries,label=P\arabic*}
-\newcommand\hyp{\item[H]}
-\newcommand\answer{\item[A]}
-\newcommand\failsf{\item[F]}
-\newcommand\fracasex[2]{\begin{lingex}\item[(#1)] \begin{subex} #2 \end{subex} \end{lingex} }
-\newcommand\jp[1]{\todo[backgroundcolor=blue!10]{JP: #1}}
-
-\newcommand\constant[1]{\mathsf{#1}}
-
-\begin{document}
-
-\title{FraCaS: Temporal Analysis}
-\author{\authorname{Jean-Philippe Bernardy} \authorname{Stergios Chatzikyriakidis}
-  \affiliation{CLASP, Deparment of Philosophy, Linguistics and Theory of Science, University of Gothenburg}
-  \email{firstname.lastname@gu.se}}
-
-\keywords{Temporal Semantics, Type-Theory, Natural Language Inference, Formal Semantics, FraCaS Test Suite, Coq}
-
-\abstract{In this paper, we propose an implementation of temporal semantics
-  which is suitable for inference problems. This implementation
-  translates syntax trees to logical formulas, suitable for
-  consumption by the Coq proof assistant. We support several phenomena
-  including: temporal references, temporal adverbs, aspectual classes
-  and progressives.  We apply these semantics to the complete FraCaS testsuite.
-  We obtain an accuracy of 81 percent overall and 73 percent for problems explicitly marked
-  as related to temporal reference.
-}
-
-\onecolumn \maketitle \normalsize \setcounter{footnote}{0} \vfill
-
-\section{Introduction}
-\label{sec:introduction}
-
-The semantics of tense and aspect has been a long standing issue in
-the study of formal semantics since the early days of Montague Grammar
-and a number of different ideas have been put forth on the table to
-deal with them throughout the years. Recent proposals include the
-works of the following authors:
-\citet{dowty:2012,prior:2003,steedman_productions_2000,higginbotham:2009,fernando:2015}.
-%
-The
-semantics of tense and aspect have been also considered in the study of
-Natural Language Inference (NLI). The various datasets
-for NLI that have been proposed by the years contain examples that
-have some implicit or explicit reliance on inferences related to tense
-and aspect. One of the early datasets used to test logical approaches,
-the FraCaS test suite~\citep{cooper:1996} contains a whole section
-dedicated to temporal and aspectual inference (section 7 of the dataset). This part
-of the FraCaS test suite has been difficult to tackle. That is,
-so far, no computational system has been capable to deal with it in its entirety: when
-authors report accuracy over the FraCaS test suite they
-skip this section.
-%
-In fact, they also often skip the anaphora and ellipsis sections, the
-exception being the system presented by
-\citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019},
-which includes support for anaphora and ellipsis but still omit the
-temporal section.\footnote{One can consider that \citet{maccartney:2007} have made a run against the whole test suite.
-However, they do not deal with multi-premise cases.
-Consequently only 36/75 cases in the temporal section are attempted.
-The general accuracy of the system is .59, and .61 for the temporal section.
-Our system, as shown \cref{tab:results}, presents considerable improvements in coverage and accuracy over that of \citeauthor{maccartney:2007}.}
-%
-In this paper, we take up the challenge of providing a
-computationally viable account of tense and aspect to deal with the
-  section 7 of the FraCaS test suite. Our account is not meant
-to be a theoretically extensive account of tense and aspect, but rather an
-account that is driven by the need to cover the test suite in a way
-that is general enough to capture the test suite examples, \emph{while still covering the rest of the FraCaS test suite}.
-%
-
-The account is evaluated on the entailment properties of various
-temporal and aspectual examples, as given by the test suite. As such,
-we are not getting into the discussion of how tense and aspect might
-affect grammaticality or infelicitousness of various sentences.
-We assume that the sentences of the FraCaS suite are syntactically and
-semantically correct, and strive to produce accurate logical
-representations given that assumption. We further assume that the
-entailment annotations of various problems are valid, and we use those
-to evaluate the correctness of the logical representations of
-sentences.
-
-The paper is structured as follows: in \cref{sec:background}, we give
-a brief summary of the computational frameworks whose various
-subsystems rely on. In particular, the Grammatical Framework is used
-to construct the syntactic parser, the Coq proof assistant checks all
-the reasoning and a monad-based dynamic semantics deals with
-Montague-style semantics, and references (anaphora).  We also provide
-some brief remarks on temporal semantics. In \cref{sec:our-semantics},
-we discuss the main aspects of the compositional semantics of our
-system, using various examples from the FraCaS suite to illustrate its
-effectiveness. In \cref{sec:results}, we evaluate how our system
-performs with respect to the FraCaS suite. We ran the system across
-the whole suite: our system is thus the first which is capable of
-handling the complete FraCaS test suite.  Yet, we are interested in
-particular in the performance on the temporal section. In
-\cref{sec:conclusion}, we conclude and discuss avenues for future
-work.
-
-\section{Temporal-Semantics in a Logic-based NLI System}
-\label{sec:background}
-
-Our temporal analysis places itself in the context of a complete NLI
-system -- which is why we can test it on the FraCaS suite. In this
-section we give a brief overview of the phases of the system,
-referring the reader to published work for details.
-
-\paragraph{GF}
-The first phase of the system, parsing, is taken care of by the 
-Grammatical Framework (GF, \citet{ranta_grammatical_2004}), which is a
-powerful parser generator for natural languages, based on
-type-theoretical abstract grammars.  The present work leverages a
-syntactic representation of the FraCaS test suite in GF abstract
-syntax~\citep{Ljunglof:2012}. Thanks to this, we skip the parsing
-phase and avoid any syntactic ambiguity.
-
-For the purpose of this paper, the important feature of GF
-syntax is that it aims at a balance of sufficient abstraction to provide a
-semantically-relevant structure, but at the same time it embeds
-sufficiently many syntactic features to be able to reconstruct
-natural-language text. That is, the parse trees generally satisfy the
-homomorphism requirement of \citet{montague_english_1970,montague_proper_1974}, and
-we can focus on the translation of syntactic trees to logical forms.
-%
-Consequently, the system presented
-here does not aim at textual natural language understanding, but rather
-provides a testable, systematic formal semantics of temporal phenomena.
-Example \ref{ex:nobel} shows an example abstract syntax tree and its
-realisation in English.
-
-\paragraph{Dynamic Semantics}
-Parse trees are then processed by a dynamic semantic component.
-Its role is essentially to support  (non-temporal) anaphora, using a
-monadic-based dynamic semantics, generally following the state of the
-art in this matter
-\citep{unger_dynamic_2011,charlow_monadic_2015,charlow_modular_2017}.
-Our particular implementation does not handle every case \ifanon
-(citations anonymized) \else \citet{bernardy_computational_2020,bernardy_wide-coverage_2019} \fi
-but non-temporal anaphoroi in the testsuite are generally resolved as
-they should be: on the whole accuracy is not affected significantly by
-issues in this subsystem.
-
-As it is the case for other basic phenomena, there is not much
-interaction between our treatment of time and non-temporal
-anaphora. Critical exceptions are discussed in
-\cref{sec:repeatable-achievements} and \cref{sec:interaction-with-anaphora}.
-
-\paragraph{Montagovian Semantics}
-Non-withstanding special support for anaphora, the core of the translation of syntax trees
-to logical form follows a largely standard
-montagovian semantics. In brief, sentences are interpreted as
-propositions, verbs and noun-phrases as predicates.  We use
-type-raising of noun-phrases, to support
-quantifiers~\citep{montague_proper_1974}.
-
-We support additionally the basic constructions and phenomena present
-in the testsuite, including adjectives, adverbs, nouns, verbs,
-anaphora, etc. The method is outlined by
-\citet{montague_english_1970,montague_proper_1973}, but we direct the
-reader to our previous work for details \ifanon (citations anonymized)
-\else \citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019} \fi, but the particular
-treatment of such phenomena is essentially independent from our
-treatment of time: in this paper we simply ignore these aspects beyond
-the fact that they are handled correctly in most of the cases of the
-FraCaS testsuite.
-
-\paragraph{Inference using Coq}
-Logical forms are then fed to the Coq
-Coq interactive theorem prover (proof assistant). Coq is based on
-the calculus of co-inductive constructions~\citep{werner_une_1994} We
-do not use any co-induction (or even induction) in this paper, relying
-on the pure lambda-calculus inner core of Coq.  Coq is a very powerful
-reasoning engine that makes it fit for implementing natural language
-semantics.  Coq also supports  dependent typing and 
- subtyping. Both concepts are instrumental in expressing NL
-semantics~\citep{chatzikyriakidis_natural_2014}.  Besides, on a more
-practical side, it works well for the the task of NLI, when the latter
-is formalised as a theorem proving task: its many tactics mean that
-many tasks in theorem proving are trivialised. In particular, all
-problems of time-intervals inclusion, which occur in every temporal
-problems, are solved with Coq's linear arithmetic tactic.
-
-\section{Our Treatment of Time}
-\label{sec:our-semantics}
-
-In montagovian semantics, (intransitive) verbs are one-place
-predicates; in types, they are functions from entities to propositions
-($e \to t$).  Our basic approach is to generalise the interpretation of 
-verbs, so that it takes two
-additional time parameters, one corresponding to the starting time
-of the action and one corresponding to its stopping time ($(e × time × time) \to t$). For example,
-if John walked between $t_0$ and $t_1$, we would have:
-$walk(john,t_0,t_1)$. From now on we will call an interval of time
-points $[t_0,t_1]$ a timespan. Every timespan $[t_0,t_1]$ has the
-property $t_0 ≤ t_1$ (it starts no later than it stops). (We are thus
-using a simple Newtonian model of time, corresponding to a layman
-intuition of a linear constant flow of time.)
-
-In principle, common nouns and adjectives should undergo the same
-procedure. For simplicity we will however only consider verbs from now
-on. (In fact, even in our implementation we chose not to extend nouns
-nor adjectives with timespan parameters. This choice limits the
-increase in complexity of the formulas compared to non-temporal
-semantics, at the expense of inaccuracy for a couple of problems in
-the FraCaS test suite: problems 271 and 272 use a an adjective as a
-copula which is subject to temporal reasoning.)
-
-\paragraph{Temporal Context}
-
-We adjust the montagovian semantics so that the interpretation of
-every category (propositions, verb phrases, etc.) takes a
-\emph{temporal context} as an additional parameter, which serves as a time
-reference for the interpretation of all time-dependent semantics
-within the phrase. (While some categories do not need this temporal context, we pass it everywhere for consistency.)  This context propagates through the compositional
-interpretation down to lexical items with atomic representation
-(verbs). By default, every interpretation passes the temporal context
-down to its components without changing it. However some key elements
-will act on it on nontrivial ways, which we proceed to detail below.
-
-This temporal context is an \emph{optional} timespan. That is, it can
-be a timespan or an explicitly unspecified context.
-\newcommand\nospan{\ensuremath{-}}
-\newcommand\anyspan{\ensuremath{\mathbf{*}}} The timespan in the context is
-optional because, in certain situations, the semantics is different
-depending on whether a timespan has been specified externally or not,
-as we explain below. A non-present timespan will be represented as
-\nospan{}. If a semantic function does not depend on the temporal
-context at all, we will write \anyspan{} instead.
-
-\paragraph{Tenses}
-
-\newcommand\varid[1]{\mathnormal{#1}}
-%
-The principal non-trivial manipulators of timespans are tense
-markers. In our syntax, inherited from GF, tenses are represented
-syntactically as an attribute of clauses. An illustration of a
-past-tense clause and its interpretation follows in
-Example \ref{ex:nobel}. Notice in particular the $\varid{past}$ argument to
-the $\varid{useCl}$ constructor.
-
-\onelingex{A scandinavian won the nobel prize.\label{ex:nobel}
-\\
-{\small\begin{multline*}
-\varid{useCl}\, \varid{past}\, \varid{pPos}\\ (\varid{predVP}\, (\varid{detCN}\, (\varid{detQuant}\, \varid{indefArt}\, \varid{numSg})\\ \varid{scandinavian\_CN})\\ (\varid{complSlash}\, (\varid{slashV2a}\, \varid{win\_V2})\\ (\varid{detCN}\, (\varid{detQuant}\, \varid{indefArt}\, \varid{numSg})\\ \varid{nobel\_prize\_CN})))
-\end{multline*}
-}}
-%
-In our semantics we deal only with present and past tenses (simple and
-continuous). Indeed we find that FraCas does not exercise additional
-specific tenses. (When a more complicated tense is used, the
-additional information is also carried by adverbs or adverbial
-phrases, in a more specific way). While we believe that many other
-tenses can be captured under the same general framework, we leave a
-detailed study to further work.
-
-Even though we discuss a refinement to handle the past continuous
-at the end of this section, the procedure to handle tense annotations
-is as follows:
-\begin{itemize}
-\item If the tense is the past, and the temporal context is
-  unspecified, then we locally quantify over a time interval
-  $[t_0,t_1]$, such that $t_1 < now$, where $now$ is a logical
-  constant representing the current timepoint. The temporal context
-  then becomes this interval.
-\item If the tense is the present and the temporal context is
-  unspecified, then the temporal context becomes the simple
-  $(now,now)$ interval.
-\item If the temporal context is specified (for example due to the presence of an adverb or an
-  adverbial clause, such as ``before James swam''), then the tense does
-  not create a new interval, but it may constrain it. Typically, a
-  past tense adds the constraint that the temporal context ends before
-  the timepoint $now$.
-\end{itemize}
-
-\paragraph{Temporal Adverbs}
-
-The other single most important source of interesting timespans are
-adverbs. Most of the temporal adverbs fall in either of the following
-categories:
-
-\begin{enumerate}[align=left,font=\itshape]
-\item[exact] For such adverbs, an exact interval is
-  provided. In fact, such adverbs typically specify a single point in
-  time (so the start and the end of the interval coincide).
-
-  \[⟦\text{at 5 pm, \(s\)}⟧(\anyspan) = ⟦s⟧(5pm,5pm)\]
-
-\item[existentially quantifying] The majority of temporal adverbs
-  existentially quantify over a timespan. Examples include ``since
-  1991'', ``in 1996'', ``for two years'', etc. The common theme is to
-  introduce the interval and then restrict its bounds or its duration
-  in some way. Sometimes the restriction is an equality, as in ``for
-  exactly two hours''. In the following example we show the inclusion
-  constraint, for ``in 1992''.
-
-  \begin{multline*}
-  ⟦\text{in 1992, \(s\)}⟧(\anyspan) = \\ ∃t_1,t_2.  [t_1,t_2] ⊆ 1992, ⟦s⟧(t_1,t_2)
-\end{multline*}
-In the FraCaS test suite, we normally do not find several
-time-modifying adverbs modifying a single verb phrase. Indeed, sentences such
-as ``in 1992, in 1991 john wrote a novel'' are infelicitous. This
-justifies ignoring the input timespan in the above interpretation --
-we are in particular not interested in modelling felicity with our
-semantics, only giving an accurate semantics when the input is
-felicitous.
-\item[universally quantifying] A few adverbs introduce intervals via a
-  universal quantification (sometimes with a constraint). Examples
-  include ``always'' and ``never''.
-
-  If there is no explicit time context, then ``always'' has no
-  constraint on the interval, otherwise the quantified interval must
-  be included in it:
-  \begin{multline*}
-  ⟦\text{always \(s\)}⟧(t_0,t_1) = \\ ∀t_0',t_1'. [t_0',t_1'] ⊆ [t_0,t_1], ⟦s⟧(t_0',t_1')
-\end{multline*}
-Note that here we \emph{do} use the input interval, resulting in a correct interpretation for
-phrases such as ``In 1994, Itel was always on time.'' .
-
-% A specifically interesting case found in FraCas is the adverbial phrase ``every month'', which we modeled
-% \begin{equation}
-%   ⟦\text{every month s}⟧(t_0,t_1) = ∀t. [t,t+1month] ⊆ [t_0,t_1], ⟦s⟧(t,t+1month)
-% \end{equation}
-\end{enumerate}
-
-\paragraph{Aside: aspectual classes in the literature}
-In this paper we borrow several notions from classical temporal
-semantics such as ``stative'', ``achievement'', ``activity'', etc.,
-even though our definitions do not perfectly match the classical
-ones. We explain our precise meaning for these terms in the body of the
-paper. Nevertheless, we refer the reader to \citet{steedman:2000} for
-an extensive review of formal temporal semantics.
-
-For the \textit{cognoscenti}, we can already point out some
-differences in terminology: we use the term activity as a general term
-which encompasses the three classical notions of activites,
-achievements and accomplishments. Indeed, insofar as the test suite is
-concerned, we find that these three categories can be collapsed into a
-single one (they are subject to \cref{eq:unicity}).  As such, we think
-it is a good classification, given that it generally affords the
-correct inferences for the testsuite. (In this paper, we always assume
-that the problems in the FraCaS testsuite are correctly annotated.)
-
-\paragraph{Time references and aspectual classes}
-
-A common theme in the testsuite is the reference to previous
-occurrences of an event:
-
-\fracasex{262}{
-\item	Smith left after Jones left.
-\item	Jones left after Anderson left.
-\hyp 	Did Smith leave after Anderson left?
-}
-
-To be able to conclude that there is entailment, as the testsuite expects, we have to make sure
-that the two occurrences of ``Jones left'' (in \textbf{P1} and \textbf{P2}) refer to the
-same time intervals.
-%
-For this purpose we postulate \emph{unicity of action} for certain time-dependent
-propositions:
-
-\begin{multline}
-  \constant{unicity}_P : P (t_1,t_2) → P (t_3,t_4) → \\ (t_1 = t_3) ∧ (t_2 = t_4)
-  \label{eq:unicity}
-\end{multline}
-
-Unicity of action holds only if the aspectual class of the proposition
-$P$ is \emph{activity}~\citep{steedman_productions_2000} (which, for our
-purposes, includes \emph{achievements} and \emph{accomplishments}).
-
-(The difference between activity and accomplishments on the one hand and
-achievement on the other hand is that for the latter, time intervals
-can be assumed to be of nil duration. In reality, this is an
-oversimplification as achievements are usually of short duration, but
-not nil. However, this plays little role in our
-analysis. As far as we can tell the FraCaS test suite does exercise temporal semantics to
-such a level of precision.)
-
-Unicity of action plays the role of event coreference in
-(neo-)Davidsonian accounts \citep{parsons1990events}. It is also a fine-grained
-principle, allowing coreference to take into account certain arguments
-when referencing. As we detail below, taking arguments into account
-yields is critical to handle repeatability of achievements.
-
-Unicity of action appears to be a non-logical principle. Indeed, it is
-quite possible that ``Jones left'' several times. However, it seems
-that this principle is never contradicted by the testsuite. As such,
-even though unicity of action is only a pragmatic rule, it can be
-taken as a valid one \emph{by default}: it is only when we have a
-sufficiently constrained situation that one should reject it. Consider
-the following discourse:
-\begin{enumerate}[parsep=0pt,itemsep=1pt,label=(\arabic*)]
-\item Smith left at 1pm.
-\item Smith went to his appointment with the lawyer.
-\item Smith left at 4pm.
-\end{enumerate}
-One would normally not say that there is contradiction. However if the
-middle sentence were not present, a contradiction should be
-flagged. We leave such discourse analysis as future work, and simply
-apply unicity of action everywhere: it is valid uniformly in the
-FraCaS test suite for activity aspect classes.
-
-\paragraph{Statives}
-\textit{A contrario}, if \(P\) is \emph{stative}, then we get a time-interval subsumption property:
-
-\begin{multline*}
-\constant {subsumption}_P :\\ [t_3,t_4] ⊆ [t_1,t_2] → P (t_1,t_2) → P (t_3,t_4)
-\end{multline*}
-This principle is used to reason about problem (314), below:
-
-\fracasex{314}
-{\item	Smith arrived in Paris on the 5th of May, 1995.
-\item	Today is the 15th of May, 1995.
-\item	She is still in Paris.
-\hyp 	Smith was in Paris on the 7th of May, 1995.
-}
-
-Indeed, from \textbf{P3} we get that Smith was in Paris between May
-5th and May 15th. Because ``being in Paris'' is stative, we also get that
-Smith was in Paris in any sub-interval. Contrary to unicity of action,
-subsumption is always valid.
-
-\paragraph{Class-modifying adverbs}
-It should be noted that some adverbs can locally disable the
-application of $\constant {subsumption}$. For example, problem 299
-features the sentence ``Smith lived in Birmingham for exactly a
-year''.  Even though ``live'' is normally stative, one can no longer
-apply subsumption in the context of ``exactly a year'' --- this can be
-done by propagating another context flag in the montagovian semantics
-(in addition to the temporal context).
-
-\paragraph{(Un)repeatable Achievements}
-\label{sec:repeatable-achievements}
-The principle of using unicity of action interacts well with the usual
-interpretation of existential quantifiers (and anaphora).  Indeed, using it, we
-can refute problem (279), as expected by the testsuite:
-
-\fracasex{279}{
-\item	Smith wrote a novel in 1991.
-\hyp 	 Smith wrote it in 1992.
-}
-Indeed, following our account, the above (contradictory) inference problem is to be
-interpreted as
-
-\begin{equation}
-  \label{eq:smith-novel-1}
-\begin{array}{l}
-∀x. novel(x) ∧ \\
-∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2) ∧ \\
-∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,x,t_3,t_4) \\
-⟶ ⊥ \\
-\end{array}
-\end{equation}
-Note here that the scope for the existential is extended beyond the
-scope of \textbf{P1}, and its polarity switched (to universal). This extension can follow the account of
-\citet{unger_dynamic_2011}, and our implemented analysis of
-anaphora\ifanon (citation anonymized)
-\else\citep{bernardy_computational_2018,bernardy_wide-coverage_2019}\fi.
-
-Thanks to the unicity of action of $write(smith,x,...)$ (the subject
-and direct object are fixed) we find \([t_1,t_2] = [t_3,t_4]\), and
-due to the years 1991 and 1992 being disjoint we obtain contradiction.
-
-However, the testsuite instructs that we should \emph{not} be able to refute
-problem (280), with the justification that ``wrote a novel'' is a repeatable
-accomplishment:
-
-\fracasex{280}{
-\item	Smith wrote a novel in 1991.
-\hyp 	Smith wrote a novel in 1992.
-}
-Here our interpretation is:
-
-\[\begin{array}{l}
-(∃x. novel(x) ∧ \\
-∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2)) ∧ \\
-(∃y. novel(y) ∧ \\
-∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,y,t_3,t_4)) \\
-⟶ ⊥ \\
-\end{array}
-\]
-Our analysis does not need to treat this last case specially. Indeed,
-even if $write(smith,x,.,.)$ is an activity and thus subject to
-unicity of action, in (280), $x$ is quantified existentially; we have
-two \emph{different} actions: $write(smith,x,t_1,t_2)$ and
-$write(smith,y,t_3,t_4)$, because $x \neq y$, and thus we cannot
-deduce equality of the intervals $t_1,t_2$ and $t_3,t_4$. In turn, the
-hypothesis cannot be refuted.
-
-\paragraph{Action-modification Verbs}
-
-The final class of lexemes carrying a temporal-dependent semantics are
-verbs taking a proposition as argument, like ``finish'', ``start'',
-etc. These verbs modify the temporal context in non-trivial
-ways. Consider for example ``finish to ...''. The timespan of the
-argument of ``finish'' should end within the timespan of the finishing
-action:
-
-\begin{multline*}
-⟦\text{finish to \(s\)}⟧(t_0,t_1) = \\ ∃(t_0',t_1'). t_1' ∈ [t_0,t_1] ∧ ⟦s⟧(t_0',t_1')
-\end{multline*}
-
-\paragraph{Progressive Aspect}
-
-We treat verbs in the progressive form as different semantically from
-the non-progressive form. For example, ``John was writing a book'' is
-encoded as
-$∃(t_1, t_2). t_1≤ t_2, t_2 ≤ now, PROG\_write(John,book,t_1,t_2)$,
-while ``John wrote a book'' is encoded as
-$∃(t_1, t_2). t_1≤ t_2, t_2 ≤ now, write(John,book,t_1,t_2)$. This distinction is
-necessary because in our analysis the progressive form ($PROG\_write$) is subject to
-$\constant{subsumption}$. That is, if John is writing in the interval
-$[t_1,t_2]$ then he is writing in any sub-interval of
-$[t_1,t_2]$. This interpretation corresponds to the idea the the action takes place continuously over the whole interval.
-However, the same cannot be said of the non-continuous form ($write$): the
-end-points of the interval indicate the time needed to complete the
-achievement. (For example, ``John wrote a book in 1993'' neither
-entails ``John wrote a book in January 1993'' nor ``John wrote a book
-in December 1993''.) (In fact, \emph{write}, in the non-progressive
-from, is on the contrary subject to $\constant{unicity}$.) Finally, we
-also have $write(x,y,t_1,t_2) → PROG\_write(x,y,t_1,t_2)$. That is,
-the achievement (or \emph{activity} in our terminology) variant implies the stative variant, for the
-same interval.  Consequently we get the entailment from ``John wrote a
-book in 1993'' to ``John was writing a book in 1993'', but not the other way around.
-
-We note however that this interpretation differs only slightly from
-the usual accounts of the progressive in the
-literature. \Citet{ogihara2007tense} summarises the position of
-\citet{bennett1978toward} as follows: a progressive sentence is true
-at an interval $[t_0,t_1]$ iff there is an interval $[t'_0,t'_1]$ such
-that $[t_0,t_1]$ is a non-final subinterval of $[t'_0,t'_1]$ and the
-progressive sentence is true at $[t'_0,t'_1]$. This is very similar to
-our approach ($\constant{subsumption}$ for the progressive form only),
-but there is a difference regarding final intervals. Yet in our view
-this difference is hard to justify: we cannot see why ``John was
-writing a book in 1993'' entails that he was writing it January,
-February, etc. but not in December.
-
-\Citet{ogihara2007tense} argues that this simple account of the
-progressive fails to reject a sentence such as ``Lee is resembling
-Terri.'' while ``Lee is walking'' is acceptable. We argue instead that
-the latter should be rejected for pragmatic reasons. Indeed, when a
-predicate holds for a very long interval, one typically uses the
-simple present tense in English. Therefore the continuous form
-pragmatically implies that the predicate holds for a limited interval.
-But, without further context, the predicate ``resemble Terri'' does
-not vary over time (while ``walk'' generally does). Therefore the
-continuous form ``Lee is resembling Terri'' is confusing: one implies
-a limited interval, but the semantics of resembling normally yield an unlimited interval.
-%
-Because we do not account for pragmatics, we prefer to retain the
-simplest account based on the subinterval property (which we call
-$\constant{subsumption}$ here).
-
-Finally we stress that not all verbs are subject to the
-stative/achievement distinction induced by the progressive. For
-example, the phrases ``John ran'' and ``John was running'' appear to
-be logically equivalent, for entailment purposes.
-
-\section{Worked out example}
-To give a sense of the additional details necessary to deal with the
-precision demanded by a proof-assistant such as Coq we show how
-problem (279) is worked out in full details.
-
-We start with input trees in GF format, given by
-\citet{Ljunglof:2012}. They can be rendered as follows:
-
-\begin{small}
-\begin{verbatim}
-s_279_1_p=
-sentence
-(useCl past pPos
- (predVP
-   (usePN (lexemePN "smith_PN"))
-   (advVP
-     (complSlash
-      (slashV2a (lexemeV2 "write_V2")))
-     (detCN (detQuant indefArt numSg)
-       (useN (lexemeN "novel_N"))))
-       (lexemeAdv "in_1991_Adv")))
-s_279_3_h=
-sentence
-(useCl past pPos
- (predVP (usePN
-   (lexemePN "smith_PN"))
-   (advVP
-     (complSlash
-      (slashV2a (lexemeV2 "write_V2"))
-      (usePron it_Pron))
-     (lexemeAdv "in_1992_Adv"))))
-\end{verbatim}
-\end{small}
-
-Of particular note is the use of the pronoun ``it'', and the fact that
-adverbials expressions such that ``in 1992'' are lexicalized.  We also
-follow the GF convention to postfix lexical items with the name of
-their category. Most of the other categories follow usual naming
-conventions. We remind the reader that ``slash'' categories are used
-to swap the order of arguments (compared to non-slashed categories of
-similar names).
-
-Our dynamic and temporal semantics gives the following interpretation
-for \verb!s_279_1_p! implies \verb!s_279_3_h!.
-{\small
-\begin{verbatim}
-FORALL (fun a=>novel_N a)
-(fun a=>(exists (b: Time),
-((exists (c: Time),
-(IS_INTERVAL Date_19910101 b /\
- IS_INTERVAL c Date_19911231 /\
- IS_INTERVAL b c /\
- appTime b c (write_V2 a)
-   (PN2object smith_PN))))) ->
-Not (exists (f: Time),
-((exists (g: Time),
- (IS_INTERVAL Date_19920101 f /\
-  IS_INTERVAL g Date_19921231 /\
-  IS_INTERVAL f g /\
-  appTime f g (write_V2 a) 
-    (PN2object smith_PN)))))).
-\end{verbatim}
-}
-In the above, one should remark the top-level quantification over the
-novel (as explained in \cref{sec:our-semantics}), the quantification
-over time intervals as individual timepoints, and the use of custom
-operators for several constructions (\verb!FORALL!, \verb!Not!, \verb!IS_INTERVAL!,
-\verb!appTime!). This use of custom operators is useful for several
-generalisations (for example, we have quantifiers such as \verb!MOST! in
-addition to \verb!FORALL! --- see \ifanon (citation anonymized)
-\else \citet{bernardy_type-theoretical_2017} \fi)
-
-Unfolding the definitions for these operators yield the following
-proposition:
-{\small
-\begin{verbatim}
-forall x : object,
-novel_N x ->
-(exists b c : Z,
-   Date_19910101 <= b /\
-   c <= Date_19911231 /\
-   b <= c /\ write_V2 x SMITH b c) ->
-(exists f g : Z,
-   Date_19920101 <= f /\
-   g <= Date_19921231 /\
-   f <= g /\ write_V2 x SMITH f g) ->
-False
-\end{verbatim}
-}This is very close to our idealised representation of the problem
-\cref{eq:smith-novel-1}. One difference is the use of abstract Coq
-integers for timepoints. Using a discrete time allows us to use
-predefined Coq tactics. The discrete nature of integers does not
-interfere with the reasoning.
-
-Finally, we can show a Coq proof for the above proposition:
-{\small
-\begin{verbatim}
-Theorem  problem279 : Problem279aFalse.
-cbv.
-intros novel isSmithsNovel P1 H.
-destruct P1 as
-   [t0 [t1 [ct1 [ct2 [ct3 P1]]]]].
-destruct H as
-   [u0 [u1 [cu1 [cu2 [cu3 H]]]]].
-specialize writeUnique
-  with (x := novel)(y := SMITH) as A.
-unfold UniqueActivity in A.
-specialize (A _ _ _ _ P1 H) as B.
-lia.
-Qed.
-\end{verbatim}
-}
-
-The intros and destruct tactics serve bookkeeping purposes. The
-critical part is the use of the \texttt{writeUnique} axiom, which
-witnesses the aspectual class of the predicate \texttt{write\_V2}.
-The proof is completed by the use of the \texttt{lia} tactic, which is
-embeds a decision procedure for linear arithmetic
-problems. Fortunately, \texttt{lia} can take care of all the problems
-which arise in the FraCaS testsuite.
-
 \begin{figure*}
@@
 \end{figure*}
-
-\section{Results and Evaluation}
-\label{sec:results}
-Our target is the FraCaS testsuite, which aims at covering a wide
-range of common natural-language phenomena. 
-%
-The suite is structured according to the semantic phenomena involved
-in the inference process for each example, and contains nine sections:
-Quantifiers, Plurals, Anaphora, Ellipsis, Adjectives, Comparatives,
-Temporal, Verbs and Attitudes.  The system described here focuses on
-the Temporal section. However, it also supports the other eight
-sections.  To our knowledge this is the first system which attempts to
-target the temporal section in full. But in fact, our system even provides
-support for all the other sections. Thus, a couple of decades after its
-formulation, we propose a first attempt at covering the whole suite.
-As such, there it is no other system to compare our system with, in
-all aspects.  We can however compare with systems which target parts
-of the FraCaS testsuite, as shown in \cref{tab:results}.
-
-\providecommand\ncases[1]{{\ensuremath{^{#1}}}}
-\begin{table}[t]
-  \centering
-  \small
-\begin{tabularx}{\columnwidth}{Xr@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r}
-Section      & {\kern -2em} \#FraCaS
-                          & This        & FC2         & FC & MINE & Nut  & LP  \\ \hline
-Quantifiers  & 75         & .93        & .96         & .96    & .77  & .53  & .93  \\
-             &            & \ncases{74}& \ncases{74} &        &      &      &     \ncases{44} \\
-Plurals      & 33         & .79        & .82         & .76    & .67  & .52  & .73 \\
-             &            &            &             &     &   &   & \ncases{24} \\
-Anaphora     & 28         & .79        & .86         &   -    & -    & -    &  -       \\
-Ellipsis     & 52         & .81        & .87         &   -    & -    & -    &  -       \\
-Adjectives   & 22         & .95        & .95         & .95    & .68  & .32  & .73 \\
-             &            & \ncases{20}&  \ncases{20}&     &   &   &  \ncases{12} \\
-Comparatives & 31         & .65        & .87         & .56    & .48  & .45  &  -       \\
-Temporal     & 75         & .73        &  -          &   -    &   -  &  -   &  -       \\
-Verbs        & 8          & .75        & .75         &   -    & -    & -    &  -       \\
-Attitudes    & 13         & .85        & .92         & .85    & .77  & .46  & .92  \\ 
-             &            &            &             &        &      &      & \ncases {9}  \\ \hline
-Total        & 337        & .81        & .89         & .83    & .69  & .50  & .85  \\
-             &            & \ncases{329}& \ncases{259}& \ncases{174}  & \ncases{174}& \ncases{174}& \ncases{89}
-  \end{tabularx}
-  \caption{Accuracy of our system compared to others.
-    ``This" refers to the approach presented in this paper. When a
-    system does not handle the nominal number of test cases (shown in
-    the second column), the actual number of test cases attempted is
-    shown below the accuracy figure, in smaller font.  ``FC''
-    refers to the work of \citet{bernardy_type-theoretical_2017}, and ``FC2'' its followup~\citep{bernardy_wide-coverage_2019}. ``MINE" refers
-    to the approach of \citet{Mineshima:2015}, ``NUT" to the CCG
-    system that utilises the first-order automated theorem prover
-    \textit{nutcracker} \cite{bos:2008}, and ``Langpro"
-    to the system presented by \citet{abzianidze_tableau_2015}. A dash
-    indicates that no attempt was made for the section. }
-  \label{tab:results}
-\end{table}
-Additionally, to provide the reader with a better sense of what the
-complete system is capable and not capable of, we list a number of
-correctly and incorrectly handled examples in \cref{fig:correct-examples} and \cref{fig:incorrect-examples}.
-\paragraph{Interaction with anaphora}
-\label{sec:interaction-with-anaphora}
-One reason explaining the lower performance of our system on some
-sections of the testsuite is that our interpretation of time interacts
-imperfectly with anaphora and ellipsis. Consider the following example:
-
-\fracasex{232}
-{\item	ITEL won more orders than APCOM did.
-\item	APCOM won ten orders.
-\hyp 	ITEL won at least eleven orders.
-}
-
-In the first premise, our system essentially resolves the ellipsis to
-get the following reading: ``ITEL won \(X\) orders and APCOM won \(Y\) orders
-and $X > Y$.''. One would need each of the verb phrases ``won \(X\) orders''
-and ``won $Y$ orders'' to introduce their own timespans with existential
-quantifiers. However, the organisation of our system is such that the
-existentials are introduced before the ellipsis is
-expanded. Consequently we get a wrong interpretation and the inference
-cannot be made.
-
-\section{Conclusions and Future Work}
-\label{sec:conclusion}
-
-We have presented a first attempt for a computational approach dealing
-with the temporal section of the FraCaS test suite. To do this, we
-have provided a new, simplified taxonomy of aspectual classes for verb
-phrases, guided by the applicability of the unicity of action and
-temporal subsumption properties. While part of this simplification is
-accidental (conflation of activity and accomplishment), we find that
-other parts (the automatic distinction between repeatable and
-unrepeatable achievements) constitute theoretical improvements.
-
-Besides inference, formal interpretation of tense is found in
-natural-language interfaces to databases. Of note is the work of
-\citet{androutsopoulos1998time}, which handles many of the time-aware
-adverbial clauses that we address. However, we cover many more
- logical aspects of inference, such as coreference via
-unity of action and interaction with quantifiers.
-
-\cite{bernardy_wide-coverage_2019} presented a
-logical system for handling 8 of the 9 sections of the FRACAS test
-suite, but excluded section 7, suggesting that it requires many
-examples that need an \textit{ad hoc} treatment. Here, we take up this
-challenge and show that a system similar to theirs can be extended to
-cover the remaining section of the test suite.
-%
-We use the same combination of a number of well-studied tools: type
-theory, parsing using the Grammatical Framework (GF), Monadic Dynamic
-Semantics and proof assistant technology (Coq). The system achieves an
-accuracy of 0.73 on the Temporal Section and 0.81 overall.
-One of the
-things to be looked at is fixing the issues associated with parts of
-the test suite that ``broke'' when the temporal analysis was
-introduced. Some of these have been already mentioned: interaction of
-the temporal variables with anaphora.
-
-Another extension of this work is to reflect more temporal semantic
-inference properties in an extended test suite. Indeed, there as properties which are
-not captured in the FraCaS test suite, such as fine-grained examples of lexical and
-grammatical aspect, as well as the interaction between those two, for example
-cases where one needs to actually distinguish between achievements and
-accomplishments on the basis of their inferential properties:
-
-\fracasex{extra1}{
-	\item	John found his keys.  
-	\hyp 	John was finding his keys  (UNK).
-}
-
-\fracasex{extra2}{
-	\item	John wrote a book.  
-	\hyp 	John was writing a book  (YES).
-}
-
-In the first of the two examples involving an achievement verb, the
-inference is UNK, since there is no guarantee that the action is
-non-instantaneous. To the contrary, for accomplishment verbs, the
-inference follows.
-
-Further cases to be included in an extended FraCaS future suite
-involve examples where the interaction between different tenses needs
-to be captured:
-
-\fracasex{extra3}{
-	\item	When the phone rang, John had entered the house.  
-	\hyp 	John entered the house before the phone rang (YES).
-}
-
-\section*{Acknowledgements}
-
-%\ifanon
-%Anonymized
-%\else
-The research reported in this paper was supported by grant 2014-39 from the
-Swedish Research Council, which funds the Centre for Linguistic Theory and
-Studies in Probability (CLASP) in the Department of Philosophy, Linguistics,
-and Theory of Science at the University of Gothenburg. We are grateful to
-our colleagues in CLASP for helpful discussion of some of the ideas presented
-here. We also thank anonymous reviewers for their useful comments on an
-earlier draft of the paper.
-%\fi
-\bibliographystyle{apalike}
-
-\bibliography{../PaperTools/bibtex/jp,paper2020,../paper2018/NLI}
-
-\end{document}
-
-% Local Variables:
-% ispell-local-dictionary: "british"
-% End:
-
-% LocalWords:  dowty steedman higginbotham fernando NLI FraCaS GF une
-% LocalWords:  aspectual dataset anaphora bernardy grammaticality NL
-% LocalWords:  infelicitousness ranta Ljunglof montague english TODO
-% LocalWords:  prover chatzikyriakidis Coq's  monadic unger
-% LocalWords:  charlow anaphoroi stative activites timespan timespans
-% LocalWords:  useCl scandinavian nobel FraCas timepoint Itel unicity
-% LocalWords:  Statives contrario th Un lexemes PROG lexicalized lia
-% LocalWords:  postfix timepoints FORALL appTime writeUnique FC CCG
-% LocalWords:  FraCoq Mineshima Langpro abzianidze premiss UNK Centre
-% LocalWords:  Stergios testsuite datasets maccartney compositional
-% LocalWords:  Montagovian montagovian coreference neo Davidsonian
-% LocalWords:  ogihara bennett iff subinterval APCOM existentials hoc
-% LocalWords:  androutsopoulos

### 2021-02-07 paper2020/paper_2021.tex
minor typos!

@@
 \paragraph{Inference using Coq}
-Logical forms are then fed to the Coq
+Logical forms are then fed to the 
 Coq interactive theorem prover (proof assistant). Coq is based on
@@
 $P$ is \emph{activity}~\citep{steedman_productions_2000} (which, for our
-purposes, includes \emph{achievements} and \emph{accomplishments}).
+purposes, includes \emph{achievements} and \emph{accomplishments as well}).
 

### 2021-02-09 paper2020/paper.tex
merge changes

@@
 
-\title{FraCaS: Temporal Analysis}
+\title{Applied Temporal Analysis: A Complete Run of the FraCaS Test Suite }
 % \author{\ifanon Anonymous for review\else
@@
 
-\begin{abstract}
-In this paper, we propose an implementation of temporal semantics
-  which is suitable for inference problems. This implementation
+\begin{abstract}In this paper, we propose an implementation of temporal semantics that
   translates syntax trees to logical formulas, suitable for
-  consumption by the Coq proof assistant. We support several phenomena
+ consumption by the Coq proof assistant. The analysis  supports a wide range of phenomena
  including: temporal references, temporal adverbs, aspectual classes
-  and progressives.  We apply these semantics to the complete FraCaS testsuite.
-  We obtain an accuracy of 81 percent overall and 73 percent for problems explicitly marked
-  as related to temporal reference.
+ and progressives.
+The new semantics are built on top of a previous system handling all sections of the FraCaS test suite except the temporal reference section, and we  obtain an accuracy of 81 percent overall and 73 percent for the problems explicitly marked
+as related to temporal reference. To the best of our knowledge, this is the best performance of a logical system on the whole of the FraCaS. 
 \end{abstract}
@@
 $P$ is \emph{activity}~\citep{steedman_productions_2000} (which, for our
-purposes, includes \emph{achievements} and \emph{accomplishments}).
+purposes, includes \emph{achievements} and \emph{accomplishments} as well).
 
@@
 \cite{bernardy_wide-coverage_2019} presented a
-logical system for handling 8 of the 9 sections of the FRACAS test
+logical system for handling 8 of the 9 sections of the FraCaS test
 suite, but excluded section 7, suggesting that it requires many
-examples that need an \textit{ad hoc} treatment. Here, we take up this
-challenge and show that a system similar to theirs can be extended to
-cover the remaining section of the test suite.
+examples that need an \textit{ad hoc} treatment. Here, we took up this
+challenge and have shown that a system similar to theirs can be extended to
+cover the remaining section of the test suite, without considerably decreasing the performance of the 
+rest of the sections.
+This is indeed a common problem with logical approaches, namely the fact that one can have theoretically motivated implementations of individual phenomena, e.g.
+anaphora, ellipsis, quantifiers, temporal reference etc., but when one tries to put all these together into a unified system, this proves to be a daunting task.
+We believe that this paper presents an exception, and provides a system that can deal with all these different semantic phenomena under a unified system with very good results.
 %

### 2021-02-09 paper2020/paper.tex
fit

@@
 provides a testable, systematic formal semantics of temporal phenomena.
-Example \ref{ex:nobel} shows an example abstract syntax tree and its
+Example \ref{ex:nobel} shows an abstract syntax tree and its
 realisation in English.
@@
 Non-withstanding special support for anaphora, the core of the translation of syntax trees
-to logical form follows a largely standard
+to logical form follows a standard
 montagovian semantics. In brief, sentences are interpreted as
@@
 \onelingex{A scandinavian won the nobel prize.\label{ex:nobel}
+  \vspace{-1.5ex}
 {\small\begin{multline*}
@@
 %
+  \vspace{-1.5ex}
   \begin{multline*}
@@
   be included in it:
+  \vspace{-1.5ex}
   \begin{multline*}
@@
 %
+  \vspace{-1.5ex}
 \begin{multline}
@@
 \textit{A contrario}, if \(P\) is \emph{stative}, then we get a time-interval subsumption property:
+\vspace{-1.5ex}
 \begin{multline*}
@@
 
+\vspace{-1.5ex}
 \begin{equation}
@@
 Here our interpretation is:
-
+\vspace{2ex}
 \[\begin{array}{l}
@@
 action:
+\vspace{-1.5ex}
 \begin{multline*}
@@
 
+\vspace{-1.5ex}
 \fracasex{232}
@@
 
-Further cases to be included in an extended FraCaS future suite
-involve examples where the interaction between different tenses needs
-to be captured:
-
-\fracasex{extra3}{
-	\item	When the phone rang, John had entered the house.  
-	\hyp 	John entered the house before the phone rang (YES).
-}
+% Further cases to be included in an extended FraCaS future suite
+% involve examples where the interaction between different tenses needs
+% to be captured:
+% \fracasex{extra3}{
+% 	\item	When the phone rang, John had entered the house.  
+% 	\hyp 	John entered the house before the phone rang (YES).
+% }
 

### 2021-03-02 paper2020/paper.tex
tweak

@@
 
-\cite{bernardy_wide-coverage_2019} presented a
+\citet{bernardy_wide-coverage_2019} presented a
 logical system for handling 8 of the 9 sections of the FraCaS test

### 2021-04-19 paper2020/paper_2021.tex
minor

@@
 the study of formal semantics since the early days of Montague Grammar
-and a number of different ideas have been put forth on the table to
+and a number of different ideas have been put forth to
 deal with them throughout the years. Recent proposals include the

### 2021-04-19 paper2020/paper_2021.tex
minor

@@
 \end{multline*}
-This principle is used to reason about problem (314), below:
+This principle is used to reason about problem (314), below (note that ``Smith'' is used as a surname in the FraCaS and can take both feminine and masculine values):
 
@@
 involve examples where the interaction between different tenses needs
-to be captured:
+to be captured:\footnote{While this work was completed, the work by \cite{vashishtha-etal-2020} was published.
+The authors present a five datasets to be used for the training of neural models' ability to capture temporal reasoning.
+It would be interesting to check the amount of data covered, most specifiaclly the level of fine-grainedness of temporal reasoning needed to capture those examples, as compared to what we have been discussing in this paper.
+We thank an anonymous reviewer for bringing this work to our attention.}
 

### 2021-04-21 paper2020/paper.tex
kill 1st todo

@@
-\newif\ifanon\anontrue
+\newif\ifanon\anonfalse
 
@@
 
+Finally it would be desirable to improve automation of the system, and
+evaluate it on a larger test set. As it stands Coq fully \emph{checks}
+the proof of entailment for each (provable) problem. However, the
+construction of such proofs has demanded human intervention. It would
+be desirable to fully automate the proof construction step. For this
+to make sense however we need a much larger test suite, properly
+separated into a development and a (secret) test set. Otherwise, only
+the limited power of the logic prevents us (or any followup work) to
+fine-tune the rules of the system until one gets full coverage. This
+kind of observation holds in general of any rule-based system, and
+thus applies not only to the proof-construction phase, but also to the
+underlying dynamic semantics and parsing phase (which is limited only
+by the power of the language and frameworks used for its
+implementation). In sum, contrary to statistical approaches to
+language understanding, the value of the present work lies not in the
+bare accuracy numbers which we are able to achieve, but in the details
+of \emph{how} we do so: the of set of rules which we use, which is
+described in detail here and in the work which we base ourselves upon
+\citep{bernardy_computational_2020,bernardy_wide-coverage_2019}.
+
 \ifanon

### 2021-04-21 paper2020/paper.tex
kill more todos

@@
 \paragraph{Dynamic Semantics}
-Parse trees are then processed by a dynamic semantic component.
-Its role is essentially to support  (non-temporal) anaphora, using a
+Parse trees are then processed by a dynamic semantic component.  Its
+role is essentially to support (non-temporal) anaphora, using a
 monadic-based dynamic semantics, generally following the state of the
@@
 \citep{unger_dynamic_2011,charlow_monadic_2015,charlow_modular_2017}.
-Our particular implementation does not handle every case \ifanon
-(citations anonymized) \else \citet{bernardy_computational_2020,bernardy_wide-coverage_2019} \fi
+Our particular implementation has weaknesses in certain areas (including group readings and counting; see \citet{bernardy_computational_2020} for details)
 but non-temporal anaphoroi in the testsuite are generally resolved as
@@
 treatment of time: in this paper we simply ignore these aspects beyond
-the fact that they are handled correctly in most of the cases of the
-FraCaS testsuite.
+the fact that they are handled correctly in the
+FraCaS testsuite, except in a few pathological cases.
 
@@
 $walk(john,t_0,t_1)$. From now on we will call an interval of time
-points $[t_0,t_1]$ a timespan. Every timespan $[t_0,t_1]$ has the
-property $t_0 ≤ t_1$ (it starts no later than it stops). (We are thus
+points $[t_0,t_1]$ a timespan, where $t_0$ and $t_1$ are elements of the $time$ type, which is represented in Coq as an abstract ordered ring. Every timespan $[t_0,t_1]$ has the
+property $t_0 ≤ t_1$: it starts no later than it stops. (We are thus
 using a simple Newtonian model of time, corresponding to a layman
@@
 copula which is subject to temporal reasoning.)
+
+\fracasex{271}{\answer unknown
+\item	Smith was present.
+\item	Jones was present.
+\item	Smith was present after Jones was present.
+\hyp 	Jones was present before Smith was present.
+}
 
@@
 concerned, we find that these three categories can be collapsed into a
-single one (they are subject to \cref{eq:unicity}).  As such, we think
-it is a good classification, given that it generally affords the
-correct inferences for the testsuite. (In this paper, we always assume
-that the problems in the FraCaS testsuite are correctly annotated.)
+single one (they are subject to \cref{eq:unicity}).  That is, it is
+sufficient for the testsuite to distinguish between events and
+states. (In this paper, we always assume that the problems in the
+FraCaS testsuite are correctly annotated.)
 
@@
 The proof is completed by the use of the \texttt{lia} tactic, which is
-embeds a decision procedure for linear arithmetic
-problems. Fortunately, \texttt{lia} can take care of all the problems
-which arise in the FraCaS testsuite.
+embeds a decision procedure for linear arithmetic problems\footnote{It
+  solves linear goals over rings by searching for linear refutations
+  and cutting planes}. Fortunately, \texttt{lia} can take care of all
+the problems which arise in the FraCaS testsuite.
 

### 2021-04-22 paper2020/paper.tex
all JP todos addressed

@@
 and direct object are fixed) we find \([t_1,t_2] = [t_3,t_4]\), and
-due to the years 1991 and 1992 being disjoint we obtain contradiction.
+due to the years 1991 and 1992 being disjoint we obtain contradiction. In sum, no special notion of accomplishment is needs to be invoked: we only need the principle of unicity of action.
 
-However, the testsuite instructs that we should \emph{not} be able to refute
+Yet, the testsuite instructs that we should \emph{not} be able to refute
 problem (280), with the justification that ``wrote a novel'' is a repeatable
@@
 $[t_1,t_2]$ then he is writing in any sub-interval of
-$[t_1,t_2]$. This interpretation corresponds to the idea the the action takes place continuously over the whole interval.
+$[t_1,t_2]$. This interpretation corresponds to the idea that the action takes place continuously over the whole interval.
 However, the same cannot be said of the non-continuous form ($write$): the
@@
 Of particular note is the use of the pronoun ``it'', and the fact that
-adverbials expressions such that ``in 1992'' are lexicalized.  We also
+adverbial expressions such that ``in 1992'' are lexicalized.  We also
 follow the GF convention to postfix lexical items with the name of

### 2021-04-26 paper2020/paper.tex
table corrections

@@
 Quantifiers  & 75         & .93        & .96         & .96    & .77  & .53  & .93  \\
-             &            & \ncases{74}& \ncases{74} &        &      &      &     \ncases{44} \\
+             &            & \ncases{74}& \ncases{74} &        &      &      &     \\
 Plurals      & 33         & .79        & .82         & .76    & .67  & .52  & .73 \\
-             &            &            &             &     &   &   & \ncases{24} \\
+             &            &            &             &     &   &   & \\
 Anaphora     & 28         & .79        & .86         &   -    & -    & -    &  -       \\
@@
 Adjectives   & 22         & .95        & .95         & .95    & .68  & .32  & .73 \\
-             &            & \ncases{20}&  \ncases{20}&     &   &   &  \ncases{12} \\
+             &            & \ncases{20}&  \ncases{20}&     &   &   &   \\
 Comparatives & 31         & .65        & .87         & .56    & .48  & .45  &  -       \\
@@
 Attitudes    & 13         & .85        & .92         & .85    & .77  & .46  & .92  \\ 
-             &            &            &             &        &      &      & \ncases {9}  \\ \hline
+             &            &            &             &        &      &      &  \\ \hline
 Total        & 337        & .81        & .89         & .83    & .69  & .50  & .85  \\
-             &            & \ncases{329}& \ncases{259}& \ncases{174}  & \ncases{174}& \ncases{174}& \ncases{89}
+             &            & \ncases{329}& \ncases{259}& \ncases{181}  & \ncases{181}& \ncases{181}& \ncases{142}
   \end{tabularx}
@@
     system that utilises the first-order automated theorem prover
-    \textit{nutcracker} \cite{bos:2008}, and ``Langpro"
+    \textit{nutcracker} \cite{bos:2008}, and ``LP"
     to the system presented by \citet{abzianidze_tableau_2015}. A dash

### 2021-04-26 paper2020/paper.tex
table corrections

@@
 Total        & 337        & .81        & .89         & .83    & .69  & .50  & .85  \\
-             &            & \ncases{329}& \ncases{259}& \ncases{181}  & \ncases{181}& \ncases{181}& \ncases{142}
+             &            & \ncases{329}& \ncases{259}& \ncases{174}  & \ncases{181}& \ncases{181}& \ncases{142}
   \end{tabularx}

### 2021-05-04 paper2020/paper.tex
deanon + citation fix

@@
 
-%\aclfinalcopy % Uncomment this line for the final submission
+\aclfinalcopy % Uncomment this line for the final submission
 %\def\aclpaperid{***} %  Enter the acl Paper ID here
@@
 
-\title{Applied Temporal Analysis: A Complete Run of the FraCaS Test Suite }
+\title{Applied Temporal Analysis: A Complete Run of the FraCaS Test Suite}
 % \author{\ifanon Anonymous for review\else
@@
 
-\author{\ifanon Anonymous for review\else
-  {Jean-Philippe Bernardy}\footnote{University of Gothenburg} ~~ {Stergios Chatzikyriakidis}\footnote{University of Gothenburg}
+\author{
+  Jean-Philippe Bernardy    \quad\quad\quad
+  Stergios Chatzikyriakidis \\
+  Centre for Linguistic Theory and Studies in Probability \\
+  Department of Philosophy, Linguistics and Theory of Science \\ University of Gothenburg \\
+  \texttt{firstname.lastname@gu.se} \\
 }
@@
 anaphora\ifanon (citation anonymized)
-\else\citep{bernardy_computational_2018,bernardy_wide-coverage_2019}\fi.
+\else\citep{bernardy_computational_2020,bernardy_wide-coverage_2019}\fi.
 

### 2021-05-04 paper2020/paper.tex
add url for source code

@@
 
-\citet{bernardy_wide-coverage_2019} presented a
-logical system for handling 8 of the 9 sections of the FraCaS test
-suite, but excluded section 7, suggesting that it requires many
-examples that need an \textit{ad hoc} treatment. Here, we took up this
-challenge and have shown that a system similar to theirs can be extended to
-cover the remaining section of the test suite, without considerably decreasing the performance of the 
-rest of the sections.
-This is indeed a common problem with logical approaches, namely the fact that one can have theoretically motivated implementations of individual phenomena, e.g.
-anaphora, ellipsis, quantifiers, temporal reference etc., but when one tries to put all these together into a unified system, this proves to be a daunting task.
-We believe that this paper presents an exception, and provides a system that can deal with all these different semantic phenomena under a unified system with very good results.
+\citet{bernardy_wide-coverage_2019} presented a logical system for
+handling 8 of the 9 sections of the FraCaS test suite, but excluded
+section 7, suggesting that it requires many examples that need an
+\textit{ad hoc} treatment. Here, we took up this challenge and have
+shown that a system similar to theirs can be extended to cover the
+remaining section of the test suite, without considerably decreasing
+the performance of the rest of the sections. This is indeed a common
+problem with logical approaches, namely the fact that one can have
+theoretically motivated implementations of individual phenomena,
+e.g. anaphora, ellipsis, quantifiers, temporal reference etc., but
+when one tries to put all these together into a unified system, this
+proves to be a daunting task. We believe that this paper presents an
+exception, and provides a system that can deal with all these
+different semantic phenomena under a unified system with very good
+results.
 %
@@
 Semantics and proof assistant technology (Coq). The system achieves an
-accuracy of 0.73 on the Temporal Section and 0.81 overall.
-One of the
-things to be looked at is fixing the issues associated with parts of
-the test suite that ``broke'' when the temporal analysis was
+accuracy of 0.73 on the Temporal Section and 0.81 overall. The whole
+system, including data sets, is available at the following url:
+\url{https://github.com/GU-CLASP/FraCoq/tree/iwcs2021}.
+
+One of the things to be looked at is fixing the issues associated with
+parts of the test suite that ``broke'' when the temporal analysis was
 introduced. Some of these have been already mentioned: interaction of

### 2021-05-04 paper2020/paper.tex
final iwcs 2021 version

@@
 syntactic representation of the FraCaS test suite in GF abstract
-syntax~\citep{Ljunglof:2012}. Thanks to this, we skip the parsing
+syntax, in effect a GF FraCaS treebank~\citep{Ljunglof:2012}. Thanks to this, we skip the parsing
 phase and avoid any syntactic ambiguity.
@@
 provides a testable, systematic formal semantics of temporal phenomena.
-Example \ref{ex:nobel} shows an abstract syntax tree and its
+Example \ref{ex:nobel} shows an example abstract syntax tree and its
 realisation in English.
@@
 Here our interpretation is:
-\vspace{2ex}
+
 \[\begin{array}{l}
@@
 action:
-\vspace{-1.5ex}
+
 \begin{multline*}
@@
 Quantifiers  & 75         & .93        & .96         & .96    & .77  & .53  & .93  \\
-             &            & \ncases{74}& \ncases{74} &        &      &      &     \\
+             &            & \ncases{74}& \ncases{74} &        &      &      &     \ncases{44} \\
 Plurals      & 33         & .79        & .82         & .76    & .67  & .52  & .73 \\
-             &            &            &             &     &   &   & \\
+             &            &            &             &     &   &   & \ncases{24} \\
 Anaphora     & 28         & .79        & .86         &   -    & -    & -    &  -       \\
@@
 Adjectives   & 22         & .95        & .95         & .95    & .68  & .32  & .73 \\
-             &            & \ncases{20}&  \ncases{20}&     &   &   &   \\
+             &            & \ncases{20}&  \ncases{20}&     &   &   &  \ncases{12} \\
 Comparatives & 31         & .65        & .87         & .56    & .48  & .45  &  -       \\
@@
 Attitudes    & 13         & .85        & .92         & .85    & .77  & .46  & .92  \\ 
-             &            &            &             &        &      &      &  \\ \hline
+             &            &            &             &        &      &      & \ncases {9}  \\ \hline
 Total        & 337        & .81        & .89         & .83    & .69  & .50  & .85  \\
-             &            & \ncases{329}& \ncases{259}& \ncases{174}  & \ncases{181}& \ncases{181}& \ncases{142}
+             &            & \ncases{329}& \ncases{259}& \ncases{174}  & \ncases{174}& \ncases{174}& \ncases{89}
   \end{tabularx}
@@
 
-\vspace{-1.5ex}
 \fracasex{232}
@@
 %
-We use the same combination of a number of well-studied tools: type
+We use the same combination of a number of well-studied tools as \citet{bernardy_wide-coverage_2019} : type
 theory, parsing using the Grammatical Framework (GF), Monadic Dynamic
@@
 % }
+Further cases to be included in an extended FraCaS future suite
+involve examples where the interaction between different tenses needs
+to be captured:\footnote{While this work was completed, the work by \cite{vashishtha-etal-2020} was published.
+The authors present a five datasets to be used for the training of neural models' ability to capture temporal reasoning.
+It would be interesting to check the amount of data covered, most specifiaclly the level of fine-grainedness of temporal reasoning needed to capture those examples, as compared to what we have been discussing in this paper.
+We thank an anonymous reviewer for bringing this work to our attention.}
+
+\fracasex{extra3}{
+	\item	When the phone rang, John had entered the house.  
+	\hyp 	John entered the house before the phone rang (YES).
+}
 
@@
 here. We also thank anonymous reviewers for their useful comments on an
-earlier draft of the paper.
-\fi
-\bibliographystyle{apalike}
+earlier draft of the paper.\fi
+
+\bibliographystyle{acl_natbib}
 

### 2021-05-18 paper2020/talk.tex
draft talk

\nonstopmode
\documentclass{beamer}
%
% Choose how your presentation looks.
%
% For more themes, color themes and font themes, see:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
%
\mode<presentation>
{
  \usetheme{default}      % or try Darmstadt, Madrid, Warsaw, ...
  \usecolortheme{default} % or try albatross, beaver, crane, ...
  \usefonttheme{default}  % or try serif, structurebold, ...
  \setbeamertemplate{navigation symbols}{}
  \setbeamertemplate{caption}[numbered]
}

\usepackage{amsmath}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{unicode-math}
\usepackage{tikz}
\usetikzlibrary{matrix}
\usetikzlibrary{calc}
\usetikzlibrary{fadings}
\usetikzlibrary{arrows,automata}
\usetikzlibrary{intersections}

\title[Temporal/FraCaS]{Applied Temporal Analysis: A Complete Run of the FraCaS Test Suite}
\author{Jean-Philippe Bernardy and Stergios Chatzikyriakidis}
\institute{CLASP, University of Gothenburg}
\date{IWCS 2021}

\newcommand\ConId[1]{\mathsf{#1}}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \begin{itemize}
  \item 
    Our starting point is the FraCoq system (Presented by the same authors at IWCS 2017 with updates at NoDaliDa 2019):

\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=4.2cm,
                    semithick,scale=0.65, every node/.style={transform shape}]
  \tikzstyle{every state}=[draw=black,text=black,shape=rectangle]

 \node[state]         (txt)                {Textual Problem};
 \node[state]         (ast) [right of=txt] {AST};
 \node[state]         (pp)  [right of=ast] {Logic Form};
 \node[state]         (math)    [right of=pp]  {Checked result};

 \path (txt) edge              node (GF)  {GF}                  (ast)
       (ast) edge              node (Hask) {\alert{Semantics}} (pp);
 \path  (pp) edge               node (Coq) {Coq} (math);

 \node[state]         (proof) [above of=Coq]  {Proof};
 \path  (proof) edge                (Coq);

\end{tikzpicture}
\item
In this paper we update the \alert{Semantics} with {Temporal Semantics}. We achieve coverage of 329 out of 337 problems in the FraCaS test suite, with an accuracy of 81\%.
  \end{itemize}
\end{frame}

\begin{frame}

\begin{itemize}
  \item (Smith wrote a novel in 1991 $⇒$ Smith wrote it in 1992) is false
    \begin{itemize}
    \item \[\begin{array}{l}
∀x. \ConId{novel}(x) ∧ \\
∃t_1,t_2.  [t_1,t_2] ⊆ 1991 ∧ \ConId {write}(\ConId{smith},x,t_1,t_2) ∧ \\
∃t_3,t_4.  [t_3,t_4] ⊆ 1992 ∧ \ConId {write}(\ConId{smith},x,t_3,t_4) \\
⟶ ⊥
\end{array}
\]
is provable
\end{itemize}
\item (Smith wrote a novel in 1991 $⇒$ Smith wrote a novel in 1992.) is neutral

  \begin{itemize}
  \item \[\begin{array}{l}
(∃x. \ConId{novel}(x) ∧ \\
∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ \ConId{write}(\ConId{smith},x,t_1,t_2)) → \\
(∃y. \ConId{novel}(y) ∧ \\
∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ \ConId{write}(\ConId{smith},y,t_3,t_4)) \\
 \\
\end{array}
\]
is not provable.
\end{itemize}
\end{itemize}

\end{frame}

\end{document}

### 2021-05-20 paper2020/talk.tex
who done what

@@
 \usetikzlibrary{intersections}
+\usetikzlibrary{positioning}
 
@@
   \item 
-    Our starting point is the FraCoq system (Presented by the same authors at IWCS 2017 with updates at NoDaliDa 2019):
-
+    Our starting point is the FraCoq system (Presented by the same authors at IWCS 2017 with updates at NoDaliDa 2019): \\
+    \bigskip
 \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=4.2cm,
-                    semithick,scale=0.65, every node/.style={transform shape}]
+                    semithick,scale=0.63, every node/.style={transform shape},inputnode/.style={draw=black,text=black,shape=rectangle,rounded corners=1mm}]
   \tikzstyle{every state}=[draw=black,text=black,shape=rectangle]
 
- \node[state]         (txt)                {Textual Problem};
+ \node[inputnode]         (txt)                {Textual Problem};
  \node[state]         (ast) [right of=txt] {AST};
@@
 
- \node[state]         (proof) [above of=Coq]  {Proof};
+ \node[inputnode]         (proof) [above of=Coq,node distance=2cm]  {Proof};
  \path  (proof) edge                (Coq);
+ \node               (peter) [above of=GF,node distance=3cm]  {[Ljunglöf 2011]};
+ \path  (peter) edge                (GF);
+ \node               (cooper) [above of=txt,node distance=2cm]  {[Cooper \textit{et al.} 1996]};
+ \path  (cooper) edge                (txt);
 
 \end{tikzpicture}
+    \medskip
 \item
